### 2022-12-12 report.Rmd
initial analysis and report

---
title: "Mapping out the punishment strategy space"
author: "Scott Claessens"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
library(targets)
library(tidyverse)
library(knitr)
opts_chunk$set(echo = FALSE)
```

This document outlines the initial data exploration and analyses for Study 1 of our project "Mapping out the punishment strategy space".

# Sample

After cleaning the data, we have data for `r nrow(tar_read(d1))` participants from Prolific. This sample is representative of the United Kingdom.

```{r sample, out.width="65%", fig.align='center'}
tar_read(plotSample)
```

# Punishment games

We asked participants to respond to six games where they had the opportunity to punish another player for their behaviour. We refer to these games as follows:

- No Disadvantageous Inequity 1
- No Disadvantageous Inequity 2
- No Disadvantageous Inequity 3 (Computer)
- No Disadvantageous Inequity 4 (1:1 Fee Fine Ratio)
- Disadvantageous Inequity
- Third-Party

In each game, participants could punish (1) when the other player chose to "take" and (2) when the other player did nothing. For more details about these games (e.g. exact payoff structures), see preregistration.

## Comprehension

Answers to the comprehension questions revealed that participants were able to understand the payoff structure of all six punishment games.

```{r comp}
kable(tar_read(tableComp), booktabs = TRUE)
```

## Punishment decisions

We can plot the proportion of participants who decided to punish in each game.

```{r decisions, out.width="75%", fig.align='center'}
tar_read(plotDecisions)
```

Participants appear more likely to punish if the other player took, compared to when they did nothing. Participants were most likely to punish when the other player took in the disadvantageous inequity game and in the third-party game.

## Reasons given for punishing in the games

At the end of the survey, we asked participants why they decided to punish (if they ever did). First, we allowed them to provide an open-ended answer to this question. The following wordcloud summarises frequently used words in these open-ended answers.

```{r wordcloud, echo=FALSE, out.width="50%", fig.align='center'}
include_graphics("figures/study1/wordcloud.png")
```

Second, we gave participants a number of sliders on which they could rate how strongly they followed different approaches to the games. Below are the raw distributions from these slider scales (0 = statement does not apply, 1000 = statement does apply).

```{r sliders1, fig.height=3, warning=FALSE, error=FALSE}
tar_read(plotSliders1)
```

We can also visualise these distributions as deviations from participant's average ratings across all sliders.

```{r sliders2, fig.height=3, warning=FALSE, error=FALSE}
tar_read(plotSliders2)
```

Participants reported being especially motivated by equality, avoiding disadvantageous inequity, and seeking advantageous inequity. People also expressed that they never punished.

# Frequencies of punishment strategies

Before data collection, we posited ten different strategies that might underlie people's punishment behaviour in the games:

- Competitive
- Avoid disadvantageous inequity
- Egalitarian
- Seek advantageous inequity
- Retributive
- Deterrent
- Norm-enforcing
- Antisocial
- Random choice
- Anti-punish

We pre-registered predictions for how these strategies would behave in the different games.

## Counts from raw data

As a first step, we can look to see how many participants fitted these strategy predictions _exactly_ across all games, by simply counting the raw data.

```{r strategyCounts}
kable(tar_read(tableStrategyCounts), booktabs = TRUE)
```

Many participants, denoted by N/A, were unable to be classified into a strategy (i.e. their pattern of behaviour across all games did not fit any of our strategy predictions). However, of the participants who could be classified, most followed the "anti-punish" strategy by never punishing. The next most common strategies were the "egalitarian" and "avoid disadvantageous inequity" strategies.

We can also look at the most common behavioural patterns across all six games (twelve punishment decisions in total). Below, we summarise the five most common patterns of behaviour as strings of twelve 0s and 1s, indicating whether participants did (1) or did not (0) punish in each game, with a brief explanation of this pattern and its frequency in the dataset.

```{r patterns}
kable(tar_read(tablePatterns), booktabs = TRUE)
```

## Bayesian modelling

We can also estimate the frequencies of different strategies using a Bayesian approach. We construct a model that contains our _apriori_ predictions for the different punishment strategies, and the feed the model our raw data to estimate the relative probabilities of following each strategy. In the model, we assume that participants sometimes make errors in converting their strategy into behaviour (5% error rate), which could explain why many of the participants were unable to be classified into a strategy type in our raw counts above. The Bayesian model is fitted in the probabilistic programming language Stan.

```{r post probs}
P <- tar_read(post1)$alpha
for (i in 1:nrow(P)) P[i,] <- rethinking::softmax(P[i,])
anti <- P[,10]
```

In line with our raw counts, the model predicts that participants follow the "anti-punish" strategy with a probability of `r round(median(anti), 2)`, 95% credible interval [`r round(quantile(anti, 0.025), 2)` `r round(quantile(anti, 0.975), 2)`]. We plot the posterior probabilities for the remaining punishment strategies below.

```{r model1, message=FALSE, out.width="80%", fig.align='center'}
tar_read(plot1)
```

Taking advantage of all available data, the model suggests that "egalitarian" is the most common punishment strategy. All other strategies have median posterior probabilities less than 10%. In order of decreasing probability, the next most common punishment strategies are "norm-enforcing", "seek advantageous inequity", "deterrent", and "avoid disadvantageous inequity", although all are similarly likely. "Competitive" and "antisocial" punishment strategies are the most unlikely.

# Predicting strategy usage

We then fit a series of models predicting the usage of different strategies from a variety of predictors. In each plot that follows, we show the posterior predictions from a model that includes a single variable as a predictor of all ten strategies simultaneously. The plots show the median regression lines with shaded 95% credible intervals. The model-estimated slopes for each strategy are included in each panel.

```{r plot2, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
tar_read(plot2_Age)
tar_read(plot2_SelfRate_1)
tar_read(plot2_SelfRate_2)
tar_read(plot2_SelfRate_3)
tar_read(plot2_SelfRate_4)
tar_read(plot2_SelfRate_5)
tar_read(plot2_SelfRate_6)
tar_read(plot2_SelfRate_7)
tar_read(plot2_SelfRate_8)
tar_read(plot2_SelfRate_9)
tar_read(plot2_SelfRate_10)
tar_read(plot2_SelfRate_11)
```

### 2022-12-12 R/functions.R

@@ -1,302 +1,0 @@
-# custom functions
-
-# simulate data
-simulateData <- function(n = 500, seed = 2113, errorRate = 0, alphas, betas) {
-  # set simulation seed
-  set.seed(seed)
-  # n = number of participants
-  # each participant has a score on an individual difference measure: x
-  x <- rnorm(n)
-  # assume that each participant follows a different punishment strategy
-  # 01. Competitive
-  # 02. Avoid disadvantageous inequality
-  # 03. Egalitarian
-  # 04. Seek advantageous inequality
-  # 05. Retributive
-  # 06. Deterrent
-  # 07. Norm-enforcing
-  # 08. Exclusively antisocial punishment
-  # 09. Random choice
-  # 10. Never punish
-  # the punishment strategy is predicted by x
-  strategies <- c()
-  for (i in 1:n) {
-    # predict probabilities with x
-    probs <- softmax(alphas + betas*x[i])
-    strategies[i] <- sample(1:10, size = 1, replace = TRUE, prob = probs)
-  }
-  # each participant plays a series of games developed to tease apart strategies
-  # we measure punishment behaviours in these games
-  # 0 = no punish, 1 = punish
-  # create function to simulate punishment behaviour based on strategy and error rate
-  simulateBehaviour <- function(punStrats = c()) {
-    out <- c()
-    for (i in 1:n) {
-      if (strategies[i] %in% punStrats) {
-        out[i] <- rbinom(1, 1, prob = 1 - errorRate) # punish (with error)
-      } else if (strategies[i] == 9) {
-        out[i] <- rbinom(1, 1, prob = 0.5)           # random choice
-      } else if (!(strategies[i] %in% punStrats)) {
-        out[i] <- rbinom(1, 1, prob = 0 + errorRate) # don't punish (with error)
-      }
-      # small chance of failed comprehension and missing data
-      out[i] <- ifelse(rbinom(1, 1, prob = 0.01) == 0, out[i], -999)
-    }
-    return(out)
-  }
-  # game 1 - No DI 1
-  pun1_1 <- simulateBehaviour(punStrats = c(1, 5, 6, 7))
-  pun1_2 <- simulateBehaviour(punStrats = c(1, 8))
[…]

### 2022-12-12 R/functionsModelFitting.R

@@ -1,0 +1,85 @@
+# custom functions
+
+# fit stan model without predictor
+fitModel1 <- function(d, compiledModel1, error = 0) {
+  # list for stan
+  dataList <-
+    list(
+      N = nrow(d),
+      pun1_1 = d$pun1_1,
+      pun1_2 = d$pun1_2,
+      pun2_1 = d$pun2_1,
+      pun2_2 = d$pun2_2,
+      pun3_1 = d$pun3_1,
+      pun3_2 = d$pun3_2,
+      pun4_1 = d$pun4_1,
+      pun4_2 = d$pun4_2,
+      pun5_1 = d$pun5_1,
+      pun5_2 = d$pun5_2,
+      pun6_1 = d$pun6_1,
+      pun6_2 = d$pun6_2,
+      error = error # assumed error rate
+    )
+  # fit model
+  out <- sampling(compiledModel1, data = dataList, cores = 4, seed = 2113)
+  return(out)
+}
+
+# fit stan model with predictor
+fitModel2 <- function(d, compiledModel2, predictor, error = 0) {
+  # listwise deletion for predictor
+  d <- d[as.vector(!is.na(d[,predictor])),]
+  # list for stan
+  dataList <-
+    list(
+      N = nrow(d),
+      pred = d %>% pull(predictor),
+      pun1_1 = d$pun1_1,
+      pun1_2 = d$pun1_2,
+      pun2_1 = d$pun2_1,
+      pun2_2 = d$pun2_2,
+      pun3_1 = d$pun3_1,
+      pun3_2 = d$pun3_2,
+      pun4_1 = d$pun4_1,
+      pun4_2 = d$pun4_2,
+      pun5_1 = d$pun5_1,
+      pun5_2 = d$pun5_2,
+      pun6_1 = d$pun6_1,
+      pun6_2 = d$pun6_2,
+      error = error # assumed error rate
+    )
+  # fit model
+  out <- sampling(compiledModel2, data = dataList, cores = 4, seed = 2113)
+  return(out)
+}
+
+# fit stan model with categorical predictor
+fitModel3 <- function(d, compiledModel3, predictor, error = 0) {
+  # listwise deletion for predictor
+  d <- d[as.vector(!is.na(d[,predictor])),]
+  # categorical predictor as numeric
+  d[[predictor]] <- as.numeric(d[[predictor]])
+  # list for stan
+  dataList <-
+    list(
[…]

### 2022-12-12 R/functionsSimulation.R

@@ -1,0 +1,136 @@
+# custom functions
+
+# simulate data
+simulateData <- function(n = 500, seed = 2113, errorRate = 0, alphas, betas) {
+  # set simulation seed
+  set.seed(seed)
+  # n = number of participants
+  # each participant has a score on an individual difference measure: x
+  x <- rnorm(n)
+  # assume that each participant follows a different punishment strategy
+  # 01. Competitive
+  # 02. Avoid disadvantageous inequality
+  # 03. Egalitarian
+  # 04. Seek advantageous inequality
+  # 05. Retributive
+  # 06. Deterrent
+  # 07. Norm-enforcing
+  # 08. Exclusively antisocial punishment
+  # 09. Random choice
+  # 10. Never punish
+  # the punishment strategy is predicted by x
+  strategies <- c()
+  for (i in 1:n) {
+    # predict probabilities with x
+    probs <- softmax(alphas + betas*x[i])
+    strategies[i] <- sample(1:10, size = 1, replace = TRUE, prob = probs)
+  }
+  # each participant plays a series of games developed to tease apart strategies
+  # we measure punishment behaviours in these games
+  # 0 = no punish, 1 = punish
+  # create function to simulate punishment behaviour based on strategy and error rate
+  simulateBehaviour <- function(punStrats = c()) {
+    out <- c()
+    for (i in 1:n) {
+      if (strategies[i] %in% punStrats) {
+        out[i] <- rbinom(1, 1, prob = 1 - errorRate) # punish (with error)
+      } else if (strategies[i] == 9) {
+        out[i] <- rbinom(1, 1, prob = 0.5)           # random choice
+      } else if (!(strategies[i] %in% punStrats)) {
+        out[i] <- rbinom(1, 1, prob = 0 + errorRate) # don't punish (with error)
+      }
+      # small chance of failed comprehension and missing data
+      out[i] <- ifelse(rbinom(1, 1, prob = 0.01) == 0, out[i], -999)
+    }
+    return(out)
+  }
+  # game 1 - No DI 1
+  pun1_1 <- simulateBehaviour(punStrats = c(1, 5, 6, 7))
+  pun1_2 <- simulateBehaviour(punStrats = c(1, 8))
+  # game 2 - No DI 2
[…]

### 2022-12-12 R/functionsStudy1.R

@@ -1,0 +1,466 @@
+# custom functions
+
+# load study 1 data
+loadData1 <- function(fileStudy1) {
+  # load data
+  read_csv(fileStudy1) %>%
+    # create variables
+    mutate(
+      # standardised variables for prediction
+      AgeStd = as.numeric(scale(Age)),
+      SelfRate1Std = as.numeric(scale(SelfRate_1)),
+      SelfRate2Std = as.numeric(scale(SelfRate_2)),
+      SelfRate3Std = as.numeric(scale(SelfRate_3)),
+      SelfRate4Std = as.numeric(scale(SelfRate_4)),
+      SelfRate5Std = as.numeric(scale(SelfRate_5)),
+      SelfRate6Std = as.numeric(scale(SelfRate_6)),
+      SelfRate7Std = as.numeric(scale(SelfRate_7)),
+      SelfRate8Std = as.numeric(scale(SelfRate_8)),
+      SelfRate9Std = as.numeric(scale(SelfRate_9)),
+      SelfRate10Std = as.numeric(scale(SelfRate_10)),
+      SelfRate11Std = as.numeric(scale(SelfRate_11)),
+      # numeric punishment vars
+      pun1_1 = ifelse(str_starts(NoDI1_Take,    "Yes"), 1, 0),
+      pun1_2 = ifelse(str_starts(NoDI1_Nothing, "Yes"), 1, 0),
+      pun2_1 = ifelse(str_starts(NoDI2_Take,    "Yes"), 1, 0),
+      pun2_2 = ifelse(str_starts(NoDI2_Nothing, "Yes"), 1, 0),
+      pun3_1 = ifelse(str_starts(NoDI3_Take,    "Yes"), 1, 0),
+      pun3_2 = ifelse(str_starts(NoDI3_Nothing, "Yes"), 1, 0),
+      pun4_1 = ifelse(str_starts(NoDI4_Take,    "Yes"), 1, 0),
+      pun4_2 = ifelse(str_starts(NoDI4_Nothing, "Yes"), 1, 0),
+      pun5_1 = ifelse(str_starts(DI_Take,       "Yes"), 1, 0),
+      pun5_2 = ifelse(str_starts(DI_Nothing,    "Yes"), 1, 0),
+      pun6_1 = ifelse(str_starts(`3PP_Take`,    "Yes"), 1, 0),
[…]

### 2022-12-12 _targets.R

@@ -1,29 +1,33 @@
 library(targets)
 library(tarchetypes)
-library(tidyverse)
-source("R/functions.R")
 options(tidyverse.quiet = TRUE)
-tar_option_set(packages = c("cowplot", "ggridges", "rethinking", 
-                            "rstan", "tidybayes", "tidytext",
-                            "wordcloud"))
+library(tidyverse)
+source("R/functionsModelFitting.R")
+source("R/functionsSimulation.R")
+source("R/functionsStudy1.R")
+tar_option_set(packages = c("cowplot", "ggridges", "rethinking", "rstan",
+                            "tidybayes", "tidytext", "tidyverse", "wordcloud"),
+               deployment = "main")
+options(
+  clustermq.scheduler = "slurm", 
+  clustermq.template = "slurm_clustermq.tmpl"
+)
 
 # full workflow
 list(
   
-  ################################
-  # Load and compile stan models #
-  ################################
+  #### Load and compile Stan models ####
   
   # model files
   tar_target(fileModel1, "stan/punishStrat_noPredictor.stan", format = "file"),
   tar_target(fileModel2, "stan/punishStrat_withPredictor.stan", format = "file"),
+  tar_target(fileModel3, "stan/punishStrat_withCategoricalPredictor.stan", format = "file"),
   # compile models
-  tar_target(compiledModel1, stan_model(fileModel1)),
-  tar_target(compiledModel2, stan_model(fileModel2)),
+  tar_target(compiledModel1, stan_model(fileModel1), deployment = "worker"),
+  tar_target(compiledModel2, stan_model(fileModel2), deployment = "worker"),
+  tar_target(compiledModel3, stan_model(fileModel3), deployment = "worker"),
   
-  ##############
-  # Simulation #
-  ##############
+  #### Simulation ####
   
   # simulate data
   tar_target(simAlphas, c(0, -0.2, 1.5, -1, -1, -0.5, -0.1, -1, -0.5, 2)),
@@ -31,22 +35,82 @@ list(
   tar_target(dSim, simulateData(n = 50, errorRate = 0, 
                                 alphas = simAlphas, betas = simBetas)),
   # fit model to simulated data
-  tar_target(fitSimModel, fitModel2(dSim, compiledModel2, predictor = "x")),
+  tar_target(fitSimModel, fitModel2(dSim, compiledModel2, predictor = "x"),
+             deployment = "worker"),
[…]

### 2022-12-12 run.R

@@ -1,0 +1,1 @@
+targets::tar_make_clustermq(workers = 13L)

### 2023-07-04 report.Rmd
add study 2 data

@@
 
+\newpage
+
 # Predicting strategy usage
@@
 The plots show the median regression lines with shaded 95% credible intervals.
-The model-estimated slopes for each strategy are included in each panel.
+For continuous predictors, the model-estimated slopes for each strategy are included in each panel.
 
-```{r plot2, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+## Demographics
+
+We find that several demographic factors predict usage of the "anti-punish" strategy.
+Specifically, people who are white, male, higher educated (PhD-level), and students are more likely to follow the anti-punish strategy.
+Age and socio-economic status are unrelated to all strategies.
+
+```{r plotPredDemographics, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
 tar_read(plot2_Age)
-tar_read(plot2_SelfRate_1)
-tar_read(plot2_SelfRate_2)
-tar_read(plot2_SelfRate_3)
-tar_read(plot2_SelfRate_4)
-tar_read(plot2_SelfRate_5)
-tar_read(plot2_SelfRate_6)
-tar_read(plot2_SelfRate_7)
-tar_read(plot2_SelfRate_8)
-tar_read(plot2_SelfRate_9)
-tar_read(plot2_SelfRate_10)
-tar_read(plot2_SelfRate_11)
+tar_read(plot3_Education)
+tar_read(plot3_Ethnicity)
+tar_read(plot3_Gender)
+tar_read(plot2_SES)
+tar_read(plot3_Student)
 ```
+
+\newpage
+
+## Self-ratings
+
+Answers to the self-rating sliders were related to strategy usage in predictable ways.
+In general, citing any reason to punish tended to be negatively related to usage of the anti-punish strategy.
+People citing egalitarian, harm-reduction, or disapproval-signalling motivations were more likely to follow egalitarian and/or norm enforcing strategies.
+In addition:
+
+- people who wanted to _have a higher final bonus than others_ or _avoid having a lower final bonus than others_ were more likely to follow the antisocial strategy
+- people who admitted that they _decided at random_ were more likely to follow the random choice strategy
+- people who _punishing those who did NOT cause harm_ were less likely to follow the egalitarian strategy and more likely to follow the avoid DI and random choice strategies
+
+```{r plotPredSelfRate, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate1)
+tar_read(plot2_SelfRate2)
+tar_read(plot2_SelfRate3)
+tar_read(plot2_SelfRate4)
+tar_read(plot2_SelfRate5)
+tar_read(plot2_SelfRate6)
+tar_read(plot2_SelfRate7)
+tar_read(plot2_SelfRate8)
+tar_read(plot2_SelfRate9)
+tar_read(plot2_SelfRate10)
+tar_read(plot2_SelfRate11)
+```
+
+\newpage
+
+# Big-6 personality
+
+Personality was largely unrelated to strategy usage.
+The only slopes with 95% credible intervals that excluded zero were for extraversion and honesty-humility.
+Extraversion negatively predicted usage of the anti-punish strategy, while honesty-humility positively predicted usage of the anti-punish strategy.
+
+```{r plotPredPersonality, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_Agree)
+tar_read(plot2_Consc)
+tar_read(plot2_Extra)
+tar_read(plot2_Honest)
+tar_read(plot2_Neur)
+tar_read(plot2_Open)
+```
+
+\newpage
+
+# Social value orientation
+
+We found that social value orientation angle positively predicted usage of the egalitarian and anti-punish strategies.
+
+```{r plotPredSVO, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SVOangle)
+```
+
+\newpage
+
+# Political views
+
+Political ideology was largely unrelated to strategy usage.
+Unidimensional political ideology was unrelated to all strategies, while SDO and RWA negatively predicted usage of the anti-punish strategy.
+Wanting to bring people up or down a peg on the SES ladder was also unrelated to strategy usage.
+
+```{r plotPredPolIdeology, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_PolSlider)
+tar_read(plot2_SDO)
+tar_read(plot2_RWA)
+tar_read(plot2_BringDown)
+tar_read(plot2_BringUp)
+```
+
+\newpage
+
+# Religion
+
+Both religiosity items were unrelated to strategy usage.
+
+```{r plotPredReligion, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_Religiosity)
+tar_read(plot2_GodC)
+```

### 2023-07-04 R/functionsModelFitting.R

@@ -1,7 +1,8 @@
 # custom functions
 
 # fit stan model without predictor
-fitModel1 <- function(d, compiledModel1, error = 0) {
+fitModel1 <- function(d, compiledModel1, error = 0, 
+                      iter = 2000,  warmup = 1000, chains = 4, cores = 4) {
   # list for stan
   dataList <-
     list(
@@ -21,12 +22,22 @@ fitModel1 <- function(d, compiledModel1, error = 0) {
       error = error # assumed error rate
     )
   # fit model
-  out <- sampling(compiledModel1, data = dataList, cores = 4, seed = 2113)
+  out <- 
+    sampling(
+      compiledModel1, 
+      data = dataList, 
+      iter = iter,
+      warmup = warmup,
+      chains = chains,
+      cores = cores,
+      seed = 2113
+      )
   return(out)
 }
 
 # fit stan model with predictor
-fitModel2 <- function(d, compiledModel2, predictor, error = 0) {
+fitModel2 <- function(d, compiledModel2, predictor, error = 0, 
+                      iter = 2000, warmup = 1000, chains = 4, cores = 4) {
   # listwise deletion for predictor
   d <- d[as.vector(!is.na(d[,predictor])),]
   # list for stan
@@ -50,12 +61,22 @@ fitModel2 <- function(d, compiledModel2, predictor, error = 0) {
       error = error # assumed error rate
     )
   # fit model
-  out <- sampling(compiledModel2, data = dataList, cores = 4, seed = 2113)
+  out <- 
+    sampling(
+      compiledModel2,
+      data = dataList,
+      iter = iter,
+      warmup = warmup,
+      chains = chains,
+      cores = cores,
+      seed = 2113
+      )
   return(out)
 }
 
 # fit stan model with categorical predictor
-fitModel3 <- function(d, compiledModel3, predictor, error = 0) {
+fitModel3 <- function(d, compiledModel3, predictor, error = 0,
+                      iter = 2000, warmup = 1000, chains = 4, cores = 4) {
   # listwise deletion for predictor
   d <- d[as.vector(!is.na(d[,predictor])),]
   # categorical predictor as numeric
@@ -81,35 +102,15 @@ fitModel3 <- function(d, compiledModel3, predictor, error = 0) {
       error = error # assumed error rate
     )
[…]

### 2023-07-04 R/functionsPowerAnalysis.R

@@ -1,0 +1,49 @@
+# custom functions
+
+# fit power model and return coefficient on x
+fitPowerModel <- function(dPower, compiledModel2, strategy) {
+  # fit model
+  m <- 
+    fitModel2(
+      dPower,
+      compiledModel2,
+      predictor = "x",
+      iter = 1000,
+      warmup = 500,
+      chains = 1,
+      cores = 1
+      )
+  # extract posterior slope of x for particular strategy
+  post <- rstan::extract(m)$beta[,strategy]
+  # return posterior summary
+  out <- 
+    tibble(
+      Estimate = median(post),
+      `Q2.5` = quantile(post, 0.025),
+      `Q97.5` = quantile(post, 0.975)
+    )
+  return(out)
+}
+
+# plot power analysis results
+plotPowerAnalysis <- function(power) {
+  # strategy vectors
+  strategies <- c("Competitive", "Avoid DI", "Egalitarian", "Seek AI",
+                  "Retributive", "Deterrent", "Norm-enforcing", 
+                  "Antisocial", "Random choice", "Anti-punish")
+  # plot
+  out <-
+    tibble(
+      Strategy = factor(strategies, levels = strategies),
+      Power = as.numeric(power)
+    ) %>%
+    ggplot(aes(x = Strategy, y = Power)) +
+    geom_point(size = 2) +
+    geom_hline(yintercept = 0.8, linetype = "dashed") +
+    lims(y = 0:1) +
+    labs(subtitle = "Power to detect medium effect size (log odds slope = 1.00)") +
+    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
+  # save
+  ggsave(out, filename = "figures/powerAnalysis/powerAnalysis.pdf", width = 5, height = 4)
+  return(out)
+}

### 2023-07-04 R/functionsStudy1.R

@@ -358,6 +358,65 @@ loadData1 <- function(fileStudy1) {
       )
 }
 
+# data exclusions from study 1
+excludeData1 <- function(d1) {
+  # labels for rwa scale
+  rwaLabels <- c(
+    "Very strongly disagree"     = 1,
+    "Strongly disagree"          = 2,
+    "Somewhat disagree"          = 3,
+    "Slightly disagree"          = 4,
+    "Neither agree nor disagree" = 5,
+    "Slightly agree"             = 6,
+    "Somewhat agree"             = 7,
+    "Strongly agree"             = 8,
+    "Very strongly agree"        = 9
+  )
+  # exclusions
+  # n = 1019
+  d1 %>%
+    # exclude participants who sped through surveys (criteria = 2 SDs below median duration)
+    filter(DurationSeconds_Part1 > median(DurationSeconds_Part1, na.rm = TRUE) - sd(DurationSeconds_Part1, na.rm = TRUE)*2) %>%
+    filter(ifelse(is.na(DurationSeconds_Part2), TRUE, 
+                  DurationSeconds_Part2 > median(DurationSeconds_Part2, na.rm = TRUE) - sd(DurationSeconds_Part2, na.rm = TRUE)*2)) %>%
+    # n = 1019
+    # exclude participants who failed "days of week" attention check
+    filter(is.na(DaysWeek_TryAgain) | DaysWeek_TryAgain == 7) %>%
+    # n = 1019
+    # exclude participants who failed "weeks in year" attention check
+    filter(is.na(WeeksYear_TryAgain) | WeeksYear_TryAgain == 52) %>%
+    # n = 1019
+    # exclude participants who provide nonsensical answer to breakfast question
+    slice(c(-21,-279,-335,-872,-912)) %>%
+    # n = 1014
+    # exclude participants who flatline (criteria = two or more matrix tables)
+    rowwise() %>%
+    mutate(
+      flatSDO   = sd(parse_number(c_across(starts_with("SDO_"))), na.rm = TRUE),
+      flatRWA   = sd(rwaLabels[c_across(starts_with("RWA_"))], na.rm = TRUE),
+      flatPers  = sd(parse_number(c_across(starts_with("Pers_"))), na.rm = TRUE),
+      flatTotal = sum(c(flatSDO == 0, flatRWA == 0, flatPers == 0), na.rm = TRUE)
+      ) %>%
+    ungroup() %>%
+    filter(!(flatTotal >= 2)) %>%
+    # n = 1014
[…]

### 2023-07-04 _targets.R

@@ -5,6 +5,7 @@ library(tidyverse)
 source("R/functionsModelFitting.R")
 source("R/functionsSimulation.R")
 source("R/functionsStudy1.R")
+source("R/functionsPowerAnalysis.R")
 tar_option_set(
   packages = c("cowplot", "ggcorrplot", "ggridges", "ltm", 
                "rethinking", "rstan", "tidybayes", "tidytext", 
@@ -16,6 +17,28 @@ options(
   clustermq.template = "slurm_clustermq.tmpl"
 )
 
+# power analysis static branching (see below)
+powerMap <-
+  tar_map(
+    values = tibble(strategy = 1:10),
+    # seeds for data simulation
+    tar_target(powerSeed, 1:100),
+    # simulate data for power analysis
+    tar_target(dPower, simulateData(n = 1019, errorRate = 0, seed = powerSeed,
+                                    # use intercepts from m1 model
+                                    alphas = c(-1.48, -0.31, 1.43, 0.22, -0.39,
+                                               -0.22, 0.22, -0.87, -0.47, 2.31),
+                                    # detecting medium effect for one strategy
+                                    # no effect for all other strategies
+                                    betas = c(rep(0, times = strategy - 1), 1, rep(0, times = 10 - strategy))),
+               pattern = map(powerSeed)),
+    # fit model to simulated data and return posterior summary
+    tar_target(powerModel, fitPowerModel(dPower, compiledModel2, strategy),
+               deployment = "worker", pattern = map(dPower)),
+    # power
+    tar_target(power, mean(powerModel$`Q2.5` > 0))
+  )
+
 # full workflow
 list(
   
@@ -25,13 +48,11 @@ list(
   tar_target(fileModel1, "stan/punishStrat_noPredictor.stan", format = "file"),
   tar_target(fileModel2, "stan/punishStrat_withPredictor.stan", format = "file"),
   tar_target(fileModel3, "stan/punishStrat_withCategoricalPredictor.stan", format = "file"),
-  tar_target(fileModel4, "stan/punishStrat_withOrdinalPredictor.stan", format = "file"),
   # compile models
   tar_target(compiledModel1, stan_model(fileModel1), deployment = "worker"),
   tar_target(compiledModel2, stan_model(fileModel2), deployment = "worker"),
   tar_target(compiledModel3, stan_model(fileModel3), deployment = "worker"),
[…]

### 2023-07-04 run.R

@@ -1,1 +1,1 @@
-targets::tar_make_clustermq(workers = 13L)
+targets::tar_make_clustermq(starts_with("power"), workers = 50L)

### 2023-07-05 _targets.R
finished study 2 data cleaning

@@ -25,7 +25,7 @@ powerMap <-
     tar_target(powerSeed, 1:100),
     # simulate data for power analysis
     tar_target(dPower, simulateData(n = 1019, errorRate = 0, seed = powerSeed,
-                                    # use intercepts from m1 model
+                                    # use intercepts from m1 model (UK only)
                                     alphas = c(-1.48, -0.31, 1.43, 0.22, -0.39,
                                                -0.22, 0.22, -0.87, -0.47, 2.31),
                                     # detecting medium effect for one strategy

### 2023-07-05 data/study1/study1_raw/study1_cleanData.R

@@ -1,0 +1,164 @@
+library(lubridate)
+library(qualtRics)
+library(tidyverse)
+
+#### Source this file to produce clean dataset and metadata for Study 1 (UK)
+
+#### Part 1 - Games ####
+
+# load completes
+part1completes <-
+  qualtRics::read_survey("study1_part1_games/Qualtrics_complete_responses_PUN_BATTERY_GAMES_November+30,+2022_15.32.csv",
+                         time_zone = "UTC") %>%
+  # remove one testing row
+  filter(!is.na(PROLIFIC_PID)) %>%
+  # remove one duplicated prolific ID (keep first submission)
+  filter(ResponseId != "R_31jE0BAgDjMSW0U")
+
+# load incompletes
+part1incompletes <-
+  qualtRics::read_survey("study1_part1_games/incomplete_responses_PUN_BATTERY_GAMES_November+29,+2022_12.24.csv",
+                         col_types = readr::cols(StartDate = readr::col_character())) %>%
+  # keep only participants that completed all the games
+  filter(!is.na(Explain)) %>%
+  # remove one duplicated prolific ID (keep first submission)
+  filter(ResponseId != "FS_1pW4oBKeoR1vtlV") %>%
+  # times as UTC
+  mutate(
+    StartDate = as.POSIXct(StartDate, tz = "UTC", format = "%d/%m/%Y %H:%M"),
+    EndDate   = as.POSIXct(EndDate,   tz = "UTC", format = "%d/%m/%Y %H:%M")
+  )
+
+# load prolific demographics
+demographics <-
+  read_csv("study1_part1_games/Demographics.csv") %>%
+  # introduce NAs for data expiry or revoked consent
+  mutate_all(function(x) ifelse(x %in% c("DATA_EXPIRED", "CONSENT_REVOKED"), NA, x)) %>%
+  # rename and keep only relevant columns
+  transmute(
+    PROLIFIC_PID = `Participant id`,
+    ProlificApprovals = `Total approvals`,
+    FluentLanguages = `Fluent languages`,
+    Age = as.numeric(Age),
+    Gender = Sex,
+    Ethnicity = `Ethnicity simplified`,
+    CountryBirth = `Country of birth`,
+    CountryResidence = `Country of residence`,
[…]

### 2023-07-05 data/study2/study2_raw/study2_cleanData.R

@@ -1,0 +1,122 @@
+library(lubridate)
+library(qualtRics)
+library(tidyverse)
+
+#### Source this file to produce clean dataset and metadata for Study 2 (US)
+
+#### Part 1 - Games ####
+
+part1 <-
+  qualtRics::read_survey("study2_part1_games/PUN_BATTERY_GAMES+-+US_June+27,+2023_10.37.csv",
+                         time_zone = "UTC") %>%
+  # remove testing rows and rows where Prolific IDs are missing
+  filter(!is.na(PROLIFIC_PID) & !is.na(prolific_id)) %>%
+  # remove rows with duplicated prolific IDs (keep more complete submission)
+  filter(!(ResponseId %in% c("R_1hzh4TPukLMGgJc", "R_bJzUSWXdJJj8VTX", "R_BG0PQUuvS5VV1T3",
+                             "R_2aQq2FpvJ7Z5NNq", "R_32PGvi2ktzIDEWM", "R_2SeipWu3jgLc0y6",
+                             "R_300O7jzueLXV1lC", "R_1memDQb8WFYYljC", "R_XTVA5YYKncUxcDD",
+                             "R_2Sw6awU7yj7VSlL", "R_28UO5K5U20rDSiE", "R_PI188DrMDAxaMnL"))) %>%
+  # keep only participants that completed all the games
+  filter(!is.na(Explain))
+
+
+# load prolific demographics (part 1)
+demographics <-
+  read_csv("study2_part1_games/Demographic data_games.csv") %>%
+  # introduce NAs for data expiry or revoked consent
+  mutate_all(function(x) ifelse(x %in% c("DATA_EXPIRED", "CONSENT_REVOKED"), NA, x)) %>%
+  # rename and keep only relevant columns
+  transmute(
+    PROLIFIC_PID = `Participant id`,
+    ProlificApprovals = `Total approvals`,
+    FluentLanguages = `Fluent languages`,
+    Age = as.numeric(Age),
+    Gender = Sex,
+    Ethnicity = `Ethnicity simplified`,
+    CountryBirth = `Country of birth`,
+    CountryResidence = `Country of residence`,
+    Nationality = Nationality,
+    Language = Language,
+    Student = `Student status`,
+    Employment = `Employment status`
+  )
+
+# combine the part 1 datasets
+part1 <-
+  part1 %>%
+  # link demographic data
+  left_join(demographics, by = "PROLIFIC_PID") %>%
[…]

### 2023-07-05 data/study1/study1_raw/study1_cleanData.R
clean up ignored files

@@ -1,164 +1,0 @@
-library(lubridate)
-library(qualtRics)
-library(tidyverse)
-
-#### Source this file to produce clean dataset and metadata for Study 1 (UK)
-
-#### Part 1 - Games ####
-
-# load completes
-part1completes <-
-  qualtRics::read_survey("study1_part1_games/Qualtrics_complete_responses_PUN_BATTERY_GAMES_November+30,+2022_15.32.csv",
-                         time_zone = "UTC") %>%
-  # remove one testing row
-  filter(!is.na(PROLIFIC_PID)) %>%
-  # remove one duplicated prolific ID (keep first submission)
-  filter(ResponseId != "R_31jE0BAgDjMSW0U")
-
-# load incompletes
-part1incompletes <-
-  qualtRics::read_survey("study1_part1_games/incomplete_responses_PUN_BATTERY_GAMES_November+29,+2022_12.24.csv",
-                         col_types = readr::cols(StartDate = readr::col_character())) %>%
-  # keep only participants that completed all the games
-  filter(!is.na(Explain)) %>%
-  # remove one duplicated prolific ID (keep first submission)
-  filter(ResponseId != "FS_1pW4oBKeoR1vtlV") %>%
-  # times as UTC
-  mutate(
-    StartDate = as.POSIXct(StartDate, tz = "UTC", format = "%d/%m/%Y %H:%M"),
-    EndDate   = as.POSIXct(EndDate,   tz = "UTC", format = "%d/%m/%Y %H:%M")
-  )
-
-# load prolific demographics
-demographics <-
-  read_csv("study1_part1_games/Demographics.csv") %>%
-  # introduce NAs for data expiry or revoked consent
-  mutate_all(function(x) ifelse(x %in% c("DATA_EXPIRED", "CONSENT_REVOKED"), NA, x)) %>%
-  # rename and keep only relevant columns
-  transmute(
-    PROLIFIC_PID = `Participant id`,
-    ProlificApprovals = `Total approvals`,
-    FluentLanguages = `Fluent languages`,
-    Age = as.numeric(Age),
-    Gender = Sex,
-    Ethnicity = `Ethnicity simplified`,
-    CountryBirth = `Country of birth`,
-    CountryResidence = `Country of residence`,
[…]

### 2023-07-05 data/study2/study2_raw/study2_cleanData.R

@@ -1,122 +1,0 @@
-library(lubridate)
-library(qualtRics)
-library(tidyverse)
-
-#### Source this file to produce clean dataset and metadata for Study 2 (US)
-
-#### Part 1 - Games ####
-
-part1 <-
-  qualtRics::read_survey("study2_part1_games/PUN_BATTERY_GAMES+-+US_June+27,+2023_10.37.csv",
-                         time_zone = "UTC") %>%
-  # remove testing rows and rows where Prolific IDs are missing
-  filter(!is.na(PROLIFIC_PID) & !is.na(prolific_id)) %>%
-  # remove rows with duplicated prolific IDs (keep more complete submission)
-  filter(!(ResponseId %in% c("R_1hzh4TPukLMGgJc", "R_bJzUSWXdJJj8VTX", "R_BG0PQUuvS5VV1T3",
-                             "R_2aQq2FpvJ7Z5NNq", "R_32PGvi2ktzIDEWM", "R_2SeipWu3jgLc0y6",
-                             "R_300O7jzueLXV1lC", "R_1memDQb8WFYYljC", "R_XTVA5YYKncUxcDD",
-                             "R_2Sw6awU7yj7VSlL", "R_28UO5K5U20rDSiE", "R_PI188DrMDAxaMnL"))) %>%
-  # keep only participants that completed all the games
-  filter(!is.na(Explain))
-
-
-# load prolific demographics (part 1)
-demographics <-
-  read_csv("study2_part1_games/Demographic data_games.csv") %>%
-  # introduce NAs for data expiry or revoked consent
-  mutate_all(function(x) ifelse(x %in% c("DATA_EXPIRED", "CONSENT_REVOKED"), NA, x)) %>%
-  # rename and keep only relevant columns
-  transmute(
-    PROLIFIC_PID = `Participant id`,
-    ProlificApprovals = `Total approvals`,
-    FluentLanguages = `Fluent languages`,
-    Age = as.numeric(Age),
-    Gender = Sex,
-    Ethnicity = `Ethnicity simplified`,
-    CountryBirth = `Country of birth`,
-    CountryResidence = `Country of residence`,
-    Nationality = Nationality,
-    Language = Language,
-    Student = `Student status`,
-    Employment = `Employment status`
-  )
-
-# combine the part 1 datasets
-part1 <-
-  part1 %>%
-  # link demographic data
-  left_join(demographics, by = "PROLIFIC_PID") %>%
[…]

### 2023-09-08 report.Rmd
add updated strategy counts

--- 2022-12-12 report.Rmd
@@
 output: pdf_document
+
 ---
@@
 ```{r setup, include=FALSE}
-library(targets)
-library(tidyverse)
-library(knitr)
 opts_chunk$set(echo = FALSE)
+tar_load(d)
 ```
 
-This document outlines the initial data exploration and analyses for Study 1 of our project "Mapping out the punishment strategy space".
+This document outlines the data exploration and analyses for our project "Mapping out the punishment strategy space".
 
@@
 
-After cleaning the data, we have data for `r nrow(tar_read(d1))` participants from Prolific. This sample is representative of the United Kingdom.
+After cleaning the data, we have data for `r nrow(d)` participants from Prolific.
+Participants are representative samples from both the United Kingdom (n = `r sum(d$Country == "United Kingdom")`) and the United States (n = `r sum(d$Country == "United States")`).
 
-```{r sample, out.width="65%", fig.align='center'}
-tar_read(plotSample)
+```{r sample, out.width="100%"}
+tar_read(plotSample_UK)
+tar_read(plotSample_US)
 ```
@@
 
-Answers to the comprehension questions revealed that participants were able to understand the payoff structure of all six punishment games.
+Answers to the comprehension questions revealed that participants were able to understand the payoff structure of all six punishment games. Here are the comprehension rates in both countries:
 
@@
 
-```{r decisions, out.width="75%", fig.align='center'}
+```{r decisions, out.width="100%"}
 tar_read(plotDecisions)
@@
 
+The pattern is very similar in both countries.
 Participants appear more likely to punish if the other player took, compared to when they did nothing.
@@
 First, we allowed them to provide an open-ended answer to this question.
-The following wordcloud summarises frequently used words in these open-ended answers.
+The following wordclouds summarise frequently used words in these open-ended answers.
 
-```{r wordcloud, echo=FALSE, out.width="50%", fig.align='center'}
-include_graphics("figures/study1/wordcloud.png")
+Here is the wordcloud for the United Kingdom sample:
+
+```{r wordcloudUK, echo=FALSE, out.width="50%", fig.align='center'}
+include_graphics("figures/dataSummaries/wordcloud_UK.png")
+```
+
+Here is the wordcloud for the United States sample:
+
+```{r wordcloudUS, echo=FALSE, out.width="50%", fig.align='center'}
+include_graphics("figures/dataSummaries/wordcloud_US.png")
 ```
@@
 Second, we gave participants a number of sliders on which they could rate how strongly they followed different approaches to the games.
-Below are the raw distributions from these slider scales (0 = statement does not apply, 1000 = statement does apply).
+Below are the raw distributions from these slider scales (0 = statement does not apply, 100 = statement does apply).
 
-```{r sliders1, fig.height=3, warning=FALSE, error=FALSE}
+```{r sliders1, fig.height=4, warning=FALSE, error=FALSE}
 tar_read(plotSliders1)
@@
 
-```{r sliders2, fig.height=3, warning=FALSE, error=FALSE}
+```{r sliders2, fig.height=4, warning=FALSE, error=FALSE}
 tar_read(plotSliders2)
@@
 
-Participants reported being especially motivated by equality, avoiding disadvantageous inequity, and seeking advantageous inequity. People also expressed that they never punished.
+In both countries, participants reported being especially motivated by equality, avoiding disadvantageous inequity, and seeking advantageous inequity. People also expressed that they never punished.
 
@@
 
-## Counts from raw data
+## Counts and proportions from raw data
 
-As a first step, we can look to see how many participants fitted these strategy predictions _exactly_ across all games, by simply counting the raw data.
+As a first step, we can look to see the proportion of participants who fitted these strategy predictions _exactly_ across all games.
 
@@
 
+We assume here that people are following a particular strategy exactly and are not comitting any errors or mistakes.
+However, it is possible that some participants intended to follow a particular strategy, but made a mistake on a particular decision.
+We repeat the counts and proportions in the table above, but include behavioural patterns that have one mistake from the exact strategy (note that the proportions will no longer sum to 1).
+
+```{r strategyCountsMistake1}
+kable(tar_read(tableStrategyCounts_Mistake1), booktabs = TRUE)
+```
+
+And including behavioural patterns that have one or two mistakes.
+
+```{r strategyCountsMistake2}
+kable(tar_read(tableStrategyCounts_Mistake2), booktabs = TRUE)
+```
+
+And finally, including behavioural patterns that have one, two, or three mistakes.
+
+```{r strategyCountsMistake3}
+kable(tar_read(tableStrategyCounts_Mistake3), booktabs = TRUE)
+```
+
 We can also look at the most common behavioural patterns across all six games (twelve punishment decisions in total).
-Below, we summarise the five most common patterns of behaviour as strings of twelve 0s and 1s, indicating whether participants did (1) or did not (0) punish in each game, with a brief explanation of this pattern and its frequency in the dataset.
+Below, we summarise the 25 most common patterns of behaviour as strings of twelve 0s and 1s, indicating whether participants did (1) or did not (0) punish in each game, with a brief explanation of this pattern and its frequency in the dataset.
 
-```{r patterns}
+```{r patternsUK}
 kable(tar_read(tablePatterns), booktabs = TRUE)
+```
+
+These behavioural patterns can be visualised, along with their frequency of usage, on a splits graph.
+This graph plots the distance between the difference strategies as proportional to the number of substitutions required to get from one to another.
+The top of the graph captures the less punitive strategies, and the strategies become more punitive towards the bottom of the graph.
+We only include behavioural patterns followed by at least two participants.
+
+```{r splitsGraph, fig.width=7, fig.height=7}
+tar_read(plotSplits)
 ```
@@
 We construct a model that contains our _apriori_ predictions for the different punishment strategies, and the feed the model our raw data to estimate the relative probabilities of following each strategy.
-In the model, we assume that participants sometimes make errors in converting their strategy into behaviour (5% error rate), which could explain why many of the participants were unable to be classified into a strategy type in our raw counts above.
+In the model, we assume that participants sometimes make errors in converting their strategy into behaviour (5% error rate), which could explain why many of the participants were unable to be classified into a strategy type in our raw counts and proportions above.
 The Bayesian model is fitted in the probabilistic programming language Stan.
 
-```{r post probs}
-P <- tar_read(post1)$alpha
-for (i in 1:nrow(P)) P[i,] <- rethinking::softmax(P[i,])
-anti <- P[,10]
-```
-
-In line with our raw counts, the model predicts that participants follow the "anti-punish" strategy with a probability of `r round(median(anti), 2)`, 95% credible interval [`r round(quantile(anti, 0.025), 2)` `r round(quantile(anti, 0.975), 2)`].
-We plot the posterior probabilities for the remaining punishment strategies below.
-
-```{r model1, message=FALSE, out.width="80%", fig.align='center'}
+```{r model1, message=FALSE, error=FALSE, warning=FALSE, out.width="100%"}
 tar_read(plot1)
@@
 
-Taking advantage of all available data, the model suggests that "egalitarian" is the most common punishment strategy.
+The pattern is similar in both countries.
+Taking advantage of all available data, the model suggests that "anti-punish" is the most common strategy.
+Of the punishment strategies, "egalitarian" is the most common.
 All other strategies have median posterior probabilities less than 10%.
-In order of decreasing probability, the next most common punishment strategies are "norm-enforcing", "seek advantageous inequity", "deterrent", and "avoid disadvantageous inequity", although all are similarly likely.
 "Competitive" and "antisocial" punishment strategies are the most unlikely.
+
+Why is egalitarian the winner in this model, considering that it wasn't an obvious winner in the raw counts?
+One explanation is that many participants punished only in the third-party punishment game (see table above).
+The model does not include this pattern of behaviour as an explicit strategy.
+This pattern of behaviour is one substitution away from egalitarian, but two substitutions away from avoid DI, so it upweights egalitarian accordingly.
 
@@
 We then fit a series of models predicting the usage of different strategies from a variety of predictors.
-In each plot that follows, we show the posterior predictions from a model that includes a single variable as a predictor of all ten strategies simultaneously.
+In each plot that follows, we show the posterior predictions from a model that includes a single variable as a predictor of all ten strategies and both countries simultaneously.
 The plots show the median regression lines with shaded 95% credible intervals.
-The model-estimated slopes for each strategy are included in each panel.
 
-```{r plot2, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+## Demographics
+
+Age, gender, and socio-economic status were unrelated to strategy usage.
+Higher education tended to predict the anti-punish strategy.
+In the United States, non-students and white/asian people were more likely to follow the anti-punish strategy.
+
+```{r plotPredDemographics, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
 tar_read(plot2_Age)
-tar_read(plot2_SelfRate_1)
-tar_read(plot2_SelfRate_2)
-tar_read(plot2_SelfRate_3)
-tar_read(plot2_SelfRate_4)
-tar_read(plot2_SelfRate_5)
-tar_read(plot2_SelfRate_6)
-tar_read(plot2_SelfRate_7)
-tar_read(plot2_SelfRate_8)
-tar_read(plot2_SelfRate_9)
-tar_read(plot2_SelfRate_10)
-tar_read(plot2_SelfRate_11)
+tar_read(plot3_Education)
+tar_read(plot3_Ethnicity)
+tar_read(plot3_Gender)
+tar_read(plot2_SES)
+tar_read(plot3_Student)
 ```
+
+\newpage
+
+## Self-ratings
+
+Participants in both countries who stated that they "punished those who harmed others" were more likely to follow egalitarian and norm-enforcing strategies and less likely to follow the anti-punish strategy.
+In the United States, this slider was additionally positively related to the seek advantageous inequity strategy.
+
+```{r plotPredSelfRate1, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate1)
+```
+
+Participants in both countries who stated that they wanted to "have a higher final bonus than others" were more likely to follow the seek advantageous inequity and antisocial strategies and less likely to follow the egalitarian and anti-punish strategies.
+In the United Kingdom, the slider was additionally related to the random choice strategy.
+In the United States, this slider was additionally related to the avoid disadvantageous inequity strategy.
+
+```{r plotPredSelfRate2, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate2)
+```
+
+Participants in both countries who stated that they wanted to "avoid having a lower final bonus than others" were less likely to follow the egalitarian and anti-punish strategies.
+In the United Kingdom, the slider was additionally related to the antisocial strategy.
+In the United States, this slider was additionally related to the avoid DI and seek AI strategies.
+
+```{r plotPredSelfRate3, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate3)
+```
+
+Participants in both countries who stated that they wanted "all players to have the same final bonus" were more likely to follow the egalitarian strategy.
+In the United Kingdom, the slider was additionally negatively related to the antisocial strategy and positively related to the anti-punish strategy.
+In the United States, this slider was additionally positively related to the avoid DI strategy and negatively related to the anti-punish strategy.
+
+```{r plotPredSelfRate4, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate4)
+```
+
+Participants in both countries who stated that they wanted to "stop others from cheating" were less likely to follow the anti-punish strategy.
+In the United Kingdom, the slider was additionally related to the egalitarian strategy.
+In the United States, this slider was additionally related to the random choice strategy.
+
+```{r plotPredSelfRate5, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate5)
+```
+
+Participants in both countries who stated that they wanted to "show that they disapproved of others' actions" were more likely to follow the egalitarian, seek AI, and norm-enforcing strategies and less likely to follow the anti-punish strategy.
+
+```{r plotPredSelfRate6, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate6)
+```
+
+Participants in both countries who stated that they "made decisions at random" were more likely to follow the random choice strategy and less likely to follow the egalitarian and anti-punish strategies.
+In the United Kingdom, the slider was additionally related to the avoid DI and seek AI strategies.
+In the United States, this slider was additionally related to the norm-enforcing strategy.
+
+```{r plotPredSelfRate7, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate7)
+```
+
+Participants in both countries who stated that they wanted to "punish people who DID NOT harm them or others" were more likely to follow the avoid DI and random choice strategies and less likely to follow the egalitarian and anti-punish strategies.
+In the United States, this slider was additionally negatively related to the deterrent strategy and positively related to the norm-enforcing strategy.
+
+```{r plotPredSelfRate8, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate8)
+```
+
+Participants in both countries who stated that they "didn't want to reduce anyone's bonus" were more likely to follow the anti-punish strategy and less likely to follow the seek AI strategy.
+In the United Kingdom, this slider was additionally negatively related to the antisocial strategy.
+
+```{r plotPredSelfRate9, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate9)
+```
+
+Participants in both countries who stated that they "didn't want to PAY to reduce bonus but would have if free" were less likely to follow the egalitarian strategy.
+In the United Kingdom, the slider was additionally positively related to the seek AI and antisocial strategies and negatively related to the anti-punish strategy.
+In the United States, this slider was additionally positively related to the anti-punish strategy.
+
+```{r plotPredSelfRate10, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate10)
+```
+
+Participants in both countries who stated that they wanted to "punish if players harmed them but not if they harmed others" were less likely to follow the anti-punish strategy.
+In the United States, this slider was additionally related to the random choice strategy.
+
+```{r plotPredSelfRate11, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SelfRate11)
+```
+
+\newpage
+
+## Big-6 personality
+
+In the United States, agreeableness positively predicts usage of the egalitarian strategy.
+
+```{r plotPredAgree, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_Agree)
+```
+
+Conscientiousness was unrelated to strategy usage in both countries.
+
+```{r plotPredConsc, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_Consc)
+```
+
+Extraversion negatively predicts usage of the anti-punish strategy in both countries.
+In the United States, extraversion positively predicts the avoid DI and seek AI strategies, and negatively predicts the egalitarian strategy.
+
+```{r plotPredExtra, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_Extra)
+```
+
+Honesty-humility positively predicts usage of the egalitarian and anti-punish strategies. In the United States, honesty-humility negatively predicts the random choice strategy.
+
+```{r plotPredHonest, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_Honest)
+```
+
+Neuroticism was unrelated to strategy usage in both countries.
+
+```{r plotPredNeur, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_Neur)
+```
+
+Openness to experience positively predicts usage of the anti-punish strategy in both countries. In the United States, openness positively predicts the egalitarian strategy.
+
+```{r plotPredOpen, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_Open)
+```
+
+\newpage
+
+## Social value orientation
+
+In both countries, social value orientation angle positively predicted usage of the egalitarian and anti-punish strategies.
+In the United States, SVO angle also negatively predicted the random choice strategy.
+
+```{r plotPredSVO, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SVOangle)
+```
+
+\newpage
+
+## Political views
+
+In both countries, Social Dominance Orientation was positively related to the seek AI and antisocial strategies and negatively related to the egalitarian and anti-punish strategies.
+In the United States, SDO was also positively related to the avoid DI and norm-enforcing strategies and negatively related to the retributive and deterrent strategies.
+
+```{r plotPredSDO, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_SDO)
+```
+
+In both countries, Right Wing Authoritarianism was negatively related to the egalitarian and anti-punish strategies.
+In the United States, RWA was also positively related to the avoid DI, seek AI, antisocial, and random choice strategies.
+
+```{r plotPredRWA, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_RWA)
+```
+
+In the United States, agreement with the motivation to "bring people above me down a peg or two" was negatively related to the anti-punish strategy.
+
+```{r plotPredBringDown, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_BringDown)
+```
+
+In the United States, agreement with the motivation to "bring people below me up a peg or two" was positively related to the egalitarian and anti-punish strategies.
+
+```{r plotPredBringUp, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_BringUp)
+```
+
+\newpage
+
+## Religion
+
+In the United States, the religiosity item positively predicted the antisocial and random choice strategies and negatively predicted the egalitarian and anti-punish strategies.
+
+```{r plotPredReligiosity, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_Religiosity)
+```
+
+In both countries, agreeing with the statement "God controls the events in the world" negatively predicts the anti-punish strategy.
+In the United States, this item also positively predicts the random choice strategy and negatively predicts the egalitarian strategy.
+
+```{r plotPredGodC, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
+tar_read(plot2_GodC)
+```

### 2023-09-08 R/functionsModelFitting.R

@@ -1,116 +1,0 @@
-# custom functions
-
-# fit stan model without predictor
-fitModel1 <- function(d, compiledModel1, error = 0, 
-                      iter = 2000,  warmup = 1000, chains = 4, cores = 4) {
-  # list for stan
-  dataList <-
-    list(
-      N = nrow(d),
-      pun1_1 = d$pun1_1,
-      pun1_2 = d$pun1_2,
-      pun2_1 = d$pun2_1,
-      pun2_2 = d$pun2_2,
-      pun3_1 = d$pun3_1,
-      pun3_2 = d$pun3_2,
-      pun4_1 = d$pun4_1,
-      pun4_2 = d$pun4_2,
-      pun5_1 = d$pun5_1,
-      pun5_2 = d$pun5_2,
-      pun6_1 = d$pun6_1,
-      pun6_2 = d$pun6_2,
-      error = error # assumed error rate
-    )
-  # fit model
-  out <- 
-    sampling(
-      compiledModel1, 
-      data = dataList, 
-      iter = iter,
-      warmup = warmup,
-      chains = chains,
-      cores = cores,
-      seed = 2113
-      )
-  return(out)
-}
-
-# fit stan model with predictor
-fitModel2 <- function(d, compiledModel2, predictor, error = 0, 
-                      iter = 2000, warmup = 1000, chains = 4, cores = 4) {
-  # listwise deletion for predictor
-  d <- d[as.vector(!is.na(d[,predictor])),]
-  # list for stan
-  dataList <-
-    list(
-      N = nrow(d),
-      # standardise predictor for model fitting
-      pred = d %>% pull(predictor) %>% scale() %>% as.numeric(),
-      pun1_1 = d$pun1_1,
-      pun1_2 = d$pun1_2,
-      pun2_1 = d$pun2_1,
-      pun2_2 = d$pun2_2,
-      pun3_1 = d$pun3_1,
-      pun3_2 = d$pun3_2,
-      pun4_1 = d$pun4_1,
-      pun4_2 = d$pun4_2,
-      pun5_1 = d$pun5_1,
-      pun5_2 = d$pun5_2,
-      pun6_1 = d$pun6_1,
-      pun6_2 = d$pun6_2,
-      error = error # assumed error rate
-    )
-  # fit model
-  out <- 
-    sampling(
-      compiledModel2,
-      data = dataList,
-      iter = iter,
-      warmup = warmup,
-      chains = chains,
-      cores = cores,
[…]

### 2023-09-08 R/functionsSimulation.R

@@ -1,136 +1,0 @@
-# custom functions
-
-# simulate data
-simulateData <- function(n = 500, seed = 2113, errorRate = 0, alphas, betas) {
-  # set simulation seed
-  set.seed(seed)
-  # n = number of participants
-  # each participant has a score on an individual difference measure: x
-  x <- rnorm(n)
-  # assume that each participant follows a different punishment strategy
-  # 01. Competitive
-  # 02. Avoid disadvantageous inequality
-  # 03. Egalitarian
-  # 04. Seek advantageous inequality
-  # 05. Retributive
-  # 06. Deterrent
-  # 07. Norm-enforcing
-  # 08. Exclusively antisocial punishment
-  # 09. Random choice
-  # 10. Never punish
-  # the punishment strategy is predicted by x
-  strategies <- c()
-  for (i in 1:n) {
-    # predict probabilities with x
-    probs <- softmax(alphas + betas*x[i])
-    strategies[i] <- sample(1:10, size = 1, replace = TRUE, prob = probs)
-  }
-  # each participant plays a series of games developed to tease apart strategies
-  # we measure punishment behaviours in these games
-  # 0 = no punish, 1 = punish
-  # create function to simulate punishment behaviour based on strategy and error rate
-  simulateBehaviour <- function(punStrats = c()) {
-    out <- c()
-    for (i in 1:n) {
-      if (strategies[i] %in% punStrats) {
-        out[i] <- rbinom(1, 1, prob = 1 - errorRate) # punish (with error)
-      } else if (strategies[i] == 9) {
-        out[i] <- rbinom(1, 1, prob = 0.5)           # random choice
-      } else if (!(strategies[i] %in% punStrats)) {
-        out[i] <- rbinom(1, 1, prob = 0 + errorRate) # don't punish (with error)
-      }
-      # small chance of failed comprehension and missing data
-      out[i] <- ifelse(rbinom(1, 1, prob = 0.01) == 0, out[i], -999)
-    }
-    return(out)
-  }
-  # game 1 - No DI 1
-  pun1_1 <- simulateBehaviour(punStrats = c(1, 5, 6, 7))
-  pun1_2 <- simulateBehaviour(punStrats = c(1, 8))
-  # game 2 - No DI 2
[…]

### 2023-09-08 R/functionsStudy1.R

@@ -1,884 +1,0 @@
-# custom functions
-
-# load study 1 data
-loadData1 <- function(fileStudy1) {
-  # labels for rwa scale
-  rwaLabels <- c(
-    "Very strongly disagree"     = 1,
-    "Strongly disagree"          = 2,
-    "Somewhat disagree"          = 3,
-    "Slightly disagree"          = 4,
-    "Neither agree nor disagree" = 5,
-    "Slightly agree"             = 6,
-    "Somewhat agree"             = 7,
-    "Strongly agree"             = 8,
-    "Very strongly agree"        = 9
-    )
-  # labels for god controls scale
-  godLabels <- c(
-    "Strongly disagree" = 1,
-    "Somewhat disagree" = 2,
-    "Slightly disagree" = 3,
-    "Neutral"           = 4,
-    "Slightly agree"    = 5,
-    "Somewhat agree"    = 6,
-    "Strongly agree"    = 7
-  )
-  # labels for education
-  educationLabels <- c(
-    "High School"                              = 1,
-    "Diploma / other professional certificate" = 2,
-    "Completed university"                     = 3,
-    "Masters degree"                           = 4,
-    "PhD or equivalent"                        = 5
-  )
-  # load data
-  read_csv(fileStudy1) %>%
-    # create variables
-    mutate(
-      
-      ### punishment games
-      # numeric punishment vars
-      pun1_1 = ifelse(str_starts(NoDI1_Take,    "Yes"), 1, 0),
-      pun1_2 = ifelse(str_starts(NoDI1_Nothing, "Yes"), 1, 0),
-      pun2_1 = ifelse(str_starts(NoDI2_Take,    "Yes"), 1, 0),
-      pun2_2 = ifelse(str_starts(NoDI2_Nothing, "Yes"), 1, 0),
-      pun3_1 = ifelse(str_starts(NoDI3_Take,    "Yes"), 1, 0),
-      pun3_2 = ifelse(str_starts(NoDI3_Nothing, "Yes"), 1, 0),
[…]

### 2023-09-08 R/functions_loadData.R

@@ -1,0 +1,427 @@
+# custom functions
+
+# load UK/US data
+loadData <- function(fileStudy1, fileStudy2) {
+  # labels for rwa scale
+  rwaLabels <- c(
+    "Very strongly disagree"     = 1,
+    "Strongly disagree"          = 2,
+    "Somewhat disagree"          = 3,
+    "Slightly disagree"          = 4,
+    "Neither agree nor disagree" = 5,
+    "Slightly agree"             = 6,
+    "Somewhat agree"             = 7,
+    "Strongly agree"             = 8,
+    "Very strongly agree"        = 9
+    )
+  # labels for god controls scale
+  godLabels <- c(
+    "Strongly disagree" = 1,
+    "Somewhat disagree" = 2,
+    "Slightly disagree" = 3,
+    "Neutral"           = 4,
+    "Slightly agree"    = 5,
+    "Somewhat agree"    = 6,
+    "Strongly agree"    = 7
+  )
+  # labels for education
+  educationLabels <- c(
+    "High School"                              = 1,
+    "Diploma / other professional certificate" = 2,
+    "Completed university"                     = 3,
+    "Masters degree"                           = 4,
+    "PhD or equivalent"                        = 5
+  )
+  # function to load data
+  loadDataFun <- function(file, country = "") {
+    read_csv(file) %>%
+      # create variables
+      mutate(
+        
+        ### punishment games
+        # numeric punishment vars
+        pun1_1 = ifelse(str_starts(NoDI1_Take,    "Yes"), 1, 0),
+        pun1_2 = ifelse(str_starts(NoDI1_Nothing, "Yes"), 1, 0),
+        pun2_1 = ifelse(str_starts(NoDI2_Take,    "Yes"), 1, 0),
+        pun2_2 = ifelse(str_starts(NoDI2_Nothing, "Yes"), 1, 0),
+        pun3_1 = ifelse(str_starts(NoDI3_Take,    "Yes"), 1, 0),
[…]

### 2023-09-08 R/functions_modelFitting.R

@@ -1,0 +1,122 @@
+# custom functions
+
+# fit stan model without predictor
+fitModel1 <- function(d, compiledModel1, error = 0, 
+                      iter = 2000,  warmup = 1000, chains = 4, cores = 4) {
+  # list for stan
+  dataList <-
+    list(
+      N = nrow(d),
+      Nc = length(unique(d$Country)),
+      countryID = ifelse(d$Country == "United Kingdom", 1L, 2L),
+      pun1_1 = d$pun1_1,
+      pun1_2 = d$pun1_2,
+      pun2_1 = d$pun2_1,
+      pun2_2 = d$pun2_2,
+      pun3_1 = d$pun3_1,
+      pun3_2 = d$pun3_2,
+      pun4_1 = d$pun4_1,
+      pun4_2 = d$pun4_2,
+      pun5_1 = d$pun5_1,
+      pun5_2 = d$pun5_2,
+      pun6_1 = d$pun6_1,
+      pun6_2 = d$pun6_2,
+      error = error # assumed error rate
+    )
+  # fit model
+  out <- 
+    sampling(
+      compiledModel1, 
+      data = dataList, 
+      iter = iter,
+      warmup = warmup,
+      chains = chains,
+      cores = cores,
+      seed = 2113
+      )
+  return(out)
+}
+
+# fit stan model with predictor
+fitModel2 <- function(d, compiledModel2, predictor, error = 0, 
+                      iter = 2000, warmup = 1000, chains = 4, cores = 4) {
+  # listwise deletion for predictor
+  d <- d[as.vector(!is.na(d[,predictor])),]
+  # list for stan
+  dataList <-
+    list(
+      N = nrow(d),
+      Nc = length(unique(d$Country)),
+      countryID = ifelse(d$Country == "United Kingdom", 1L, 2L),
+      # standardise predictor for model fitting
+      pred = d %>% pull(predictor) %>% scale() %>% as.numeric(),
+      pun1_1 = d$pun1_1,
+      pun1_2 = d$pun1_2,
+      pun2_1 = d$pun2_1,
+      pun2_2 = d$pun2_2,
+      pun3_1 = d$pun3_1,
+      pun3_2 = d$pun3_2,
+      pun4_1 = d$pun4_1,
+      pun4_2 = d$pun4_2,
+      pun5_1 = d$pun5_1,
+      pun5_2 = d$pun5_2,
+      pun6_1 = d$pun6_1,
+      pun6_2 = d$pun6_2,
[…]

### 2023-09-08 R/functions_modelResults.R

@@ -1,0 +1,156 @@
+# custom functions
+
+# plot model results
+plotModel1 <- function(post) {
+  # strategy vector
+  strategies <- c("Competitive", "Avoid DI", "Egalitarian", "Seek AI",
+                  "Retributive", "Deterrent", "Norm-enforcing", 
+                  "Antisocial", "Random choice", "No punish")
+  # calculate probabilities
+  P <- post$alpha
+  for (i in 1:nrow(P)) # iterations
+    for (j in 1:2)     # countries
+      P[i,j,] <- softmax(P[i,j,])
+  # plot
+  out <- 
+    tibble(
+      p = c(P[,1,1], P[,1,2], P[,1,3], P[,1,4], P[,1,5], 
+            P[,1,6], P[,1,7], P[,1,8], P[,1,9], P[,1,10],
+            P[,2,1], P[,2,2], P[,2,3], P[,2,4], P[,2,5], 
+            P[,2,6], P[,2,7], P[,2,8], P[,2,9], P[,2,10]),
+      strategy = rep(rep(strategies, each = length(P[,1,1])), times = 2),
+      Country = rep(c("United Kingdom", "United States"), each = length(P[,1,1]) * 10)
+    ) %>%
+    mutate(strategy = factor(strategy, levels = strategies)) %>%
+    ggplot(aes(x = p, y = fct_rev(strategy), colour = Country)) +
+    #geom_density_ridges(rel_min_height = 0.02, colour = "white", scale = 1) +
+    stat_pointinterval(position = position_dodge(width = 0.5)) +
+    scale_x_continuous(name = "Probability of using punishment strategy",
+                       limits = c(0, 0.55), breaks = seq(0, 0.5, by = 0.1)) +
+    ylab(NULL) +
+    theme_classic()
+  # save
+  ggsave(out, filename = "figures/modelResults/model1.pdf", height = 5, width = 7)
+  return(out)
+}
+
+# plot effect of predictor on strategy usage
+plotModelPred <- function(d, post, pred, xlab, xBreaks, file) {
+  # strategy vector
+  strategies <- c("Competitive", "Avoid DI", "Egalitarian", "Seek AI",
+                  "Retributive", "Deterrent", "Norm-enforcing", 
+                  "Antisocial", "Random choice", "Anti-punish")
+  # mean, sd, min, and max for predictor
[…]

### 2023-09-08 R/functions_simulation.R

@@ -1,0 +1,144 @@
+# custom functions
+
+# simulate data
+simulateData <- function(n = 500, seed = 2113, errorRate = 0, alphas, betas) {
+  # set simulation seed
+  set.seed(seed)
+  # n = number of participants
+  # each participant has a score on an individual difference measure: x
+  x <- rnorm(n)
+  # assume that each participant follows a different punishment strategy
+  # 01. Competitive
+  # 02. Avoid disadvantageous inequality
+  # 03. Egalitarian
+  # 04. Seek advantageous inequality
+  # 05. Retributive
+  # 06. Deterrent
+  # 07. Norm-enforcing
+  # 08. Exclusively antisocial punishment
+  # 09. Random choice
+  # 10. Never punish
+  # the punishment strategy is predicted by x
+  strategies <- c()
+  for (i in 1:n) {
+    # predict probabilities with x
+    probs <- softmax(alphas + betas*x[i])
+    strategies[i] <- sample(1:10, size = 1, replace = TRUE, prob = probs)
+  }
+  # each participant plays a series of games developed to tease apart strategies
+  # we measure punishment behaviours in these games
+  # 0 = no punish, 1 = punish
+  # create function to simulate punishment behaviour based on strategy and error rate
+  simulateBehaviour <- function(punStrats = c()) {
+    out <- c()
+    for (i in 1:n) {
+      if (strategies[i] %in% punStrats) {
+        out[i] <- rbinom(1, 1, prob = 1 - errorRate) # punish (with error)
+      } else if (strategies[i] == 9) {
+        out[i] <- rbinom(1, 1, prob = 0.5)           # random choice
+      } else if (!(strategies[i] %in% punStrats)) {
+        out[i] <- rbinom(1, 1, prob = 0 + errorRate) # don't punish (with error)
+      }
+      # small chance of failed comprehension and missing data
+      out[i] <- ifelse(rbinom(1, 1, prob = 0.01) == 0, out[i], -999)
+    }
+    return(out)
+  }
+  # game 1 - No DI 1
+  pun1_1 <- simulateBehaviour(punStrats = c(1, 5, 6, 7))
+  pun1_2 <- simulateBehaviour(punStrats = c(1, 8))
+  # game 2 - No DI 2
[…]

### 2023-09-08 R/functions_summariseData.R

@@ -1,0 +1,512 @@
+# custom functions
+
+# plot sample demographics
+plotSampleStudy <- function(d, country = "") {
+  # subset to country
+  d <- filter(d, Country == country)
+  # age
+  pA <- 
+    ggplot(d, aes(x = Age)) +
+    geom_histogram(fill = "lightsteelblue2") +
+    scale_y_continuous(name = "Count") +
+    theme_classic()
+  # gender
+  pB <-
+    ggplot(d, aes(x = Gender)) +
+    geom_bar(fill = "lightsteelblue2") +
+    scale_y_continuous(name = "Count") +
+    theme_classic()
+  # ethnicity
+  pC <-
+    ggplot(d, aes(x = Ethnicity)) +
+    geom_bar(fill = "lightsteelblue2") +
+    scale_y_continuous(name = "Count") +
+    theme_classic()
+  # student
+  pD <-
+    ggplot(d, aes(x = Student)) +
+    geom_bar(fill = "lightsteelblue2") +
+    scale_y_continuous(name = "Count") +
+    theme_classic()
+  # employment
+  pE <-
+    d %>%
+    mutate(
+      Employment = ifelse(str_starts(Employment, fixed("Not")), "Not in paid work", Employment),
+      Employment = ifelse(str_starts(Employment, fixed("Un")), "Unemployed", Employment),
+      Employment = ifelse(str_starts(Employment, fixed("Due")), "Starting new job", Employment)
+    ) %>%
+    ggplot(aes(x = Employment)) +
+    geom_bar(fill = "lightsteelblue2") +
+    scale_y_continuous(name = "Count") +
+    theme_classic() +
+    theme(axis.text.x = element_text(angle = 30, hjust = 1, size = 7))
+  # title
+  title <- ggdraw() + draw_label(paste0("Sample from ", country), hjust = 1.2)
+  # put together and save
+  out <- 
+    plot_grid(
+      title,
+      plot_grid(pA, pC, nrow = 1), 
+      plot_grid(pD, pB, pE, nrow = 1), 
+      nrow = 3,
+      rel_heights = c(0.15, 1, 1)
+    )
+  # save
+  file <- 
+    paste0(
+      "figures/dataSummaries/",
+      ifelse(
+        country == "United Kingdom",
+        "sample_UK.pdf",
+        "sample_US.pdf"
+        )
+      )
+  ggsave(out, filename = file, width = 7, height = 5)
+  return(out)
+}
+
[…]

### 2023-09-08 _targets.R

@@ -2,14 +2,17 @@ library(targets)
 library(tarchetypes)
 options(tidyverse.quiet = TRUE)
 library(tidyverse)
-source("R/functionsModelFitting.R")
-source("R/functionsSimulation.R")
-source("R/functionsStudy1.R")
-source("R/functionsPowerAnalysis.R")
+source("R/functions_modelFitting.R")
+source("R/functions_simulation.R")
+source("R/functions_powerAnalysis.R")
+source("R/functions_loadData.R")
+source("R/functions_summariseData.R")
+source("R/functions_modelResults.R")
 tar_option_set(
-  packages = c("cowplot", "ggcorrplot", "ggridges", "ltm", 
-               "rethinking", "rstan", "tidybayes", "tidytext", 
-               "tidyverse", "wordcloud"),
+  packages = c("ape", "cowplot", "ggcorrplot", "ggrepel", 
+               "ggridges", "knitr", "ltm", "phangorn", 
+               "rethinking", "rstan", "tanggle", "tidybayes", 
+               "tidytext", "tidyverse", "wordcloud"),
   deployment = "main"
 )
 options(
@@ -17,27 +20,28 @@ options(
   clustermq.template = "slurm_clustermq.tmpl"
 )
 
-# power analysis static branching (see below)
-powerMap <-
-  tar_map(
-    values = tibble(strategy = 1:10),
-    # seeds for data simulation
-    tar_target(powerSeed, 1:100),
-    # simulate data for power analysis
-    tar_target(dPower, simulateData(n = 1019, errorRate = 0, seed = powerSeed,
-                                    # use intercepts from m1 model (UK only)
-                                    alphas = c(-1.48, -0.31, 1.43, 0.22, -0.39,
-                                               -0.22, 0.22, -0.87, -0.47, 2.31),
-                                    # detecting medium effect for one strategy
-                                    # no effect for all other strategies
-                                    betas = c(rep(0, times = strategy - 1), 1, rep(0, times = 10 - strategy))),
-               pattern = map(powerSeed)),
-    # fit model to simulated data and return posterior summary
-    tar_target(powerModel, fitPowerModel(dPower, compiledModel2, strategy),
-               deployment = "worker", pattern = map(dPower)),
-    # power
[…]

### 2023-09-08 run.R

@@ -1,1 +1,1 @@
-targets::tar_make_clustermq(starts_with("power"), workers = 50L)
+targets::tar_make()

### 2023-09-08 testNetwork.R

@@ -1,0 +1,189 @@
+library(ape)
+library(tanggle)
+library(phangorn)
+library(targets)
+library(tidyverse)
+library(ggrepel)
+
+d <-
+  tar_read(d) %>%
+  # subset to behavioural decisions
+  dplyr::select(starts_with("pun"), strategy) %>%
+  # reduce to 403 unique vectors with counts
+  group_by_at(vars(starts_with("pun"))) %>%
+  summarise(
+    count = n(),
+    strategy = unique(strategy),
+    .groups = "drop"
+    ) %>%
+  # keep groups with more than 1 count
+  filter(count > 1)
+# n is number of unique patterns of behaviour
+n <- nrow(d)
+# calculate hamming distance matrix
+dist <- matrix(NA, nrow = n, ncol = n)
+for (i in 1:n) {
+  for (j in 1:n) {
+    dist[i,j] <- sum(d[i,1:12] != d[j,1:12])
+  }
+}
+rownames(dist) <- colnames(dist) <- 1:n
+# create neighbour net
+net <- neighborNet(dist)
+# plot splits graph
+p <- 
+  ggsplitnet(
+    net,
+    colour = "grey90",
+    linewidth = 0.05
+    )
+# labels for strategies
+labelStrategies <- 
+  c(
+    "Anti-punish"                    = "Anti-punish\n(000000000000)",
+    "Seek advantageous inequity"     = "Seek AI\n(001010000000)",
+    "Egalitarian"                    = "Egalitarian\n(000000001010)",
+    "Avoid disadvantageous inequity" = "Avoid DI\n(000000001000)",
+    "Retributive"                    = "Retributive\n(101010101000)",
+    "Deterrent"                      = "Deterrent\n(101000101000)",
+    "Norm-enforcing"                 = "Norm-enforcing\n(101000101010)",
+    "Competitive"                    = "Competitive\n(111111001111)"
+  )
+# add the data
+pd <- 
+  tibble(
+    tip = as.character(1:n),
+    Count = d$count,
+    strategy = ifelse(d$strategy == "N/A", "", labelStrategies[as.character(d$strategy)]),
+    exact = strategy != ""
+  ) %>%
+  left_join(p$data, by = c("tip" = "label"))
+p <-
+  p +
+  geom_point(
+    data = pd,
[…]

### 2023-10-20 appendix.Rmd
starting writing manuscript

\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\renewcommand{\thefigure}{\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}

# **Supplementary Information**

\setcounter{page}{1}
\centering
\noindent \hspace*{10mm} \small Mapping the punishment strategy space \newline
\hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, Nichola Raihani\textsuperscript{1,2} \newline
\raggedright
\noindent \footnotesize \textsuperscript{1} School of Psychology, University of Auckland, Auckland, New Zealand \newline
\noindent \footnotesize \textsuperscript{2} Department of Experimental Psychology, University College London, London, United Kingdom \newline
\normalsize
\newpage

## Supplementary Figures

## Supplementary Tables

## Supplementary References

### 2023-10-20 manuscript.Rmd
starting writing manuscript

---
title             : "Mapping the punishment strategy space"
shorttitle        : "Punishment strategies"

author: 
  - name          : "Scott Claessens"
    affiliation   : "*1"
  - name          : "Quentin D. Atkinson"
    affiliation   : "1"
  - name          : "Nichola Raihani"
    affiliation   : "1,2"

affiliation:
  - id            : "1"
    institution   : "\\footnotesize School of Psychology, University of Auckland, Auckland, New Zealand"
  - id            : "2"
    institution   : "\\footnotesize Department of Experimental Psychology, University College London, London, United Kingdom"

abstract: |
 x

keywords          : "x"
wordcount         : "xxxx words"
bibliography      : ["r-references.bib"]
floatsintext      : yes
figurelist        : no
figsintext        : yes
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
numbersections    : false
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes: 
  - \usepackage{array}
  - \usepackage{caption}
  - \usepackage{graphicx}
  - \usepackage{siunitx}
  - \usepackage[normalem]{ulem}
  - \usepackage{colortbl}
  - \usepackage{multirow}
  - \usepackage{hhline}
  - \usepackage{calc}
  - \usepackage{tabularx}
  - \usepackage{threeparttable}
  - \usepackage{wrapfig}
  - \usepackage{adjustbox}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \usepackage{amsmath}
  - \usepackage{xcolor}
  - \usepackage{makecell}
  - \renewcommand\theadfont{}
  - \raggedbottom
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
  - \captionsetup[table]{font={stretch=1,small}}
  - \captionsetup[figure]{font={stretch=1,small}}
  - \nolinenumbers
  - \note{\raggedright * Correspondence concerning this article should be addressed to Scott Claessens, Level 2, Building 302, 23 Symonds Street, Auckland, New Zealand. \text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This working paper has not yet been peer-reviewed.}
appendix          : "appendix.Rmd"
csl               : nature.csl
---

\linenumbers

# Introduction

Humans cooperate on a scale unparalleled in the animal kingdom. One mechanism thought to sustain this level of cooperation is costly punishment, whereby individuals harm others at a personal cost[@CluttonBrock1995], ostensibly encouraging cooperative behaviour from the target (or bystanders[@DosSantos2011;@DosSantos2013;@Raihani2015]) in the future. Punishment therefore offers a route to maintaining or increasing cooperation by changing the payoff structure of social interactions such that it no longer pays to cheat or exploit social partners[@CluttonBrock1995;@Ostrom1990].

In humans, many studies of punishment have been carried out in laboratory settings using economic games[@Balliet2011;@Balliet2013;@Chaudhuri2011;@Dreber2008;@Fehr2000;@Fehr2002;@Henrich2006;@Nikiforakis2008;@Raihani2012a;@Raihani2019]. In these games, participants are usually given a sum of money that they can use to invest in collective action or to help others. Alternatively, participants can 'cheat' by keeping the money for themselves or by exploiting the contributions of others. Punishment is introduced into such games by giving participants the option to pay a small 'fee' to impose a greater 'fine' on their co-players. Several lines of experimental evidence indicate that people use this punishment option[@Henrich2006], that they enjoy punishing[@DeQuervain2004], and that they frequently, though not always[@Herrmann2008], punish cheating or exploitative co-players[@Fehr2000;@Fehr2002].

Evidence from these experiments suggests that the threat of costly punishment plays an important role in promoting human cooperation. People tend to cooperate more in games where punishment is possible compared to those where it is not[@Balliet2011;@Balliet2013;@Raihani2019]. The effect that the threat of punishment has on cooperation is also evident in the higher contributions typically observed in the Ultimatum Game (where punishment is possible) compared to the structurally-similar Dictator Game (where it is not)[@Camerer2003]. This typical cooperation-enhancing effect of punishment has also been observed across societies[@Balliet2013], leading some to suggest that costly punishment has played a key role in the cultural evolution of cooperation in humans[@Bowles2013;@Boyd2003;@Chudek2011;@Henrich2017].

Nevertheless, it remains unclear whether individuals playing economic games use punishment as a behaviour-change tool to enforce cooperation or as a means to achieve other ends. Some have argued that punishment is primarily used to shape future behaviour, either to deter personal harm[@Delton2017;@Dreber2008;@DosSantos2013] or to uphold normative standards of cooperative behaviour[@Boyd2003;@Chudek2011;@Fehr2018;@Mathew2011;@Mathew2014;@Richerson2016]. But while the _threat_ of punishment can have a cooperation-enhancing effect, the _enactment_ of this punishment seldom changes the future behaviour of deviant targets[@Raihani2019]. This calls into question whether punishment primarily operates as a behaviour-change tool or whether it is used to achieve other goals.

Beyond behaviour-shaping concerns, there are a host of other reasons that people may want to punish in economic games. Punishers might be motivated by a desire for retribution rather than deterrence, punishing in proportion to the amount of harm that was personally caused[@Carlsmith2002]. Punishment might be driven by concerns about relative payoffs, such as disadvantageous inequity aversion (i.e., avoiding having less than others[@Raihani2012b;@Raihani2019]) and/or general egalitarian preferences (i.e., wanting all participants to receive the same payoffs[@Dawes2007]). Such concerns about relative payoffs may be activated when participants earn less than cheats in economic games or when there are income disparities in these settings. People might also use punishment for competitive purposes, seeking advantageous inequity for themselves (i.e., having more than others) and/or improving their relative position[@Raihani2019].

Common economic game designs have been unable to tease apart these different motives for punishment because participants who interact with cheats in these games experience both losses _and_ lower relative payoffs. The typical 1:3 fee-to-fine ratio of punishment in economic games compounds this issue. With this setup, people can simultaneously use punishment to reciprocate losses, to deter others from cheating, and to reduce or reverse disparities in payoffs between themselves and targets. To add to this complexity, it is evident that people use punishment in seemingly disparate ways: punishing when no behaviour change is possible, such as in one-shot games[@Bone2015;@Henrich2006;@Raihani2012b;@Walker2004] or in games where the target never learns about the punishment[@Crockett2014]; punishing those who did not cheat or who over-contributed to collective action (antisocial punishment[@Herrmann2008;@Sylwester2013]); punishing in scenarios where they were not personally harmed (third-party punishment[@Fehr2004]); and punishing in scenarios where disparities in payoffs did not arise from participants' actions[@Dawes2007;@Johnson2009;@Fowler2005].

The general conclusion from this research is that there is no one unifying function of costly punishment in humans. Instead, punishment should be thought of as a flexible behavioural tool that serves a variety of functions that are not mutually exclusive[@Raihani2019]. Due to its multipurpose nature, we should therefore expect variation in punishment strategies in the population, much like the observed variation in social learning strategies[@Molleman2014]. Some individuals may use punishment as a behaviour shaping tool, for example, while others may use it to reduce or reverse payoff differentials.

This insight raises several underexplored questions. First, which punishment strategies are more frequent in human populations? Second, what traits predict adherence to a particular strategy? Previous work has reported that personality is related to cooperative behaviour[@Thielmann2020] and demographics, political ideology, and religiosity are related to punitive behaviour[@Hofmann2018], but no research has related these variables to specific punishment strategies. Third, do people have insight into their own punishment strategy? Previous work has argued that people are often unaware of the underlying function of their punitive behaviour, yet they feel compelled to enact it anyway[@Carlsmith2002;@Nisbett1977].

Here, we aim to delineate the many possible punishment strategies by asking whether people punish in a manner consistent with a specific strategy and, if so, what other characteristics (personality, social preferences, political orientation) predict the use of different punishment strategies. Table \@ref(tab:tableStrategies) summarises the potential functions for costly punishment in economic games and the behavioural strategies they predict. Note that we do not include reputational functions of punishment in this table, such as signalling trustworthiness[@Barclay2006;@Batistoni2022;@Jordan2016;@Jordan2017;@Raihani2015], because our focus is on punishment strategies in anonymous economic games without reputational incentives (but see ref[@Jordan2020]).

(ref:tableStrategiesCaption) xxx

```{r tableStrategies, echo=FALSE}
apa_table(
  tar_read(tableStrategies),
  escape = FALSE,
  landscape = TRUE,
  align = c("p{1.7cm}", "p{3.3cm}", rep("c", 12)),
  col_spanners = list(
    "\\thead{No DI 1 \\\\ 70-10}"   = 3:4,
    "\\thead{No DI 2 \\\\ 70-30}"   = 5:6,
    "\\thead{No DI 3 \\\\ (Computer) \\\\ 70-30}"   = 7:8,
    "\\thead{No DI 4 \\\\ (1:1 Fee-Fine) \\\\ 70-30}"   = 9:10,
    "\\thead{DI \\\\ 70-50}"        = 11:12,
    "\\thead{3PP \\\\ 70-70 [100]}" = 13:14
    ),
  caption = "(ref:tableStrategiesCaption)",
  font_size = "scriptsize"
)
```

Building on previous designs[@Bone2015;@Deutchman2021;@Marczyk2017;@Raihani2012b], we employ a suite of one-shot economic games where individuals are given the opportunity to punish targets at a personal cost. We carefully designed this suite of games to tease apart the proposed punishment strategies in Table 1, such that each strategy predicts a different pattern of behaviour across all the games. We use the resulting behavioural patterns to discern which punishment strategy participants are employing. We then combine these behavioural patterns with data on demographics, personality, political ideology, religiosity, and self-reported strategy usage.

# Results

x

# Discussion

x

# Methods

x

\newpage
\nolinenumbers

# Acknowledgements

x

# Author Contributions

x

# Competing Interests

The authors declare no competing interests.

# Data Availability

All data used in this study are publicly available on GitHub: https://github.com/ScottClaessens/punishStrategies

# Code Availability

All code to reproduce the analyses in this study are publicly available on GitHub: https://github.com/ScottClaessens/punishStrategies

\newpage

# References

\begingroup

<div id="refs" custom-style="Bibliography"></div>
\endgroup

\newpage
\vspace*{60mm}

```{r child = "appendix.Rmd"}
```

### 2023-10-20 manuscript.tex
starting writing manuscript

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  man,floatsintext]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{english}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\centering\begin{threeparttable}}
%   {\end{threeparttable}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother

\usepackage{xpatch}
\makeatletter
\xapptocmd\appendix
  {\xapptocmd\section
    {\addcontentsline{toc}{section}{\appendixname\ifoneappendix\else~\theappendix\fi\\: #1}}
    {}{\InnerPatchFailed}%
  }
{}{\PatchFailed}
\keywords{x\newline\indent Word count: xxxx words}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\usepackage{array}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[normalem]{ulem}
\usepackage{colortbl}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{calc}
\usepackage{tabularx}
\usepackage{threeparttable}
\usepackage{wrapfig}
\usepackage{adjustbox}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{makecell}
\renewcommand\theadfont{}
\raggedbottom
\AtBeginEnvironment{tabular}{\singlespacing}
\AtBeginEnvironment{lltable}{\singlespacing}
\AtBeginEnvironment{tablenotes}{\doublespacing}
\captionsetup[table]{font={stretch=1,small}}
\captionsetup[figure]{font={stretch=1,small}}
\nolinenumbers
\note{\raggedright * Correspondence concerning this article should be addressed to Scott Claessens, Level 2, Building 302, 23 Symonds Street, Auckland, New Zealand. \text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This working paper has not yet been peer-reviewed.}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Mapping the punishment strategy space},
  pdfauthor={Scott Claessens*1, Quentin D. Atkinson1, \& Nichola Raihani1,2},
  pdflang={en-EN},
  pdfkeywords={x},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Mapping the punishment strategy space}
\author{Scott Claessens\textsuperscript{*1}, Quentin D. Atkinson\textsuperscript{1}, \& Nichola Raihani\textsuperscript{1,2}}
\date{}

\shorttitle{Punishment strategies}

\affiliation{\vspace{0.5cm}\textsuperscript{1} \footnotesize School of Psychology, University of Auckland, Auckland, New Zealand\\\textsuperscript{2} \footnotesize Department of Experimental Psychology, University College London, London, United Kingdom}

\abstract{%
x
}

\begin{document}
\maketitle

\linenumbers

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Humans cooperate on a scale unparalleled in the animal kingdom. One mechanism thought to sustain this level of cooperation is costly punishment, whereby individuals harm others at a personal cost\textsuperscript{1}, ostensibly encouraging cooperative behaviour from the target (or bystanders\textsuperscript{2--4}) in the future. Punishment therefore offers a route to maintaining or increasing cooperation by changing the payoff structure of social interactions such that it no longer pays to cheat or exploit social partners\textsuperscript{1,5}.

In humans, many studies of punishment have been carried out in laboratory settings using economic games\textsuperscript{6--15}. In these games, participants are usually given a sum of money that they can use to invest in collective action or to help others. Alternatively, participants can `cheat' by keeping the money for themselves or by exploiting the contributions of others. Punishment is introduced into such games by giving participants the option to pay a small `fee' to impose a greater `fine' on their co-players. Several lines of experimental evidence indicate that people use this punishment option\textsuperscript{12}, that they enjoy punishing\textsuperscript{16}, and that they frequently, though not always\textsuperscript{17}, punish cheating or exploitative co-players\textsuperscript{10,11}.

Evidence from these experiments suggests that the threat of costly punishment plays an important role in promoting human cooperation. People tend to cooperate more in games where punishment is possible compared to those where it is not\textsuperscript{6,7,15}. The effect that the threat of punishment has on cooperation is also evident in the higher contributions typically observed in the Ultimatum Game (where punishment is possible) compared to the structurally-similar Dictator Game (where it is not)\textsuperscript{18}. This typical cooperation-enhancing effect of punishment has also been observed across societies\textsuperscript{7}, leading some to suggest that costly punishment has played a key role in the cultural evolution of cooperation in humans\textsuperscript{19--22}.

Nevertheless, it remains unclear whether individuals playing economic games use punishment as a behaviour-change tool to enforce cooperation or as a means to achieve other ends. Some have argued that punishment is primarily used to shape future behaviour, either to deter personal harm\textsuperscript{3,9,23} or to uphold normative standards of cooperative behaviour\textsuperscript{20,21,24--27}. But while the \emph{threat} of punishment can have a cooperation-enhancing effect, the \emph{enactment} of this punishment seldom changes the future behaviour of deviant targets\textsuperscript{15}. This calls into question whether punishment primarily operates as a behaviour-change tool or whether it is used to achieve other goals.

Beyond behaviour-shaping concerns, there are a host of other reasons that people may want to punish in economic games. Punishers might be motivated by a desire for retribution rather than deterrence, punishing in proportion to the amount of harm that was personally caused\textsuperscript{28}. Punishment might be driven by concerns about relative payoffs, such as disadvantageous inequity aversion (i.e., avoiding having less than others\textsuperscript{15,29}) and/or general egalitarian preferences (i.e., wanting all participants to receive the same payoffs\textsuperscript{30}). Such concerns about relative payoffs may be activated when participants earn less than cheats in economic games or when there are income disparities in these settings. People might also use punishment for competitive purposes, seeking advantageous inequity for themselves (i.e., having more than others) and/or improving their relative position\textsuperscript{15}.

Common economic game designs have been unable to tease apart these different motives for punishment because participants who interact with cheats in these games experience both losses \emph{and} lower relative payoffs. The typical 1:3 fee-to-fine ratio of punishment in economic games compounds this issue. With this setup, people can simultaneously use punishment to reciprocate losses, to deter others from cheating, and to reduce or reverse disparities in payoffs between themselves and targets. To add to this complexity, it is evident that people use punishment in seemingly disparate ways: punishing when no behaviour change is possible, such as in one-shot games\textsuperscript{12,29,31,32} or in games where the target never learns about the punishment\textsuperscript{33}; punishing those who did not cheat or who over-contributed to collective action (antisocial punishment\textsuperscript{17,34}); punishing in scenarios where they were not personally harmed (third-party punishment\textsuperscript{35}); and punishing in scenarios where disparities in payoffs did not arise from participants' actions\textsuperscript{30,36,37}.

The general conclusion from this research is that there is no one unifying function of costly punishment in humans. Instead, punishment should be thought of as a flexible behavioural tool that serves a variety of functions that are not mutually exclusive\textsuperscript{15}. Due to its multipurpose nature, we should therefore expect variation in punishment strategies in the population, much like the observed variation in social learning strategies\textsuperscript{38}. Some individuals may use punishment as a behaviour shaping tool, for example, while others may use it to reduce or reverse payoff differentials.

This insight raises several underexplored questions. First, which punishment strategies are more frequent in human populations? Second, what traits predict adherence to a particular strategy? Previous work has reported that personality is related to cooperative behaviour\textsuperscript{39} and demographics, political ideology, and religiosity are related to punitive behaviour\textsuperscript{40}, but no research has related these variables to specific punishment strategies. Third, do people have insight into their own punishment strategy? Previous work has argued that people are often unaware of the underlying function of their punitive behaviour, yet they feel compelled to enact it anyway\textsuperscript{28,41}.

Here, we aim to delineate the many possible punishment strategies by asking whether people punish in a manner consistent with a specific strategy and, if so, what other characteristics (personality, social preferences, political orientation) predict the use of different punishment strategies. Table \ref{tab:tableStrategies} summarises the potential functions for costly punishment in economic games and the behavioural strategies they predict. Note that we do not include reputational functions of punishment in this table, such as signalling trustworthiness\textsuperscript{4,42--45}, because our focus is on punishment strategies in anonymous economic games without reputational incentives (but see ref\textsuperscript{46}).

\begin{lltable}

\scriptsize{

\begin{longtable}{p{1.7cm}p{3.3cm}cccccccccccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
\caption{\label{tab:tableStrategies}xxx}\\
\toprule
 &  & \multicolumn{2}{c}{\thead{No DI 1 \\ 70-10}} & \multicolumn{2}{c}{\thead{No DI 2 \\ 70-30}} & \multicolumn{2}{c}{\thead{No DI 3 \\ (Computer) \\ 70-30}} & \multicolumn{2}{c}{\thead{No DI 4 \\ (1:1 Fee-Fine) \\ 70-30}} & \multicolumn{2}{c}{\thead{DI \\ 70-50}} & \multicolumn{2}{c}{\thead{3PP \\ 70-70 [100]}} \\
\cmidrule(r){3-4} \cmidrule(r){5-6} \cmidrule(r){7-8} \cmidrule(r){9-10} \cmidrule(r){11-12} \cmidrule(r){13-14}
Function & \multicolumn{1}{c}{Behavioural strategy} & \multicolumn{1}{c}{\thead{Take \\ 50-30}} & \multicolumn{1}{c}{\thead{No take \\ 70-10}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-70}} & \multicolumn{1}{c}{\thead{No take \\ 70-50}} & \multicolumn{1}{c}{\thead{Take \\ 50-90 [100]}} & \multicolumn{1}{c}{\thead{No take \\ 70-70 [100]}}\\
\midrule
\endfirsthead
\caption*{\normalfont{Table \ref{tab:tableStrategies} continued}}\\
\toprule
 &  & \multicolumn{2}{c}{\thead{No DI 1 \\ 70-10}} & \multicolumn{2}{c}{\thead{No DI 2 \\ 70-30}} & \multicolumn{2}{c}{\thead{No DI 3 \\ (Computer) \\ 70-30}} & \multicolumn{2}{c}{\thead{No DI 4 \\ (1:1 Fee-Fine) \\ 70-30}} & \multicolumn{2}{c}{\thead{DI \\ 70-50}} & \multicolumn{2}{c}{\thead{3PP \\ 70-70 [100]}} \\
\cmidrule(r){3-4} \cmidrule(r){5-6} \cmidrule(r){7-8} \cmidrule(r){9-10} \cmidrule(r){11-12} \cmidrule(r){13-14}
Function & \multicolumn{1}{c}{Behavioural strategy} & \multicolumn{1}{c}{\thead{Take \\ 50-30}} & \multicolumn{1}{c}{\thead{No take \\ 70-10}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-70}} & \multicolumn{1}{c}{\thead{No take \\ 70-50}} & \multicolumn{1}{c}{\thead{Take \\ 50-90 [100]}} & \multicolumn{1}{c}{\thead{No take \\ 70-70 [100]}}\\
\midrule
\endhead
Deterrent & Punish to deter another who has harmed you from harming you again in the future & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize\\
Norm-enforcing & Punish to enforce a shared anti-harm norm and encourage future norm compliance, even amongst third parties & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize\\
Retributive & Punish if doing so harms another who has harmed you & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize\\
Avoid DI & Punish if doing so avoids disadvantageous inequity for self & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize\\
Egalitarian & Punish if doing so makes payoffs for all more equal & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize\\
Seek AI & Punish if doing so produces advantageous inequity for self & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize\\
Competitive & Punish if doing so improves your relative position & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize\\
Antisocial & Punish exclusively those who do not cause harm & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{green} \checkmark \normalcolor \scriptsize\\
Never punish & Never punish others & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize & \Large \color{red}         x     \normalcolor \scriptsize\\
\bottomrule
\end{longtable}

}

\end{lltable}

Building on previous designs\textsuperscript{29,31,47,48}, we employ a suite of one-shot economic games where individuals are given the opportunity to punish targets at a personal cost. We carefully designed this suite of games to tease apart the proposed punishment strategies in Table 1, such that each strategy predicts a different pattern of behaviour across all the games. We use the resulting behavioural patterns to discern which punishment strategy participants are employing. We then combine these behavioural patterns with data on demographics, personality, political ideology, religiosity, and self-reported strategy usage.

\hypertarget{results}{%
\section{Results}\label{results}}

x

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

x

\hypertarget{methods}{%
\section{Methods}\label{methods}}

x

\newpage
\nolinenumbers

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

x

\hypertarget{author-contributions}{%
\section{Author Contributions}\label{author-contributions}}

x

\hypertarget{competing-interests}{%
\section{Competing Interests}\label{competing-interests}}

The authors declare no competing interests.

\hypertarget{data-availability}{%
\section{Data Availability}\label{data-availability}}

All data used in this study are publicly available on GitHub: \url{https://github.com/ScottClaessens/punishStrategies}

\hypertarget{code-availability}{%
\section{Code Availability}\label{code-availability}}

All code to reproduce the analyses in this study are publicly available on GitHub: \url{https://github.com/ScottClaessens/punishStrategies}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\vadjust pre{\hypertarget{ref-CluttonBrock1995}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{Clutton-Brock, T. H. \& Parker, G. A. \href{https://doi.org/10.1038/373209a0}{Punishment in animal societies}. \emph{Nature} \textbf{373}, 209--216 (1995).}

\leavevmode\vadjust pre{\hypertarget{ref-DosSantos2011}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{dos Santos, M., Rankin, D. J. \& Wedekind, C. \href{https://doi.org/10.1098/rspb.2010.1275}{The evolution of punishment through reputation}. \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{278}, 371--377 (2011).}

\leavevmode\vadjust pre{\hypertarget{ref-DosSantos2013}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{dos Santos, M., Rankin, D. J. \& Wedekind, C. \href{https://www.jstor.org/stable/24032701}{Human cooperation based on punishment reputation}. \emph{Evolution} \textbf{67}, 2446--2450 (2013).}

\leavevmode\vadjust pre{\hypertarget{ref-Raihani2015}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{Raihani, N. J. \& Bshary, R. \href{https://doi.org/10.1016/j.tree.2014.12.003}{The reputation of punishers}. \emph{Trends in Ecology \& Evolution} \textbf{30}, 98--103 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-Ostrom1990}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{Ostrom, E. \emph{Governing the commons: The evolution of institutions for collective action}. (Cambridge University Press, 1990).}

\leavevmode\vadjust pre{\hypertarget{ref-Balliet2011}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{Balliet, D., Mulder, L. B. \& Van Lange, P. A. M. \href{https://doi.org/10.1037/a0023489}{Reward, punishment, and cooperation: A meta-analysis}. \emph{Psychological Bulletin} \textbf{137}, 594--615 (2011).}

\leavevmode\vadjust pre{\hypertarget{ref-Balliet2013}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{Balliet, D. \& Lange, P. A. M. V. \href{https://doi.org/10.1177/1745691613488533}{Trust, punishment, and cooperation across 18 societies: A meta-analysis}. \emph{Perspectives on Psychological Science} \textbf{8}, 363--379 (2013).}

\leavevmode\vadjust pre{\hypertarget{ref-Chaudhuri2011}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{Chaudhuri, A. \href{https://doi.org/10.1007/s10683-010-9257-1}{Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature}. \emph{Experimental Economics} \textbf{14}, 47--83 (2011).}

\leavevmode\vadjust pre{\hypertarget{ref-Dreber2008}{}}%
\CSLLeftMargin{9. }%
\CSLRightInline{Dreber, A., Rand, D. G., Fudenberg, D. \& Nowak, M. A. \href{https://doi.org/10.1038/nature06723}{Winners don't punish}. \emph{Nature} \textbf{452}, 348--351 (2008).}

\leavevmode\vadjust pre{\hypertarget{ref-Fehr2000}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{Fehr, E. \& Gächter, S. \href{https://www.jstor.org/stable/117319}{Cooperation and punishment in public goods experiments}. \emph{American Economic Review} \textbf{90}, 980--994 (2000).}

\leavevmode\vadjust pre{\hypertarget{ref-Fehr2002}{}}%
\CSLLeftMargin{11. }%
\CSLRightInline{Fehr, E. \& Gächter, S. \href{https://doi.org/10.1038/415137a}{Altruistic punishment in humans}. \emph{Nature} \textbf{415}, 137--140 (2002).}

\leavevmode\vadjust pre{\hypertarget{ref-Henrich2006}{}}%
\CSLLeftMargin{12. }%
\CSLRightInline{Henrich, J. \emph{et al.} \href{https://doi.org/10.1126/science.1127333}{Costly punishment across human societies}. \emph{Science} \textbf{312}, 1767--1770 (2006).}

\leavevmode\vadjust pre{\hypertarget{ref-Nikiforakis2008}{}}%
\CSLLeftMargin{13. }%
\CSLRightInline{Nikiforakis, N. \& Normann, H.-T. A comparative statics analysis of punishment in public-good experiments. \emph{Experimental Economics} \textbf{11}, 358--369 (2008).}

\leavevmode\vadjust pre{\hypertarget{ref-Raihani2012a}{}}%
\CSLLeftMargin{14. }%
\CSLRightInline{Raihani, N. J., Thornton, A. \& Bshary, R. \href{https://doi.org/10.1016/j.tree.2011.12.004}{Punishment and cooperation in nature}. \emph{Trends in Ecology \& Evolution} \textbf{27}, 288--295 (2012).}

\leavevmode\vadjust pre{\hypertarget{ref-Raihani2019}{}}%
\CSLLeftMargin{15. }%
\CSLRightInline{Raihani, N. J. \& Bshary, R. \href{https://doi.org/10.1017/ehs.2019.12}{Punishment: One tool, many uses}. \emph{Evolutionary Human Sciences} \textbf{1}, e12 (2019).}

\leavevmode\vadjust pre{\hypertarget{ref-DeQuervain2004}{}}%
\CSLLeftMargin{16. }%
\CSLRightInline{de Quervain, D. J.-F. \emph{et al.} \href{https://doi.org/10.1126/science.1100735}{The neural basis of altruistic punishment}. \emph{Science} \textbf{305}, 1254--1258 (2004).}

\leavevmode\vadjust pre{\hypertarget{ref-Herrmann2008}{}}%
\CSLLeftMargin{17. }%
\CSLRightInline{Herrmann, B., Thöni, C. \& Gächter, S. \href{https://doi.org/10.1126/science.1153808}{Antisocial punishment across societies}. \emph{Science} \textbf{319}, 1362--1367 (2008).}

\leavevmode\vadjust pre{\hypertarget{ref-Camerer2003}{}}%
\CSLLeftMargin{18. }%
\CSLRightInline{Camerer, C. F. \emph{Behavioral game theory: Experiments in strategic interaction}. (Russell Sage Foundation, 2003).}

\leavevmode\vadjust pre{\hypertarget{ref-Bowles2013}{}}%
\CSLLeftMargin{19. }%
\CSLRightInline{Bowles, S. \& Gintis, H. \emph{A cooperative species: Human reciprocity and its evolution}. (Princeton University Press, 2013).}

\leavevmode\vadjust pre{\hypertarget{ref-Boyd2003}{}}%
\CSLLeftMargin{20. }%
\CSLRightInline{Boyd, R., Gintis, H., Bowles, S. \& Richerson, P. J. The evolution of altruistic punishment. \emph{Proceedings of the National Academy of Sciences} \textbf{100}, 3531--3535 (2003).}

\leavevmode\vadjust pre{\hypertarget{ref-Chudek2011}{}}%
\CSLLeftMargin{21. }%
\CSLRightInline{Chudek, M. \& Henrich, J. \href{https://doi.org/10.1016/j.tics.2011.03.003}{Culture-gene coevolution, norm-psychology and the emergence of human prosociality}. \emph{Trends in Cognitive Sciences} \textbf{15}, 218--226 (2011).}

\leavevmode\vadjust pre{\hypertarget{ref-Henrich2017}{}}%
\CSLLeftMargin{22. }%
\CSLRightInline{Henrich, J. \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter}. (Princeton University Press, 2017).}

\leavevmode\vadjust pre{\hypertarget{ref-Delton2017}{}}%
\CSLLeftMargin{23. }%
\CSLRightInline{Delton, A. W. \& Krasnow, M. M. \href{https://doi.org/10.1016/j.evolhumbehav.2017.07.003}{The psychology of deterrence explains why group membership matters for third-party punishment}. \emph{Evolution and Human Behavior} \textbf{38}, 734--743 (2017).}

\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018}{}}%
\CSLLeftMargin{24. }%
\CSLRightInline{Fehr, E. \& Schurtenberger, I. \href{https://doi.org/10.1038/s41562-018-0385-5}{Normative foundations of human cooperation}. \emph{Nature Human Behaviour} \textbf{2}, 458--468 (2018).}

\leavevmode\vadjust pre{\hypertarget{ref-Mathew2011}{}}%
\CSLLeftMargin{25. }%
\CSLRightInline{Mathew, S. \& Boyd, R. Punishment sustains large-scale cooperation in prestate warfare. \emph{Proceedings of the National Academy of Sciences} \textbf{108}, 11375--11380 (2011).}

\leavevmode\vadjust pre{\hypertarget{ref-Mathew2014}{}}%
\CSLLeftMargin{26. }%
\CSLRightInline{Mathew, S. \& Boyd, R. \href{https://doi.org/10.1016/j.evolhumbehav.2013.10.001}{The cost of cowardice: Punitive sentiments towards free riders in {Turkana} raids}. \emph{Evolution and Human Behavior} \textbf{35}, 58--64 (2014).}

\leavevmode\vadjust pre{\hypertarget{ref-Richerson2016}{}}%
\CSLLeftMargin{27. }%
\CSLRightInline{Richerson, P. \emph{et al.} \href{https://doi.org/10.1017/S0140525X1400106X}{Cultural group selection plays an essential role in explaining human cooperation: A sketch of the evidence}. \emph{Behavioral and Brain Sciences} \textbf{39}, e30 (2016).}

\leavevmode\vadjust pre{\hypertarget{ref-Carlsmith2002}{}}%
\CSLLeftMargin{28. }%
\CSLRightInline{Carlsmith, K. M., Darley, J. M. \& Robinson, P. H. \href{https://doi.org/10.1037/0022-3514.83.2.284}{Why do we punish? Deterrence and just deserts as motives for punishment}. \emph{Journal of Personality and Social Psychology} \textbf{83}, 284--299 (2002).}

\leavevmode\vadjust pre{\hypertarget{ref-Raihani2012b}{}}%
\CSLLeftMargin{29. }%
\CSLRightInline{Raihani, N. J. \& McAuliffe, K. \href{https://doi.org/10.1098/rsbl.2012.0470}{Human punishment is motivated by inequity aversion, not a desire for reciprocity}. \emph{Biology Letters} \textbf{8}, 802--804 (2012).}

\leavevmode\vadjust pre{\hypertarget{ref-Dawes2007}{}}%
\CSLLeftMargin{30. }%
\CSLRightInline{Dawes, C. T., Fowler, J. H., Johnson, T., McElreath, R. \& Smirnov, O. Egalitarian motives in humans. \emph{Nature} \textbf{446}, 794--796 (2007).}

\leavevmode\vadjust pre{\hypertarget{ref-Bone2015}{}}%
\CSLLeftMargin{31. }%
\CSLRightInline{Bone, J. E. \& Raihani, N. J. Human punishment is motivated by both a desire for revenge and a desire for equality. \emph{Evolution and Human Behavior} \textbf{36}, 323--330 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-Walker2004}{}}%
\CSLLeftMargin{32. }%
\CSLRightInline{Walker, J. M. \& Halloran, M. A. Rewards and sanctions and the provision of public goods in one-shot settings. \emph{Experimental Economics} \textbf{7}, 235--247 (2004).}

\leavevmode\vadjust pre{\hypertarget{ref-Crockett2014}{}}%
\CSLLeftMargin{33. }%
\CSLRightInline{Crockett, M. J., Özdemir, Y. \& Fehr, E. \href{https://doi.org/10.1037/xge0000018}{The value of vengeance and the demand for deterrence.} \emph{Journal of Experimental Psychology: General} \textbf{143}, 2279--2286 (2014).}

\leavevmode\vadjust pre{\hypertarget{ref-Sylwester2013}{}}%
\CSLLeftMargin{34. }%
\CSLRightInline{Sylwester, K., Herrmann, B. \& Bryson, J. J. \href{https://doi.org/10.1037/npe0000009}{Homo homini lupus? Explaining antisocial punishment}. \emph{Journal of Neuroscience, Psychology, and Economics} \textbf{6}, 167--188 (2013).}

\leavevmode\vadjust pre{\hypertarget{ref-Fehr2004}{}}%
\CSLLeftMargin{35. }%
\CSLRightInline{Fehr, E. \& Fischbacher, U. Third-party punishment and social norms. \emph{Evolution and Human Behavior} \textbf{25}, 63--87 (2004).}

\leavevmode\vadjust pre{\hypertarget{ref-Johnson2009}{}}%
\CSLLeftMargin{36. }%
\CSLRightInline{Johnson, T., Dawes, C. T., Fowler, J. H., McElreath, R. \& Smirnov, O. \href{https://doi.org/10.1016/j.econlet.2009.01.003}{The role of egalitarian motives in altruistic punishment}. \emph{Economics Letters} \textbf{102}, 192--194 (2009).}

\leavevmode\vadjust pre{\hypertarget{ref-Fowler2005}{}}%
\CSLLeftMargin{37. }%
\CSLRightInline{Fowler, J. H., Johnson, T. \& Smirnov, O. \href{https://doi.org/10.1038/nature03256}{Egalitarian motive and altruistic punishment}. \emph{Nature} \textbf{433}, E1 (2005).}

\leavevmode\vadjust pre{\hypertarget{ref-Molleman2014}{}}%
\CSLLeftMargin{38. }%
\CSLRightInline{Molleman, L., Van den Berg, P. \& Weissing, F. J. \href{https://doi.org/10.1038/ncomms4570}{Consistent individual differences in human social learning strategies}. \emph{Nature Communications} \textbf{5}, 3570 (2014).}

\leavevmode\vadjust pre{\hypertarget{ref-Thielmann2020}{}}%
\CSLLeftMargin{39. }%
\CSLRightInline{Thielmann, I., Spadaro, G. \& Balliet, D. \href{https://doi.org/10.1037/bul0000217}{Personality and prosocial behavior: A theoretical framework and meta-analysis.} \emph{Psychological Bulletin} \textbf{146}, 30--90 (2020).}

\leavevmode\vadjust pre{\hypertarget{ref-Hofmann2018}{}}%
\CSLLeftMargin{40. }%
\CSLRightInline{Hofmann, W., Brandt, M. J., Wisneski, D. C., Rockenbach, B. \& Skitka, L. J. \href{https://doi.org/10.1177/0146167218775075}{Moral punishment in everyday life}. \emph{Personality and Social Psychology Bulletin} \textbf{44}, 1697--1711 (2018).}

\leavevmode\vadjust pre{\hypertarget{ref-Nisbett1977}{}}%
\CSLLeftMargin{41. }%
\CSLRightInline{Nisbett, R. E. \& Wilson, T. D. \href{https://doi.org/10.1037/0033-295X.84.3.231}{Telling more than we can know: Verbal reports on mental processes}. \emph{Psychological Review} \textbf{84}, 231--259 (1977).}

\leavevmode\vadjust pre{\hypertarget{ref-Barclay2006}{}}%
\CSLLeftMargin{42. }%
\CSLRightInline{Barclay, P. \href{https://doi.org/10.1016/j.evolhumbehav.2006.01.003}{Reputational benefits for altruistic punishment}. \emph{Evolution and Human Behavior} \textbf{27}, 325--344 (2006).}

\leavevmode\vadjust pre{\hypertarget{ref-Batistoni2022}{}}%
\CSLLeftMargin{43. }%
\CSLRightInline{Batistoni, T., Barclay, P. \& Raihani, N. J. \href{https://doi.org/10.1098/rspb.2021.1773}{Third-party punishers do not compete to be chosen as partners in an experimental game}. \emph{Proceedings of the Royal Society B} \textbf{289}, 20211773 (2022).}

\leavevmode\vadjust pre{\hypertarget{ref-Jordan2016}{}}%
\CSLLeftMargin{44. }%
\CSLRightInline{Jordan, J. J., Hoffman, M., Bloom, P. \& Rand, D. G. \href{https://doi.org/10.1038/nature16981}{Third-party punishment as a costly signal of trustworthiness}. \emph{Nature} \textbf{530}, 473--476 (2016).}

\leavevmode\vadjust pre{\hypertarget{ref-Jordan2017}{}}%
\CSLLeftMargin{45. }%
\CSLRightInline{Jordan, J. J. \& Rand, D. G. Third-party punishment as a costly signal of high continuation probabilities in repeated games. \emph{Journal of Theoretical Biology} \textbf{421}, 189--202 (2017).}

\leavevmode\vadjust pre{\hypertarget{ref-Jordan2020}{}}%
\CSLLeftMargin{46. }%
\CSLRightInline{Jordan, J. J. \& Rand, D. G. \href{https://doi.org/10.1037/pspi0000186}{Signaling when no one is watching: A reputation heuristics account of outrage and punishment in one-shot anonymous interactions}. \emph{Journal of Personality and Social Psychology} \textbf{118}, 57--88 (2020).}

\leavevmode\vadjust pre{\hypertarget{ref-Deutchman2021}{}}%
\CSLLeftMargin{47. }%
\CSLRightInline{Deutchman, P., Bračič, M., Raihani, N. J. \& McAuliffe, K. \href{https://doi.org/10.1016/j.evolhumbehav.2020.06.001}{Punishment is strongly motivated by revenge and weakly motivated by inequity aversion}. \emph{Evolution and Human Behavior} \textbf{42}, 12--20 (2021).}

\leavevmode\vadjust pre{\hypertarget{ref-Marczyk2017}{}}%
\CSLLeftMargin{48. }%
\CSLRightInline{Marczyk, J. \href{https://doi.org/10.1371/journal.pone.0171298}{Human punishment is not primarily motivated by inequality}. \emph{PLoS ONE} \textbf{12}, e0171298 (2017).}

\end{CSLReferences}

\endgroup

\newpage
\vspace*{60mm}

\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\renewcommand{\thefigure}{\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}

\hypertarget{supplementary-information}{%
\section{\texorpdfstring{\textbf{Supplementary Information}}{Supplementary Information}}\label{supplementary-information}}

\setcounter{page}{1}
\centering

\noindent \hspace*{10mm} \small Mapping the punishment strategy space \newline
\hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, Nichola Raihani\textsuperscript{1,2} \newline

\raggedright

\noindent \footnotesize \textsuperscript{1} School of Psychology, University of Auckland, Auckland, New Zealand \newline
\noindent \footnotesize \textsuperscript{2} Department of Experimental Psychology, University College London, London, United Kingdom \newline
\normalsize
\newpage

\hypertarget{supplementary-figures}{%
\subsection{Supplementary Figures}\label{supplementary-figures}}

\hypertarget{supplementary-tables}{%
\subsection{Supplementary Tables}\label{supplementary-tables}}

\hypertarget{supplementary-references}{%
\subsection{Supplementary References}\label{supplementary-references}}

\end{document}

### 2023-10-20 report.Rmd
starting writing manuscript

--- 2022-12-12 report.Rmd
@@
 
+We assume here that people are following a particular strategy exactly and are not comitting any errors or mistakes.
+However, it is possible that some participants intended to follow a particular strategy, but made a mistake on a particular decision.
+We repeat the counts and proportions in the table above, but include behavioural patterns that have one mistake from the exact strategy (note that the proportions will no longer sum to 1).
+
+```{r strategyCountsMistake1}
+kable(tar_read(tableStrategyCounts_Mistake1), booktabs = TRUE)
+```
+
+And including behavioural patterns that have one or two mistakes.
+
+```{r strategyCountsMistake2}
+kable(tar_read(tableStrategyCounts_Mistake2), booktabs = TRUE)
+```
+
+And finally, including behavioural patterns that have one, two, or three mistakes.
+
+```{r strategyCountsMistake3}
+kable(tar_read(tableStrategyCounts_Mistake3), booktabs = TRUE)
+```
+
 We can also look at the most common behavioural patterns across all six games (twelve punishment decisions in total).
-Below, we summarise the five most common patterns of behaviour as strings of twelve 0s and 1s, indicating whether participants did (1) or did not (0) punish in each game, with a brief explanation of this pattern and its frequency in the dataset.
+Below, we summarise the 25 most common patterns of behaviour as strings of twelve 0s and 1s, indicating whether participants did (1) or did not (0) punish in each game, with a brief explanation of this pattern and its frequency in the dataset.
 
-```{r patterns}
+```{r patternsUK}
 kable(tar_read(tablePatterns), booktabs = TRUE)
+```
+
+These behavioural patterns can be visualised, along with their frequency of usage, on a splits graph.
+This graph plots the distance between the difference strategies as proportional to the number of substitutions required to get from one to another.
+The top of the graph captures the less punitive strategies, and the strategies become more punitive towards the bottom of the graph.
+We only include behavioural patterns followed by at least three participants.
+
+```{r splitsGraph, fig.width=7, fig.height=7}
+tar_read(plotSplits)
 ```
@@
 We construct a model that contains our _apriori_ predictions for the different punishment strategies, and the feed the model our raw data to estimate the relative probabilities of following each strategy.
-In the model, we assume that participants sometimes make errors in converting their strategy into behaviour (5% error rate), which could explain why many of the participants were unable to be classified into a strategy type in our raw counts above.
+In the model, we assume that participants sometimes make errors in converting their strategy into behaviour (5% error rate), which could explain why many of the participants were unable to be classified into a strategy type in our raw counts and proportions above.
 The Bayesian model is fitted in the probabilistic programming language Stan.
 
-```{r post probs}
-P <- tar_read(post1)$alpha
-for (i in 1:nrow(P)) P[i,] <- rethinking::softmax(P[i,])
-anti <- P[,10]
+```{r model1, message=FALSE, error=FALSE, warning=FALSE, out.width="100%"}
+tar_read(plot1.1)
 ```
 
-In line with our raw counts, the model predicts that participants follow the "anti-punish" strategy with a probability of `r round(median(anti), 2)`, 95% credible interval [`r round(quantile(anti, 0.025), 2)` `r round(quantile(anti, 0.975), 2)`].
-We plot the posterior probabilities for the remaining punishment strategies below.
+The pattern is similar in both countries.
+Taking advantage of all available data, the model suggests that "anti-punish" is the most common strategy.
+Of the punishment strategies, "egalitarian" is the most common.
+All other strategies have median posterior probabilities less than 10%.
+"Competitive" and "antisocial" punishment strategies are the most unlikely.
 
-```{r model1, message=FALSE, out.width="80%", fig.align='center'}
-tar_read(plot1)
-```
-
-Taking advantage of all available data, the model suggests that "egalitarian" is the most common punishment strategy.
-All other strategies have median posterior probabilities less than 10%.
-In order of decreasing probability, the next most common punishment strategies are "norm-enforcing", "seek advantageous inequity", "deterrent", and "avoid disadvantageous inequity", although all are similarly likely.
-"Competitive" and "antisocial" punishment strategies are the most unlikely.
+Why is egalitarian the winner in this model, considering that it wasn't an obvious winner in the raw counts?
+One explanation is that many participants punished only in the third-party punishment game (see table above).
+The model does not include this pattern of behaviour as an explicit strategy.
+This pattern of behaviour is one substitution away from egalitarian, but two substitutions away from avoid DI, so it upweights egalitarian accordingly.
 
@@
 We then fit a series of models predicting the usage of different strategies from a variety of predictors.
-In each plot that follows, we show the posterior predictions from a model that includes a single variable as a predictor of all ten strategies simultaneously.
-The plots show the median regression lines with shaded 95% credible intervals.
-The model-estimated slopes for each strategy are included in each panel.
+In each plot that follows, we show the posterior slopes from several models that include different variables as predictors of all ten strategies in both countries.
+Each predictor is included separately, apart from SDO and RWA where we include both in the model simultaneously to control for one another.
 
-```{r plot2, fig.width=7, fig.height=4, fig.align='center', out.width="100%"}
-tar_read(plot2_Age)
-tar_read(plot2_SelfRate_1)
-tar_read(plot2_SelfRate_2)
-tar_read(plot2_SelfRate_3)
-tar_read(plot2_SelfRate_4)
-tar_read(plot2_SelfRate_5)
-tar_read(plot2_SelfRate_6)
-tar_read(plot2_SelfRate_7)
-tar_read(plot2_SelfRate_8)
-tar_read(plot2_SelfRate_9)
-tar_read(plot2_SelfRate_10)
-tar_read(plot2_SelfRate_11)
+## Demographics
+
+```{r echo=F}
+tar_read(plotDemsNoTPP)
 ```
+
+\newpage
+
+## Self-ratings
+
+```{r echo=F}
+tar_read(plotSlidersNoTPP)
+```
+
+\newpage
+
+## Big-6 personality and SVO
+
+```{r echo=F}
+tar_read(plotPersNoTPP)
+```
+
+\newpage
+
+## Politics and religion
+
+```{r echo=F}
+tar_read(plotPolRelNoTPP)
+```

### 2023-10-20 R/functions_modelFitting.R

@@ -78,8 +78,51 @@ fitModel2 <- function(d, compiledModel2, predictor, error = 0,
   return(out)
 }
 
+# fit stan model with two predictors
+fitModel3 <- function(d, compiledModel3, predictor1, predictor2, error = 0, 
+                      iter = 2000, warmup = 1000, chains = 4, cores = 4) {
+  # listwise deletion for predictors
+  d <- d[as.vector(!is.na(d[,predictor1])),]
+  d <- d[as.vector(!is.na(d[,predictor2])),]
+  # list for stan
+  dataList <-
+    list(
+      N = nrow(d),
+      Nc = length(unique(d$Country)),
+      countryID = ifelse(d$Country == "United Kingdom", 1L, 2L),
+      # standardise predictor for model fitting
+      pred1 = d %>% pull(predictor1) %>% scale() %>% as.numeric(),
+      pred2 = d %>% pull(predictor2) %>% scale() %>% as.numeric(),
+      pun1_1 = d$pun1_1,
+      pun1_2 = d$pun1_2,
+      pun2_1 = d$pun2_1,
+      pun2_2 = d$pun2_2,
+      pun3_1 = d$pun3_1,
+      pun3_2 = d$pun3_2,
+      pun4_1 = d$pun4_1,
+      pun4_2 = d$pun4_2,
+      pun5_1 = d$pun5_1,
+      pun5_2 = d$pun5_2,
+      pun6_1 = d$pun6_1,
+      pun6_2 = d$pun6_2,
+      error = error # assumed error rate
+    )
+  # fit model
+  out <- 
+    sampling(
+      compiledModel3,
+      data = dataList,
+      iter = iter,
+      warmup = warmup,
+      chains = chains,
+      cores = cores,
+      seed = 2113
+    )
+  return(out)
+}
+
 # fit stan model with categorical predictor
-fitModel3 <- function(d, compiledModel3, predictor, error = 0,
+fitModel4 <- function(d, compiledModel4, predictor, error = 0,
                       iter = 2000, warmup = 1000, chains = 4, cores = 4) {
   # listwise deletion for predictor
   d <- d[as.vector(!is.na(d[,predictor])),]
@@ -110,7 +153,7 @@ fitModel3 <- function(d, compiledModel3, predictor, error = 0,
   # fit model
   out <- 
     sampling(
-      compiledModel3,
+      compiledModel4,
       data = dataList,
[…]

### 2023-10-20 R/functions_modelResults.R

@@ -1,7 +1,7 @@
 # custom functions
 
 # plot model results
-plotModel1 <- function(post) {
+plotModel1.1 <- function(post) {
   # strategy vector
   strategies <- c("Competitive", "Avoid DI", "Egalitarian", "Seek AI",
                   "Retributive", "Deterrent", "Norm-enforcing", 
@@ -30,12 +30,12 @@ plotModel1 <- function(post) {
     ylab(NULL) +
     theme_classic()
   # save
-  ggsave(out, filename = "figures/modelResults/model1.pdf", height = 5, width = 7)
+  ggsave(out, filename = "figures/modelResults/noTPP/model1.pdf", height = 5, width = 7)
   return(out)
 }
 
 # plot effect of predictor on strategy usage
-plotModelPred <- function(d, post, pred, xlab, xBreaks, file) {
+plotModel2.1 <- function(d, post, pred, xlab, xBreaks, file) {
   # strategy vector
   strategies <- c("Competitive", "Avoid DI", "Egalitarian", "Seek AI",
                   "Retributive", "Deterrent", "Norm-enforcing", 
@@ -99,7 +99,7 @@ plotModelPred <- function(d, post, pred, xlab, xBreaks, file) {
 }
 
 # plot effect of categorical predictor on strategy usage
-plotModelPredCat <- function(d, post, pred, xlab, file) {
+plotModel4.1 <- function(d, post, pred, xlab, file) {
   # posterior predictions
   postPred <- tibble()
   # categories
@@ -156,7 +156,7 @@ plotModelPredCat <- function(d, post, pred, xlab, file) {
 }
 
 # plot model results with third-party strategy
-plotModel4 <- function(post) {
+plotModel1.2 <- function(post) {
   # strategy vector
   strategies <- c("Competitive", "Avoid DI", "Egalitarian", "Seek AI",
                   "Retributive", "Deterrent", "Norm-enforcing", 
@@ -185,8 +185,689 @@ plotModel4 <- function(post) {
     ylab(NULL) +
     theme_classic()
   # save
-  ggsave(out, filename = "figures/modelResults/model4.pdf", height = 5, width = 7)
+  ggsave(out, filename = "figures/modelResults/withTPP/model1.pdf", height = 5, width = 7)
   return(out)
 }
 
+# plot effect of predictor on strategy usage
+plotModel2.2 <- function(d, post, pred, xlab, xBreaks, file) {
+  # strategy vector
[…]

### 2023-10-20 R/functions_summariseData.R

@@ -426,7 +426,7 @@ makePatternsTable <- function(d) {
 # plot splits graph of behavioural patterns
 plotSplitsGraph <- function(d) {
   d <-
-    tar_read(d) %>%
+    d %>%
     # subset to behavioural decisions
     dplyr::select(starts_with("pun"), strategy) %>%
     # unite behavioural decisions
@@ -445,9 +445,15 @@ plotSplitsGraph <- function(d) {
       .groups = "drop"
     ) %>%
     # add tpp strategy
-    mutate(strategy = ifelse(pattern == "000000000010", "Third-party", strategy)) %>%
-    # keep only behavioural patterns with certain number of counts
-    filter(count >= 2)
+    #mutate(strategy = ifelse(pattern == "000000000010", "Third-party", strategy)) %>%
+    # add antisocial strategy
+    bind_rows(
+      tibble(pun1_1 = 0, pun1_2 = 1, pun2_1 = 0, pun2_2 = 1, pun3_1 = 0, pun3_2 = 1,
+             pun4_1 = 0, pun4_2 = 1, pun5_1 = 0, pun5_2 = 1, pun6_1 = 0, pun6_2 = 1,
+             count = 0, pattern = "010101010101", strategy = "Antisocial")
+    ) %>%
+    # keep only other behavioural patterns with certain number of counts
+    filter(strategy != "N/A" | (strategy == "N/A" & count >= 3))
   # n is number of unique patterns of behaviour
   n <- nrow(d)
   # calculate Hamming distance matrix for behavioural patterns
@@ -467,7 +473,7 @@ plotSplitsGraph <- function(d) {
   p <- 
     ggsplitnet(
       net,
-      colour = "grey90",
+      colour = "grey70",
       linewidth = 0.05
     )
   # labels for strategies
@@ -481,7 +487,8 @@ plotSplitsGraph <- function(d) {
       "Deterrent"                      = "Deterrent\n(101000101000)",
       "Norm-enforcing"                 = "Norm-enforcing\n(101000101010)",
       "Competitive"                    = "Competitive\n(111111001111)",
-      "Third-party"                    = "Third-party\n(000000000010)"
+      #"Third-party"                    = "Third-party\n(000000000010)",
+      "Antisocial"                     = "Antisocial\n(010101010101)"
     )
   # additional data for plot (counts, labels)
   pd <- 
@@ -510,8 +517,9 @@ plotSplitsGraph <- function(d) {
       data = pd,
[…]

### 2023-10-20 _targets.R

@@ -10,7 +10,7 @@ source("R/functions_summariseData.R")
 source("R/functions_modelResults.R")
 tar_option_set(
   packages = c("ape", "cowplot", "ggcorrplot", "ggrepel", 
-               "ggridges", "knitr", "ltm", "phangorn", 
+               "ggridges", "knitr", "ltm", "papaja", "phangorn", 
                "rethinking", "rstan", "tanggle", "tidybayes", 
                "tidytext", "tidyverse", "wordcloud"),
   deployment = "main"
@@ -49,26 +49,32 @@ list(
   #### Load and compile Stan models ####
   
   # model files
-  tar_target(fileModel1, "stan/noTPP/punishStrat_noTPP_noPredictor.stan", format = "file"),
-  tar_target(fileModel2, "stan/noTPP/punishStrat_noTPP_withPredictor.stan", format = "file"),
-  tar_target(fileModel3, "stan/noTPP/punishStrat_noTPP_withCategoricalPredictor.stan", format = "file"),
-  tar_target(fileModel4, "stan/withTPP/punishStrat_withTPP_noPredictor.stan", format = "file"),
+  tar_target(fileModel1.1, "stan/noTPP/punishStrat_noTPP_noPredictor.stan", format = "file"),
+  tar_target(fileModel2.1, "stan/noTPP/punishStrat_noTPP_withPredictor.stan", format = "file"),
+  tar_target(fileModel3.1, "stan/noTPP/punishStrat_noTPP_withTwoPredictors.stan", format = "file"),
+  tar_target(fileModel4.1, "stan/noTPP/punishStrat_noTPP_withCategoricalPredictor.stan", format = "file"),
+  tar_target(fileModel1.2, "stan/withTPP/punishStrat_withTPP_noPredictor.stan", format = "file"),
+  tar_target(fileModel2.2, "stan/withTPP/punishStrat_withTPP_withPredictor.stan", format = "file"),
+  tar_target(fileModel3.2, "stan/withTPP/punishStrat_withTPP_withTwoPredictors.stan", format = "file"),
+  tar_target(fileModel4.2, "stan/withTPP/punishStrat_withTPP_withCategoricalPredictor.stan", format = "file"),
   # compile models
-  tar_target(compiledModel1, stan_model(fileModel1), deployment = "worker"),
-  tar_target(compiledModel2, stan_model(fileModel2), deployment = "worker"),
-  tar_target(compiledModel3, stan_model(fileModel3), deployment = "worker"),
[…]

### 2023-11-24 appendix.Rmd
finished manuscript first draft

@@
 \renewcommand{\tablename}{Supplementary Table}
-\renewcommand{\thefigure}{\arabic{figure}} \setcounter{figure}{0}
-\renewcommand{\thetable}{\arabic{table}} \setcounter{table}{0}
+\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
+\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
 \renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}
 
-# **Supplementary Information**
+# **Supplementary Material**
 
@@
 \centering
-\noindent \hspace*{10mm} \small Mapping the punishment strategy space \newline
+\noindent \hspace*{10mm} \small Why do people punish? Evidence for a range of strategic concerns \newline
 \hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, Nichola Raihani\textsuperscript{1,2} \newline
@@
 
+(ref:plotModel1aCaption) *Posterior estimates of the probabilities of following
+different punishment strategies from the Bayesian latent state model fitted to
+the full dataset without pre-registered exclusions.* The model assumes an 
+implementation error rate of 5%. Points represent posterior medians, line ranges 
+represent 50% and 95% credible intervals.
+
+```{r plotModel1a, fig.width=7, fig.height=5, fig.cap="(ref:plotModel1aCaption)", warning=FALSE}
+tar_read(plot1.1)
+```
+
+\newpage
+
+(ref:plotAllDems2Caption) *Posterior slopes from models including age, 
+socio-economic status, gender, and student status, fitted to the subsetted 
+dataset with pre-registered exclusions.* Each row represents a separate model. 
+Points represent posterior medians, line ranges represent 95% credible intervals.
+
+```{r plotAllDems2, fig.width=7.5, fig.height=3, fig.cap="(ref:plotAllDems2Caption)", warning=FALSE}
+tar_read(plotAllDems2)
+```
+
+\newpage
+
+(ref:plotAllDems1Caption) *Posterior slopes from models including age, 
+socio-economic status, gender, and student status, fitted to the full dataset 
+without pre-registered exclusions.* Each row represents a separate model. Points 
+represent posterior medians, line ranges represent 95% credible intervals.
+
+```{r plotAllDems1, fig.width=7.5, fig.height=3, fig.cap="(ref:plotAllDems1Caption)", warning=FALSE}
+tar_read(plotAllDems1)
+```
+
+\newpage
+
+(ref:plotAllPers1Caption) *Posterior slopes from models including Big-6 
+personality dimensions and Social Value Orientation, fitted to the full dataset 
+without pre-registered exclusions.* Each row represents a separate model. Points 
+represent posterior medians, line ranges represent 95% credible intervals.
+
+```{r plotAllPers1, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotAllPers1Caption)", warning=FALSE}
+tar_read(plotAllPers1)
+```
+
+\newpage
+
+(ref:plotAllPolRel1Caption) *Posterior slopes from models including political 
+ideology, views about social inequality, and religiosity, fitted to the full 
+dataset without pre-registered exclusions.* Each row represents a separate model 
+aside from Social Dominance Orientation and Right Wing Authoritarianism, which 
+control for one another within the same model. Points represent posterior 
+medians, line ranges represent 95% credible intervals.
+
+```{r plotAllPolRel1, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotAllPolRel1Caption)", warning=FALSE}
+tar_read(plotAllPolRel1)
+```
+
+\newpage
+
+(ref:plotSliders1Caption) *Boxplots showing the distribution of responses to each
+self-report question about the reasons for participants' behaviour in the games.*
+Boxplots represent medians and interquartile ranges.
+
+```{r plotSliders1, fig.width=7, fig.height=4, fig.cap="(ref:plotSliders1Caption)", warning=FALSE}
+tar_read(plotSliders1)
+```
+
+\newpage
+
+(ref:plotSliders2Caption) *Boxplots showing the distribution of responses to each
+self-report question about the reasons for participants' behaviour in the games,
+presented as deviations from participants' average rating across all questions.*
+Boxplots represent medians and interquartile ranges.
+
+```{r plotSliders2, fig.width=7, fig.height=4, fig.cap="(ref:plotSliders2Caption)", warning=FALSE}
+tar_read(plotSliders2)
+```
+
+\newpage
+
+(ref:plotAllSliders1Caption) *Posterior slopes from models including 
+self-reported strategy usage, fitted to the full dataset without pre-registered 
+exclusions.* Each row represents a separate model. Highlighted estimates 
+represent combinations where the self-report slider matched the behavioural 
+strategy. Each strategy had an associated self-report slider except for the 
+competitive strategy. Points represent posterior medians, line ranges represent 
+95% credible intervals.
+
+```{r plotAllSliders1, fig.width=7.5, fig.height=4, fig.cap="(ref:plotAllSliders1Caption)", warning=FALSE}
+tar_read(plotAllSliders1)
+```
+
+\newpage
+
+(ref:plotSampleUKCaption) *Sample characteristics in the United Kingdom.*
+
+```{r plotSampleUK, fig.width=7, fig.height=5, fig.cap="(ref:plotSampleUKCaption)", warning=FALSE}
+tar_read(plotSample_UK)
+```
+
+\newpage
+
+(ref:plotSampleUSCaption) *Sample characteristics in the United States.*
+
+```{r plotSampleUS, fig.width=7, fig.height=5, fig.cap="(ref:plotSampleUSCaption)", warning=FALSE}
+tar_read(plotSample_US)
+```
+
+\newpage
+
+(ref:plotTraceCaption) *Trace plots for parameter values from the Bayesian
+latent state model fitted to data with exclusions.*
+
+```{r plotTrace, fig.width=7, fig.height=6, fig.cap="(ref:plotTraceCaption)", warning=FALSE}
+tar_read(plotTrace1.2)
+```
+
+\newpage
+
+(ref:plotSimCaption) *Results of Bayesian latent state model fitted to simulated
+data (n = 100) with known strategy frequencies in the population.* Blue points
+represent known strategy frequencies, grey densities represent posterior
+estimates of strategy frequencies.
+
+```{r plotSim, fig.width=6, fig.height=5, fig.cap="(ref:plotSimCaption)", warning=FALSE}
+tar_read(plotSim1)
+```
+
+\newpage
+
 ## Supplementary Tables
 
-## Supplementary References
+(ref:tablePatternsCaption) Counts and proportions of the 25 most common patterns
+of punitive behaviour across all twelve decisions, split by country. *Binary 
+strings represent punishment (1) or no punishment (0) in each decision, aligning 
+with the order of game decision columns in Table \@ref(tab:tableStrategies).*
+
+```{r tablePatterns}
+apa_table(
+  tar_read(tablePatterns),
+  escape = FALSE,
+  booktabs = TRUE,
+  align = c(rep("l", 2), rep("c", 4)),
+  placement = "H",
+  col_spanners = list(
+    "\\thead{United Kingdom \\\\ (N = 1014)}" = 3:4,
+    "\\thead{United States \\\\ (N = 996)}" = 5:6
+    ),
+  caption = "(ref:tablePatternsCaption)",
+  font_size = "footnotesize"
+)
+```
+
+\newpage
+
+(ref:tableSliderWordingsCaption) Wordings for 11 self-report slider questions
+asking participants to report the reasons for their behaviour in the six games.
+*Participants were prompted with the following text: "We would now like you to 
+answer a few questions about your main motivation in the games. Please answer 
+truthfully - there is no right or wrong answer and your first answer is probably 
+best. Please rate the extent to which the following statements apply to your 
+decisions to reduce or not to reduce other players' bonuses in the games."*
+
+```{r tableSliderWordings}
+apa_table(
+  tar_read(tableSliderWordings),
+  caption = "(ref:tableSliderWordingsCaption)",
+  booktabs = TRUE,
+  placement = "H",
+  font_size = "footnotesize"
+)
+```
+
+\newpage
+
+(ref:tableSurveyWordingsCaption) Wordings for survey questions in the study.
+
+```{r tableSurveyWordings}
+apa_table(
+  tar_read(tableSurveyWordings),
+  landscape = TRUE,
+  longtable = TRUE,
+  align = c("p{6cm}", "p{11cm}", "p{3cm}"),
+  caption = "(ref:tableSurveyWordingsCaption)",
+  booktabs = TRUE,
+  placement = "H",
+  fontsize = "scriptsize"
+)
+```
+
+\newpage
+
+(ref:tableCompCaption) Proportions of correct answers to comprehension 
+questions for all six economic games, split by country.
+
+```{r tableComp}
+apa_table(
+  tar_read(tableComp),
+  caption = "(ref:tableCompCaption)",
+  booktabs = TRUE,
+  placement = "H"
+)
+```

### 2023-11-24 manuscript.Rmd
finished manuscript first draft

@@
 ---
-title             : "Mapping the punishment strategy space"
+title             : "Why do people punish? Evidence for a range of strategic concerns"
 shorttitle        : "Punishment strategies"
@@
 abstract: |
- x
+ Costly punishment is thought to be one of the key mechanisms sustaining cooperation in humans.
+However, the motives for punitive behaviour remain unclear.
+Punishment is often assumed to be motivated by a desire to convert free riders into cooperators, but it is also consistent with a host of other functions, such as levelling payoffs or increasing one’s relative position.
+We used a suite of six economic games to tease apart the different motives for punishment.
+Across representative samples from the United Kingdom and the United States, we estimated the frequency of different punishment strategies in the population, finding that egalitarian motives for punishment are more common than behaviour-change motives.
+Moreover, different punishment strategies were differentially predicted by personality, social preferences, political ideology, and religious views.
+Self-reports of behaviour in the games suggested that people have some degree of insight into their punishment strategy.
+These findings highlight the multipurpose nature of human punishment.
 
-keywords          : "x"
-wordcount         : "xxxx words"
+keywords          : "punishment; cooperation; economic games"
+wordcount         : "5185 words"
 bibliography      : ["r-references.bib"]
@@
   - \usepackage{makecell}
-  - \renewcommand\theadfont{}
   - \raggedbottom
@@
 
+```{r options, echo=FALSE}
+opts_chunk$set(echo = FALSE)
+```
+
 \linenumbers
@@
 
-Humans cooperate on a scale unparalleled in the animal kingdom.
-One mechanism thought to sustain this level of cooperation is costly punishment, whereby individuals harm others at a personal cost[@CluttonBrock1995], ostensibly encouraging cooperative behaviour from the target (or bystanders[@DosSantos2011;@DosSantos2013;@Raihani2015]) in the future.
-Punishment therefore offers a route to maintaining or increasing cooperation by changing the payoff structure of social interactions such that it no longer pays to cheat or exploit social partners[@CluttonBrock1995;@Ostrom1990].
+Humans cooperate on a scale that is unparalleled in the animal kingdom. One 
+mechanism thought to sustain this level of cooperation is costly punishment, 
+whereby individuals harm others at a personal cost[@CluttonBrock1995], 
+ostensibly encouraging cooperative behaviour from the target (or bystanders [@DosSantos2011;@DosSantos2013;@Raihani2015]) in the future. Punishment 
+therefore offers a route to maintaining or increasing cooperation by changing 
+the payoff structure of social interactions such that it no longer pays to cheat 
+or exploit social partners[@CluttonBrock1995;@Ostrom1990].
 
-In humans, many studies of punishment have been carried out in laboratory settings using economic games[@Balliet2011;@Balliet2013;@Chaudhuri2011;@Dreber2008;@Fehr2000;@Fehr2002;@Henrich2006;@Nikiforakis2008;@Raihani2012a;@Raihani2019].
-In these games, participants are usually given a sum of money that they can use to invest in collective action or to help others.
-Alternatively, participants can 'cheat' by keeping the money for themselves or by exploiting the contributions of others.
-Punishment is introduced into such games by giving participants the option to pay a small 'fee' to impose a greater 'fine' on their co-players.
-Several lines of experimental evidence indicate that people use this punishment option[@Henrich2006], that they enjoy punishing[@DeQuervain2004], and that they frequently, though not always[@Herrmann2008], punish cheating or exploitative co-players[@Fehr2000;@Fehr2002].
+In humans, many studies of punishment have been carried out in laboratory 
+settings using economic games[@Balliet2011;@Balliet2013;@Chaudhuri2011;
+@Dreber2008;@Fehr2000;@Fehr2002;@Henrich2006;@Nikiforakis2008a;@Raihani2012a;
+@Raihani2019]. In these games, participants are usually given a sum of money 
+that they can use to invest in collective action or to help others. 
+Alternatively, participants can 'cheat' by keeping the money for themselves or 
+by exploiting the contributions of others. Punishment is introduced into such 
+games by giving participants the option to pay a small 'fee' to impose a greater 
+'fine' on their co-players. Several lines of experimental evidence indicate that 
+people use this punishment option[@Henrich2006], that they enjoy 
+punishing[@DeQuervain2004], and that they frequently, though not 
+always[@Herrmann2008], punish cheating or exploitative 
+co-players[@Fehr2000;@Fehr2002].
 
-Evidence from these experiments suggests that the threat of costly punishment plays an important role in promoting human cooperation.
-People tend to cooperate more in games where punishment is possible compared to those where it is not[@Balliet2011;@Balliet2013;@Raihani2019].
-The effect that the threat of punishment has on cooperation is also evident in the higher contributions typically observed in the Ultimatum Game (where punishment is possible) compared to the structurally-similar Dictator Game (where it is not)[@Camerer2003].
-This typical cooperation-enhancing effect of punishment has also been observed across societies[@Balliet2013], leading some to suggest that costly punishment has played a key role in the cultural evolution of cooperation in humans[@Bowles2013;@Boyd2003;@Chudek2011;@Henrich2017].
+Evidence from these experiments suggests that the threat of costly punishment 
+plays an important role in promoting human cooperation. People tend to cooperate 
+more in games where punishment is possible compared to those where it is not
+[@Balliet2011;@Balliet2013;@Raihani2019]. The effect that the threat of 
+punishment has on cooperation is also evident in the higher contributions 
+typically observed in the Ultimatum Game (where punishment is possible) compared 
+to the structurally-similar Dictator Game (where it is not)[@Camerer2003]. This 
+typical cooperation-enhancing effect of punishment has also been observed across 
+societies[@Balliet2013], leading some to suggest that costly punishment has 
+played a key role in the cultural evolution of cooperation in humans
+[@Bowles2013;@Boyd2003;@Chudek2011;@Henrich2017].
 
-Nevertheless, it remains unclear whether individuals playing economic games use punishment as a behaviour-change tool to enforce cooperation or as a means to achieve other ends.
-Some have argued that punishment is primarily used to shape future behaviour, either to deter personal harm[@Delton2017;@Dreber2008;@DosSantos2013] or to uphold normative standards of cooperative behaviour[@Boyd2003;@Chudek2011;@Fehr2018;@Mathew2011;@Mathew2014;@Richerson2016].
-But while the _threat_ of punishment can have a cooperation-enhancing effect, the _enactment_ of this punishment seldom changes the future behaviour of deviant targets[@Raihani2019].
-This calls into question whether punishment primarily operates as a behaviour-change tool or whether it is used to achieve other goals.
+Nevertheless, it remains unclear whether individuals playing economic games use
+punishment as a behaviour-change tool to enforce cooperation or as a means to 
+achieve other ends. Some have argued that punishment is primarily used to shape 
+future behaviour, either to deter personal harm[@Delton2017;@Dreber2008;
+@DosSantos2013] or to uphold normative standards of cooperative behaviour 
+[@Boyd2003;@Chudek2011;@Fehr2018;@Mathew2011;@Mathew2014;@Richerson2016]. 
+But while the _threat_ of punishment can have a cooperation-enhancing effect, 
+the _enactment_ of this punishment does not consistently deter targets from 
+cheating in the future[@Raihani2019]. This calls into question whether 
+punishment primarily operates as a behaviour-change tool or whether it is used 
+to achieve other goals.
 
-Beyond behaviour-shaping concerns, there are a host of other reasons that people may want to punish in economic games.
-Punishers might be motivated by a desire for retribution rather than deterrence, punishing in proportion to the amount of harm that was personally caused[@Carlsmith2002].
-Punishment might be driven by concerns about relative payoffs, such as disadvantageous inequity aversion (i.e., avoiding having less than others[@Raihani2012b;@Raihani2019]) and/or general egalitarian preferences (i.e., wanting all participants to receive the same payoffs[@Dawes2007]).
-Such concerns about relative payoffs may be activated when participants earn less than cheats in economic games or when there are income disparities in these settings.
-People might also use punishment for competitive purposes, seeking advantageous inequity for themselves (i.e., having more than others) and/or improving their relative position[@Raihani2019].
+Beyond behaviour-shaping concerns, there are a host of other reasons why people 
+may want to punish in economic games. Punishers might be motivated by a desire 
+for retribution rather than deterrence, punishing in proportion to the amount 
+of harm that was personally caused[@Carlsmith2002]. Punishment might be driven 
+by concerns about relative payoffs, such as disadvantageous inequity aversion 
+(i.e., avoiding having less than others[@Raihani2012b;@Raihani2019]) and/or 
+general egalitarian preferences (i.e., wanting all participants to receive the 
+same payoffs[@Dawes2007]). Such concerns about relative payoffs may be activated 
+when participants earn less than cheats in economic games or when there are 
+income disparities in these settings. People might also use punishment for 
+competitive purposes, seeking advantageous inequity for themselves (i.e., having 
+more than others) and/or improving their relative position[@Raihani2019].
 
-Common economic game designs have been unable to tease apart these different motives for punishment because participants who interact with cheats in these games experience both losses _and_ lower relative payoffs.
-The typical 1:3 fee-to-fine ratio of punishment in economic games compounds this issue.
-With this setup, people can simultaneously use punishment to reciprocate losses, to deter others from cheating, and to reduce or reverse disparities in payoffs between themselves and targets.
-To add to this complexity, it is evident that people use punishment in seemingly disparate ways: punishing when no behaviour change is possible, such as in one-shot games[@Bone2015;@Henrich2006;@Raihani2012b;@Walker2004] or in games where the target never learns about the punishment[@Crockett2014]; punishing those who did not cheat or who over-contributed to collective action (antisocial punishment[@Herrmann2008;@Sylwester2013]); punishing in scenarios where they were not personally harmed (third-party punishment[@Fehr2004]); and punishing in scenarios where disparities in payoffs did not arise from participants' actions[@Dawes2007;@Johnson2009;@Fowler2005].
+Common economic game designs have been unable to tease apart these different 
+motives for punishment because participants who interact with cheats in these 
+games experience both losses _and_ lower relative payoffs. The typical 1:3 
+fee-fine ratio of punishment in economic games compounds this issue. With 
+this setup, people can simultaneously use punishment to reciprocate losses, to 
+deter others from cheating, and to reduce or reverse disparities in payoffs 
+between themselves and targets. To add to this complexity, it is evident that 
+people use punishment in seemingly disparate ways: punishing when no behaviour 
+change is possible, such as in one-shot games[@Bone2015;@Henrich2006;
+@Raihani2012b;@Walker2004], on the very last round of repeated games
+[@Barclay2016], or in games where the target never learns about the 
+punishment[@Crockett2014]; punishing those who did not cheat or who 
+over-contributed to collective action (antisocial punishment
+[@Herrmann2008;@Sylwester2013]); punishing in scenarios where they were not 
+personally harmed (third-party punishment[@Fehr2004]); and punishing in 
+scenarios where disparities in payoffs did not arise from participants' 
+actions[@Dawes2007;@Johnson2009;@Fowler2005].
 
-The general conclusion from this research is that there is no one unifying function of costly punishment in humans.
-Instead, punishment should be thought of as a flexible behavioural tool that serves a variety of functions that are not mutually exclusive[@Raihani2019].
-Due to its multipurpose nature, we should therefore expect variation in punishment strategies in the population, much like the observed variation in social learning strategies[@Molleman2014].
-Some individuals may use punishment as a behaviour shaping tool, for example, while others may use it to reduce or reverse payoff differentials.
+The general conclusion from this research is that there is no one unifying 
+function of costly punishment in humans. Instead, punishment should be thought 
+of as a flexible behavioural tool that serves a variety of functions that are 
+not mutually exclusive[@Raihani2019]. Due to its multipurpose nature, we should 
+therefore expect variation in punishment strategies in the population, much like 
+the observed variation in social learning strategies[@Molleman2014]. Some 
+individuals may use punishment as a behaviour shaping tool, for example, while 
+others may use it to reduce or reverse payoff differentials.
 
-This insight raises several underexplored questions.
-First, which punishment strategies are more frequent in human populations?
-Second, what traits predict adherence to a particular strategy?
-Previous work has reported that personality is related to cooperative behaviour[@Thielmann2020] and demographics, political ideology, and religiosity are related to punitive behaviour[@Hofmann2018], but no research has related these variables to specific punishment strategies.
-Third, do people have insight into their own punishment strategy?
-Previous work has argued that people are often unaware of the underlying function of their punitive behaviour, yet they feel compelled to enact it anyway[@Carlsmith2002;@Nisbett1977].
+This insight raises several underexplored questions. First, which punishment 
+strategies are more frequent in human populations? Second, what traits predict 
+adherence to a particular strategy? Previous work has reported that personality 
+is related to cooperative behaviour[@Thielmann2020] and demographics, political 
+ideology, and religiosity are related to punitive behaviour[@Hofmann2018], but 
+no research has related these variables to specific punishment strategies. 
+Third, do people have insight into their own punishment strategy? Previous work 
+has argued that people are often unaware of the underlying function of their 
+punitive behaviour, yet they feel compelled to enact it anyway
+[@Carlsmith2002;@Nisbett1977].
 
-Here, we aim to delineate the many possible punishment strategies by asking whether people punish in a manner consistent with a specific strategy and, if so, what other characteristics (personality, social preferences, political orientation) predict the use of different punishment strategies.
-Table \@ref(tab:tableStrategies) summarises the potential functions for costly punishment in economic games and the behavioural strategies they predict.
-Note that we do not include reputational functions of punishment in this table, such as signalling trustworthiness[@Barclay2006;@Batistoni2022;@Jordan2016;@Jordan2017;@Raihani2015], because our focus is on punishment strategies in anonymous economic games without reputational incentives (but see ref[@Jordan2020]).
+Here, we aim to delineate nine possible punishment strategies by asking 
+whether people punish in a manner consistent with a specific strategy and, if 
+so, what other characteristics (personality, social preferences, political 
+orientation) predict the use of different punishment strategies. Table 
+\@ref(tab:tableStrategies) summarises the potential functions for costly 
+punishment in the economic games that we considered, and the behavioural 
+strategies they predict. Note that Table \@ref(tab:tableStrategies) is not 
+an exhaustive list of all possible punishment strategies: we do not include 
+reputational functions of punishment in this table, such as signalling 
+trustworthiness[@Barclay2006;@Batistoni2022;@Jordan2016;@Jordan2017;@Raihani2015], 
+because our focus is on punishment strategies in anonymous economic games without 
+reputational incentives (but see ref[@Jordan2020]).
 
-(ref:tableStrategiesCaption) xxx
+Building on previous designs[@Bone2015;@Deutchman2021;@Marczyk2017;
+@Raihani2012b], we employ a suite of one-shot economic games where individuals 
+are given the opportunity to punish targets at a personal cost (Figure 
+\@ref(fig:plotGames)). In each game, targets either steal from another 
+individual or do nothing. Representative samples of participants from the United 
+Kingdom (*n* = `r sum(tar_read(dExc)$Country == "United Kingdom")`) and the 
+United States (*n* = `r sum(tar_read(dExc)$Country == "United States")`) 
+completed all six games on the online platform Prolific. We carefully designed 
+the suite of games to tease apart the proposed punishment strategies in Table 
+\@ref(tab:tableStrategies), such that each strategy predicts a different pattern 
+of behaviour across all the games (see Methods for more detail about the six 
+games). We use the resulting behavioural patterns to discern which punishment 
+strategy participants are employing. We then combine these behavioural patterns 
+with data on demographics, personality, social preferences, political ideology,
+religiosity, and self-reported strategy usage.
 
-```{r tableStrategies, echo=FALSE}
+(ref:plotGamesCaption) *Visual summary of the six economic games.* In all
+games, Player 2 either steals £0.20 from Player 1 (the focal player) or does 
+nothing. Player 1 is then given the option to punish by paying a certain amount 
+of money to remove money from Player 2 (this money is destroyed). The six games
+are variants on this general setup, creating situations where (A) Player 2 is
+still worse off by stealing, (B) Player 2 creates equality by stealing, (C) the
+computer "decides" whether Player 2 steals, (D) the fee-fine ratio is 1:1, (E)
+Player 2 is better off by stealing, and (F) Player 2 steals instead from a 
+third-party.
+
+```{r plotGames, out.width="100%", fig.cap="(ref:plotGamesCaption)"}
+include_graphics("images/games.png")
+```
+
+(ref:tableStrategiesCaption) \scriptsize Summary of the different functions for 
+punishment and the behavioural strategies they predict. *Games A-F are 
+the games employed in the current study (see Methods for more details). In each 
+of the six games, participants are given the opportunity to punish players who 
+"steal" and those who do not, meaning that participants make twelve punishment 
+decisions in total. Each behavioural strategy implies a unique pattern of 
+punishment across all decisions. Green ticks reflect decisions to punish, red 
+crosses reflect decisions to not punish. In column headers, payoffs at the first 
+stage (above) and the second stage (below) are denoted as P1-P2 (or P2-P3 [P1] 
+for Game F) where participants take the role of P1 and P2 is the target of 
+punishment. AI = advantageous inequity, DI = disadvantageous inequity.*
+
+```{r tableStrategies}
 apa_table(
@@
   col_spanners = list(
-    "\\thead{No DI 1 \\\\ 70-10}"   = 3:4,
-    "\\thead{No DI 2 \\\\ 70-30}"   = 5:6,
-    "\\thead{No DI 3 \\\\ (Computer) \\\\ 70-30}"   = 7:8,
-    "\\thead{No DI 4 \\\\ (1:1 Fee-Fine) \\\\ 70-30}"   = 9:10,
-    "\\thead{DI \\\\ 70-50}"        = 11:12,
-    "\\thead{3PP \\\\ 70-70 [100]}" = 13:14
+    "\\thead{\\scriptsize Game A \\\\ (AI) \\\\ 70-10}"   = 3:4,
+    "\\thead{\\scriptsize Game B \\\\ (Equal) \\\\ 70-30}"   = 5:6,
+    "\\thead{\\scriptsize Game C \\\\ (Computer) \\\\ 70-30}"   = 7:8,
+    "\\thead{\\scriptsize Game D \\\\ (1:1 Fee-Fine) \\\\ 70-30}"   = 9:10,
+    "\\thead{\\scriptsize Game E \\\\ (DI) \\\\ 70-50}"        = 11:12,
+    "\\thead{\\scriptsize Game F \\\\ (Third-Party) \\\\ 70-70 [100]}" = 13:14
     ),
@@
 
-Building on previous designs[@Bone2015;@Deutchman2021;@Marczyk2017;@Raihani2012b], we employ a suite of one-shot economic games where individuals are given the opportunity to punish targets at a personal cost.
-We carefully designed this suite of games to tease apart the proposed punishment strategies in Table 1, such that each strategy predicts a different pattern of behaviour across all the games.
-We use the resulting behavioural patterns to discern which punishment strategy participants are employing.
-We then combine these behavioural patterns with data on demographics, personality, political ideology, religiosity, and self-reported strategy usage.
-
 # Results
 
-x
+```{r MLMs}
+mlm1 <- coef(summary(tar_read(mlm1)))
+mlm2 <- coef(summary(tar_read(mlm2)))
+```
+
+The overall pattern of punitive behaviour in the six economic games was in line 
+with previous research and very similar across both countries (Figure 
+\@ref(fig:plotDecisions)). Participants were generally more likely to punish 
+targets who stole from another individual compared to targets who did not steal 
+(multilevel logistic regression; _b_ = 
+`r printnum(mlm1["otherTook","Estimate"])`, standard error = 
+`r printnum(mlm1["otherTook","Std. Error"])`, _p_ 
+`r printp(mlm1["otherTook","Pr(>|z|)"])`). Participants were also more likely 
+to punish when targets' stealing behaviour generated inequalities, 
+specifically in Games E and F (_b_ = 
+`r printnum(mlm2["DI","Estimate"])`, SE = `r printnum(mlm2["DI","Std. Error"])`, 
+_p_ `r printp(mlm2["DI","Pr(>|z|)"])`).
+
+(ref:plotDecisionsCaption) *Overall pattern of punitive behaviour across all six 
+economic games, split by country.* AI = advantageous inequity, DI = 
+disadvantageous inequity.
+
+```{r plotDecisions, fig.width=8, fig.height=5, fig.cap="(ref:plotDecisionsCaption)"}
+tar_read(plotDecisions)
+```
+
+We classified participants into a particular strategy if their behaviour across 
+all twelve decisions matched our behavioural predictions shown in Table 
+\@ref(tab:tableStrategies) exactly. Table \@ref(tab:tableStrategyCounts) shows 
+the proportion of participants following each strategy, with N/A used to 
+represent participants who did not fit exactly into any particular strategy type.
+Overall, `r printnum(mean(tar_read(dExc)$strategy != "N/A")*100, digits = 0)`%
+of our participants could be classified exactly into one of the strategies. The
+most common strategy in both countries was to never punish across any of the
+games. The next most common strategies were those that care about
+minimising payoff differences (avoid disadvantageous inequity, egalitarian). 
+Less common were the behaviour-shaping strategies (deterrent, norm-enforcing),
+the retributive strategy, and the competitive strategies (seek AI, competitive). 
+Although participants often punished targets who did not steal 
+in the six games (Figure \@ref(fig:plotDecisions)), no participants followed the 
+antisocial strategy by exclusively punishing targets who did not steal across 
+_all_ games.
+
+(ref:tableStrategyCountsCaption) Counts and proportions of participants 
+following each punishment strategy exactly, split by country. N/A implies that
+participants were unable to be classified exactly into any of the punishment
+strategies.
+
+```{r tableStrategyCounts}
+apa_table(
+  tar_read(tableStrategyCounts),
+  escape = FALSE,
+  align = c("l", rep("c", 4)),
+  col_spanners = list(
+    "\\thead{United Kingdom \\\\ (N = 1014)}" = 2:3,
+    "\\thead{United States \\\\ (N = 996)}" = 4:5
+    ),
+  caption = "(ref:tableStrategyCountsCaption)"
+)
+```
+
+To further investigate the strategies that participants were following, we 
+examined the most common patterns of punitive behaviour across all twelve 
+decisions. Supplementary Table \@ref(tab:tablePatterns) shows the proportion of 
+participants following the 25 most common behavioural patterns, including, 
+where appropriate, the predetermined strategies from Table 
+\@ref(tab:tableStrategies). In both countries, a common pattern of behaviour 
+not captured by any of the strategies was punishing only when the target stole
+in the third-party game (Game F). Punishment in this game is consistent with an 
+egalitarian motive, as stealing produces unequal outcomes, but third-party 
+punishment here is also consistent with norm-enforcing and competitive motives 
+(see Table \@ref(tab:tableStrategies)). Other common behavioural patterns not 
+captured by our strategies included punishing whenever the target stole across 
+all games and always punishing in every game irrespective of the targets' 
+behaviour.
+
+While it is useful to look at exact patterns of behaviour, participants may not 
+have implemented their chosen punishment strategy with exact precision.
+In reality, strategies may have been implemented probabilistically for each 
+punishment decision. There is also the possibility of implementation errors, 
+whereby participants occasionally "slip up" and make decisions that are 
+incongruent with a particular strategy. This may explain why some participants 
+were unable to be classified exactly into a single punishment strategy.
+
+To deal with this complexity and include all observed data in our frequency 
+estimates, we fitted a Bayesian latent state model to the data. This model 
+assumes that the nine strategies in Table \@ref(tab:tableStrategies) (plus a 
+"random choice" strategy that chooses randomly for each decision) are the only 
+latent strategies and that these are instantiated into observed behaviour 
+according to the logic in Table \@ref(tab:tableStrategies) with some 
+probability of implementation error (i.e., an intention to punish is implemented 
+as non-punishment and vice versa). Averaging over all strategies and 
+incorporating the possibility of implementation errors, the model estimates the 
+probability of participants following any particular strategy, given the 
+observed data.
+
+The posterior estimates from the model are presented in Figure 
+\@ref(fig:plotModel1b). The posterior probabilities for each strategy did not
+differ between the two countries. In both countries, the never punish strategy
+had the highest probability, followed by the egalitarian strategy. The
+norm-enforcing and seek AI strategies were the next most likely, with higher
+posterior estimates than the competitive and antisocial strategies. None of the
+other strategies differed in their posterior estimates. The same general pattern
+emerged when we analysed the full dataset without pre-registered exclusions 
+(Supplementary Figure \@ref(fig:plotModel1a)).
+
+(ref:plotModel1bCaption) *Posterior estimates of the probabilities of following
+different punishment strategies from the Bayesian latent state model.* The model 
+assumes an implementation error rate of 5%. Points represent posterior medians, 
+line ranges represent 50% and 95% credible intervals.
+
+```{r plotModel1b, fig.width=7, fig.height=5, fig.cap="(ref:plotModel1bCaption)", warning=FALSE}
+tar_read(plot1.2)
+```
+
+Next, we explored which traits predicted adherence to different punishment 
+strategies. To answer this question, we included variables capturing 
+demographics, personality, social preferences, political views, and religious 
+views as predictors in our Bayesian latent state model. We included each 
+variable in a separate model, predicting all ten punishment strategies 
+(the nine from Table 1, plus the 'random choice' strategy) simultaneously.
+
+```{r sesSlope}
+ses <- tar_read(post2.2_SES)$beta[,2,10]
+```
+
+Demographic variables tended to be unrelated to strategy usage: age and gender
+did not predict adherence to a particular punishment strategy (Supplementary 
+Figures \@ref(fig:plotAllDems2) and \@ref(fig:plotAllDems1)). In the United 
+States, the never punish strategy was slightly more common among participants 
+lower in socio-economic status (median posterior slope = 
+`r printnum(median(ses))`, 95% CI [`r printnum(quantile(ses, 0.025))` 
+`r printnum(quantile(ses, 0.975))`]) but this effect was small.
+
+Conversely, personality and social preferences were linked to variation in 
+punishment strategies. When including the Big-6 personality dimensions and 
+Social Value Orientation (SVO) in the model, we found associations with the 
+egalitarian, never punish, and random choice strategies (Figure
+\@ref(fig:plotAllPers2)). Participants higher in SVO were more likely to follow 
+the egalitarian and the never punish strategies, while those with lower SVO 
+scores were more likely to enact the random choice strategy. The personality 
+dimensions of honesty-humility and openness to experience were both positively 
+associated with following the never punish strategy, while extraversion 
+negatively predicted this strategy. The effects were mostly similar across
+countries, but occasionally differed: for example, in the United States,
+but not in the United Kingdom, honesty-humility was positively associated with 
+following the egalitarian strategy and negatively associated with following 
+the random choice strategy. Overall, the same pattern of results emerged when 
+analysing the full dataset without exclusions (Supplementary Figure 
+\@ref(fig:plotAllPers1)).
+
+(ref:plotAllPers2Caption) *Posterior slopes from Bayesian latent state models 
+including Big-6 personality dimensions and Social Value Orientation.* Each row 
+represents a separate model. All models assume an implementation error rate of 
+5%. Points represent posterior medians, line ranges represent 95% credible 
+intervals.
+
+```{r plotAllPers2, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotAllPers2Caption)", warning=FALSE}
+tar_read(plotAllPers2)
+```
+
+Political and religious variables were also associated with punishment strategy 
+(Figure \@ref(fig:plotAllPolRel2)). These effects tended to be more pronounced 
+in the United States. Controlling for Social Dominance Orientation, American 
+participants higher in Right Wing Authoritarianism were more likely to follow 
+the strategies avoiding disadvantageous inequity and seeking advantageous 
+inequity. Participants who stated that they would like to "bring those below 
+them [on the socio-economic status ladder] up a peg" were more likely to follow 
+the egalitarian strategy, while American participants higher in Social Dominance 
+Orientation, Right Wing Authoritarianism, and believing that God controls events 
+in the world were less likely to follow the egalitarian strategy. In general, 
+religious and conservative participants were less likely to follow the never 
+punish strategy. This general pattern of results was replicated with the full 
+dataset (Supplementary Figure \@ref(fig:plotAllPolRel1)).
+
+(ref:plotAllPolRel2Caption) *Posterior slopes from Bayesian latent state models 
+including political ideology, views about social inequality, and religiosity.* 
+Each row represents a separate model aside from Social Dominance Orientation and 
+Right Wing Authoritarianism, which control for one another within the same 
+model. All models assume an implementation error rate of 5%. Points represent 
+posterior medians, line ranges represent 95% credible intervals.
+
+```{r plotAllPolRel2, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotAllPolRel2Caption)", warning=FALSE}
+tar_read(plotAllPolRel2)
+```
+
+Finally, we asked whether participants had insight into their own punishment 
+strategy. In other words, could participants self-report the strategy that they 
+were following during the games? To answer this question, we included 
+participants' responses to post-game questions about their strategy as 
+predictors in the model. As before, each predictor was included in a separate 
+model, predicting all ten strategies simultaneously.
+
+In general, we found that self-reported strategy usage was positively associated
+with the behavioural strategy that participants employed (see Supplementary 
+Figures \@ref(fig:plotSliders1) and \@ref(fig:plotSliders2) for the distribution 
+of responses to self-report questions). Figure \@ref(fig:plotAllSliders2) shows 
+the relationships between self-report questions and the different punishment 
+strategies, highlighting the combinations where the question matched the 
+behavioural strategy. We found positive relationships between the self-report 
+questions and strategy usage for the norm-enforcing, egalitarian, seek 
+advantageous inequity, never punish, and random choice  strategies. The 95% 
+credible intervals for other estimates included zero, though these estimates 
+often trended in a positive direction. The same pattern of results was found 
+when analysing the full dataset without exclusions  (Supplementary Figure 
+\@ref(fig:plotAllSliders1)).
+
+(ref:plotAllSliders2Caption) *Posterior slopes from models including 
+self-reported strategy usage.* Each row represents a separate model. All models 
+assume an implementation error rate of 5%. Highlighted estimates represent 
+combinations where the self-report question matched the behavioural strategy. 
+Points represent posterior medians, line ranges represent 95% credible 
+intervals.
+
+```{r plotAllSliders2, fig.width=7.5, fig.height=4, fig.cap="(ref:plotAllSliders2Caption)", warning=FALSE}
+tar_read(plotAllSliders2)
+```
 
@@
 
-x
+Using a suite of economic games measuring punishment in different situations, 
+we have shown that punishment does not serve just one function, but instead is 
+a flexible tool that can be and is used for different purposes[@Raihani2019]. 
+Punishment is more akin to a swiss army knife than a hammer, used by some to 
+enforce norms of cooperation and by others to reduce or even create inequality 
+between individuals. We found that people's punishment strategy can, to some 
+extent, be predicted by individual differences in personality, social 
+preferences, and political and religious views. Moreover, contrary to the view 
+that people are often unable to articulate the reasons for their punitive 
+behaviour[@Carlsmith2002], people seem to have some degree of insight into the 
+strategy they are using. Despite small differences, these general patterns 
+replicated in samples from both the United Kingdom and the United States.
+
+Among the punitive strategies, the most common were particularly sensitive to 
+inequality in payoffs, either from a self-referential perspective (i.e., avoid 
+disadvantageous inequity) or more generally (i.e., egalitarian). This is in 
+line with previous studies which have highlighted inequity aversion as an 
+important motivation for punishment in economic games 
+[@Dawes2007;@Raihani2012b]. Personality and social preference variables mapped 
+onto these strategies in expected ways. Traits associated with other-regarding 
+concern, such as SVO and honesty-humility, predicted following the egalitarian 
+strategy, whereas religious and conservative individuals were less likely to 
+follow this strategy, especially in the United States. Moreover, participants 
+following the egalitarian strategy were able to self-report this strategy, 
+though the same was not true for the avoid disadvantageous inequity strategy.
+
+Behaviour shaping strategies, such as deterrence and norm-enforcement, were less 
+common than strategies sensitive to inequality in our set of games. This was 
+reflected both in participants’ elicited punishment behaviour (Figure 
+\@ref(fig:plotModel1b)) and in their self-reports of their own strategy 
+(Supplementary Figures \@ref(fig:plotSliders1) and \@ref(fig:plotSliders2)). 
+Regarding the predictors of these strategies, we found that demographic, 
+personality, political, and religious variables tended to be unrelated to 
+deterrent and norm-enforcing punishment strategies. We also found that 
+participants had insight into the norm-enforcing strategy, but not the deterrent 
+strategy. This finding is in line with previous research showing that people 
+struggle to accurately report the deterrent motivations for their punitive 
+behaviour[@Carlsmith2002].
+
+Other punitive strategies were less common in our dataset, but some were more 
+prevalent than others. For example, participants were more likely to use 
+punishment to seek advantageous inequity than to exclusively harm those who did 
+not steal (i.e., antisocial punishment). The existence of the "seek AI" strategy 
+in our dataset supports the claim that punishment can also be used as a tool to 
+increase one’s own relative position[@Raihani2019]. While generally rare, we 
+found that this motive for punishment was more common among authoritarian 
+participants, at least in the United States, potentially providing an 
+explanation for why peer punishment has been found to be more common among 
+conservatives in previous work[@Hofmann2018]. Moreover, the fact that no 
+participants in our sample punished non-stealing across all games suggests that 
+antisocial punishment does not function to harm cooperators specifically, as has 
+been previously suggested[@Herrmann2008]. Instead, antisocial punishment appears 
+to be motivated by improving one's relative position in general, which is in 
+line with work showing that antisocial punishment disappears with a 1:1 fee-fine 
+ratio[@Sylwester2013].
+
+The fact that people use punishment for many different reasons poses problems 
+for the way that punishment is operationalised in classic behavioural economic 
+game studies. In these studies, a common assumption is that participants will 
+punish to change the behaviour of cheats. But in reality, people may be choosing 
+the punishment option to achieve a variety of different goals. The targets of 
+punishment in these studies are likely well aware that punishment could be 
+levied for these different reasons and this knowledge may impact their 
+responses. For example, if cheating targets interpret punishment as serving a 
+competitive motive, it may elicit retaliation rather than encourage 
+cooperation[@Dreber2008;@Nikiforakis2008b;@Raihani2019]. This might help to 
+explain the mixed findings in the field as to whether punishment actually 
+motivates cheating targets to cooperate in the future[@Raihani2019].
+
+It is striking that the most common strategy in our dataset was to never punish. 
+This is partly because punishment in these games imposes an economic cost for no 
+tangible benefit. If the fee-fine ratio had been lower such that it was cheaper 
+to punish, we may have seen more punishment from participants. Indeed, 
+`r printnum(mean(tar_read(dExc)$SelfRate10[tar_read(dExc)$strategy == "Never punish"] > 0, na.rm = TRUE)*100, digits = 0)`% 
+of participants following the never punish strategy positively stated that they 
+didn't want to pay to reduce anyone's bonus but would have done so if it were 
+free. But the frequency of the never punish strategy perhaps also reflects a 
+more general aversion to peer punishment, an aversion that has been highlighted 
+in both WEIRD (Western, educated, industrialised, rich, and democratic) samples
+[@Balafoutas2014;@Balafoutas2016] and in small-scale societies[@Baumard2010]. 
+One reason that people may be averse to peer punishment is that, due to its 
+multipurpose nature, it may be interpreted as a competitive challenge by targets 
+and trigger retaliation[@Raihani2019]. In situations that lack clear 
+institutional norms to legitimise punishment, such as our economic games and 
+some situations in the real world, people might abstain from peer punishment to 
+avoid such retaliation, regardless of whether retaliation is actually possible. 
+By contrast, institutionalised punishment in small-scale societies often 
+functions to compensate victims adequately while limiting the potential for 
+feuds and cycles of retaliation[@Fitouchi2023;@Singh2022]. Future research 
+should uncover whether people are more willing to punish in these 
+conventionalised contexts.
+
+There are several limitations with our study design that can guide future 
+research. First, we used one-shot economic games to measure punishment 
+strategies, which may have led us to underestimate behaviour-change strategies 
+like deterrence. Our inclusion of Game C somewhat mitigated this issue by 
+manipulating whether stealing behaviour was intentional vs. unintentional and 
+thus whether there was any behaviour to be deterred. Due to limits on our 
+within-subjects design and the complexity of the strategy space, it was not 
+feasible for us to expand our study to include additional contexts to elicit 
+behaviour-change strategies (e.g., repeated games, games where targets are not 
+made aware of the punishment, games where targets can retaliate). Future work 
+could study these contexts separately.
+
+A second limitation is that some strategies required more punishment than others 
+to be met. For example, the competitive strategy required punishment in ten of 
+twelve decisions, compared to the avoid disadvantageous inequity strategy which 
+required only one instance of punishment (Table \@ref(tab:tableStrategies)). 
+Strategies thus differed in how "expensive" they were to implement, perhaps 
+explaining why some strategies were more common than others. This issue is 
+largely unavoidable in our design since strategies, by their very nature, differ 
+in how punitive they are. To partially mitigate this issue, we employed the 
+strategy method to incentivise participants, such that payoffs were calculated 
+from a randomly chosen game instead of summed across all games.
+
+Another limitation is that our results may be contingent on the particular suite 
+of anonymous stealing games that we used. With our anonymous design, we were 
+unable to study other potential reputational strategies underlying punishment, 
+such as signalling trustworthiness[@Jordan2020]. Moreover, stealing may be 
+evaluated differently to other forms of cheating, such as not contributing to 
+public goods, and other negative behaviours, such as lying or breaking taboos. 
+Finally, we were unable to include all permutations of situational features in 
+the games (e.g., second-party vs. third-party, equal vs. unequal) making it 
+difficult to interpret some patterns of behaviour. For example, many 
+participants punished only in the third-party game (Game F), but it is not clear 
+whether these participants were driven by the third-party nature of the game or 
+simply by the fact that stealing in that game generated inequality. Future work 
+should determine whether different reputational contexts, target behaviours, and 
+combinations of situational features elicit different punishment strategies.
+
+In sum, we have shown that while many people choose not to punish peers, those 
+who do are motivated by a variety of different concerns, including behaviour 
+shaping, egalitarianism, and competition. Much like the observed variation in 
+human social learning strategies[@Molleman2014], humans thus also exhibit 
+variation in their punishment strategies. These individual differences map onto 
+personality dimensions, social preferences, political and religious views, and 
+self-reports of behaviour. We hope that future work will continue to unpack the 
+multifaceted nature of human punishment.
 
@@
 
-x
+## Ethical approval
+
+Ethical approval was granted by the University College London Ethics Board 
+(project: 3720/002). The study was performed in accordance with all the relevant 
+guidelines and regulations. Informed consent was obtained from all participants 
+prior to the study.
+
+## Pre-registration
+
+We pre-registered the study on the Open Science Framework before collecting data 
+in the United Kingdom (11^th^ November 2022; https://osf.io/k75fc). We submitted 
+another pre-registration before collecting data in the United States (20^th^
+June 2023; https://osf.io/q4hdy). In the pre-registrations, we outlined our 
+study design, exclusion criteria, and analysis plan. As the study was 
+exploratory, we did not pre-register any explicit hypotheses. We did not deviate 
+from the pre-registrations.
+
+## Exclusion criteria
+
+We pre-registered that we would exclude participants who failed any of the 
+attention checks, sped through the surveys (i.e., two standard deviations below 
+the median duration), or flatlined (i.e., provided identical responses to matrix 
+questions). We also stated that we would exclude data for particular games if 
+participants failed the comprehension question for that game. We followed our 
+pre-registered plan of conducting analyses with and without these exclusions 
+(analyses without exclusions are reported in the Supplementary Material).
+
+## Participants
+
+```{r}
+tar_load(c(d, dExc))
+uk_n_full <- sum(d$Country == "United Kingdom")
+uk_n_rtrn <- sum(d$Country == "United Kingdom" & !is.na(d$StartDate_Part2))
+uk_n_excl <- sum(dExc$Country == "United Kingdom")
+us_n_full <- sum(d$Country == "United States")
+us_n_rtrn <- sum(d$Country == "United States" & !is.na(d$StartDate_Part2))
+us_n_excl <- sum(dExc$Country == "United States")
+```
+
+We collected a representative sample of `r printnum(uk_n_full)` participants 
+from the United Kingdom through the online platform Prolific 
+(https://www.prolific.com/). All of these participants completed the economic 
+games and `r printnum(uk_n_rtrn)` returned to complete the follow-up survey a 
+week later (`r printnum((uk_n_rtrn / uk_n_full)*100, digits = 0)`% retention 
+rate). After exclusions, we were left with `r printnum(uk_n_excl)` participants 
+overall (see Supplementary Figure \@ref(fig:plotSampleUK) for sample 
+characteristics).
+
+We later collected a representative sample of `r printnum(us_n_full)` 
+participants from the United States through Prolific. All of these participants 
+completed the economic games and `r printnum(us_n_rtrn)` returned to complete 
+the follow-up survey (`r printnum((us_n_rtrn / us_n_full) * 100, digits = 0)`% 
+retention rate). After exclusions, we were left with `r printnum(us_n_excl)` 
+participants overall (see Supplementary Figure \@ref(fig:plotSampleUS) for 
+sample characteristics).
+
+## Materials
+
+### Economic games
+
+In the first part of the study, participants completed six economic games, each 
+with slight variations. In all games, there are multiple players and the 
+participant takes the role of P1. P2 either (a) steals £0.20 from another player 
+and adds it to their payoff or (b) does nothing. For each of these cases, 
+participants are asked whether they would like to pay money to reduce P2’s 
+payoff. Games A-E have two players and Game F has three players.
+
+The six games are as follows (variations bolded; see Figure \@ref(fig:plotGames)
+for a visual representation of the games):
+
+1. *Game A (Advantageous Inequity)*. P1 starts with £0.70 and P2 starts with 
+   £0.10. P2 is given the option to either steal £0.20 from P1 or do nothing. P1 
+   can then pay £0.10 to reduce P2’s payoff by £0.30.
+2. *Game B (Equal)*. P1 starts with £0.70 and P2 starts with **£0.30**. P2 is 
+   given the option to either steal £0.20 from P1 or do nothing. P1 can then pay 
+   £0.10 to reduce P2’s payoff by £0.30.
+3. *Game C (Computer)*. P1 starts with £0.70 and P2 starts with £0.30. 
+   Participants are told that **"the computer will decide"** whether P2 steals
+   £0.20 from P1 or does nothing. P1 can then pay £0.10 to reduce P2’s payoff by 
+   £0.30.
+4. *Game D (1:1 Fee-Fine)*. P1 starts with £0.70 and P2 starts with £0.30. P2 is 
+   given the option to either steal £0.20 from P1 or do nothing. P1 can then pay 
+   £0.10 to reduce P2’s payoff by **£0.10**.
+5. *Game E (Disadvantageous Inequity)*. P1 starts with £0.70 and P2 starts with 
+   **£0.50**. P2 is given the option to either steal £0.20 from P1 or do 
+   nothing. P1 can then pay £0.10 to reduce P2’s payoff by £0.30.
+6. *Game F (Third-Party)*. P1 starts with £1.00, P2 and P3 start with £0.70. P2 
+   is given the option to either steal £0.20 **from P3** or do nothing. P1 can 
+   then pay £0.10 to reduce P2’s payoff by £0.30.
+
+For each game, participants saw the game instructions and answered a 
+comprehension question before providing their decisions. After completing all 
+the games, participants were asked to give an open-ended response explaining 
+their behaviour in the games, and then responded to several slider questions 
+capturing the different reasons for their decisions (for full wordings, see 
+Supplementary Table \@ref(tab:tableSliderWordings)).
+
+### Survey questions
+
+In a follow-up survey, we collected the following data on participants (for 
+wordings of all questions, see Supplementary Table 
+\@ref(tab:tableSurveyWordings)):
+
+- *Demographics*. In the survey, we collected information on participants' 
+  education level and self-reported socio-economic status (MacArthur ladder
+  [@Adler2000]). We also collected additional demographic data from Prolific 
+  (e.g., age, gender, student status).
+- *Personality*. We used the Mini-IPIP scale[@Sibley2011] to measure the Big 6 
+  personality dimensions of agreeableness, conscientiousness, extraversion, 
+  honesty-humility, openness to experience, and neuroticism (four items each).
+- *Social Value Orientation*. We used the Social Value Orientation Slider 
+  Measure to measure other-regarding preferences[@Murphy2011]. Across fifteen 
+  items, participants made decisions on how to allocate different amounts of 
+  money between themselves and another anonymous individual. From these 
+  decisions, we calculated participants' Social Value Orientation "angle" as a 
+  measure of their other-regarding preference, following the steps outlined in 
+  ref[@Murphy2011].
+- *Political ideology*. We included several measures of political ideology, 
+  including left-right conservatism, Social Dominance Orientation[@Ho2015] 
+  (eight items), and Right Wing Authoritarianism[@Bizumic2018] (six items). 
+  We also probed participants' views on social inequality by asking them whether 
+  they would like to bring people above (below) them on the MacArthur 
+  socio-economic status ladder down (up) a peg or two.
+- *Religious views*. We asked participants how religious they consider 
+  themselves and whether they believe that God or another spiritual non-human 
+  entity controls the events in the world[@Laurin2012].
+
+## Procedure
+
+We began data collection in the United Kingdom on 28^th^ November 2022, with 
+participants returning to complete the follow-up survey on 5^th^ December 2022. 
+We then ran a second wave of data collection in the United States on 20^th^ June 
+2023, with participants returning to complete the follow-up survey on 27^th^ 
+June 2023. Our surveys were designed through the online survey platform 
+Qualtrics (https://www.qualtrics.com/).
+
+In the initial games survey, participants completed all six economic games in a 
+random order, with punishment decisions (whether to punish a stealing target and 
+whether to punish a target who did nothing) randomised within games. Responses 
+to comprehension questions suggested that participants understood the six 
+economic games (Supplementary Table \@ref(tab:tableComp)). We used the strategy 
+method to incentivise the economic games, choosing a random game to determine 
+bonus payment. After all games, 
+`r printnum((sum(dExc$DecisionConsequences == "Yes", na.rm = TRUE) / nrow(dExc)) * 100, digits = 0)`% 
+of participants stated that they believed that their decisions had real 
+consequences for others.
+
+In the follow-up survey, participants completed blocks of questions on 
+demographics, personality, Social Value Orientation, political ideology, and 
+religious views in a random order, with questions randomised within blocks. A 
+random decision from the Social Value Orientation Slider Measure was chosen to 
+determine bonus payment.
+
+Participants were paid £1.80 for completing the games survey, plus a bonus 
+payment from the six economic games (between £0.40 -- £0.70 depending on their 
+decision). Participants were paid £1.50 for completing the follow-up survey, 
+plus a bonus payment from the Social Value Orientation Slider Measure (between 
+£0.50 -- £0.85 depending on their decision).
+
+## Statistical analysis
+
+We pre-registered that we would use a Bayesian latent state model to infer 
+unobserved punishment strategies from the observed data (for a similar version 
+of this model, see ref[@McElreath2020]). In this model, participants $i$ in 
+countries $c$ make binary punishment decisions across twelve decisions $j$. We 
+assume that the probability of the observed data $y_{i,j}$ is the weighted 
+average of the probability of the observed data conditional on each of the ten 
+punishment strategies $s$. From this logic, the model estimates the probability 
+of each strategy $p_{s}$. The full model is as follows:
+
+\begin{align}
+y_{i,j} &\sim \text{Bernoulli}(\theta_{j}) \\
+\theta_{j} &= \sum_{s=1}^{10} p_{s} \text{Pr}(\text{punish}|s,j) \nonumber \\
+p &= \text{softmax}(\alpha_{c[i]}) \nonumber \\
+\alpha_{s,c} &\sim \text{Normal(0, 1)} \nonumber
+\end{align}
+
+The conditional probabilities $\text{Pr}(\text{punish}|s,j)$ are hard coded in 
+the model as outlined in Table \@ref(tab:tableStrategies). We incorporate an 
+implementation error rate $\delta$ into these conditional probabilities by coding 
+green ticks in Table \@ref(tab:tableStrategies) with a conditional probability 
+of $1 - \delta$ and coding red crosses with a conditional probability of 
+$0 + \delta$. The random choice is consistently coded with a conditional 
+probability of ½ across all decisions.
+
+To include a categorical predictor in the model, we estimate a different 
+$\alpha_{s,c}$ for each categorical level. To include a continuous predictor 
+$x$ in the model, we include a slope $\beta$ in the linear model for $p$:
+
+\begin{align}
+y_{i,j} &\sim \text{Bernoulli}(\theta_{j}) \\
+\theta_{j} &= \sum_{s=1}^{10} p_{s} \text{Pr}(\text{punish}|s,j) \nonumber \\
+p &= \text{softmax}(\alpha_{c[i]} + \beta_{c[i]}x_{i}) \nonumber \\
+\alpha_{s,c} &\sim \text{Normal(0, 1)} \nonumber \\
+\beta_{s,c} &\sim \text{Normal(0, 0.2)} \nonumber
+\end{align}
+
+We estimated the posterior distributions of these models using Hamiltonian Monte 
+Carlo as implemented in Stan version 2.26.1[@Stan2020]. We ran each model for 
+2000 samples, with 1000 warmup samples. R-hat values and effective sample sizes 
+suggested that all models converged normally. Trace plots are reported in 
+Supplementary Figure \@ref(fig:plotTrace).
+
+We validated the model by simulating observed data (n = 100) from a known 
+frequency of strategies. The model was successfully able to recover the known 
+frequency of strategies from the simulated data (Supplementary Figure 
+\@ref(fig:plotSim)).
+
+## Reproducibility
+
+All data and code are accessible on GitHub: 
+https://github.com/ScottClaessens/punishStrategies. All analyses were conducted 
+in R version 4.2.1[@RCoreTeam]. Visualisations were created with the *ggplot2*
+[@Wickham2016] and *cowplot*[@Wilke2020] R packages. We used the *targets*
+[@Landau2021] R package to create a reproducible data analysis pipeline and the 
+*papaja*[@Aust2022] R package to reproducibly generate the manuscript.
 
@@
 
-x
+This work was supported by a Royal Society of New Zealand Catalyst Leaders Grant 
+to Q.D.A and N.R. (ref: ILFUOA2002).
 
@@
 
-x
+All authors conceptualised the research, designed the study, and developed the
+surveys. N.J. conducted data collection on Prolific. S.C. conducted all analyses
+and visualisation of the data. All authors wrote the manuscript.
 
@@
 
-All data used in this study are publicly available on GitHub: https://github.com/ScottClaessens/punishStrategies
+All data used in this study are publicly available on GitHub: 
+https://github.com/ScottClaessens/punishStrategies
 
@@
 
-All code to reproduce the analyses in this study are publicly available on GitHub: https://github.com/ScottClaessens/punishStrategies
+All code to reproduce the analyses in this study are publicly available on 
+GitHub: https://github.com/ScottClaessens/punishStrategies
 

### 2023-11-24 manuscript.tex
finished manuscript first draft

@@
 {}{\PatchFailed}
-\keywords{x\newline\indent Word count: xxxx words}
+\keywords{punishment; cooperation; economic games\newline\indent Word count: 5185 words}
 \usepackage{lineno}
@@
 \usepackage{makecell}
-\renewcommand\theadfont{}
 \raggedbottom
@@
 \hypersetup{
-  pdftitle={Mapping the punishment strategy space},
+  pdftitle={Why do people punish? Evidence for a range of strategic concerns},
   pdfauthor={Scott Claessens*1, Quentin D. Atkinson1, \& Nichola Raihani1,2},
   pdflang={en-EN},
-  pdfkeywords={x},
+  pdfkeywords={punishment; cooperation; economic games},
   hidelinks,
@@
 
-\title{Mapping the punishment strategy space}
+\title{Why do people punish? Evidence for a range of strategic concerns}
 \author{Scott Claessens\textsuperscript{*1}, Quentin D. Atkinson\textsuperscript{1}, \& Nichola Raihani\textsuperscript{1,2}}
@@
 \abstract{%
-x
+Costly punishment is thought to be one of the key mechanisms sustaining cooperation in humans.
+However, the motives for punitive behaviour remain unclear.
+Punishment is often assumed to be motivated by a desire to convert free riders into cooperators, but it is also consistent with a host of other functions, such as levelling payoffs or increasing one's relative position.
+We used a suite of six economic games to tease apart the different motives for punishment.
+Across representative samples from the United Kingdom and the United States, we estimated the frequency of different punishment strategies in the population, finding that egalitarian motives for punishment are more common than behaviour-change motives.
+Moreover, different punishment strategies were differentially predicted by personality, social preferences, political ideology, and religious views.
+Self-reports of behaviour in the games suggested that people have some degree of insight into their punishment strategy.
+These findings highlight the multipurpose nature of human punishment.
 }
@@
 
-Humans cooperate on a scale unparalleled in the animal kingdom.
-One mechanism thought to sustain this level of cooperation is costly punishment, whereby individuals harm others at a personal cost\textsuperscript{1}, ostensibly encouraging cooperative behaviour from the target (or bystanders\textsuperscript{2--4}) in the future.
-Punishment therefore offers a route to maintaining or increasing cooperation by changing the payoff structure of social interactions such that it no longer pays to cheat or exploit social partners\textsuperscript{1,5}.
-
-In humans, many studies of punishment have been carried out in laboratory settings using economic games\textsuperscript{6--15}.
-In these games, participants are usually given a sum of money that they can use to invest in collective action or to help others.
-Alternatively, participants can `cheat' by keeping the money for themselves or by exploiting the contributions of others.
-Punishment is introduced into such games by giving participants the option to pay a small `fee' to impose a greater `fine' on their co-players.
-Several lines of experimental evidence indicate that people use this punishment option\textsuperscript{12}, that they enjoy punishing\textsuperscript{16}, and that they frequently, though not always\textsuperscript{17}, punish cheating or exploitative co-players\textsuperscript{10,11}.
-
-Evidence from these experiments suggests that the threat of costly punishment plays an important role in promoting human cooperation.
-People tend to cooperate more in games where punishment is possible compared to those where it is not\textsuperscript{6,7,15}.
-The effect that the threat of punishment has on cooperation is also evident in the higher contributions typically observed in the Ultimatum Game (where punishment is possible) compared to the structurally-similar Dictator Game (where it is not)\textsuperscript{18}.
-This typical cooperation-enhancing effect of punishment has also been observed across societies\textsuperscript{7}, leading some to suggest that costly punishment has played a key role in the cultural evolution of cooperation in humans\textsuperscript{19--22}.
-
-Nevertheless, it remains unclear whether individuals playing economic games use punishment as a behaviour-change tool to enforce cooperation or as a means to achieve other ends.
-Some have argued that punishment is primarily used to shape future behaviour, either to deter personal harm\textsuperscript{3,9,23} or to uphold normative standards of cooperative behaviour\textsuperscript{20,21,24--27}.
-But while the \emph{threat} of punishment can have a cooperation-enhancing effect, the \emph{enactment} of this punishment seldom changes the future behaviour of deviant targets\textsuperscript{15}.
-This calls into question whether punishment primarily operates as a behaviour-change tool or whether it is used to achieve other goals.
-
-Beyond behaviour-shaping concerns, there are a host of other reasons that people may want to punish in economic games.
-Punishers might be motivated by a desire for retribution rather than deterrence, punishing in proportion to the amount of harm that was personally caused\textsuperscript{28}.
-Punishment might be driven by concerns about relative payoffs, such as disadvantageous inequity aversion (i.e., avoiding having less than others\textsuperscript{15,29}) and/or general egalitarian preferences (i.e., wanting all participants to receive the same payoffs\textsuperscript{30}).
-Such concerns about relative payoffs may be activated when participants earn less than cheats in economic games or when there are income disparities in these settings.
-People might also use punishment for competitive purposes, seeking advantageous inequity for themselves (i.e., having more than others) and/or improving their relative position\textsuperscript{15}.
-
-Common economic game designs have been unable to tease apart these different motives for punishment because participants who interact with cheats in these games experience both losses \emph{and} lower relative payoffs.
-The typical 1:3 fee-to-fine ratio of punishment in economic games compounds this issue.
-With this setup, people can simultaneously use punishment to reciprocate losses, to deter others from cheating, and to reduce or reverse disparities in payoffs between themselves and targets.
-To add to this complexity, it is evident that people use punishment in seemingly disparate ways: punishing when no behaviour change is possible, such as in one-shot games\textsuperscript{12,29,31,32} or in games where the target never learns about the punishment\textsuperscript{33}; punishing those who did not cheat or who over-contributed to collective action (antisocial punishment\textsuperscript{17,34}); punishing in scenarios where they were not personally harmed (third-party punishment\textsuperscript{35}); and punishing in scenarios where disparities in payoffs did not arise from participants' actions\textsuperscript{30,36,37}.
-
-The general conclusion from this research is that there is no one unifying function of costly punishment in humans.
-Instead, punishment should be thought of as a flexible behavioural tool that serves a variety of functions that are not mutually exclusive\textsuperscript{15}.
-Due to its multipurpose nature, we should therefore expect variation in punishment strategies in the population, much like the observed variation in social learning strategies\textsuperscript{38}.
-Some individuals may use punishment as a behaviour shaping tool, for example, while others may use it to reduce or reverse payoff differentials.
-
-This insight raises several underexplored questions.
-First, which punishment strategies are more frequent in human populations?
-Second, what traits predict adherence to a particular strategy?
-Previous work has reported that personality is related to cooperative behaviour\textsuperscript{39} and demographics, political ideology, and religiosity are related to punitive behaviour\textsuperscript{40}, but no research has related these variables to specific punishment strategies.
-Third, do people have insight into their own punishment strategy?
-Previous work has argued that people are often unaware of the underlying function of their punitive behaviour, yet they feel compelled to enact it anyway\textsuperscript{28,41}.
-
-Here, we aim to delineate the many possible punishment strategies by asking whether people punish in a manner consistent with a specific strategy and, if so, what other characteristics (personality, social preferences, political orientation) predict the use of different punishment strategies.
-Table \ref{tab:tableStrategies} summarises the potential functions for costly punishment in economic games and the behavioural strategies they predict.
-Note that we do not include reputational functions of punishment in this table, such as signalling trustworthiness\textsuperscript{4,42--45}, because our focus is on punishment strategies in anonymous economic games without reputational incentives (but see ref\textsuperscript{46}).
+Humans cooperate on a scale that is unparalleled in the animal kingdom. One
+mechanism thought to sustain this level of cooperation is costly punishment,
+whereby individuals harm others at a personal cost\textsuperscript{1},
+ostensibly encouraging cooperative behaviour from the target (or bystanders\textsuperscript{2--4}) in the future. Punishment
+therefore offers a route to maintaining or increasing cooperation by changing
+the payoff structure of social interactions such that it no longer pays to cheat
+or exploit social partners\textsuperscript{1,5}.
+
+In humans, many studies of punishment have been carried out in laboratory
+settings using economic games\textsuperscript{6--15}. In these games, participants are usually given a sum of money
+that they can use to invest in collective action or to help others.
+Alternatively, participants can `cheat' by keeping the money for themselves or
+by exploiting the contributions of others. Punishment is introduced into such
+games by giving participants the option to pay a small `fee' to impose a greater
+`fine' on their co-players. Several lines of experimental evidence indicate that
+people use this punishment option\textsuperscript{12}, that they enjoy
+punishing\textsuperscript{16}, and that they frequently, though not
+always\textsuperscript{17}, punish cheating or exploitative
+co-players\textsuperscript{10,11}.
+
+Evidence from these experiments suggests that the threat of costly punishment
+plays an important role in promoting human cooperation. People tend to cooperate
+more in games where punishment is possible compared to those where it is not\textsuperscript{6,7,15}. The effect that the threat of
+punishment has on cooperation is also evident in the higher contributions
+typically observed in the Ultimatum Game (where punishment is possible) compared
+to the structurally-similar Dictator Game (where it is not)\textsuperscript{18}. This
+typical cooperation-enhancing effect of punishment has also been observed across
+societies\textsuperscript{7}, leading some to suggest that costly punishment has
+played a key role in the cultural evolution of cooperation in humans\textsuperscript{19--22}.
+
+Nevertheless, it remains unclear whether individuals playing economic games use
+punishment as a behaviour-change tool to enforce cooperation or as a means to
+achieve other ends. Some have argued that punishment is primarily used to shape
+future behaviour, either to deter personal harm\textsuperscript{3,9,23} or to uphold normative standards of cooperative behaviour\textsuperscript{20,21,24--27}.
+But while the \emph{threat} of punishment can have a cooperation-enhancing effect,
+the \emph{enactment} of this punishment does not consistently deter targets from
+cheating in the future\textsuperscript{15}. This calls into question whether
+punishment primarily operates as a behaviour-change tool or whether it is used
+to achieve other goals.
+
+Beyond behaviour-shaping concerns, there are a host of other reasons why people
+may want to punish in economic games. Punishers might be motivated by a desire
+for retribution rather than deterrence, punishing in proportion to the amount
+of harm that was personally caused\textsuperscript{28}. Punishment might be driven
+by concerns about relative payoffs, such as disadvantageous inequity aversion
+(i.e., avoiding having less than others\textsuperscript{15,29}) and/or
+general egalitarian preferences (i.e., wanting all participants to receive the
+same payoffs\textsuperscript{30}). Such concerns about relative payoffs may be activated
+when participants earn less than cheats in economic games or when there are
+income disparities in these settings. People might also use punishment for
+competitive purposes, seeking advantageous inequity for themselves (i.e., having
+more than others) and/or improving their relative position\textsuperscript{15}.
+
+Common economic game designs have been unable to tease apart these different
+motives for punishment because participants who interact with cheats in these
+games experience both losses \emph{and} lower relative payoffs. The typical 1:3
+fee-fine ratio of punishment in economic games compounds this issue. With
+this setup, people can simultaneously use punishment to reciprocate losses, to
+deter others from cheating, and to reduce or reverse disparities in payoffs
+between themselves and targets. To add to this complexity, it is evident that
+people use punishment in seemingly disparate ways: punishing when no behaviour
+change is possible, such as in one-shot games\textsuperscript{12,29,31,32}, on the very last round of repeated games\textsuperscript{33}, or in games where the target never learns about the
+punishment\textsuperscript{34}; punishing those who did not cheat or who
+over-contributed to collective action (antisocial punishment\textsuperscript{17,35}); punishing in scenarios where they were not
+personally harmed (third-party punishment\textsuperscript{36}); and punishing in
+scenarios where disparities in payoffs did not arise from participants'
+actions\textsuperscript{30,37,38}.
+
+The general conclusion from this research is that there is no one unifying
+function of costly punishment in humans. Instead, punishment should be thought
+of as a flexible behavioural tool that serves a variety of functions that are
+not mutually exclusive\textsuperscript{15}. Due to its multipurpose nature, we should
+therefore expect variation in punishment strategies in the population, much like
+the observed variation in social learning strategies\textsuperscript{39}. Some
+individuals may use punishment as a behaviour shaping tool, for example, while
+others may use it to reduce or reverse payoff differentials.
+
+This insight raises several underexplored questions. First, which punishment
+strategies are more frequent in human populations? Second, what traits predict
+adherence to a particular strategy? Previous work has reported that personality
+is related to cooperative behaviour\textsuperscript{40} and demographics, political
+ideology, and religiosity are related to punitive behaviour\textsuperscript{41}, but
+no research has related these variables to specific punishment strategies.
+Third, do people have insight into their own punishment strategy? Previous work
+has argued that people are often unaware of the underlying function of their
+punitive behaviour, yet they feel compelled to enact it anyway\textsuperscript{28,42}.
+
+Here, we aim to delineate nine possible punishment strategies by asking
+whether people punish in a manner consistent with a specific strategy and, if
+so, what other characteristics (personality, social preferences, political
+orientation) predict the use of different punishment strategies. Table
+\ref{tab:tableStrategies} summarises the potential functions for costly
+punishment in the economic games that we considered, and the behavioural
+strategies they predict. Note that Table \ref{tab:tableStrategies} is not
+an exhaustive list of all possible punishment strategies: we do not include
+reputational functions of punishment in this table, such as signalling
+trustworthiness\textsuperscript{4,43--46},
+because our focus is on punishment strategies in anonymous economic games without
+reputational incentives (but see ref\textsuperscript{47}).
+
+Building on previous designs\textsuperscript{29,31,48,49}, we employ a suite of one-shot economic games where individuals
+are given the opportunity to punish targets at a personal cost (Figure
+\ref{fig:plotGames}). In each game, targets either steal from another
+individual or do nothing. Representative samples of participants from the United
+Kingdom (\emph{n} = 1014) and the
+United States (\emph{n} = 996)
+completed all six games on the online platform Prolific. We carefully designed
+the suite of games to tease apart the proposed punishment strategies in Table
+\ref{tab:tableStrategies}, such that each strategy predicts a different pattern
+of behaviour across all the games (see Methods for more detail about the six
+games). We use the resulting behavioural patterns to discern which punishment
+strategy participants are employing. We then combine these behavioural patterns
+with data on demographics, personality, social preferences, political ideology,
+religiosity, and self-reported strategy usage.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{images/games} \caption{\emph{Visual summary of the six economic games.} In all
+games, Player 2 either steals £0.20 from Player 1 (the focal player) or does
+nothing. Player 1 is then given the option to punish by paying a certain amount
+of money to remove money from Player 2 (this money is destroyed). The six games
+are variants on this general setup, creating situations where (A) Player 2 is
+still worse off by stealing, (B) Player 2 creates equality by stealing, (C) the
+computer ``decides'' whether Player 2 steals, (D) the fee-fine ratio is 1:1, (E)
+Player 2 is better off by stealing, and (F) Player 2 steals instead from a
+third-party.}\label{fig:plotGames}
+\end{figure}
 
@@
 \begin{longtable}{p{1.7cm}p{3.3cm}cccccccccccc}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:tableStrategies}xxx}\\
+\caption{\label{tab:tableStrategies}\scriptsize Summary of the different functions for
+punishment and the behavioural strategies they predict. \emph{Games A-F are
+the games employed in the current study (see Methods for more details). In each
+of the six games, participants are given the opportunity to punish players who
+``steal'' and those who do not, meaning that participants make twelve punishment
+decisions in total. Each behavioural strategy implies a unique pattern of
+punishment across all decisions. Green ticks reflect decisions to punish, red
+crosses reflect decisions to not punish. In column headers, payoffs at the first
+stage (above) and the second stage (below) are denoted as P1-P2 (or P2-P3 {[}P1{]}
+for Game F) where participants take the role of P1 and P2 is the target of
+punishment. AI = advantageous inequity, DI = disadvantageous inequity.}}\\
 \toprule
- &  & \multicolumn{2}{c}{\thead{No DI 1 \\ 70-10}} & \multicolumn{2}{c}{\thead{No DI 2 \\ 70-30}} & \multicolumn{2}{c}{\thead{No DI 3 \\ (Computer) \\ 70-30}} & \multicolumn{2}{c}{\thead{No DI 4 \\ (1:1 Fee-Fine) \\ 70-30}} & \multicolumn{2}{c}{\thead{DI \\ 70-50}} & \multicolumn{2}{c}{\thead{3PP \\ 70-70 [100]}} \\
+ &  & \multicolumn{2}{c}{\thead{\scriptsize Game A \\ (AI) \\ 70-10}} & \multicolumn{2}{c}{\thead{\scriptsize Game B \\ (Equal) \\ 70-30}} & \multicolumn{2}{c}{\thead{\scriptsize Game C \\ (Computer) \\ 70-30}} & \multicolumn{2}{c}{\thead{\scriptsize Game D \\ (1:1 Fee-Fine) \\ 70-30}} & \multicolumn{2}{c}{\thead{\scriptsize Game E \\ (DI) \\ 70-50}} & \multicolumn{2}{c}{\thead{\scriptsize Game F \\ (Third-Party) \\ 70-70 [100]}} \\
 \cmidrule(r){3-4} \cmidrule(r){5-6} \cmidrule(r){7-8} \cmidrule(r){9-10} \cmidrule(r){11-12} \cmidrule(r){13-14}
-Function & \multicolumn{1}{c}{Behavioural strategy} & \multicolumn{1}{c}{\thead{Take \\ 50-30}} & \multicolumn{1}{c}{\thead{No take \\ 70-10}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-70}} & \multicolumn{1}{c}{\thead{No take \\ 70-50}} & \multicolumn{1}{c}{\thead{Take \\ 50-90 [100]}} & \multicolumn{1}{c}{\thead{No take \\ 70-70 [100]}}\\
+Function & \multicolumn{1}{c}{Behavioural strategy} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-30}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-10}} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-50}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-30}} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-50}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-30}} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-50}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-30}} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-70}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-50}} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-90 [100]}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-70 [100]}}\\
 \midrule
@@
 \toprule
- &  & \multicolumn{2}{c}{\thead{No DI 1 \\ 70-10}} & \multicolumn{2}{c}{\thead{No DI 2 \\ 70-30}} & \multicolumn{2}{c}{\thead{No DI 3 \\ (Computer) \\ 70-30}} & \multicolumn{2}{c}{\thead{No DI 4 \\ (1:1 Fee-Fine) \\ 70-30}} & \multicolumn{2}{c}{\thead{DI \\ 70-50}} & \multicolumn{2}{c}{\thead{3PP \\ 70-70 [100]}} \\
+ &  & \multicolumn{2}{c}{\thead{\scriptsize Game A \\ (AI) \\ 70-10}} & \multicolumn{2}{c}{\thead{\scriptsize Game B \\ (Equal) \\ 70-30}} & \multicolumn{2}{c}{\thead{\scriptsize Game C \\ (Computer) \\ 70-30}} & \multicolumn{2}{c}{\thead{\scriptsize Game D \\ (1:1 Fee-Fine) \\ 70-30}} & \multicolumn{2}{c}{\thead{\scriptsize Game E \\ (DI) \\ 70-50}} & \multicolumn{2}{c}{\thead{\scriptsize Game F \\ (Third-Party) \\ 70-70 [100]}} \\
 \cmidrule(r){3-4} \cmidrule(r){5-6} \cmidrule(r){7-8} \cmidrule(r){9-10} \cmidrule(r){11-12} \cmidrule(r){13-14}
-Function & \multicolumn{1}{c}{Behavioural strategy} & \multicolumn{1}{c}{\thead{Take \\ 50-30}} & \multicolumn{1}{c}{\thead{No take \\ 70-10}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-50}} & \multicolumn{1}{c}{\thead{No take \\ 70-30}} & \multicolumn{1}{c}{\thead{Take \\ 50-70}} & \multicolumn{1}{c}{\thead{No take \\ 70-50}} & \multicolumn{1}{c}{\thead{Take \\ 50-90 [100]}} & \multicolumn{1}{c}{\thead{No take \\ 70-70 [100]}}\\
+Function & \multicolumn{1}{c}{Behavioural strategy} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-30}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-10}} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-50}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-30}} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-50}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-30}} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-50}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-30}} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-70}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-50}} & \multicolumn{1}{c}{\thead{\scriptsize Steal \\ 50-90 [100]}} & \multicolumn{1}{c}{\thead{\scriptsize No steal \\ 70-70 [100]}}\\
 \midrule
@@
 
-Building on previous designs\textsuperscript{29,31,47,48}, we employ a suite of one-shot economic games where individuals are given the opportunity to punish targets at a personal cost.
-We carefully designed this suite of games to tease apart the proposed punishment strategies in Table 1, such that each strategy predicts a different pattern of behaviour across all the games.
-We use the resulting behavioural patterns to discern which punishment strategy participants are employing.
-We then combine these behavioural patterns with data on demographics, personality, political ideology, religiosity, and self-reported strategy usage.
-
 \hypertarget{results}{%
@@
 
-x
+The overall pattern of punitive behaviour in the six economic games was in line
+with previous research and very similar across both countries (Figure
+\ref{fig:plotDecisions}). Participants were generally more likely to punish
+targets who stole from another individual compared to targets who did not steal
+(multilevel logistic regression; \emph{b} =
+1.93, standard error =
+0.27, \emph{p}
+\textless{} .001). Participants were also more likely
+to punish when targets' stealing behaviour generated inequalities,
+specifically in Games E and F (\emph{b} =
+2.42, SE = 0.44,
+\emph{p} \textless{} .001).
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotDecisions-1.pdf}
+\caption{\label{fig:plotDecisions}\emph{Overall pattern of punitive behaviour across all six
+economic games, split by country.} AI = advantageous inequity, DI =
+disadvantageous inequity.}
+\end{figure}
+
+We classified participants into a particular strategy if their behaviour across
+all twelve decisions matched our behavioural predictions shown in Table
+\ref{tab:tableStrategies} exactly. Table \ref{tab:tableStrategyCounts} shows
+the proportion of participants following each strategy, with N/A used to
+represent participants who did not fit exactly into any particular strategy type.
+Overall, 59\%
+of our participants could be classified exactly into one of the strategies. The
+most common strategy in both countries was to never punish across any of the
+games. The next most common strategies were those that care about
+minimising payoff differences (avoid disadvantageous inequity, egalitarian).
+Less common were the behaviour-shaping strategies (deterrent, norm-enforcing),
+the retributive strategy, and the competitive strategies (seek AI, competitive).
+Although participants often punished targets who did not steal
+in the six games (Figure \ref{fig:plotDecisions}), no participants followed the
+antisocial strategy by exclusively punishing targets who did not steal across
+\emph{all} games.
+
+\begin{table}[tbp]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:tableStrategyCounts}Counts and proportions of participants
+following each punishment strategy exactly, split by country. N/A implies that
+participants were unable to be classified exactly into any of the punishment
+strategies.}
+
+\begin{tabular}{lcccc}
+\toprule
+ & \multicolumn{2}{c}{\thead{United Kingdom \\ (N = 1014)}} & \multicolumn{2}{c}{\thead{United States \\ (N = 996)}} \\
+\cmidrule(r){2-3} \cmidrule(r){4-5}
+Strategy & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Prop} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Prop}\\
+\midrule
+Deterrent & 9 & 0.009 & 6 & 0.006\\
+Norm-enforcing & 8 & 0.008 & 16 & 0.016\\
+Retributive & 6 & 0.006 & 5 & 0.005\\
+Avoid DI & 67 & 0.066 & 62 & 0.062\\
+Egalitarian & 65 & 0.064 & 71 & 0.071\\
+Seek AI & 2 & 0.002 & 0 & 0.000\\
+Competitive & 3 & 0.003 & 1 & 0.001\\
+Antisocial & 0 & 0.000 & 0 & 0.000\\
+Never punish & 426 & 0.420 & 447 & 0.449\\
+N/A & 428 & 0.422 & 388 & 0.390\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+To further investigate the strategies that participants were following, we
+examined the most common patterns of punitive behaviour across all twelve
+decisions. Supplementary Table \ref{tab:tablePatterns} shows the proportion of
+participants following the 25 most common behavioural patterns, including,
+where appropriate, the predetermined strategies from Table
+\ref{tab:tableStrategies}. In both countries, a common pattern of behaviour
+not captured by any of the strategies was punishing only when the target stole
+in the third-party game (Game F). Punishment in this game is consistent with an
+egalitarian motive, as stealing produces unequal outcomes, but third-party
+punishment here is also consistent with norm-enforcing and competitive motives
+(see Table \ref{tab:tableStrategies}). Other common behavioural patterns not
+captured by our strategies included punishing whenever the target stole across
+all games and always punishing in every game irrespective of the targets'
+behaviour.
+
+While it is useful to look at exact patterns of behaviour, participants may not
+have implemented their chosen punishment strategy with exact precision.
+In reality, strategies may have been implemented probabilistically for each
+punishment decision. There is also the possibility of implementation errors,
+whereby participants occasionally ``slip up'' and make decisions that are
+incongruent with a particular strategy. This may explain why some participants
+were unable to be classified exactly into a single punishment strategy.
+
+To deal with this complexity and include all observed data in our frequency
+estimates, we fitted a Bayesian latent state model to the data. This model
+assumes that the nine strategies in Table \ref{tab:tableStrategies} (plus a
+``random choice'' strategy that chooses randomly for each decision) are the only
+latent strategies and that these are instantiated into observed behaviour
+according to the logic in Table \ref{tab:tableStrategies} with some
+probability of implementation error (i.e., an intention to punish is implemented
+as non-punishment and vice versa). Averaging over all strategies and
+incorporating the possibility of implementation errors, the model estimates the
+probability of participants following any particular strategy, given the
+observed data.
+
+The posterior estimates from the model are presented in Figure
+\ref{fig:plotModel1b}. The posterior probabilities for each strategy did not
+differ between the two countries. In both countries, the never punish strategy
+had the highest probability, followed by the egalitarian strategy. The
+norm-enforcing and seek AI strategies were the next most likely, with higher
+posterior estimates than the competitive and antisocial strategies. None of the
+other strategies differed in their posterior estimates. The same general pattern
+emerged when we analysed the full dataset without pre-registered exclusions
+(Supplementary Figure \ref{fig:plotModel1a}).
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotModel1b-1.pdf}
+\caption{\label{fig:plotModel1b}\emph{Posterior estimates of the probabilities of following
+different punishment strategies from the Bayesian latent state model.} The model
+assumes an implementation error rate of 5\%. Points represent posterior medians,
+line ranges represent 50\% and 95\% credible intervals.}
+\end{figure}
+
+Next, we explored which traits predicted adherence to different punishment
+strategies. To answer this question, we included variables capturing
+demographics, personality, social preferences, political views, and religious
+views as predictors in our Bayesian latent state model. We included each
+variable in a separate model, predicting all ten punishment strategies
+(the nine from Table 1, plus the `random choice' strategy) simultaneously.
+
+Demographic variables tended to be unrelated to strategy usage: age and gender
+did not predict adherence to a particular punishment strategy (Supplementary
+Figures \ref{fig:plotAllDems2} and \ref{fig:plotAllDems1}). In the United
+States, the never punish strategy was slightly more common among participants
+lower in socio-economic status (median posterior slope =
+-0.20, 95\% CI {[}-0.38
+-0.02{]}) but this effect was small.
+
+Conversely, personality and social preferences were linked to variation in
+punishment strategies. When including the Big-6 personality dimensions and
+Social Value Orientation (SVO) in the model, we found associations with the
+egalitarian, never punish, and random choice strategies (Figure
+\ref{fig:plotAllPers2}). Participants higher in SVO were more likely to follow
+the egalitarian and the never punish strategies, while those with lower SVO
+scores were more likely to enact the random choice strategy. The personality
+dimensions of honesty-humility and openness to experience were both positively
+associated with following the never punish strategy, while extraversion
+negatively predicted this strategy. The effects were mostly similar across
+countries, but occasionally differed: for example, in the United States,
+but not in the United Kingdom, honesty-humility was positively associated with
+following the egalitarian strategy and negatively associated with following
+the random choice strategy. Overall, the same pattern of results emerged when
+analysing the full dataset without exclusions (Supplementary Figure
+\ref{fig:plotAllPers1}).
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAllPers2-1.pdf}
+\caption{\label{fig:plotAllPers2}\emph{Posterior slopes from Bayesian latent state models
+including Big-6 personality dimensions and Social Value Orientation.} Each row
+represents a separate model. All models assume an implementation error rate of
+5\%. Points represent posterior medians, line ranges represent 95\% credible
+intervals.}
+\end{figure}
+
+Political and religious variables were also associated with punishment strategy
+(Figure \ref{fig:plotAllPolRel2}). These effects tended to be more pronounced
+in the United States. Controlling for Social Dominance Orientation, American
+participants higher in Right Wing Authoritarianism were more likely to follow
+the strategies avoiding disadvantageous inequity and seeking advantageous
+inequity. Participants who stated that they would like to ``bring those below
+them {[}on the socio-economic status ladder{]} up a peg'' were more likely to follow
+the egalitarian strategy, while American participants higher in Social Dominance
+Orientation, Right Wing Authoritarianism, and believing that God controls events
+in the world were less likely to follow the egalitarian strategy. In general,
+religious and conservative participants were less likely to follow the never
+punish strategy. This general pattern of results was replicated with the full
+dataset (Supplementary Figure \ref{fig:plotAllPolRel1}).
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAllPolRel2-1.pdf}
+\caption{\label{fig:plotAllPolRel2}\emph{Posterior slopes from Bayesian latent state models
+including political ideology, views about social inequality, and religiosity.}
+Each row represents a separate model aside from Social Dominance Orientation and
+Right Wing Authoritarianism, which control for one another within the same
+model. All models assume an implementation error rate of 5\%. Points represent
+posterior medians, line ranges represent 95\% credible intervals.}
+\end{figure}
+
+Finally, we asked whether participants had insight into their own punishment
+strategy. In other words, could participants self-report the strategy that they
+were following during the games? To answer this question, we included
+participants' responses to post-game questions about their strategy as
+predictors in the model. As before, each predictor was included in a separate
+model, predicting all ten strategies simultaneously.
+
+In general, we found that self-reported strategy usage was positively associated
+with the behavioural strategy that participants employed (see Supplementary
+Figures \ref{fig:plotSliders1} and \ref{fig:plotSliders2} for the distribution
+of responses to self-report questions). Figure \ref{fig:plotAllSliders2} shows
+the relationships between self-report questions and the different punishment
+strategies, highlighting the combinations where the question matched the
+behavioural strategy. We found positive relationships between the self-report
+questions and strategy usage for the norm-enforcing, egalitarian, seek
+advantageous inequity, never punish, and random choice strategies. The 95\%
+credible intervals for other estimates included zero, though these estimates
+often trended in a positive direction. The same pattern of results was found
+when analysing the full dataset without exclusions (Supplementary Figure
+\ref{fig:plotAllSliders1}).
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAllSliders2-1.pdf}
+\caption{\label{fig:plotAllSliders2}\emph{Posterior slopes from models including
+self-reported strategy usage.} Each row represents a separate model. All models
+assume an implementation error rate of 5\%. Highlighted estimates represent
+combinations where the self-report question matched the behavioural strategy.
+Points represent posterior medians, line ranges represent 95\% credible
+intervals.}
+\end{figure}
 
@@
 
-x
+Using a suite of economic games measuring punishment in different situations,
+we have shown that punishment does not serve just one function, but instead is
+a flexible tool that can be and is used for different purposes\textsuperscript{15}.
+Punishment is more akin to a swiss army knife than a hammer, used by some to
+enforce norms of cooperation and by others to reduce or even create inequality
+between individuals. We found that people's punishment strategy can, to some
+extent, be predicted by individual differences in personality, social
+preferences, and political and religious views. Moreover, contrary to the view
+that people are often unable to articulate the reasons for their punitive
+behaviour\textsuperscript{28}, people seem to have some degree of insight into the
+strategy they are using. Despite small differences, these general patterns
+replicated in samples from both the United Kingdom and the United States.
+
+Among the punitive strategies, the most common were particularly sensitive to
+inequality in payoffs, either from a self-referential perspective (i.e., avoid
+disadvantageous inequity) or more generally (i.e., egalitarian). This is in
+line with previous studies which have highlighted inequity aversion as an
+important motivation for punishment in economic games\textsuperscript{29,30}. Personality and social preference variables mapped
+onto these strategies in expected ways. Traits associated with other-regarding
+concern, such as SVO and honesty-humility, predicted following the egalitarian
+strategy, whereas religious and conservative individuals were less likely to
+follow this strategy, especially in the United States. Moreover, participants
+following the egalitarian strategy were able to self-report this strategy,
+though the same was not true for the avoid disadvantageous inequity strategy.
+
+Behaviour shaping strategies, such as deterrence and norm-enforcement, were less
+common than strategies sensitive to inequality in our set of games. This was
+reflected both in participants' elicited punishment behaviour (Figure
+\ref{fig:plotModel1b}) and in their self-reports of their own strategy
+(Supplementary Figures \ref{fig:plotSliders1} and \ref{fig:plotSliders2}).
+Regarding the predictors of these strategies, we found that demographic,
+personality, political, and religious variables tended to be unrelated to
+deterrent and norm-enforcing punishment strategies. We also found that
+participants had insight into the norm-enforcing strategy, but not the deterrent
+strategy. This finding is in line with previous research showing that people
+struggle to accurately report the deterrent motivations for their punitive
+behaviour\textsuperscript{28}.
+
+Other punitive strategies were less common in our dataset, but some were more
+prevalent than others. For example, participants were more likely to use
+punishment to seek advantageous inequity than to exclusively harm those who did
+not steal (i.e., antisocial punishment). The existence of the ``seek AI'' strategy
+in our dataset supports the claim that punishment can also be used as a tool to
+increase one's own relative position\textsuperscript{15}. While generally rare, we
+found that this motive for punishment was more common among authoritarian
+participants, at least in the United States, potentially providing an
+explanation for why peer punishment has been found to be more common among
+conservatives in previous work\textsuperscript{41}. Moreover, the fact that no
+participants in our sample punished non-stealing across all games suggests that
+antisocial punishment does not function to harm cooperators specifically, as has
+been previously suggested\textsuperscript{17}. Instead, antisocial punishment appears
+to be motivated by improving one's relative position in general, which is in
+line with work showing that antisocial punishment disappears with a 1:1 fee-fine
+ratio\textsuperscript{35}.
+
+The fact that people use punishment for many different reasons poses problems
+for the way that punishment is operationalised in classic behavioural economic
+game studies. In these studies, a common assumption is that participants will
+punish to change the behaviour of cheats. But in reality, people may be choosing
+the punishment option to achieve a variety of different goals. The targets of
+punishment in these studies are likely well aware that punishment could be
+levied for these different reasons and this knowledge may impact their
+responses. For example, if cheating targets interpret punishment as serving a
+competitive motive, it may elicit retaliation rather than encourage
+cooperation\textsuperscript{9,15,50}. This might help to
+explain the mixed findings in the field as to whether punishment actually
+motivates cheating targets to cooperate in the future\textsuperscript{15}.
+
+It is striking that the most common strategy in our dataset was to never punish.
+This is partly because punishment in these games imposes an economic cost for no
+tangible benefit. If the fee-fine ratio had been lower such that it was cheaper
+to punish, we may have seen more punishment from participants. Indeed,
+72\%
+of participants following the never punish strategy positively stated that they
+didn't want to pay to reduce anyone's bonus but would have done so if it were
+free. But the frequency of the never punish strategy perhaps also reflects a
+more general aversion to peer punishment, an aversion that has been highlighted
+in both WEIRD (Western, educated, industrialised, rich, and democratic) samples\textsuperscript{51,52} and in small-scale societies\textsuperscript{53}.
+One reason that people may be averse to peer punishment is that, due to its
+multipurpose nature, it may be interpreted as a competitive challenge by targets
+and trigger retaliation\textsuperscript{15}. In situations that lack clear
+institutional norms to legitimise punishment, such as our economic games and
+some situations in the real world, people might abstain from peer punishment to
+avoid such retaliation, regardless of whether retaliation is actually possible.
+By contrast, institutionalised punishment in small-scale societies often
+functions to compensate victims adequately while limiting the potential for
+feuds and cycles of retaliation\textsuperscript{54,55}. Future research
+should uncover whether people are more willing to punish in these
+conventionalised contexts.
+
+There are several limitations with our study design that can guide future
+research. First, we used one-shot economic games to measure punishment
+strategies, which may have led us to underestimate behaviour-change strategies
+like deterrence. Our inclusion of Game C somewhat mitigated this issue by
+manipulating whether stealing behaviour was intentional vs.~unintentional and
+thus whether there was any behaviour to be deterred. Due to limits on our
+within-subjects design and the complexity of the strategy space, it was not
+feasible for us to expand our study to include additional contexts to elicit
+behaviour-change strategies (e.g., repeated games, games where targets are not
+made aware of the punishment, games where targets can retaliate). Future work
+could study these contexts separately.
+
+A second limitation is that some strategies required more punishment than others
+to be met. For example, the competitive strategy required punishment in ten of
+twelve decisions, compared to the avoid disadvantageous inequity strategy which
+required only one instance of punishment (Table \ref{tab:tableStrategies}).
+Strategies thus differed in how ``expensive'' they were to implement, perhaps
+explaining why some strategies were more common than others. This issue is
+largely unavoidable in our design since strategies, by their very nature, differ
+in how punitive they are. To partially mitigate this issue, we employed the
+strategy method to incentivise participants, such that payoffs were calculated
+from a randomly chosen game instead of summed across all games.
+
+Another limitation is that our results may be contingent on the particular suite
+of anonymous stealing games that we used. With our anonymous design, we were
+unable to study other potential reputational strategies underlying punishment,
+such as signalling trustworthiness\textsuperscript{47}. Moreover, stealing may be
+evaluated differently to other forms of cheating, such as not contributing to
+public goods, and other negative behaviours, such as lying or breaking taboos.
+Finally, we were unable to include all permutations of situational features in
+the games (e.g., second-party vs.~third-party, equal vs.~unequal) making it
+difficult to interpret some patterns of behaviour. For example, many
+participants punished only in the third-party game (Game F), but it is not clear
+whether these participants were driven by the third-party nature of the game or
+simply by the fact that stealing in that game generated inequality. Future work
+should determine whether different reputational contexts, target behaviours, and
+combinations of situational features elicit different punishment strategies.
+
+In sum, we have shown that while many people choose not to punish peers, those
+who do are motivated by a variety of different concerns, including behaviour
+shaping, egalitarianism, and competition. Much like the observed variation in
+human social learning strategies\textsuperscript{39}, humans thus also exhibit
+variation in their punishment strategies. These individual differences map onto
+personality dimensions, social preferences, political and religious views, and
+self-reports of behaviour. We hope that future work will continue to unpack the
+multifaceted nature of human punishment.
 
@@
 
-x
+\hypertarget{ethical-approval}{%
+\subsection{Ethical approval}\label{ethical-approval}}
+
+Ethical approval was granted by the University College London Ethics Board
+(project: 3720/002). The study was performed in accordance with all the relevant
+guidelines and regulations. Informed consent was obtained from all participants
+prior to the study.
+
+\hypertarget{pre-registration}{%
+\subsection{Pre-registration}\label{pre-registration}}
+
+We pre-registered the study on the Open Science Framework before collecting data
+in the United Kingdom (11\textsuperscript{th} November 2022; \url{https://osf.io/k75fc}). We submitted
+another pre-registration before collecting data in the United States (20\textsuperscript{th}
+June 2023; \url{https://osf.io/q4hdy}). In the pre-registrations, we outlined our
+study design, exclusion criteria, and analysis plan. As the study was
+exploratory, we did not pre-register any explicit hypotheses. We did not deviate
+from the pre-registrations.
+
+\hypertarget{exclusion-criteria}{%
+\subsection{Exclusion criteria}\label{exclusion-criteria}}
+
+We pre-registered that we would exclude participants who failed any of the
+attention checks, sped through the surveys (i.e., two standard deviations below
+the median duration), or flatlined (i.e., provided identical responses to matrix
+questions). We also stated that we would exclude data for particular games if
+participants failed the comprehension question for that game. We followed our
+pre-registered plan of conducting analyses with and without these exclusions
+(analyses without exclusions are reported in the Supplementary Material).
+
+\hypertarget{participants}{%
+\subsection{Participants}\label{participants}}
+
+We collected a representative sample of 1019 participants
+from the United Kingdom through the online platform Prolific
+(\url{https://www.prolific.com/}). All of these participants completed the economic
+games and 973 returned to complete the follow-up survey a
+week later (95\% retention
+rate). After exclusions, we were left with 1014 participants
+overall (see Supplementary Figure \ref{fig:plotSampleUK} for sample
+characteristics).
+
+We later collected a representative sample of 1005
+participants from the United States through Prolific. All of these participants
+completed the economic games and 957 returned to complete
+the follow-up survey (95\%
+retention rate). After exclusions, we were left with 996
+participants overall (see Supplementary Figure \ref{fig:plotSampleUS} for
+sample characteristics).
+
+\hypertarget{materials}{%
+\subsection{Materials}\label{materials}}
+
+\hypertarget{economic-games}{%
+\subsubsection{Economic games}\label{economic-games}}
+
+In the first part of the study, participants completed six economic games, each
+with slight variations. In all games, there are multiple players and the
+participant takes the role of P1. P2 either (a) steals £0.20 from another player
+and adds it to their payoff or (b) does nothing. For each of these cases,
+participants are asked whether they would like to pay money to reduce P2's
+payoff. Games A-E have two players and Game F has three players.
+
+The six games are as follows (variations bolded; see Figure \ref{fig:plotGames}
+for a visual representation of the games):
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  \emph{Game A (Advantageous Inequity)}. P1 starts with £0.70 and P2 starts with
+  £0.10. P2 is given the option to either steal £0.20 from P1 or do nothing. P1
+  can then pay £0.10 to reduce P2's payoff by £0.30.
+\item
+  \emph{Game B (Equal)}. P1 starts with £0.70 and P2 starts with \textbf{£0.30}. P2 is
+  given the option to either steal £0.20 from P1 or do nothing. P1 can then pay
+  £0.10 to reduce P2's payoff by £0.30.
+\item
+  \emph{Game C (Computer)}. P1 starts with £0.70 and P2 starts with £0.30.
+  Participants are told that \textbf{``the computer will decide''} whether P2 steals
+  £0.20 from P1 or does nothing. P1 can then pay £0.10 to reduce P2's payoff by
+  £0.30.
+\item
+  \emph{Game D (1:1 Fee-Fine)}. P1 starts with £0.70 and P2 starts with £0.30. P2 is
+  given the option to either steal £0.20 from P1 or do nothing. P1 can then pay
+  £0.10 to reduce P2's payoff by \textbf{£0.10}.
+\item
+  \emph{Game E (Disadvantageous Inequity)}. P1 starts with £0.70 and P2 starts with
+  \textbf{£0.50}. P2 is given the option to either steal £0.20 from P1 or do
+  nothing. P1 can then pay £0.10 to reduce P2's payoff by £0.30.
+\item
+  \emph{Game F (Third-Party)}. P1 starts with £1.00, P2 and P3 start with £0.70. P2
+  is given the option to either steal £0.20 \textbf{from P3} or do nothing. P1 can
+  then pay £0.10 to reduce P2's payoff by £0.30.
+\end{enumerate}
+
+For each game, participants saw the game instructions and answered a
+comprehension question before providing their decisions. After completing all
+the games, participants were asked to give an open-ended response explaining
+their behaviour in the games, and then responded to several slider questions
+capturing the different reasons for their decisions (for full wordings, see
+Supplementary Table \ref{tab:tableSliderWordings}).
+
+\hypertarget{survey-questions}{%
+\subsubsection{Survey questions}\label{survey-questions}}
+
+In a follow-up survey, we collected the following data on participants (for
+wordings of all questions, see Supplementary Table
+\ref{tab:tableSurveyWordings}):
+
+\begin{itemize}
+\tightlist
+\item
+  \emph{Demographics}. In the survey, we collected information on participants'
+  education level and self-reported socio-economic status (MacArthur ladder\textsuperscript{56}). We also collected additional demographic data from Prolific
+  (e.g., age, gender, student status).
+\item
+  \emph{Personality}. We used the Mini-IPIP scale\textsuperscript{57} to measure the Big 6
+  personality dimensions of agreeableness, conscientiousness, extraversion,
+  honesty-humility, openness to experience, and neuroticism (four items each).
+\item
+  \emph{Social Value Orientation}. We used the Social Value Orientation Slider
+  Measure to measure other-regarding preferences\textsuperscript{58}. Across fifteen
+  items, participants made decisions on how to allocate different amounts of
+  money between themselves and another anonymous individual. From these
+  decisions, we calculated participants' Social Value Orientation ``angle'' as a
+  measure of their other-regarding preference, following the steps outlined in
+  ref\textsuperscript{58}.
+\item
+  \emph{Political ideology}. We included several measures of political ideology,
+  including left-right conservatism, Social Dominance Orientation\textsuperscript{59}
+  (eight items), and Right Wing Authoritarianism\textsuperscript{60} (six items).
+  We also probed participants' views on social inequality by asking them whether
+  they would like to bring people above (below) them on the MacArthur
+  socio-economic status ladder down (up) a peg or two.
+\item
+  \emph{Religious views}. We asked participants how religious they consider
+  themselves and whether they believe that God or another spiritual non-human
+  entity controls the events in the world\textsuperscript{61}.
+\end{itemize}
+
+\hypertarget{procedure}{%
+\subsection{Procedure}\label{procedure}}
+
+We began data collection in the United Kingdom on 28\textsuperscript{th} November 2022, with
+participants returning to complete the follow-up survey on 5\textsuperscript{th} December 2022.
+We then ran a second wave of data collection in the United States on 20\textsuperscript{th} June
+2023, with participants returning to complete the follow-up survey on 27\textsuperscript{th}
+June 2023. Our surveys were designed through the online survey platform
+Qualtrics (\url{https://www.qualtrics.com/}).
+
+In the initial games survey, participants completed all six economic games in a
+random order, with punishment decisions (whether to punish a stealing target and
+whether to punish a target who did nothing) randomised within games. Responses
+to comprehension questions suggested that participants understood the six
+economic games (Supplementary Table \ref{tab:tableComp}). We used the strategy
+method to incentivise the economic games, choosing a random game to determine
+bonus payment. After all games,
+62\%
+of participants stated that they believed that their decisions had real
+consequences for others.
+
+In the follow-up survey, participants completed blocks of questions on
+demographics, personality, Social Value Orientation, political ideology, and
+religious views in a random order, with questions randomised within blocks. A
+random decision from the Social Value Orientation Slider Measure was chosen to
+determine bonus payment.
+
+Participants were paid £1.80 for completing the games survey, plus a bonus
+payment from the six economic games (between £0.40 -- £0.70 depending on their
+decision). Participants were paid £1.50 for completing the follow-up survey,
+plus a bonus payment from the Social Value Orientation Slider Measure (between
+£0.50 -- £0.85 depending on their decision).
+
+\hypertarget{statistical-analysis}{%
+\subsection{Statistical analysis}\label{statistical-analysis}}
+
+We pre-registered that we would use a Bayesian latent state model to infer
+unobserved punishment strategies from the observed data (for a similar version
+of this model, see ref\textsuperscript{62}). In this model, participants \(i\) in
+countries \(c\) make binary punishment decisions across twelve decisions \(j\). We
+assume that the probability of the observed data \(y_{i,j}\) is the weighted
+average of the probability of the observed data conditional on each of the ten
+punishment strategies \(s\). From this logic, the model estimates the probability
+of each strategy \(p_{s}\). The full model is as follows:
+
+\begin{align}
+y_{i,j} &\sim \text{Bernoulli}(\theta_{j}) \\
+\theta_{j} &= \sum_{s=1}^{10} p_{s} \text{Pr}(\text{punish}|s,j) \nonumber \\
+p &= \text{softmax}(\alpha_{c[i]}) \nonumber \\
+\alpha_{s,c} &\sim \text{Normal(0, 1)} \nonumber
+\end{align}
+
+The conditional probabilities \(\text{Pr}(\text{punish}|s,j)\) are hard coded in
+the model as outlined in Table \ref{tab:tableStrategies}. We incorporate an
+implementation error rate \(\delta\) into these conditional probabilities by coding
+green ticks in Table \ref{tab:tableStrategies} with a conditional probability
+of \(1 - \delta\) and coding red crosses with a conditional probability of
+\(0 + \delta\). The random choice is consistently coded with a conditional
+probability of ½ across all decisions.
+
+To include a categorical predictor in the model, we estimate a different
+\(\alpha_{s,c}\) for each categorical level. To include a continuous predictor
+\(x\) in the model, we include a slope \(\beta\) in the linear model for \(p\):
+
+\begin{align}
+y_{i,j} &\sim \text{Bernoulli}(\theta_{j}) \\
+\theta_{j} &= \sum_{s=1}^{10} p_{s} \text{Pr}(\text{punish}|s,j) \nonumber \\
+p &= \text{softmax}(\alpha_{c[i]} + \beta_{c[i]}x_{i}) \nonumber \\
+\alpha_{s,c} &\sim \text{Normal(0, 1)} \nonumber \\
+\beta_{s,c} &\sim \text{Normal(0, 0.2)} \nonumber
+\end{align}
+
+We estimated the posterior distributions of these models using Hamiltonian Monte
+Carlo as implemented in Stan version 2.26.1\textsuperscript{63}. We ran each model for
+2000 samples, with 1000 warmup samples. R-hat values and effective sample sizes
+suggested that all models converged normally. Trace plots are reported in
+Supplementary Figure \ref{fig:plotTrace}.
+
+We validated the model by simulating observed data (n = 100) from a known
+frequency of strategies. The model was successfully able to recover the known
+frequency of strategies from the simulated data (Supplementary Figure
+\ref{fig:plotSim}).
+
+\hypertarget{reproducibility}{%
+\subsection{Reproducibility}\label{reproducibility}}
+
+All data and code are accessible on GitHub:
+\url{https://github.com/ScottClaessens/punishStrategies}. All analyses were conducted
+in R version 4.2.1\textsuperscript{64}.
+Visualisations were created with the \emph{ggplot2}\textsuperscript{65} and \emph{cowplot}\textsuperscript{66} R packages.
+We used the \emph{targets}\textsuperscript{67} R package to create a reproducible data analysis pipeline and the
+\emph{papaja}\textsuperscript{68} R package to reproducibly generate the manuscript.
 
@@
 
-x
+This work was supported by a Royal Society of New Zealand Catalyst Leaders Grant
+to Q.D.A and N.R. (ref: ILFUOA2002).
 
@@
 
-All data used in this study are publicly available on GitHub: \url{https://github.com/ScottClaessens/punishStrategies}
+All data used in this study are publicly available on GitHub:
+\url{https://github.com/ScottClaessens/punishStrategies}
 
@@
 
-All code to reproduce the analyses in this study are publicly available on GitHub: \url{https://github.com/ScottClaessens/punishStrategies}
+All code to reproduce the analyses in this study are publicly available on
+GitHub: \url{https://github.com/ScottClaessens/punishStrategies}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-Nikiforakis2008}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-Nikiforakis2008a}{}}%
 \CSLLeftMargin{13. }%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Barclay2016}{}}%
+\CSLLeftMargin{33. }%
+\CSLRightInline{Barclay, P.
+\& Raihani, N.
+\href{https://doi.org/10.1016/j.evolhumbehav.2015.12.004}{Partner choice versus punishment in human prisoner's dilemmas}.
+\emph{Evolution and Human Behavior} \textbf{37}, 263--271 (2016).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Crockett2014}{}}%
-\CSLLeftMargin{33. }%
+\CSLLeftMargin{34. }%
 \CSLRightInline{Crockett, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Sylwester2013}{}}%
-\CSLLeftMargin{34. }%
+\CSLLeftMargin{35. }%
 \CSLRightInline{Sylwester, K., Herrmann, B.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Fehr2004}{}}%
-\CSLLeftMargin{35. }%
+\CSLLeftMargin{36. }%
 \CSLRightInline{Fehr, E. \& Fischbacher, U. Third-party punishment and social norms. \emph{Evolution and Human Behavior} \textbf{25}, 63--87 (2004).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Johnson2009}{}}%
-\CSLLeftMargin{36. }%
+\CSLLeftMargin{37. }%
 \CSLRightInline{Johnson, T., Dawes, C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Fowler2005}{}}%
-\CSLLeftMargin{37. }%
+\CSLLeftMargin{38. }%
 \CSLRightInline{Fowler, J. H., Johnson, T. \& Smirnov, O. \href{https://doi.org/10.1038/nature03256}{Egalitarian motive and altruistic punishment}. \emph{Nature} \textbf{433}, E1 (2005).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Molleman2014}{}}%
-\CSLLeftMargin{38. }%
+\CSLLeftMargin{39. }%
 \CSLRightInline{Molleman, L., Van den Berg, P.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Thielmann2020}{}}%
-\CSLLeftMargin{39. }%
+\CSLLeftMargin{40. }%
 \CSLRightInline{Thielmann, I., Spadaro, G.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Hofmann2018}{}}%
-\CSLLeftMargin{40. }%
+\CSLLeftMargin{41. }%
 \CSLRightInline{Hofmann, W., Brandt, M.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Nisbett1977}{}}%
-\CSLLeftMargin{41. }%
+\CSLLeftMargin{42. }%
 \CSLRightInline{Nisbett, R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Barclay2006}{}}%
-\CSLLeftMargin{42. }%
+\CSLLeftMargin{43. }%
 \CSLRightInline{Barclay, P.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Batistoni2022}{}}%
-\CSLLeftMargin{43. }%
+\CSLLeftMargin{44. }%
 \CSLRightInline{Batistoni, T., Barclay, P.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Jordan2016}{}}%
-\CSLLeftMargin{44. }%
+\CSLLeftMargin{45. }%
 \CSLRightInline{Jordan, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Jordan2017}{}}%
-\CSLLeftMargin{45. }%
+\CSLLeftMargin{46. }%
 \CSLRightInline{Jordan, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Jordan2020}{}}%
-\CSLLeftMargin{46. }%
+\CSLLeftMargin{47. }%
 \CSLRightInline{Jordan, J.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Deutchman2021}{}}%
-\CSLLeftMargin{47. }%
+\CSLLeftMargin{48. }%
 \CSLRightInline{Deutchman, P., Bračič, M., Raihani, N.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Marczyk2017}{}}%
-\CSLLeftMargin{48. }%
+\CSLLeftMargin{49. }%
 \CSLRightInline{Marczyk, J. \href{https://doi.org/10.1371/journal.pone.0171298}{Human punishment is not primarily motivated by inequality}. \emph{PLoS ONE} \textbf{12}, e0171298 (2017).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Nikiforakis2008b}{}}%
+\CSLLeftMargin{50. }%
+\CSLRightInline{Nikiforakis, N.
+\href{https://doi.org/10.1016/j.jpubeco.2007.04.008}{Punishment and counter-punishment in public good games: Can we really govern ourselves?} \emph{Journal of Public Economics} \textbf{92}, 91--112 (2008).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Balafoutas2014}{}}%
+\CSLLeftMargin{51. }%
+\CSLRightInline{Balafoutas, L., Nikiforakis, N.
+\& Rockenbach, B.
+\href{https://doi.org/10.1073/pnas.1413170111}{Direct and indirect punishment among strangers in the field}.
+\emph{Proceedings of the National Academy of Sciences} \textbf{111}, 15924--15927 (2014).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Balafoutas2016}{}}%
+\CSLLeftMargin{52. }%
+\CSLRightInline{Balafoutas, L., Nikiforakis, N.
+\& Rockenbach, B.
+\href{https://doi.org/10.1038/ncomms13327}{Altruistic punishment does not increase with the severity of norm violations in the field}.
+\emph{Nature Communications} \textbf{7}, 13327 (2016).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Baumard2010}{}}%
+\CSLLeftMargin{53. }%
+\CSLRightInline{Baumard, N.
+\href{https://doi.org/10.1007/s11299-010-0079-9}{Has punishment played a role in the evolution of cooperation?
+A critical review}.
+\emph{Mind \& Society} \textbf{9}, 171--192 (2010).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Fitouchi2023}{}}%
+\CSLLeftMargin{54. }%
+\CSLRightInline{Fitouchi, L.
+\& Singh, M.
+\href{https://doi.org/10.1016/j.evolhumbehav.2023.03.001}{Punitive justice serves to restore reciprocal cooperation in three small-scale societies}.
+\emph{Evolution and Human Behavior} \textbf{44}, 502--514 (2023).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Singh2022}{}}%
+\CSLLeftMargin{55. }%
+\CSLRightInline{Singh, M. \& Garfield, Z. H. Evidence for third-party mediation but not punishment in {M}entawai justice. \emph{Nature Human Behaviour} \textbf{6}, 930--940 (2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Adler2000}{}}%
+\CSLLeftMargin{56. }%
+\CSLRightInline{Adler, N.
+E., Epel, E.
+S., Castellazzo, G.
+\& Ickovics, J.
+R.
+\href{https://doi.org/10.1037/0278-6133.19.6.586}{Relationship of subjective and objective social status with psychological and physiological functioning: Preliminary data in healthy, {W}hite women}.
+\emph{Health Psychology} \textbf{19}, 586--592 (2000).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Sibley2011}{}}%
+\CSLLeftMargin{57. }%
+\CSLRightInline{Sibley, C.
+G.
+\emph{et al.} The {M}ini-{IPIP}6: Validation and extension of a short measure of the {B}ig-{S}ix factors of personality in {N}ew {Z}ealand.
+\emph{New Zealand Journal of Psychology (Online)} \textbf{40}, 142 (2011).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Murphy2011}{}}%
+\CSLLeftMargin{58. }%
+\CSLRightInline{Murphy, R.
+O., Ackermann, K.
+A.
+\& Handgraaf, M.
+J.
+J.
+\href{https://doi.org/10.1017/S1930297500004204}{Measuring social value orientation}.
+\emph{Judgment and Decision Making} \textbf{6}, 771--781 (2011).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Ho2015}{}}%
+\CSLLeftMargin{59. }%
+\CSLRightInline{Ho, A.
+K.
+\emph{et al.} \href{https://doi.org/10.1037/pspi0000033}{The nature of social dominance orientation: Theorizing and measuring preferences for intergroup inequality using the new SDO7 scale}.
+\emph{Journal of Personality and Social Psychology} \textbf{109}, 1003 (2015).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Bizumic2018}{}}%
+\CSLLeftMargin{60. }%
+\CSLRightInline{Bizumic, B.
+\& Duckitt, J.
+\href{https://doi.org/10.5964/jspp.v6i1.835}{Investigating right wing authoritarianism with a {V}ery {S}hort {A}uthoritarianism {S}cale}.
+\emph{Journal of Social and Political Psychology} \textbf{6}, 129--150 (2018).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Laurin2012}{}}%
+\CSLLeftMargin{61. }%
+\CSLRightInline{Laurin, K., Shariff, A.
+F., Henrich, J.
+\& Kay, A.
+C.
+\href{https://doi.org/10.1098/rspb.2012.0615}{Outsourcing punishment to god: Beliefs in divine control reduce earthly punishment}.
+\emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{279}, 3272--3281 (2012).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-McElreath2020}{}}%
+\CSLLeftMargin{62. }%
+\CSLRightInline{McElreath, R. \emph{\href{http://xcelab.net/rm/statistical-rethinking/}{Statistical rethinking: A {Bayesian} course with examples in {R} and {Stan}, 2nd edition}}. (CRC Press, 2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Stan2020}{}}%
+\CSLLeftMargin{63. }%
+\CSLRightInline{Stan Development Team. \href{http://mc-stan.org/}{{RStan}: The {R} interface to {Stan}}. (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
+\CSLLeftMargin{64. }%
+\CSLRightInline{R Core Team. \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}}. (R Foundation for Statistical Computing, 2022).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
+\CSLLeftMargin{65. }%
+\CSLRightInline{Wickham, H. \emph{\href{https://ggplot2.tidyverse.org}{{ggplot2}: Elegant graphics for data analysis}}. (Springer-Verlag New York, 2016).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
+\CSLLeftMargin{66. }%
+\CSLRightInline{Wilke, C. O. \emph{\href{https://CRAN.R-project.org/package=cowplot}{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}}. (2020).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
+\CSLLeftMargin{67. }%
+\CSLRightInline{Landau, W.
+M.
+\href{https://doi.org/10.21105/joss.02959}{The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing}.
+\emph{Journal of Open Source Software} \textbf{6}, 2959 (2021).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
+\CSLLeftMargin{68. }%
+\CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}}. (2022).}
 
@@
 \renewcommand{\tablename}{Supplementary Table}
-\renewcommand{\thefigure}{\arabic{figure}} \setcounter{figure}{0}
-\renewcommand{\thetable}{\arabic{table}} \setcounter{table}{0}
+\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
+\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
 \renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}
 
-\hypertarget{supplementary-information}{%
-\section{\texorpdfstring{\textbf{Supplementary Information}}{Supplementary Information}}\label{supplementary-information}}
+\hypertarget{supplementary-material}{%
+\section{\texorpdfstring{\textbf{Supplementary Material}}{Supplementary Material}}\label{supplementary-material}}
 
@@
 
-\noindent \hspace*{10mm} \small Mapping the punishment strategy space \newline
+\noindent \hspace*{10mm} \small Why do people punish? Evidence for a range of strategic concerns \newline
 \hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, Nichola Raihani\textsuperscript{1,2} \newline
@@
 
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotModel1a-1.pdf}
+\caption{\label{fig:plotModel1a}\emph{Posterior estimates of the probabilities of following
+different punishment strategies from the Bayesian latent state model fitted to
+the full dataset without pre-registered exclusions.} The model assumes an
+implementation error rate of 5\%. Points represent posterior medians, line ranges
+represent 50\% and 95\% credible intervals.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAllDems2-1.pdf}
+\caption{\label{fig:plotAllDems2}\emph{Posterior slopes from models including age,
+socio-economic status, gender, and student status, fitted to the subsetted
+dataset with pre-registered exclusions.} Each row represents a separate model.
+Points represent posterior medians, line ranges represent 95\% credible intervals.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAllDems1-1.pdf}
+\caption{\label{fig:plotAllDems1}\emph{Posterior slopes from models including age,
+socio-economic status, gender, and student status, fitted to the full dataset
+without pre-registered exclusions.} Each row represents a separate model. Points
+represent posterior medians, line ranges represent 95\% credible intervals.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAllPers1-1.pdf}
+\caption{\label{fig:plotAllPers1}\emph{Posterior slopes from models including Big-6
+personality dimensions and Social Value Orientation, fitted to the full dataset
+without pre-registered exclusions.} Each row represents a separate model. Points
+represent posterior medians, line ranges represent 95\% credible intervals.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAllPolRel1-1.pdf}
+\caption{\label{fig:plotAllPolRel1}\emph{Posterior slopes from models including political
+ideology, views about social inequality, and religiosity, fitted to the full
+dataset without pre-registered exclusions.} Each row represents a separate model
+aside from Social Dominance Orientation and Right Wing Authoritarianism, which
+control for one another within the same model. Points represent posterior
+medians, line ranges represent 95\% credible intervals.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotSliders1-1.pdf}
+\caption{\label{fig:plotSliders1}\emph{Boxplots showing the distribution of responses to each
+self-report question about the reasons for participants' behaviour in the games.}
+Boxplots represent medians and interquartile ranges.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotSliders2-1.pdf}
+\caption{\label{fig:plotSliders2}\emph{Boxplots showing the distribution of responses to each
+self-report question about the reasons for participants' behaviour in the games,
+presented as deviations from participants' average rating across all questions.}
+Boxplots represent medians and interquartile ranges.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotAllSliders1-1.pdf}
+\caption{\label{fig:plotAllSliders1}\emph{Posterior slopes from models including
+self-reported strategy usage, fitted to the full dataset without pre-registered
+exclusions.} Each row represents a separate model. Highlighted estimates
+represent combinations where the self-report slider matched the behavioural
+strategy. Each strategy had an associated self-report slider except for the
+competitive strategy. Points represent posterior medians, line ranges represent
+95\% credible intervals.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotSampleUK-1.pdf}
+\caption{\label{fig:plotSampleUK}\emph{Sample characteristics in the United Kingdom.}}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotSampleUS-1.pdf}
+\caption{\label{fig:plotSampleUS}\emph{Sample characteristics in the United States.}}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotTrace-1.pdf}
+\caption{\label{fig:plotTrace}\emph{Trace plots for parameter values from the Bayesian
+latent state model fitted to data with exclusions.}}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotSim-1.pdf}
+\caption{\label{fig:plotSim}\emph{Results of Bayesian latent state model fitted to simulated
+data (n = 100) with known strategy frequencies in the population.} Blue points
+represent known strategy frequencies, grey densities represent posterior
+estimates of strategy frequencies.}
+\end{figure}
+
+\newpage
+
 \hypertarget{supplementary-tables}{%
@@
 
-\hypertarget{supplementary-references}{%
-\subsection{Supplementary References}\label{supplementary-references}}
+\begin{table}[H]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:tablePatterns}Counts and proportions of the 25 most common patterns
+of punitive behaviour across all twelve decisions, split by country. \emph{Binary
+strings represent punishment (1) or no punishment (0) in each decision, aligning
+with the order of game decision columns in Table \ref{tab:tableStrategies}.}}
+
+\footnotesize{
+
+\begin{tabular}{llcccc}
+\toprule
+ &  & \multicolumn{2}{c}{\thead{United Kingdom \\ (N = 1014)}} & \multicolumn{2}{c}{\thead{United States \\ (N = 996)}} \\
+\cmidrule(r){3-4} \cmidrule(r){5-6}
+Pattern & \multicolumn{1}{c}{Explanation} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Prop} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Prop}\\
+\midrule
+000000000000 & \textit{Never punish strategy (exact)} & 426 & 0.420 & 447 & 0.449\\
+000000001000 & \textit{Avoid DI strategy (exact)} & 67 & 0.066 & 62 & 0.062\\
+000000001010 & \textit{Egalitarian strategy (exact)} & 65 & 0.064 & 71 & 0.071\\
+000000000010 & Punish when take in Game F & 55 & 0.054 & 49 & 0.049\\
+001000001000 & Punish when take in Games B and E & 14 & 0.014 & 11 & 0.011\\
+101000001010 & Punish when take in Games A, B, E, and F & 11 & 0.011 & 4 & 0.004\\
+100000000000 & Punish when take in Game A & 10 & 0.010 & 2 & 0.002\\
+000000100000 & Punish when take in Game D & 9 & 0.009 & 3 & 0.003\\
+001000001010 & Punish when take in Games B, E, and F & 9 & 0.009 & 17 & 0.017\\
+101000101000 & \textit{Deterrent strategy (exact)} & 9 & 0.009 & 6 & 0.006\\
+101010101010 & Punish when take in all games & 9 & 0.009 & 15 & 0.015\\
+101000101010 & \textit{Norm-enforcing strategy (exact)} & 8 & 0.008 & 16 & 0.016\\
+001000000000 & Punish when take in Game B & 7 & 0.007 & 4 & 0.004\\
+001010101000 & Punish when take in Games B, C, D, and E & 7 & 0.007 & 0 & 0.000\\
+100000001000 & Punish when take in Games A and E & 6 & 0.006 & 5 & 0.005\\
+101000001000 & Punish when take in Games A, B, and E & 6 & 0.006 & 7 & 0.007\\
+101010101000 & \textit{Retributive strategy (exact)} & 6 & 0.006 & 5 & 0.005\\
+111111111111 & Always punish & 6 & 0.006 & 16 & 0.016\\
+000000101000 & Punish when take in Games D and E & 5 & 0.005 & 2 & 0.002\\
+000000101010 & Punish when take in Games D, E, and F & 5 & 0.005 & 3 & 0.003\\
+101010001010 & Punish when take in all games except Game D & 5 & 0.005 & 2 & 0.002\\
+001000101000 & Punish when take in Games B, D, and E & 4 & 0.004 & 2 & 0.002\\
+001000101010 & Punish when take in Games B, D, E, and F & 4 & 0.004 & 6 & 0.006\\
+101000000000 & Punish when take in Games A and B & 4 & 0.004 & 2 & 0.002\\
+101010001000 & Punish when take in Games A, B, C, and E & 4 & 0.004 & 0 & 0.000\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\newpage
+
+\begin{table}[H]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:tableSliderWordings}Wordings for 11 self-report slider questions
+asking participants to report the reasons for their behaviour in the six games.
+\emph{Participants were prompted with the following text: ``We would now like you to
+answer a few questions about your main motivation in the games. Please answer
+truthfully - there is no right or wrong answer and your first answer is probably
+best. Please rate the extent to which the following statements apply to your
+decisions to reduce or not to reduce other players' bonuses in the games.''}}
+
+\footnotesize{
+
+\begin{tabular}{ll}
+\toprule
+Slider & \multicolumn{1}{c}{Wording}\\
+\midrule
+1 & I wanted to punish people who harmed others\\
+2 & I wanted to have a higher final bonus than others\\
+3 & I wanted to avoid having a lower final bonus than others\\
+4 & I wanted all players to have the same final bonus\\
+5 & I wanted to stop others from cheating\\
+6 & I wanted to show that I disapproved of others' actions\\
+7 & I made decisions at random\\
+8 & I wanted to punish people who DID NOT harm me or others\\
+9 & I didn't want to reduce anyone's bonus, no matter what they did\\
+10 & I didn't want to PAY to reduce anyone's bonus but I would have done so if it were free\\
+11 & I wanted to punish people if they harmed me but not if they harmed others\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\newpage
+
+\begin{lltable}
+
+\begin{longtable}{p{6cm}p{11cm}p{3cm}}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:tableSurveyWordings}Wordings for survey questions in the study.}\\
+\toprule
+Measure & \multicolumn{1}{c}{Wording} & \multicolumn{1}{c}{Scale}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:tableSurveyWordings} continued}}\\
+\toprule
+Measure & \multicolumn{1}{c}{Wording} & \multicolumn{1}{c}{Scale}\\
+\midrule
+\endhead
+Demographics & What is your highest level of education? & \\
+ & Where would you place yourself on this ladder? Please indicate which number on the rung best represents where you stand at this time in your life, relative to other people in your country & \\
+ & Please could you tell us roughly how many years have you lived in your current country of residence? & \\
+Big 6 Extraversion & I am the life of the party & 1-7\\
+ & I don't talk a lot (reversed) & 1-7\\
+ & I keep in the background (reversed) & 1-7\\
+ & I talk to a lot of different people at parties & 1-7\\
+Big 6 Agreeableness & I sympathise with others' feelings & 1-7\\
+ & I am not interested in other people's problems (reversed) & 1-7\\
+ & I feel others' emotions & 1-7\\
+ & I am not really interested in others (reversed) & 1-7\\
+Big 6 Conscientiousness & I get chores done right away & 1-7\\
+ & I like order & 1-7\\
+ & I make a mess of things (reversed) & 1-7\\
+ & I often forget to put things back in their proper place (reversed) & 1-7\\
+Big 6 Neuroticism & I have frequent mood swings & 1-7\\
+ & I am relaxed most of the time (reversed) & 1-7\\
+ & I get upset easily & 1-7\\
+ & I seldom feel blue (reversed) & 1-7\\
+Big 6 Openness to experience & I have a vivid imagination & 1-7\\
+ & I have difficulty understanding abstract ideas & 1-7\\
+ & I do not have a good imagination (reversed) & 1-7\\
+ & I am not interested in abstract ideas (reversed) & 1-7\\
+Big 6 Honesty-humility & I feel entitled to more of everything (reversed) & 1-7\\
+ & I deserve more things in life (reversed) & 1-7\\
+ & I would like to be seen driving around in a very expensive car (reversed) & 1-7\\
+ & I would get a lot of pleasure from owning expensive luxury goods (reversed) & 1-7\\
+Social Value Orientation & Please indicate how you would like to distribute money between yourself and the other player & 9 choices\\
+Left-right political ideology & Political views are often organised on a single scale from left to right.
+For example, in the United States, the Democratic Party is described as more to the left and the Republican Party is described as more to the right.
+If you had to place your political views on this left-right scale, generally speaking, where would you put yourself?
+& 0-100 slider\\
+Social Dominance Orientation & An ideal society requires some groups to be on top and others to be on the bottom & 1-7\\
+ & Some groups of people are simply inferior to other groups & 1-7\\
+ & No one group should dominate in society (reversed) & 1-7\\
+ & Groups at the bottom are just as deserving as groups at the top (reversed) & 1-7\\
+ & Group equality should not be our primary goal & 1-7\\
+ & It is unjust to try to make groups equal & 1-7\\
+ & We should do what we can to equalize conditions for different groups (reversed) & 1-7\\
+ & We should work to give all groups an equal chance to succeed (reversed) & 1-7\\
+Right Wing Authoritarianism & It's great that many young people today are prepared to defy authority (reversed) & 1-9\\
+ & What our country needs most is discipline, with everyone following our leaders in unity & 1-9\\
+ & God's laws about abortion, pornography, and marriage must be strictly followed before it is too late & 1-9\\
+ & There is nothing wrong with premarital sexual intercourse (reversed) & 1-9\\
+ & Our society does NOT need tougher government and stricter laws (reversed) & 1-9\\
+ & The facts on crime and the recent public disorders show we have to crack down harder on troublemakers, if we are going to preserve law and order & 1-9\\
+Views on social inequality & I would like to bring the people above me on the ladder down a peg or two & 1-7\\
+ & I would like to bring the people below me on the ladder up a peg or two & 1-7\\
+Religious views & How religious are you? & 1-5\\
+ & It is likely that God, or some other type of spiritual non-human entity, controls the events in the world & 1-7\\
+\bottomrule
+\end{longtable}
+
+\end{lltable}
+
+\newpage
+
+\begin{table}[H]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:tableComp}Proportions of correct answers to comprehension
+questions for all six economic games, split by country.}
+
+\begin{tabular}{lll}
+\toprule
+Game & \multicolumn{1}{c}{United Kingdom} & \multicolumn{1}{c}{United States}\\
+\midrule
+Game A (AI) & 0.96 & 0.94\\
+Game B (Equal) & 0.95 & 0.93\\
+Game C (Computer) & 0.95 & 0.95\\
+Game D (1:1 Fee-Fine) & 0.95 & 0.94\\
+Game E (DI) & 0.96 & 0.94\\
+Game F (Third-Party) & 0.95 & 0.94\\
+\bottomrule
+\end{tabular}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 

### 2023-11-24 report.Rmd
finished manuscript first draft

@@
 
-We assume here that people are following a particular strategy exactly and are not comitting any errors or mistakes.
-However, it is possible that some participants intended to follow a particular strategy, but made a mistake on a particular decision.
-We repeat the counts and proportions in the table above, but include behavioural patterns that have one mistake from the exact strategy (note that the proportions will no longer sum to 1).
-
-```{r strategyCountsMistake1}
-kable(tar_read(tableStrategyCounts_Mistake1), booktabs = TRUE)
-```
-
-And including behavioural patterns that have one or two mistakes.
-
-```{r strategyCountsMistake2}
-kable(tar_read(tableStrategyCounts_Mistake2), booktabs = TRUE)
-```
-
-And finally, including behavioural patterns that have one, two, or three mistakes.
-
-```{r strategyCountsMistake3}
-kable(tar_read(tableStrategyCounts_Mistake3), booktabs = TRUE)
-```
-
 We can also look at the most common behavioural patterns across all six games (twelve punishment decisions in total).
@@
 ```{r echo=F}
-tar_read(plotDemsNoTPP)
+tar_read(plotAllDems2)
 ```
@@
 ```{r echo=F}
-tar_read(plotSlidersNoTPP)
+tar_read(plotAllSliders2)
 ```
@@
 ```{r echo=F}
-tar_read(plotPersNoTPP)
+tar_read(plotAllPers2)
 ```
@@
 ```{r echo=F}
-tar_read(plotPolRelNoTPP)
+tar_read(plotAllPolRel2)
 ```

### 2023-11-24 R/functions_loadData.R

@@ -325,20 +325,18 @@ loadData <- function(fileStudy1, fileStudy2) {
                                             pun3_1 == 0, pun3_2 == 0, pun4_1 == 0, pun4_2 == 0,
                                             pun5_1 == 0, pun5_2 == 0, pun6_1 == 0, pun6_2 == 0)),
         # exact strategy classification (i.e. if any of the above == 1)
-        strategy = ifelse(propStrategy_Competitive   == 1, "Competitive",                       "N/A"),
-        strategy = ifelse(propStrategy_AvoidDI       == 1, "Avoid disadvantageous inequity",    strategy),
-        strategy = ifelse(propStrategy_Egalitarian   == 1, "Egalitarian",                       strategy),
-        strategy = ifelse(propStrategy_SeekAI        == 1, "Seek advantageous inequity",        strategy),
-        strategy = ifelse(propStrategy_Retributive   == 1, "Retributive",                       strategy),
-        strategy = ifelse(propStrategy_Deterrent     == 1, "Deterrent",                         strategy),
-        strategy = ifelse(propStrategy_NormEnforcing == 1, "Norm-enforcing",                    strategy),
-        strategy = ifelse(propStrategy_Antisocial    == 1, "Exclusively antisocial punishment", strategy),
-        strategy = ifelse(propStrategy_NoPunish      == 1, "Anti-punish",                       strategy),
-        strategy = factor(strategy, levels = c("Competitive", "Avoid disadvantageous inequity",
-                                               "Egalitarian", "Seek advantageous inequity",
-                                               "Retributive", "Deterrent", "Norm-enforcing",
-                                               "Exclusively antisocial punishment",
-                                               "Anti-punish", "N/A"))
+        strategy = ifelse(propStrategy_Competitive   == 1, "Competitive",    "N/A"),
+        strategy = ifelse(propStrategy_AvoidDI       == 1, "Avoid DI",       strategy),
+        strategy = ifelse(propStrategy_Egalitarian   == 1, "Egalitarian",    strategy),
[…]

### 2023-11-24 R/functions_modelFitting.R

@@ -2,25 +2,25 @@
 
 # fit stan model without predictor
 fitModel1 <- function(d, compiledModel1, error = 0, 
-                      iter = 2000,  warmup = 1000, chains = 4, cores = 4) {
+                      iter = 2000, warmup = 1000, chains = 4, cores = 4) {
   # list for stan
   dataList <-
     list(
       N = nrow(d),
       Nc = length(unique(d$Country)),
       countryID = ifelse(d$Country == "United Kingdom", 1L, 2L),
-      pun1_1 = d$pun1_1,
-      pun1_2 = d$pun1_2,
-      pun2_1 = d$pun2_1,
-      pun2_2 = d$pun2_2,
-      pun3_1 = d$pun3_1,
-      pun3_2 = d$pun3_2,
-      pun4_1 = d$pun4_1,
-      pun4_2 = d$pun4_2,
-      pun5_1 = d$pun5_1,
-      pun5_2 = d$pun5_2,
-      pun6_1 = d$pun6_1,
-      pun6_2 = d$pun6_2,
+      pun1_1 = ifelse(!is.na(d$NoDI1_Take),    d$pun1_1, -999),
+      pun1_2 = ifelse(!is.na(d$NoDI1_Nothing), d$pun1_2, -999),
+      pun2_1 = ifelse(!is.na(d$NoDI2_Take),    d$pun2_1, -999),
+      pun2_2 = ifelse(!is.na(d$NoDI2_Nothing), d$pun2_2, -999),
+      pun3_1 = ifelse(!is.na(d$NoDI3_Take),    d$pun3_1, -999),
+      pun3_2 = ifelse(!is.na(d$NoDI3_Nothing), d$pun3_2, -999),
+      pun4_1 = ifelse(!is.na(d$NoDI4_Take),    d$pun4_1, -999),
+      pun4_2 = ifelse(!is.na(d$NoDI4_Nothing), d$pun4_2, -999),
+      pun5_1 = ifelse(!is.na(d$DI_Take),       d$pun5_1, -999),
+      pun5_2 = ifelse(!is.na(d$DI_Nothing),    d$pun5_2, -999),
+      pun6_1 = ifelse(!is.na(d$`3PP_Take`),    d$pun6_1, -999),
+      pun6_2 = ifelse(!is.na(d$`3PP_Nothing`), d$pun6_2, -999),
       error = error # assumed error rate
     )
   # fit model
[…]

### 2023-11-24 R/functions_modelResults.R

@@ -1,11 +1,11 @@
 # custom functions
 
 # plot model results
-plotModel1.1 <- function(post) {
+plotModel1 <- function(post, file) {
   # strategy vector
   strategies <- c("Competitive", "Avoid DI", "Egalitarian", "Seek AI",
                   "Retributive", "Deterrent", "Norm-enforcing", 
-                  "Antisocial", "Random choice", "No punish")
+                  "Antisocial", "Random choice", "Never punish")
   # calculate probabilities
   P <- post$alpha
   for (i in 1:nrow(P)) # iterations
@@ -21,7 +21,11 @@ plotModel1.1 <- function(post) {
       strategy = rep(rep(strategies, each = length(P[,1,1])), times = 2),
       Country = rep(c("United Kingdom", "United States"), each = length(P[,1,1]) * 10)
     ) %>%
-    mutate(strategy = factor(strategy, levels = strategies)) %>%
+    mutate(strategy = factor(strategy, 
+                             levels = c("Deterrent", "Norm-enforcing", "Retributive", 
+                                        "Avoid DI", "Egalitarian", "Seek AI",
+                                        "Competitive", "Antisocial", 
+                                        "Never punish", "Random choice"))) %>%
     ggplot(aes(x = p, y = fct_rev(strategy), colour = Country)) +
     #geom_density_ridges(rel_min_height = 0.02, colour = "white", scale = 1) +
     stat_pointinterval(position = position_dodge(width = 0.5)) +
@@ -30,16 +34,16 @@ plotModel1.1 <- function(post) {
     ylab(NULL) +
     theme_classic()
   # save
-  ggsave(out, filename = "figures/modelResults/noTPP/model1.pdf", height = 5, width = 7)
+  ggsave(out, filename = file, height = 5, width = 7)
   return(out)
 }
 
 # plot effect of predictor on strategy usage
-plotModel2.1 <- function(d, post, pred, xlab, xBreaks, file) {
+plotModel2 <- function(d, post, pred, xlab, xBreaks, file) {
   # strategy vector
   strategies <- c("Competitive", "Avoid DI", "Egalitarian", "Seek AI",
                   "Retributive", "Deterrent", "Norm-enforcing", 
-                  "Antisocial", "Random choice", "Anti-punish")
+                  "Antisocial", "Random choice", "Never punish")
   # mean, sd, min, and max for predictor
[…]

### 2023-11-24 R/functions_simulation.R

@@ -67,6 +67,20 @@ simulateData <- function(n = 500, seed = 2113, errorRate = 0, alphas, betas) {
                 pun1_1, pun1_2, pun2_1, pun2_2, 
                 pun3_1, pun3_2, pun4_1, pun4_2,
                 pun5_1, pun5_2, pun6_1, pun6_2,
+                # empty strings for raw behaviours
+                # (necessary to avoid missing data deletion in model fitting function)
+                NoDI1_Take    = " ", 
+                NoDI1_Nothing = " ",
+                NoDI2_Take    = " ",
+                NoDI2_Nothing = " ",
+                NoDI3_Take    = " ",
+                NoDI3_Nothing = " ",
+                NoDI4_Take    = " ",
+                NoDI4_Nothing = " ",
+                DI_Take       = " ",
+                DI_Nothing    = " ",
+                `3PP_Take`    = " ",
+                `3PP_Nothing` = " ",
                 # country ID necessary for stan...
                 Country = "United Kingdom")
   return(out)

### 2023-11-24 R/functions_summariseData.R

@@ -119,12 +119,12 @@ plotWordcloud <- function(d, country = "") {
 plotPunDecisions <- function(d) {
   # games vector
   games <- c(
-    "No Disadvantageous\nInequity 1",
-    "No Disadvantageous\nInequity 2",
-    "No Disadvantageous\nInequity 3 (Computer)",
-    "No Disadvantageous\nInequity 4 (1:1 Fee Fine)",
-    "Disadvantageous\nInequity",
-    "Third-Party"
+    "Game A\n(AI)",
+    "Game B\n(Equal)",
+    "Game C\n(Computer)",
+    "Game D\n(1:1 Fee Fine)",
+    "Game E\n(DI)",
+    "Game F\n(Third-Party)"
   )
   # plot
   out <-
@@ -148,13 +148,13 @@ plotPunDecisions <- function(d) {
       Game = ifelse(Game == "pun6", games[6], Game),
       Game = factor(Game, levels = games),
       # other labels
-      Other = factor(ifelse(Other == 1, "Take", "Do nothing"), levels = c("Take", "Do nothing")),
+      Other = factor(ifelse(Other == 1, "Steal", "No steal"), levels = c("Steal", "No steal")),
       # punishment
       Punishment = Punishment
     ) %>%
     # summarise
     group_by(Country, Game, Other) %>%
-    summarise(Punishment = mean(Punishment), .groups = "drop") %>%
+    summarise(Punishment = mean(Punishment, na.rm = TRUE), .groups = "drop") %>%
     # plot
     ggplot(aes(x = Game, y = Punishment, fill = Other)) +
     geom_col(position = "dodge") +
@@ -261,17 +261,17 @@ plotSurveyCorrelations <- function(d) {
   out <-
     d %>%
     transmute(
-      PolIdeology = PolSlider,
+      `Political ideology` = PolSlider,
       SDO = SDO,
       RWA = RWA,
       Religiosity = Religiosity,
-      OpennessToExperience = Open,
+      `Openness to experience` = Open,
       Conscientiousness = Consc,
       Extraversion = Extra,
       Agreeableness = Agree,
       Neuroticism = Neur,
-      HonestyHumility = Honest,
-      SVOAngle = SVOangle
+      `Honesty-humility` = Honest,
+      `SVO angle` = SVOangle
       ) %>%
     cor(
       method = "spearman",
@@ -291,12 +291,12 @@ plotSurveyCorrelations <- function(d) {
[…]

### 2023-11-24 _targets.R

@@ -10,7 +10,7 @@ source("R/functions_summariseData.R")
 source("R/functions_modelResults.R")
 tar_option_set(
   packages = c("ape", "cowplot", "ggcorrplot", "ggrepel", 
-               "ggridges", "knitr", "ltm", "papaja", "phangorn", 
+               "ggridges", "knitr", "lme4", "ltm", "papaja", "phangorn", 
                "rethinking", "rstan", "tanggle", "tidybayes", 
                "tidytext", "tidyverse", "wordcloud"),
   deployment = "main"
@@ -49,23 +49,15 @@ list(
   #### Load and compile Stan models ####
   
   # model files
-  tar_target(fileModel1.1, "stan/noTPP/punishStrat_noTPP_noPredictor.stan", format = "file"),
-  tar_target(fileModel2.1, "stan/noTPP/punishStrat_noTPP_withPredictor.stan", format = "file"),
-  tar_target(fileModel3.1, "stan/noTPP/punishStrat_noTPP_withTwoPredictors.stan", format = "file"),
-  tar_target(fileModel4.1, "stan/noTPP/punishStrat_noTPP_withCategoricalPredictor.stan", format = "file"),
-  tar_target(fileModel1.2, "stan/withTPP/punishStrat_withTPP_noPredictor.stan", format = "file"),
-  tar_target(fileModel2.2, "stan/withTPP/punishStrat_withTPP_withPredictor.stan", format = "file"),
-  tar_target(fileModel3.2, "stan/withTPP/punishStrat_withTPP_withTwoPredictors.stan", format = "file"),
-  tar_target(fileModel4.2, "stan/withTPP/punishStrat_withTPP_withCategoricalPredictor.stan", format = "file"),
+  tar_target(fileModel1, "stan/punishStrat_noPredictor.stan", format = "file"),
+  tar_target(fileModel2, "stan/punishStrat_withPredictor.stan", format = "file"),
+  tar_target(fileModel3, "stan/punishStrat_withTwoPredictors.stan", format = "file"),
+  tar_target(fileModel4, "stan/punishStrat_withCategoricalPredictor.stan", format = "file"),
   # compile models
-  tar_target(compiledModel1.1, stan_model(fileModel1.1), deployment = "worker"),
-  tar_target(compiledModel2.1, stan_model(fileModel2.1), deployment = "worker"),
[…]

### 2023-12-13 appendix.Rmd
update manuscript

@@
 implementation error rate of 5%. Points represent posterior medians, line ranges 
-represent 50% and 95% credible intervals.
+represent 50% and 95% credible intervals. AI = advantageous inequity, DI = 
+disadvantageous inequity.
 
@@
 dataset with pre-registered exclusions.* Each row represents a separate model. 
-Points represent posterior medians, line ranges represent 95% credible intervals.
+Points represent posterior medians, line ranges represent 95% credible
+intervals. AI = advantageous inequity, DI = disadvantageous inequity.
 
@@
 without pre-registered exclusions.* Each row represents a separate model. Points 
-represent posterior medians, line ranges represent 95% credible intervals.
+represent posterior medians, line ranges represent 95% credible intervals. AI = 
+advantageous inequity, DI = disadvantageous inequity.
 
@@
 control for one another within the same model. Points represent posterior 
-medians, line ranges represent 95% credible intervals.
+medians, line ranges represent 95% credible intervals. AI = advantageous 
+inequity, DI = disadvantageous inequity.
 
@@
 competitive strategy. Points represent posterior medians, line ranges represent 
-95% credible intervals.
+95% credible intervals. AI = advantageous inequity, DI = disadvantageous 
+inequity.
 
@@
 
+(ref:plotModel5Caption) *Posterior estimates of the probabilities of following
+different punishment strategies from the Bayesian latent state model that
+estimates the implementation error rate as a free parameter.* The model
+estimated the implementation error rate to be `r printnum(median(error))`
+(95% CI [`r printnum(quantile(error, 0.025))`
+`r printnum(quantile(error, 0.975))`]). Points represent posterior medians, line 
+ranges represent 50% and 95% credible intervals. AI = advantageous inequity, DI
+= disadvantageous inequity.
+
+```{r plotModel5, fig.width=7, fig.height=5, fig.cap="(ref:plotModel5Caption)", warning=FALSE}
+tar_read(plot5)
+```
+
+\newpage
+
 (ref:plotTraceCaption) *Trace plots for parameter values from the Bayesian
@@
 represent known strategy frequencies, grey densities represent posterior
-estimates of strategy frequencies.
+estimates of strategy frequencies. AI = advantageous inequity, DI = 
+disadvantageous inequity.
 

### 2023-12-13 manuscript.Rmd
update manuscript

@@
 abstract: |
- Costly punishment is thought to be one of the key mechanisms sustaining cooperation in humans.
+ Costly punishment is thought to be a key mechanism sustaining human cooperation.
 However, the motives for punitive behaviour remain unclear.
-Punishment is often assumed to be motivated by a desire to convert free riders into cooperators, but it is also consistent with a host of other functions, such as levelling payoffs or increasing one’s relative position.
-We used a suite of six economic games to tease apart the different motives for punishment.
-Across representative samples from the United Kingdom and the United States, we estimated the frequency of different punishment strategies in the population, finding that egalitarian motives for punishment are more common than behaviour-change motives.
-Moreover, different punishment strategies were differentially predicted by personality, social preferences, political ideology, and religious views.
-Self-reports of behaviour in the games suggested that people have some degree of insight into their punishment strategy.
-These findings highlight the multipurpose nature of human punishment.
+Although often assumed to be motivated by a desire to convert cheats into cooperators, punishment is also consistent with other functions, such as levelling payoffs or improving one’s relative position.
+We used six economic games to tease apart different motives for punishment and to explore whether different punishment strategies were associated with personality variables, political ideology, and religiosity.
+We used representative samples from the United Kingdom and the United States (N = 2010) to estimate the frequency of different punishment strategies in the population.
+The most common strategy was to never punish: prosocial individuals and those scoring high in honesty-humility were among the least punitive in our sample, while religious individuals and those scoring high for right-wing authoritarianism and social dominance orientation were more punitive.
+For people who did punish, strategy use was more consistent with egalitarian motives than behaviour-change motives.
+Nevertheless, different punishment strategies were also associated with personality, social preferences, political ideology, and religiosity.
+Self-reports of behaviour in the games suggested that people have some insight into their punishment strategy.
+These findings highlight the multipurpose nature of human punishment and show how the different motives underpinning punishment decisions are linked with core character traits.
 
 keywords          : "punishment; cooperation; economic games"
-wordcount         : "5185 words"
+wordcount         : "xxxx words"
 bibliography      : ["r-references.bib"]
@@
 whereby individuals harm others at a personal cost[@CluttonBrock1995], 
-ostensibly encouraging cooperative behaviour from the target (or bystanders [@DosSantos2011;@DosSantos2013;@Raihani2015]) in the future. Punishment 
-therefore offers a route to maintaining or increasing cooperation by changing 
-the payoff structure of social interactions such that it no longer pays to cheat 
-or exploit social partners[@CluttonBrock1995;@Ostrom1990].
+ostensibly encouraging cooperative behaviour from the target (or bystanders 
+[@DosSantos2011;@DosSantos2013;@Raihani2015]) in the future. Punishment offers a 
+route to maintaining or increasing cooperation by changing the payoff structure 
+of social interactions such that it no longer pays to cheat or exploit social 
+partners[@CluttonBrock1995;@Ostrom1990]. Yet, despite its theoretical 
+importance, the question of why people choose to punish others is still hotly 
+contested[@Raihani2019]. In this study, we use a battery of economic games to 
+disentangle the different motives underpinning punishment and explore how these 
+motives vary across individuals.
 
@@
 
-Beyond behaviour-shaping concerns, there are a host of other reasons why people 
+Beyond behaviour shaping concerns, there are a host of other reasons why people 
 may want to punish in economic games. Punishers might be motivated by a desire 
@@
 so, what other characteristics (personality, social preferences, political 
-orientation) predict the use of different punishment strategies. Table 
-\@ref(tab:tableStrategies) summarises the potential functions for costly 
+orientation, religious views) predict the use of different punishment strategies.
+Table \@ref(tab:tableStrategies) summarises the potential functions for costly 
 punishment in the economic games that we considered, and the behavioural 
@@
 are variants on this general setup, creating situations where (A) Player 2 is
-still worse off by stealing, (B) Player 2 creates equality by stealing, (C) the
-computer "decides" whether Player 2 steals, (D) the fee-fine ratio is 1:1, (E)
-Player 2 is better off by stealing, and (F) Player 2 steals instead from a 
-third-party.
+still worse off after stealing, (B) Player 2 creates equality by stealing, (C) 
+the computer "decides" whether Player 2 steals, (D) the fee-fine ratio is 1:1, 
+(E) Player 2 is better off after stealing, and (F) Player 2 steals instead from 
+a third-party.
 
@@
 
-The overall pattern of punitive behaviour in the six economic games was in line 
-with previous research and very similar across both countries (Figure 
-\@ref(fig:plotDecisions)). Participants were generally more likely to punish 
-targets who stole from another individual compared to targets who did not steal 
-(multilevel logistic regression; _b_ = 
+The overall pattern of punitive behaviour in the six economic games was very 
+similar across both countries (Figure \@ref(fig:plotDecisions)). Participants 
+were generally more likely to punish targets who stole compared to targets who 
+did not steal (multilevel logistic regression; _b_ = 
 `r printnum(mlm1["otherTook","Estimate"])`, standard error = 
@@
 minimising payoff differences (avoid disadvantageous inequity, egalitarian). 
-Less common were the behaviour-shaping strategies (deterrent, norm-enforcing),
-the retributive strategy, and the competitive strategies (seek AI, competitive). 
-Although participants often punished targets who did not steal 
-in the six games (Figure \@ref(fig:plotDecisions)), no participants followed the 
-antisocial strategy by exclusively punishing targets who did not steal across 
-_all_ games.
+Less common were the behaviour shaping strategies (deterrent, norm-enforcing),
+the retributive strategy, and the competitive strategies (seek advantageous 
+inequity, competitive). Although participants often punished targets who did not 
+steal in the six games (Figure \@ref(fig:plotDecisions)), no participants 
+followed the antisocial strategy by exclusively punishing targets who did not 
+steal across _all_ games.
 
 (ref:tableStrategyCountsCaption) Counts and proportions of participants 
-following each punishment strategy exactly, split by country. N/A implies that
+following each punishment strategy exactly, split by country. *N/A implies that
 participants were unable to be classified exactly into any of the punishment
-strategies.
+strategies. AI = advantageous inequity, DI = disadvantageous inequity.*
 
@@
 had the highest probability, followed by the egalitarian strategy. The
-norm-enforcing and seek AI strategies were the next most likely, with higher
-posterior estimates than the competitive and antisocial strategies. None of the
-other strategies differed in their posterior estimates. The same general pattern
-emerged when we analysed the full dataset without pre-registered exclusions 
-(Supplementary Figure \@ref(fig:plotModel1a)).
+norm-enforcing and seek advantageous inequity strategies were the next most 
+likely, with higher posterior estimates than the competitive and antisocial 
+strategies. None of the other strategies differed in their posterior estimates. 
+The same general pattern emerged when we analysed the full dataset without 
+pre-registered exclusions (Supplementary Figure \@ref(fig:plotModel1a)).
 
@@
 assumes an implementation error rate of 5%. Points represent posterior medians, 
-line ranges represent 50% and 95% credible intervals.
+line ranges represent 50% and 95% credible intervals. AI = advantageous 
+inequity, DI = disadvantageous inequity.
 
@@
 Social Value Orientation (SVO) in the model, we found associations with the 
-egalitarian, never punish, and random choice strategies (Figure
-\@ref(fig:plotAllPers2)). Participants higher in SVO were more likely to follow 
-the egalitarian and the never punish strategies, while those with lower SVO 
-scores were more likely to enact the random choice strategy. The personality 
-dimensions of honesty-humility and openness to experience were both positively 
-associated with following the never punish strategy, while extraversion 
-negatively predicted this strategy. The effects were mostly similar across
-countries, but occasionally differed: for example, in the United States,
-but not in the United Kingdom, honesty-humility was positively associated with 
-following the egalitarian strategy and negatively associated with following 
-the random choice strategy. Overall, the same pattern of results emerged when 
-analysing the full dataset without exclusions (Supplementary Figure 
-\@ref(fig:plotAllPers1)).
+egalitarian, never punish, and random choice strategies, with small-to-medium 
+effect sizes (Figure \@ref(fig:plotAllPers2)). More prosocial participants 
+(those with higher SVO scores) were more likely to follow the egalitarian and 
+the never punish strategies, while those with lower SVO scores were more likely 
+to enact the random choice strategy. The personality dimensions of 
+honesty-humility and openness to experience were both positively associated with 
+following the never punish strategy, while extraversion negatively predicted 
+this strategy. The effects were mostly similar across countries, but 
+occasionally differed: for example, in the United States, but not in the United 
+Kingdom, honesty-humility was positively associated with following the 
+egalitarian strategy and negatively associated with following the random 
+choice strategy. Overall, the same pattern of results emerged when analysing the 
+full dataset without exclusions (Supplementary Figure \@ref(fig:plotAllPers1)).
 
@@
 5%. Points represent posterior medians, line ranges represent 95% credible 
-intervals.
+intervals. AI = advantageous inequity, DI = disadvantageous inequity.
 
@@
 Political and religious variables were also associated with punishment strategy 
-(Figure \@ref(fig:plotAllPolRel2)). These effects tended to be more pronounced 
-in the United States. Controlling for Social Dominance Orientation, American 
-participants higher in Right Wing Authoritarianism were more likely to follow 
-the strategies avoiding disadvantageous inequity and seeking advantageous 
-inequity. Participants who stated that they would like to "bring those below 
-them [on the socio-economic status ladder] up a peg" were more likely to follow 
-the egalitarian strategy, while American participants higher in Social Dominance 
-Orientation, Right Wing Authoritarianism, and believing that God controls events 
-in the world were less likely to follow the egalitarian strategy. In general, 
-religious and conservative participants were less likely to follow the never 
-punish strategy. This general pattern of results was replicated with the full 
-dataset (Supplementary Figure \@ref(fig:plotAllPolRel1)).
+(Figure \@ref(fig:plotAllPolRel2)). These effects were small-to-medium in size 
+and tended to be more pronounced in the United States. Controlling for Social 
+Dominance Orientation, American participants higher in Right Wing 
+Authoritarianism were more likely to follow the strategies avoiding 
+disadvantageous inequity and seeking advantageous inequity. Participants who 
+stated that they would like to "bring those below them [on the socio-economic 
+status ladder] up a peg" were more likely to follow the egalitarian strategy, 
+while American participants higher in Social Dominance Orientation, Right Wing 
+Authoritarianism, and believing that God controls events in the world were less 
+likely to follow the egalitarian strategy. In general, religious and 
+conservative participants were less likely to follow the never punish strategy. 
+This general pattern of results was replicated with the full dataset 
+(Supplementary Figure \@ref(fig:plotAllPolRel1)).
 
@@
 model. All models assume an implementation error rate of 5%. Points represent 
-posterior medians, line ranges represent 95% credible intervals.
+posterior medians, line ranges represent 95% credible intervals. AI = 
+advantageous inequity, DI = disadvantageous inequity.
 
@@
 In general, we found that self-reported strategy usage was positively associated
-with the behavioural strategy that participants employed (see Supplementary 
-Figures \@ref(fig:plotSliders1) and \@ref(fig:plotSliders2) for the distribution 
-of responses to self-report questions). Figure \@ref(fig:plotAllSliders2) shows 
-the relationships between self-report questions and the different punishment 
-strategies, highlighting the combinations where the question matched the 
-behavioural strategy. We found positive relationships between the self-report 
-questions and strategy usage for the norm-enforcing, egalitarian, seek 
-advantageous inequity, never punish, and random choice  strategies. The 95% 
-credible intervals for other estimates included zero, though these estimates 
-often trended in a positive direction. The same pattern of results was found 
-when analysing the full dataset without exclusions  (Supplementary Figure 
-\@ref(fig:plotAllSliders1)).
+with the behavioural strategy that participants employed, with effects ranging 
+from small to large in size (see Supplementary Figures \@ref(fig:plotSliders1) 
+and \@ref(fig:plotSliders2) for the distribution of responses to self-report 
+questions). Figure \@ref(fig:plotAllSliders2) shows the relationships between 
+self-report questions and the different punishment strategies, highlighting the 
+combinations where the question matched the behavioural strategy. We found 
+positive relationships between the self-report questions and strategy usage for 
+the norm-enforcing, egalitarian, seek advantageous inequity, never punish, and 
+random choice  strategies. The 95% credible intervals for other estimates 
+included zero, though these estimates often trended in a positive direction. The 
+same pattern of results was found when analysing the full dataset without 
+exclusions  (Supplementary Figure \@ref(fig:plotAllSliders1)).
 
@@
 Points represent posterior medians, line ranges represent 95% credible 
-intervals.
+intervals. AI = advantageous inequity, DI = disadvantageous inequity.
 
@@
 a flexible tool that can be and is used for different purposes[@Raihani2019]. 
-Punishment is more akin to a swiss army knife than a hammer, used by some to 
-enforce norms of cooperation and by others to reduce or even create inequality 
-between individuals. We found that people's punishment strategy can, to some 
-extent, be predicted by individual differences in personality, social 
-preferences, and political and religious views. Moreover, contrary to the view 
-that people are often unable to articulate the reasons for their punitive 
-behaviour[@Carlsmith2002], people seem to have some degree of insight into the 
-strategy they are using. Despite small differences, these general patterns 
-replicated in samples from both the United Kingdom and the United States.
+While some use punishment to enforce norms of cooperation, others use it to 
+reduce or even create inequality between individuals. We found that people's 
+punishment strategy can, to some extent, be predicted by individual differences 
+in personality, social preferences, and political and religious views. Moreover, 
+contrary to the view that people are often unable to articulate the reasons for 
+their punitive behaviour[@Carlsmith2002], people seem to have some degree of 
+insight into the strategy they are using. Despite small differences, these 
+general patterns replicated in samples from both the United Kingdom and the 
+United States, providing further confidence in the results.
 
@@
 important motivation for punishment in economic games 
-[@Dawes2007;@Raihani2012b]. Personality and social preference variables mapped 
-onto these strategies in expected ways. Traits associated with other-regarding 
-concern, such as SVO and honesty-humility, predicted following the egalitarian 
-strategy, whereas religious and conservative individuals were less likely to 
-follow this strategy, especially in the United States. Moreover, participants 
-following the egalitarian strategy were able to self-report this strategy, 
-though the same was not true for the avoid disadvantageous inequity strategy.
+[@Bone2015;@Bone2016;@Dawes2007;@Deutchman2021;@Johnson2009;@Raihani2012b]. 
+Personality, social preferences, and political and religious views were related 
+to these strategies. Traits associated with other-regarding concern, such as SVO 
+and honesty-humility, predicted following the egalitarian strategy. By contrast,
+religious and conservative individuals, including individuals higher in SDO and
+RWA, were less likely to follow the egalitarian strategy and more likely to 
+follow the avoid disadvantageous inequity strategy, especially in the United 
+States. Authoritarian participants in the United States were also more likely to 
+actively seek advantageous inequity with their punishment. These findings 
+provide insight into the motives that may have driven higher levels of peer 
+punishment among conservatives in previous work[@Hofmann2018].
 
@@
 (Supplementary Figures \@ref(fig:plotSliders1) and \@ref(fig:plotSliders2)). 
-Regarding the predictors of these strategies, we found that demographic, 
-personality, political, and religious variables tended to be unrelated to 
-deterrent and norm-enforcing punishment strategies. We also found that 
-participants had insight into the norm-enforcing strategy, but not the deterrent 
+Although our design did not explicitly allow for behaviour shaping as the 
+interactions were all one-shot, we did manipulate whether the target's stealing 
+behaviour was intentional or not (Game C), an approach which has been used in 
+previous vignette studies to identify behaviour shaping motives[@Carlsmith2002].
+The lower prevalence of behaviour shaping strategies in our study is consistent 
+with prior work showing that punishment often continues to be used when 
+behaviour shaping is impossible, such as when the target will never find out 
+that they have been punished[@Crockett2014] or on the last round of repeated 
+interactions[@Barclay2016]. We found that demographic, personality, political, 
+and religious variables tended to be unrelated to behaviour shaping strategies, 
+though this could reflect low power to detect such associations given the low 
+prevalence of these strategies in our sample. We also found that participants 
+accurately reported using the norm-enforcing strategy, but not the deterrent 
 strategy. This finding is in line with previous research showing that people 
@@
 
-Other punitive strategies were less common in our dataset, but some were more 
-prevalent than others. For example, participants were more likely to use 
-punishment to seek advantageous inequity than to exclusively harm those who did 
-not steal (i.e., antisocial punishment). The existence of the "seek AI" strategy 
-in our dataset supports the claim that punishment can also be used as a tool to 
-increase one’s own relative position[@Raihani2019]. While generally rare, we 
-found that this motive for punishment was more common among authoritarian 
-participants, at least in the United States, potentially providing an 
-explanation for why peer punishment has been found to be more common among 
-conservatives in previous work[@Hofmann2018]. Moreover, the fact that no 
-participants in our sample punished non-stealing across all games suggests that 
-antisocial punishment does not function to harm cooperators specifically, as has 
-been previously suggested[@Herrmann2008]. Instead, antisocial punishment appears 
-to be motivated by improving one's relative position in general, which is in 
-line with work showing that antisocial punishment disappears with a 1:1 fee-fine 
-ratio[@Sylwester2013].
+We defined antisocial punishment as occurring when a participant exclusively 
+punished those who did not steal. There were no participants in our sample who 
+followed this strategy, although some players did punish non-stealing
+co-players. Harming non-stealing individuals was also consistent with the 
+competitive strategies, which did appear in our sample albeit at low frequencies.
+The fact that no participants in our sample punished non-stealing across all 
+games suggests that the antisocial punishment that has been observed in other 
+studies[@Herrmann2008;@Pleasant2018] is not aimed at harming cooperators 
+specifically, as has been previously suggested[@Herrmann2008]. Instead, 
+antisocial punishment is more likely to be motivated by improving one's 
+relative position, which is in line with work showing that antisocial punishment 
+disappears when relative payoffs cannot be changed (e.g., when punishment is 
+only available with a 1:1 fee-fine ratio[@Sylwester2013]).
 
@@
 for the way that punishment is operationalised in classic behavioural economic 
-game studies. In these studies, a common assumption is that participants will 
-punish to change the behaviour of cheats. But in reality, people may be choosing 
-the punishment option to achieve a variety of different goals. The targets of 
-punishment in these studies are likely well aware that punishment could be 
-levied for these different reasons and this knowledge may impact their 
-responses. For example, if cheating targets interpret punishment as serving a 
-competitive motive, it may elicit retaliation rather than encourage 
-cooperation[@Dreber2008;@Nikiforakis2008b;@Raihani2019]. This might help to 
-explain the mixed findings in the field as to whether punishment actually 
-motivates cheating targets to cooperate in the future[@Raihani2019].
+game studies. In these studies, a common assumption is that participants punish 
+to change the behaviour of cheats[@Fehr2000;@Fehr2002]. But in reality, people 
+may be choosing the punishment option to achieve a variety of different goals. 
+This has implications for how people respond to being targeted by punishers in 
+these games. Targets of punishment in these studies may know that punishment 
+reflects different motives and can respond accordingly. For example, if targets
+interpret punishment as serving a competitive motive, it may elicit retaliation
+rather than encourage cooperation[@Dreber2008;@Nikiforakis2008b;@Raihani2019].
+As punishers’ motives must be inferred (and such inferences likely depend on 
+character traits of the target, as well as the context in which punishment 
+occurs), there is likely to be some variation and error in attributing motives 
+to punishers. To the extent that inferred motives affect target responses, this 
+might help to explain the mixed findings in the field as to whether punishment 
+actually motivates cheating targets to cooperate in the future[@Raihani2019].
 
@@
 This is partly because punishment in these games imposes an economic cost for no 
-tangible benefit. If the fee-fine ratio had been lower such that it was cheaper 
+tangible benefit. If the fee-fine ratio had been lower, such that it was cheaper 
 to punish, we may have seen more punishment from participants. Indeed, 
@@
 [@Balafoutas2014;@Balafoutas2016] and in small-scale societies[@Baumard2010]. 
-One reason that people may be averse to peer punishment is that, due to its 
-multipurpose nature, it may be interpreted as a competitive challenge by targets 
-and trigger retaliation[@Raihani2019]. In situations that lack clear 
-institutional norms to legitimise punishment, such as our economic games and 
-some situations in the real world, people might abstain from peer punishment to 
-avoid such retaliation, regardless of whether retaliation is actually possible. 
-By contrast, institutionalised punishment in small-scale societies often 
-functions to compensate victims adequately while limiting the potential for 
-feuds and cycles of retaliation[@Fitouchi2023;@Singh2022]. Future research 
-should uncover whether people are more willing to punish in these 
-conventionalised contexts.
+One reason that people may be averse to peer punishment is that, because it is 
+a fundamentally harmful act, punishment can reflect badly on the punisher and 
+people might therefore refrain from punishing others to avoid reputational 
+damage[@Raihani2015]. People frequently avoid taking actions that could harm 
+their reputation, even when they don’t know if reputation is at stake (as in the 
+one-shot anonymous settings used here). Another reason that people may be averse
+to peer punishment is that it can trigger retaliation[@Raihani2019]. This may be
+especially likely in situations that lack clear institutional norms to 
+legitimise punishment, such as our economic games. As with reputation damage, 
+people might abstain from peer punishment to avoid retaliation, regardless of 
+whether retaliation is actually possible. By contrast, institutionalised 
+punishment in small-scale societies often functions to compensate victims while 
+limiting the potential for feuds and cycles of retaliation
+[@Fitouchi2023;@Singh2022]. Future research should uncover whether people are 
+more willing to punish in these conventionalised contexts (e.g., see
+[@Molleman2019]).
 
-There are several limitations with our study design that can guide future 
-research. First, we used one-shot economic games to measure punishment 
-strategies, which may have led us to underestimate behaviour-change strategies 
-like deterrence. Our inclusion of Game C somewhat mitigated this issue by 
-manipulating whether stealing behaviour was intentional vs. unintentional and 
-thus whether there was any behaviour to be deterred. Due to limits on our 
-within-subjects design and the complexity of the strategy space, it was not 
-feasible for us to expand our study to include additional contexts to elicit 
-behaviour-change strategies (e.g., repeated games, games where targets are not 
-made aware of the punishment, games where targets can retaliate). Future work 
-could study these contexts separately.
-
-A second limitation is that some strategies required more punishment than others 
-to be met. For example, the competitive strategy required punishment in ten of 
-twelve decisions, compared to the avoid disadvantageous inequity strategy which 
-required only one instance of punishment (Table \@ref(tab:tableStrategies)). 
+We finally highlight some key limitations that can guide future research. One 
+limitation with our design is that some strategies required more punishment than 
+others to be met. For example, the competitive strategy required punishment in 
+ten of twelve decisions, compared to the avoid disadvantageous inequity strategy
+which required only one instance of punishment (Table \@ref(tab:tableStrategies)). 
 Strategies thus differed in how "expensive" they were to implement, perhaps 
@@
 guidelines and regulations. Informed consent was obtained from all participants 
-prior to the study.
+prior to the study and participation was voluntary.
 
@@
 
+Our punishment strategies make different behavioural predictions across these 
+different games (Table \@ref(tab:tableStrategies)). The deterrent strategy
+punishes whenever it is intentionally personally harmed (when the target steals
+in Games A, B, D, and E). The norm-enforcing strategy is the same, except that 
+it also punishes when _others_ are intentionally harmed (when the target steals
+in Game F). The retributive strategy punishes whenever it is personally harmed,
+whether the harm was intentional or not (when the target steals in Games A-E).
+The avoid disadvantageous inequity strategy only punishes when it currently 
+experiences disadvantageous inequity and can alleviate it by punishing (when the
+target steals in Game E). The egalitarian strategy punishes when punishment
+makes payoffs for all players more equal than they currently are (when the
+target steals in Games E and F). The seek advantageous inequity strategy
+punishes when it currently does not experience advantageous inequity and 
+punishment would produce advantageous inequity for itself (when the target
+steals in Games B and C). The competitive strategy punishes whenever the
+fee-fine ratio is 1:3, as it can improve its relative position by punishing, but
+does not punish when the fee-fine ratio is 1:1 (Game D). The antisocial strategy
+punishes whenever the target does not steal across all games. The never punish
+strategy never punishes in any game.
+
 For each game, participants saw the game instructions and answered a 
@@
 - *Personality*. We used the Mini-IPIP scale[@Sibley2011] to measure the Big 6 
-  personality dimensions of agreeableness, conscientiousness, extraversion, 
-  honesty-humility, openness to experience, and neuroticism (four items each).
+  personality dimensions of agreeableness ($\alpha$ = 
+  `r printnum(tar_read(alphaAgree)$alpha)`), conscientiousness ($\alpha$ = 
+  `r printnum(tar_read(alphaConsc)$alpha)`), extraversion ($\alpha$ = 
+  `r printnum(tar_read(alphaExtra)$alpha)`), honesty-humility ($\alpha$ = 
+  `r printnum(tar_read(alphaHonest)$alpha)`), openness to experience ($\alpha$ = 
+  `r printnum(tar_read(alphaOpen)$alpha)`), and neuroticism ($\alpha$ = 
+  `r printnum(tar_read(alphaNeur)$alpha)`). Four items were used for each 
+  personality dimension.
 - *Social Value Orientation*. We used the Social Value Orientation Slider 
@@
   including left-right conservatism, Social Dominance Orientation[@Ho2015] 
-  (eight items), and Right Wing Authoritarianism[@Bizumic2018] (six items). 
-  We also probed participants' views on social inequality by asking them whether 
-  they would like to bring people above (below) them on the MacArthur 
-  socio-economic status ladder down (up) a peg or two.
+  ($\alpha$ = `r printnum(tar_read(alphaSDO)$alpha)`; eight items), and Right 
+  Wing Authoritarianism[@Bizumic2018] ($\alpha$ = 
+  `r printnum(tar_read(alphaRWA)$alpha)`; six items). We also probed 
+  participants' views on social inequality by asking them whether they would 
+  like to bring people above (below) them on the MacArthur socio-economic status 
+  ladder down (up) a peg or two.
 - *Religious views*. We asked participants how religious they consider 
@@
 to comprehension questions suggested that participants understood the six 
-economic games (Supplementary Table \@ref(tab:tableComp)). We used the strategy 
-method to incentivise the economic games, choosing a random game to determine 
-bonus payment. After all games, 
+economic games (Supplementary Table \@ref(tab:tableComp)). In order to partially 
+mitigate the fact that some punishment strategies predict more punishment than 
+others and are thus more "expensive" to implement (Table 
+\@ref(tab:tableStrategies)), we used the strategy method to incentivise the 
+economic games, choosing a random game to determine bonus payments rather than 
+summing participants' earnings across all games. After all games, 
 `r printnum((sum(dExc$DecisionConsequences == "Yes", na.rm = TRUE) / nrow(dExc)) * 100, digits = 0)`% 
@@
 payment from the six economic games (between £0.40 -- £0.70 depending on their 
-decision). Participants were paid £1.50 for completing the follow-up survey, 
+decisions). Participants were paid £1.50 for completing the follow-up survey, 
 plus a bonus payment from the Social Value Orientation Slider Measure (between 
-£0.50 -- £0.85 depending on their decision).
+£0.50 -- £0.85 depending on their decisions).
 
@@
 
+```{r}
+error <- tar_read(post5)$error
+```
+
 The conditional probabilities $\text{Pr}(\text{punish}|s,j)$ are hard coded in 
@@
 of $1 - \delta$ and coding red crosses with a conditional probability of 
-$0 + \delta$. The random choice is consistently coded with a conditional 
-probability of ½ across all decisions.
+$0 + \delta$. We set $\delta$ to 0.05 in all models, which is similar to its 
+value when we estimate it as a free parameter in an additional model (median 
+posterior $\delta$ = `r printnum(median(error))`, 95% CI
+[`r printnum(quantile(error, 0.025))` `r printnum(quantile(error, 0.975))`];
+Supplementary Figure \@ref(fig:plotModel5)). The random choice strategy is 
+consistently coded with a conditional probability of ½ across all decisions.
 
@@
 
+These models control for multiple comparisons across strategies by estimating 
+the effects of the predictor on all strategies simultaneously.
+
 We estimated the posterior distributions of these models using Hamiltonian Monte 
@@
 
-We validated the model by simulating observed data (n = 100) from a known 
+We validated the model by simulating observed data (_n_ = 100) from a known 
 frequency of strategies. The model was successfully able to recover the known 
@@
 
-All authors conceptualised the research, designed the study, and developed the
-surveys. N.J. conducted data collection on Prolific. S.C. conducted all analyses
-and visualisation of the data. All authors wrote the manuscript.
+All authors (S.C., Q.D.A., N.R.) conceptualised the research, designed the 
+study, and developed the surveys. N.R. conducted data collection on Prolific.
+S.C. conducted all analyses and visualisation of the data. All authors wrote the 
+manuscript.
 

### 2023-12-13 manuscript.tex
update manuscript

@@
 {}{\PatchFailed}
-\keywords{punishment; cooperation; economic games\newline\indent Word count: 5185 words}
+\keywords{punishment; cooperation; economic games\newline\indent Word count: xxxx words}
 \usepackage{lineno}
@@
 \abstract{%
-Costly punishment is thought to be one of the key mechanisms sustaining cooperation in humans.
+Costly punishment is thought to be a key mechanism sustaining human cooperation.
 However, the motives for punitive behaviour remain unclear.
-Punishment is often assumed to be motivated by a desire to convert free riders into cooperators, but it is also consistent with a host of other functions, such as levelling payoffs or increasing one's relative position.
-We used a suite of six economic games to tease apart the different motives for punishment.
-Across representative samples from the United Kingdom and the United States, we estimated the frequency of different punishment strategies in the population, finding that egalitarian motives for punishment are more common than behaviour-change motives.
-Moreover, different punishment strategies were differentially predicted by personality, social preferences, political ideology, and religious views.
-Self-reports of behaviour in the games suggested that people have some degree of insight into their punishment strategy.
-These findings highlight the multipurpose nature of human punishment.
+Although often assumed to be motivated by a desire to convert cheats into cooperators, punishment is also consistent with other functions, such as levelling payoffs or improving one's relative position.
+We used six economic games to tease apart different motives for punishment and to explore whether different punishment strategies were associated with personality variables, political ideology, and religiosity.
+We used representative samples from the United Kingdom and the United States (N = 2010) to estimate the frequency of different punishment strategies in the population.
+The most common strategy was to never punish: prosocial individuals and those scoring high in honesty-humility were among the least punitive in our sample, while religious individuals and those scoring high for right-wing authoritarianism and social dominance orientation were more punitive.
+For people who did punish, strategy use was more consistent with egalitarian motives than behaviour-change motives.
+Nevertheless, different punishment strategies were also associated with personality, social preferences, political ideology, and religiosity.
+Self-reports of behaviour in the games suggested that people have some insight into their punishment strategy.
+These findings highlight the multipurpose nature of human punishment and show how the different motives underpinning punishment decisions are linked with core character traits.
 }
@@
 whereby individuals harm others at a personal cost\textsuperscript{1},
-ostensibly encouraging cooperative behaviour from the target (or bystanders\textsuperscript{2--4}) in the future. Punishment
-therefore offers a route to maintaining or increasing cooperation by changing
-the payoff structure of social interactions such that it no longer pays to cheat
-or exploit social partners\textsuperscript{1,5}.
+ostensibly encouraging cooperative behaviour from the target (or bystanders\textsuperscript{2--4}) in the future. Punishment offers a
+route to maintaining or increasing cooperation by changing the payoff structure
+of social interactions such that it no longer pays to cheat or exploit social
+partners\textsuperscript{1,5}. Yet, despite its theoretical
+importance, the question of why people choose to punish others is still hotly
+contested\textsuperscript{6}. In this study, we use a battery of economic games to
+disentangle the different motives underpinning punishment and explore how these
+motives vary across individuals.
 
@@
 `fine' on their co-players. Several lines of experimental evidence indicate that
-people use this punishment option\textsuperscript{12}, that they enjoy
+people use this punishment option\textsuperscript{13}, that they enjoy
 punishing\textsuperscript{16}, and that they frequently, though not
 always\textsuperscript{17}, punish cheating or exploitative
-co-players\textsuperscript{10,11}.
+co-players\textsuperscript{11,12}.
 
@@
 plays an important role in promoting human cooperation. People tend to cooperate
-more in games where punishment is possible compared to those where it is not\textsuperscript{6,7,15}. The effect that the threat of
+more in games where punishment is possible compared to those where it is not\textsuperscript{6--8}. The effect that the threat of
 punishment has on cooperation is also evident in the higher contributions
@@
 typical cooperation-enhancing effect of punishment has also been observed across
-societies\textsuperscript{7}, leading some to suggest that costly punishment has
+societies\textsuperscript{8}, leading some to suggest that costly punishment has
 played a key role in the cultural evolution of cooperation in humans\textsuperscript{19--22}.
@@
 achieve other ends. Some have argued that punishment is primarily used to shape
-future behaviour, either to deter personal harm\textsuperscript{3,9,23} or to uphold normative standards of cooperative behaviour\textsuperscript{20,21,24--27}.
+future behaviour, either to deter personal harm\textsuperscript{3,10,23} or to uphold normative standards of cooperative behaviour\textsuperscript{20,21,24--27}.
 But while the \emph{threat} of punishment can have a cooperation-enhancing effect,
 the \emph{enactment} of this punishment does not consistently deter targets from
-cheating in the future\textsuperscript{15}. This calls into question whether
+cheating in the future\textsuperscript{6}. This calls into question whether
 punishment primarily operates as a behaviour-change tool or whether it is used
@@
 
-Beyond behaviour-shaping concerns, there are a host of other reasons why people
+Beyond behaviour shaping concerns, there are a host of other reasons why people
 may want to punish in economic games. Punishers might be motivated by a desire
@@
 by concerns about relative payoffs, such as disadvantageous inequity aversion
-(i.e., avoiding having less than others\textsuperscript{15,29}) and/or
+(i.e., avoiding having less than others\textsuperscript{6,29}) and/or
 general egalitarian preferences (i.e., wanting all participants to receive the
@@
 competitive purposes, seeking advantageous inequity for themselves (i.e., having
-more than others) and/or improving their relative position\textsuperscript{15}.
+more than others) and/or improving their relative position\textsuperscript{6}.
 
@@
 people use punishment in seemingly disparate ways: punishing when no behaviour
-change is possible, such as in one-shot games\textsuperscript{12,29,31,32}, on the very last round of repeated games\textsuperscript{33}, or in games where the target never learns about the
+change is possible, such as in one-shot games\textsuperscript{13,29,31,32}, on the very last round of repeated games\textsuperscript{33}, or in games where the target never learns about the
 punishment\textsuperscript{34}; punishing those who did not cheat or who
@@
 of as a flexible behavioural tool that serves a variety of functions that are
-not mutually exclusive\textsuperscript{15}. Due to its multipurpose nature, we should
+not mutually exclusive\textsuperscript{6}. Due to its multipurpose nature, we should
 therefore expect variation in punishment strategies in the population, much like
@@
 so, what other characteristics (personality, social preferences, political
-orientation) predict the use of different punishment strategies. Table
-\ref{tab:tableStrategies} summarises the potential functions for costly
+orientation, religious views) predict the use of different punishment strategies.
+Table \ref{tab:tableStrategies} summarises the potential functions for costly
 punishment in the economic games that we considered, and the behavioural
@@
 are variants on this general setup, creating situations where (A) Player 2 is
-still worse off by stealing, (B) Player 2 creates equality by stealing, (C) the
-computer ``decides'' whether Player 2 steals, (D) the fee-fine ratio is 1:1, (E)
-Player 2 is better off by stealing, and (F) Player 2 steals instead from a
-third-party.}\label{fig:plotGames}
+still worse off after stealing, (B) Player 2 creates equality by stealing, (C)
+the computer ``decides'' whether Player 2 steals, (D) the fee-fine ratio is 1:1,
+(E) Player 2 is better off after stealing, and (F) Player 2 steals instead from
+a third-party.}\label{fig:plotGames}
 \end{figure}
@@
 
-The overall pattern of punitive behaviour in the six economic games was in line
-with previous research and very similar across both countries (Figure
-\ref{fig:plotDecisions}). Participants were generally more likely to punish
-targets who stole from another individual compared to targets who did not steal
-(multilevel logistic regression; \emph{b} =
+The overall pattern of punitive behaviour in the six economic games was very
+similar across both countries (Figure \ref{fig:plotDecisions}). Participants
+were generally more likely to punish targets who stole compared to targets who
+did not steal (multilevel logistic regression; \emph{b} =
 1.93, standard error =
@@
 minimising payoff differences (avoid disadvantageous inequity, egalitarian).
-Less common were the behaviour-shaping strategies (deterrent, norm-enforcing),
-the retributive strategy, and the competitive strategies (seek AI, competitive).
-Although participants often punished targets who did not steal
-in the six games (Figure \ref{fig:plotDecisions}), no participants followed the
-antisocial strategy by exclusively punishing targets who did not steal across
-\emph{all} games.
+Less common were the behaviour shaping strategies (deterrent, norm-enforcing),
+the retributive strategy, and the competitive strategies (seek advantageous
+inequity, competitive). Although participants often punished targets who did not
+steal in the six games (Figure \ref{fig:plotDecisions}), no participants
+followed the antisocial strategy by exclusively punishing targets who did not
+steal across \emph{all} games.
 
@@
 \caption{\label{tab:tableStrategyCounts}Counts and proportions of participants
-following each punishment strategy exactly, split by country. N/A implies that
+following each punishment strategy exactly, split by country. \emph{N/A implies that
 participants were unable to be classified exactly into any of the punishment
-strategies.}
+strategies. AI = advantageous inequity, DI = disadvantageous inequity.}}
 
@@
 had the highest probability, followed by the egalitarian strategy. The
-norm-enforcing and seek AI strategies were the next most likely, with higher
-posterior estimates than the competitive and antisocial strategies. None of the
-other strategies differed in their posterior estimates. The same general pattern
-emerged when we analysed the full dataset without pre-registered exclusions
-(Supplementary Figure \ref{fig:plotModel1a}).
+norm-enforcing and seek advantageous inequity strategies were the next most
+likely, with higher posterior estimates than the competitive and antisocial
+strategies. None of the other strategies differed in their posterior estimates.
+The same general pattern emerged when we analysed the full dataset without
+pre-registered exclusions (Supplementary Figure \ref{fig:plotModel1a}).
 
@@
 assumes an implementation error rate of 5\%. Points represent posterior medians,
-line ranges represent 50\% and 95\% credible intervals.}
+line ranges represent 50\% and 95\% credible intervals. AI = advantageous
+inequity, DI = disadvantageous inequity.}
 \end{figure}
@@
 Social Value Orientation (SVO) in the model, we found associations with the
-egalitarian, never punish, and random choice strategies (Figure
-\ref{fig:plotAllPers2}). Participants higher in SVO were more likely to follow
-the egalitarian and the never punish strategies, while those with lower SVO
-scores were more likely to enact the random choice strategy. The personality
-dimensions of honesty-humility and openness to experience were both positively
-associated with following the never punish strategy, while extraversion
-negatively predicted this strategy. The effects were mostly similar across
-countries, but occasionally differed: for example, in the United States,
-but not in the United Kingdom, honesty-humility was positively associated with
-following the egalitarian strategy and negatively associated with following
-the random choice strategy. Overall, the same pattern of results emerged when
-analysing the full dataset without exclusions (Supplementary Figure
-\ref{fig:plotAllPers1}).
+egalitarian, never punish, and random choice strategies, with small-to-medium
+effect sizes (Figure \ref{fig:plotAllPers2}). More prosocial participants
+(those with higher SVO scores) were more likely to follow the egalitarian and
+the never punish strategies, while those with lower SVO scores were more likely
+to enact the random choice strategy. The personality dimensions of
+honesty-humility and openness to experience were both positively associated with
+following the never punish strategy, while extraversion negatively predicted
+this strategy. The effects were mostly similar across countries, but
+occasionally differed: for example, in the United States, but not in the United
+Kingdom, honesty-humility was positively associated with following the
+egalitarian strategy and negatively associated with following the random
+choice strategy. Overall, the same pattern of results emerged when analysing the
+full dataset without exclusions (Supplementary Figure \ref{fig:plotAllPers1}).
 
@@
 5\%. Points represent posterior medians, line ranges represent 95\% credible
-intervals.}
+intervals. AI = advantageous inequity, DI = disadvantageous inequity.}
 \end{figure}
@@
 Political and religious variables were also associated with punishment strategy
-(Figure \ref{fig:plotAllPolRel2}). These effects tended to be more pronounced
-in the United States. Controlling for Social Dominance Orientation, American
-participants higher in Right Wing Authoritarianism were more likely to follow
-the strategies avoiding disadvantageous inequity and seeking advantageous
-inequity. Participants who stated that they would like to ``bring those below
-them {[}on the socio-economic status ladder{]} up a peg'' were more likely to follow
-the egalitarian strategy, while American participants higher in Social Dominance
-Orientation, Right Wing Authoritarianism, and believing that God controls events
-in the world were less likely to follow the egalitarian strategy. In general,
-religious and conservative participants were less likely to follow the never
-punish strategy. This general pattern of results was replicated with the full
-dataset (Supplementary Figure \ref{fig:plotAllPolRel1}).
+(Figure \ref{fig:plotAllPolRel2}). These effects were small-to-medium in size
+and tended to be more pronounced in the United States. Controlling for Social
+Dominance Orientation, American participants higher in Right Wing
+Authoritarianism were more likely to follow the strategies avoiding
+disadvantageous inequity and seeking advantageous inequity. Participants who
+stated that they would like to ``bring those below them {[}on the socio-economic
+status ladder{]} up a peg'' were more likely to follow the egalitarian strategy,
+while American participants higher in Social Dominance Orientation, Right Wing
+Authoritarianism, and believing that God controls events in the world were less
+likely to follow the egalitarian strategy. In general, religious and
+conservative participants were less likely to follow the never punish strategy.
+This general pattern of results was replicated with the full dataset
+(Supplementary Figure \ref{fig:plotAllPolRel1}).
 
@@
 model. All models assume an implementation error rate of 5\%. Points represent
-posterior medians, line ranges represent 95\% credible intervals.}
+posterior medians, line ranges represent 95\% credible intervals. AI =
+advantageous inequity, DI = disadvantageous inequity.}
 \end{figure}
@@
 In general, we found that self-reported strategy usage was positively associated
-with the behavioural strategy that participants employed (see Supplementary
-Figures \ref{fig:plotSliders1} and \ref{fig:plotSliders2} for the distribution
-of responses to self-report questions). Figure \ref{fig:plotAllSliders2} shows
-the relationships between self-report questions and the different punishment
-strategies, highlighting the combinations where the question matched the
-behavioural strategy. We found positive relationships between the self-report
-questions and strategy usage for the norm-enforcing, egalitarian, seek
-advantageous inequity, never punish, and random choice strategies. The 95\%
-credible intervals for other estimates included zero, though these estimates
-often trended in a positive direction. The same pattern of results was found
-when analysing the full dataset without exclusions (Supplementary Figure
-\ref{fig:plotAllSliders1}).
+with the behavioural strategy that participants employed, with effects ranging
+from small to large in size (see Supplementary Figures \ref{fig:plotSliders1}
+and \ref{fig:plotSliders2} for the distribution of responses to self-report
+questions). Figure \ref{fig:plotAllSliders2} shows the relationships between
+self-report questions and the different punishment strategies, highlighting the
+combinations where the question matched the behavioural strategy. We found
+positive relationships between the self-report questions and strategy usage for
+the norm-enforcing, egalitarian, seek advantageous inequity, never punish, and
+random choice strategies. The 95\% credible intervals for other estimates
+included zero, though these estimates often trended in a positive direction. The
+same pattern of results was found when analysing the full dataset without
+exclusions (Supplementary Figure \ref{fig:plotAllSliders1}).
 
@@
 Points represent posterior medians, line ranges represent 95\% credible
-intervals.}
+intervals. AI = advantageous inequity, DI = disadvantageous inequity.}
 \end{figure}
@@
 we have shown that punishment does not serve just one function, but instead is
-a flexible tool that can be and is used for different purposes\textsuperscript{15}.
-Punishment is more akin to a swiss army knife than a hammer, used by some to
-enforce norms of cooperation and by others to reduce or even create inequality
-between individuals. We found that people's punishment strategy can, to some
-extent, be predicted by individual differences in personality, social
-preferences, and political and religious views. Moreover, contrary to the view
-that people are often unable to articulate the reasons for their punitive
-behaviour\textsuperscript{28}, people seem to have some degree of insight into the
-strategy they are using. Despite small differences, these general patterns
-replicated in samples from both the United Kingdom and the United States.
+a flexible tool that can be and is used for different purposes\textsuperscript{6}.
+While some use punishment to enforce norms of cooperation, others use it to
+reduce or even create inequality between individuals. We found that people's
+punishment strategy can, to some extent, be predicted by individual differences
+in personality, social preferences, and political and religious views. Moreover,
+contrary to the view that people are often unable to articulate the reasons for
+their punitive behaviour\textsuperscript{28}, people seem to have some degree of
+insight into the strategy they are using. Despite small differences, these
+general patterns replicated in samples from both the United Kingdom and the
+United States, providing further confidence in the results.
 
@@
 line with previous studies which have highlighted inequity aversion as an
-important motivation for punishment in economic games\textsuperscript{29,30}. Personality and social preference variables mapped
-onto these strategies in expected ways. Traits associated with other-regarding
-concern, such as SVO and honesty-humility, predicted following the egalitarian
-strategy, whereas religious and conservative individuals were less likely to
-follow this strategy, especially in the United States. Moreover, participants
-following the egalitarian strategy were able to self-report this strategy,
-though the same was not true for the avoid disadvantageous inequity strategy.
+important motivation for punishment in economic games\textsuperscript{29--31,37,48,50}.
+Personality, social preferences, and political and religious views were related
+to these strategies. Traits associated with other-regarding concern, such as SVO
+and honesty-humility, predicted following the egalitarian strategy. By contrast,
+religious and conservative individuals, including individuals higher in SDO and
+RWA, were less likely to follow the egalitarian strategy and more likely to
+follow the avoid disadvantageous inequity strategy, especially in the United
+States. Authoritarian participants in the United States were also more likely to
+actively seek advantageous inequity with their punishment. These findings
+provide insight into the motives that may have driven higher levels of peer
+punishment among conservatives in previous work\textsuperscript{41}.
 
@@
 (Supplementary Figures \ref{fig:plotSliders1} and \ref{fig:plotSliders2}).
-Regarding the predictors of these strategies, we found that demographic,
-personality, political, and religious variables tended to be unrelated to
-deterrent and norm-enforcing punishment strategies. We also found that
-participants had insight into the norm-enforcing strategy, but not the deterrent
+Although our design did not explicitly allow for behaviour shaping as the
+interactions were all one-shot, we did manipulate whether the target's stealing
+behaviour was intentional or not (Game C), an approach which has been used in
+previous vignette studies to identify behaviour shaping motives\textsuperscript{28}.
+The lower prevalence of behaviour shaping strategies in our study is consistent
+with prior work showing that punishment often continues to be used when
+behaviour shaping is impossible, such as when the target will never find out
+that they have been punished\textsuperscript{34} or on the last round of repeated
+interactions\textsuperscript{33}. We found that demographic, personality, political,
+and religious variables tended to be unrelated to behaviour shaping strategies,
+though this could reflect low power to detect such associations given the low
+prevalence of these strategies in our sample. We also found that participants
+accurately reported using the norm-enforcing strategy, but not the deterrent
 strategy. This finding is in line with previous research showing that people
@@
 
-Other punitive strategies were less common in our dataset, but some were more
-prevalent than others. For example, participants were more likely to use
-punishment to seek advantageous inequity than to exclusively harm those who did
-not steal (i.e., antisocial punishment). The existence of the ``seek AI'' strategy
-in our dataset supports the claim that punishment can also be used as a tool to
-increase one's own relative position\textsuperscript{15}. While generally rare, we
-found that this motive for punishment was more common among authoritarian
-participants, at least in the United States, potentially providing an
-explanation for why peer punishment has been found to be more common among
-conservatives in previous work\textsuperscript{41}. Moreover, the fact that no
-participants in our sample punished non-stealing across all games suggests that
-antisocial punishment does not function to harm cooperators specifically, as has
-been previously suggested\textsuperscript{17}. Instead, antisocial punishment appears
-to be motivated by improving one's relative position in general, which is in
-line with work showing that antisocial punishment disappears with a 1:1 fee-fine
-ratio\textsuperscript{35}.
+We defined antisocial punishment as occurring when a participant exclusively
+punished those who did not steal. There were no participants in our sample who
+followed this strategy, although some players did punish non-stealing
+co-players. Harming non-stealing individuals was also consistent with the
+competitive strategies, which did appear in our sample albeit at low frequencies.
+The fact that no participants in our sample punished non-stealing across all
+games suggests that the antisocial punishment that has been observed in other
+studies\textsuperscript{17,51} is not aimed at harming cooperators
+specifically, as has been previously suggested\textsuperscript{17}. Instead,
+antisocial punishment is more likely to be motivated by improving one's
+relative position, which is in line with work showing that antisocial punishment
+disappears when relative payoffs cannot be changed (e.g., when punishment is
+only available with a 1:1 fee-fine ratio\textsuperscript{35}).
 
@@
 for the way that punishment is operationalised in classic behavioural economic
-game studies. In these studies, a common assumption is that participants will
-punish to change the behaviour of cheats. But in reality, people may be choosing
-the punishment option to achieve a variety of different goals. The targets of
-punishment in these studies are likely well aware that punishment could be
-levied for these different reasons and this knowledge may impact their
-responses. For example, if cheating targets interpret punishment as serving a
-competitive motive, it may elicit retaliation rather than encourage
-cooperation\textsuperscript{9,15,50}. This might help to
-explain the mixed findings in the field as to whether punishment actually
-motivates cheating targets to cooperate in the future\textsuperscript{15}.
+game studies. In these studies, a common assumption is that participants punish
+to change the behaviour of cheats\textsuperscript{11,12}. But in reality, people
+may be choosing the punishment option to achieve a variety of different goals.
+This has implications for how people respond to being targeted by punishers in
+these games. Targets of punishment in these studies may know that punishment
+reflects different motives and can respond accordingly. For example, if targets
+interpret punishment as serving a competitive motive, it may elicit retaliation
+rather than encourage cooperation\textsuperscript{6,10,52}.
+As punishers' motives must be inferred (and such inferences likely depend on
+character traits of the target, as well as the context in which punishment
+occurs), there is likely to be some variation and error in attributing motives
+to punishers. To the extent that inferred motives affect target responses, this
+might help to explain the mixed findings in the field as to whether punishment
+actually motivates cheating targets to cooperate in the future\textsuperscript{6}.
 
@@
 This is partly because punishment in these games imposes an economic cost for no
-tangible benefit. If the fee-fine ratio had been lower such that it was cheaper
+tangible benefit. If the fee-fine ratio had been lower, such that it was cheaper
 to punish, we may have seen more punishment from participants. Indeed,
@@
 more general aversion to peer punishment, an aversion that has been highlighted
-in both WEIRD (Western, educated, industrialised, rich, and democratic) samples\textsuperscript{51,52} and in small-scale societies\textsuperscript{53}.
-One reason that people may be averse to peer punishment is that, due to its
-multipurpose nature, it may be interpreted as a competitive challenge by targets
-and trigger retaliation\textsuperscript{15}. In situations that lack clear
-institutional norms to legitimise punishment, such as our economic games and
-some situations in the real world, people might abstain from peer punishment to
-avoid such retaliation, regardless of whether retaliation is actually possible.
-By contrast, institutionalised punishment in small-scale societies often
-functions to compensate victims adequately while limiting the potential for
-feuds and cycles of retaliation\textsuperscript{54,55}. Future research
-should uncover whether people are more willing to punish in these
-conventionalised contexts.
-
-There are several limitations with our study design that can guide future
-research. First, we used one-shot economic games to measure punishment
-strategies, which may have led us to underestimate behaviour-change strategies
-like deterrence. Our inclusion of Game C somewhat mitigated this issue by
-manipulating whether stealing behaviour was intentional vs.~unintentional and
-thus whether there was any behaviour to be deterred. Due to limits on our
-within-subjects design and the complexity of the strategy space, it was not
-feasible for us to expand our study to include additional contexts to elicit
-behaviour-change strategies (e.g., repeated games, games where targets are not
-made aware of the punishment, games where targets can retaliate). Future work
-could study these contexts separately.
-
-A second limitation is that some strategies required more punishment than others
-to be met. For example, the competitive strategy required punishment in ten of
-twelve decisions, compared to the avoid disadvantageous inequity strategy which
-required only one instance of punishment (Table \ref{tab:tableStrategies}).
+in both WEIRD (Western, educated, industrialised, rich, and democratic) samples\textsuperscript{53,54} and in small-scale societies\textsuperscript{55}.
+One reason that people may be averse to peer punishment is that, because it is
+a fundamentally harmful act, punishment can reflect badly on the punisher and
+people might therefore refrain from punishing others to avoid reputational
+damage\textsuperscript{4}. People frequently avoid taking actions that could harm
+their reputation, even when they don't know if reputation is at stake (as in the
+one-shot anonymous settings used here). Another reason that people may be averse
+to peer punishment is that it can trigger retaliation\textsuperscript{6}. This may be
+especially likely in situations that lack clear institutional norms to
+legitimise punishment, such as our economic games. As with reputation damage,
+people might abstain from peer punishment to avoid retaliation, regardless of
+whether retaliation is actually possible. By contrast, institutionalised
+punishment in small-scale societies often functions to compensate victims while
+limiting the potential for feuds and cycles of retaliation\textsuperscript{56,57}. Future research should uncover whether people are
+more willing to punish in these conventionalised contexts (e.g., see\textsuperscript{58}).
+
+We finally highlight some key limitations that can guide future research. One
+limitation with our design is that some strategies required more punishment than
+others to be met. For example, the competitive strategy required punishment in
+ten of twelve decisions, compared to the avoid disadvantageous inequity strategy
+which required only one instance of punishment (Table \ref{tab:tableStrategies}).
 Strategies thus differed in how ``expensive'' they were to implement, perhaps
@@
 guidelines and regulations. Informed consent was obtained from all participants
-prior to the study.
+prior to the study and participation was voluntary.
 
@@
 
+Our punishment strategies make different behavioural predictions across these
+different games (Table \ref{tab:tableStrategies}). The deterrent strategy
+punishes whenever it is intentionally personally harmed (when the target steals
+in Games A, B, D, and E). The norm-enforcing strategy is the same, except that
+it also punishes when \emph{others} are intentionally harmed (when the target steals
+in Game F). The retributive strategy punishes whenever it is personally harmed,
+whether the harm was intentional or not (when the target steals in Games A-E).
+The avoid disadvantageous inequity strategy only punishes when it currently
+experiences disadvantageous inequity and can alleviate it by punishing (when the
+target steals in Game E). The egalitarian strategy punishes when punishment
+makes payoffs for all players more equal than they currently are (when the
+target steals in Games E and F). The seek advantageous inequity strategy
+punishes when it currently does not experience advantageous inequity and
+punishment would produce advantageous inequity for itself (when the target
+steals in Games B and C). The competitive strategy punishes whenever the
+fee-fine ratio is 1:3, as it can improve its relative position by punishing, but
+does not punish when the fee-fine ratio is 1:1 (Game D). The antisocial strategy
+punishes whenever the target does not steal across all games. The never punish
+strategy never punishes in any game.
+
 For each game, participants saw the game instructions and answered a
@@
   \emph{Demographics}. In the survey, we collected information on participants'
-  education level and self-reported socio-economic status (MacArthur ladder\textsuperscript{56}). We also collected additional demographic data from Prolific
+  education level and self-reported socio-economic status (MacArthur ladder\textsuperscript{59}). We also collected additional demographic data from Prolific
   (e.g., age, gender, student status).
 \item
-  \emph{Personality}. We used the Mini-IPIP scale\textsuperscript{57} to measure the Big 6
-  personality dimensions of agreeableness, conscientiousness, extraversion,
-  honesty-humility, openness to experience, and neuroticism (four items each).
+  \emph{Personality}. We used the Mini-IPIP scale\textsuperscript{60} to measure the Big 6
+  personality dimensions of agreeableness (\(\alpha\) =
+  0.83), conscientiousness (\(\alpha\) =
+  0.75), extraversion (\(\alpha\) =
+  0.83), honesty-humility (\(\alpha\) =
+  0.77), openness to experience (\(\alpha\) =
+  0.79), and neuroticism (\(\alpha\) =
+  0.79). Four items were used for each
+  personality dimension.
 \item
   \emph{Social Value Orientation}. We used the Social Value Orientation Slider
-  Measure to measure other-regarding preferences\textsuperscript{58}. Across fifteen
+  Measure to measure other-regarding preferences\textsuperscript{61}. Across fifteen
   items, participants made decisions on how to allocate different amounts of
@@
   measure of their other-regarding preference, following the steps outlined in
-  ref\textsuperscript{58}.
+  ref\textsuperscript{61}.
 \item
   \emph{Political ideology}. We included several measures of political ideology,
-  including left-right conservatism, Social Dominance Orientation\textsuperscript{59}
-  (eight items), and Right Wing Authoritarianism\textsuperscript{60} (six items).
-  We also probed participants' views on social inequality by asking them whether
-  they would like to bring people above (below) them on the MacArthur
-  socio-economic status ladder down (up) a peg or two.
+  including left-right conservatism, Social Dominance Orientation\textsuperscript{62}
+  (\(\alpha\) = 0.91; eight items), and Right
+  Wing Authoritarianism\textsuperscript{63} (\(\alpha\) =
+  0.82; six items). We also probed
+  participants' views on social inequality by asking them whether they would
+  like to bring people above (below) them on the MacArthur socio-economic status
+  ladder down (up) a peg or two.
 \item
@@
   themselves and whether they believe that God or another spiritual non-human
-  entity controls the events in the world\textsuperscript{61}.
+  entity controls the events in the world\textsuperscript{64}.
 \end{itemize}
@@
 to comprehension questions suggested that participants understood the six
-economic games (Supplementary Table \ref{tab:tableComp}). We used the strategy
-method to incentivise the economic games, choosing a random game to determine
-bonus payment. After all games,
+economic games (Supplementary Table \ref{tab:tableComp}). In order to partially
+mitigate the fact that some punishment strategies predict more punishment than
+others and are thus more ``expensive'' to implement (Table
+\ref{tab:tableStrategies}), we used the strategy method to incentivise the
+economic games, choosing a random game to determine bonus payments rather than
+summing participants' earnings across all games. After all games,
 62\%
@@
 payment from the six economic games (between £0.40 -- £0.70 depending on their
-decision). Participants were paid £1.50 for completing the follow-up survey,
+decisions). Participants were paid £1.50 for completing the follow-up survey,
 plus a bonus payment from the Social Value Orientation Slider Measure (between
-£0.50 -- £0.85 depending on their decision).
+£0.50 -- £0.85 depending on their decisions).
 
@@
 unobserved punishment strategies from the observed data (for a similar version
-of this model, see ref\textsuperscript{62}). In this model, participants \(i\) in
+of this model, see ref\textsuperscript{65}). In this model, participants \(i\) in
 countries \(c\) make binary punishment decisions across twelve decisions \(j\). We
@@
 of \(1 - \delta\) and coding red crosses with a conditional probability of
-\(0 + \delta\). The random choice is consistently coded with a conditional
-probability of ½ across all decisions.
+\(0 + \delta\). We set \(\delta\) to 0.05 in all models, which is similar to its
+value when we estimate it as a free parameter in an additional model (median
+posterior \(\delta\) = 0.03, 95\% CI
+{[}0.00 0.06{]};
+Supplementary Figure \ref{fig:plotModel5}). The random choice strategy is
+consistently coded with a conditional probability of ½ across all decisions.
 
@@
 
+These models control for multiple comparisons across strategies by estimating
+the effects of the predictor on all strategies simultaneously.
+
 We estimated the posterior distributions of these models using Hamiltonian Monte
-Carlo as implemented in Stan version 2.26.1\textsuperscript{63}. We ran each model for
+Carlo as implemented in Stan version 2.26.1\textsuperscript{66}. We ran each model for
 2000 samples, with 1000 warmup samples. R-hat values and effective sample sizes
@@
 
-We validated the model by simulating observed data (n = 100) from a known
+We validated the model by simulating observed data (\emph{n} = 100) from a known
 frequency of strategies. The model was successfully able to recover the known
@@
 \url{https://github.com/ScottClaessens/punishStrategies}. All analyses were conducted
-in R version 4.2.1\textsuperscript{64}.
-Visualisations were created with the \emph{ggplot2}\textsuperscript{65} and \emph{cowplot}\textsuperscript{66} R packages.
-We used the \emph{targets}\textsuperscript{67} R package to create a reproducible data analysis pipeline and the
-\emph{papaja}\textsuperscript{68} R package to reproducibly generate the manuscript.
+in R version 4.2.1\textsuperscript{67}.
+Visualisations were created with the \emph{ggplot2}\textsuperscript{68} and \emph{cowplot}\textsuperscript{69} R packages.
+We used the \emph{targets}\textsuperscript{70} R package to create a reproducible data analysis pipeline and the
+\emph{papaja}\textsuperscript{71} R package to reproducibly generate the manuscript.
 
@@
 
-All authors conceptualised the research, designed the study, and developed the
-surveys. N.J. conducted data collection on Prolific. S.C. conducted all analyses
-and visualisation of the data. All authors wrote the manuscript.
+All authors (S.C., Q.D.A., N.R.) conceptualised the research, designed the
+study, and developed the surveys. N.R. conducted data collection on Prolific.
+S.C. conducted all analyses and visualisation of the data. All authors wrote the
+manuscript.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Raihani2019}{}}%
+\CSLLeftMargin{6. }%
+\CSLRightInline{Raihani, N. J. \& Bshary, R. \href{https://doi.org/10.1017/ehs.2019.12}{Punishment: One tool, many uses}. \emph{Evolutionary Human Sciences} \textbf{1}, e12 (2019).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Balliet2011}{}}%
-\CSLLeftMargin{6. }%
+\CSLLeftMargin{7. }%
 \CSLRightInline{Balliet, D., Mulder, L.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Balliet2013}{}}%
-\CSLLeftMargin{7. }%
+\CSLLeftMargin{8. }%
 \CSLRightInline{Balliet, D.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Chaudhuri2011}{}}%
-\CSLLeftMargin{8. }%
+\CSLLeftMargin{9. }%
 \CSLRightInline{Chaudhuri, A.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Dreber2008}{}}%
-\CSLLeftMargin{9. }%
+\CSLLeftMargin{10. }%
 \CSLRightInline{Dreber, A., Rand, D. G., Fudenberg, D. \& Nowak, M. A. \href{https://doi.org/10.1038/nature06723}{Winners don't punish}. \emph{Nature} \textbf{452}, 348--351 (2008).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Fehr2000}{}}%
-\CSLLeftMargin{10. }%
+\CSLLeftMargin{11. }%
 \CSLRightInline{Fehr, E.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Fehr2002}{}}%
-\CSLLeftMargin{11. }%
+\CSLLeftMargin{12. }%
 \CSLRightInline{Fehr, E. \& Gächter, S. \href{https://doi.org/10.1038/415137a}{Altruistic punishment in humans}. \emph{Nature} \textbf{415}, 137--140 (2002).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Henrich2006}{}}%
-\CSLLeftMargin{12. }%
+\CSLLeftMargin{13. }%
 \CSLRightInline{Henrich, J. \emph{et al.} \href{https://doi.org/10.1126/science.1127333}{Costly punishment across human societies}. \emph{Science} \textbf{312}, 1767--1770 (2006).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Nikiforakis2008a}{}}%
-\CSLLeftMargin{13. }%
+\CSLLeftMargin{14. }%
 \CSLRightInline{Nikiforakis, N. \& Normann, H.-T. A comparative statics analysis of punishment in public-good experiments. \emph{Experimental Economics} \textbf{11}, 358--369 (2008).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Raihani2012a}{}}%
-\CSLLeftMargin{14. }%
+\CSLLeftMargin{15. }%
 \CSLRightInline{Raihani, N.
@@
 \emph{Trends in Ecology \& Evolution} \textbf{27}, 288--295 (2012).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Raihani2019}{}}%
-\CSLLeftMargin{15. }%
-\CSLRightInline{Raihani, N. J. \& Bshary, R. \href{https://doi.org/10.1017/ehs.2019.12}{Punishment: One tool, many uses}. \emph{Evolutionary Human Sciences} \textbf{1}, e12 (2019).}
 
@@
 \& Fehr, E.
-\href{https://doi.org/10.1037/xge0000018}{The value of vengeance and the demand for deterrence.} \emph{Journal of Experimental Psychology: General} \textbf{143}, 2279--2286 (2014).}
+\href{https://doi.org/10.1037/xge0000018}{The value of vengeance and the demand for deterrence}.
+\emph{Journal of Experimental Psychology: General} \textbf{143}, 2279--2286 (2014).}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Bone2016}{}}%
+\CSLLeftMargin{50. }%
+\CSLRightInline{Bone, J. E., McAuliffe, K. \& Raihani, N. J. Exploring the motivations for punishment: Framing and country-level effects. \emph{PLoS One} \textbf{11}, e0159769 (2016).}
+
+\leavevmode\vadjust pre{\hypertarget{ref-Pleasant2018}{}}%
+\CSLLeftMargin{51. }%
+\CSLRightInline{Pleasant, A.
+\& Barclay, P.
+Why hate the good guy?
+Antisocial punishment of high cooperators is greater when people compete to be chosen.
+\emph{Psychological Science} \textbf{29}, 868--876 (2018).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Nikiforakis2008b}{}}%
-\CSLLeftMargin{50. }%
+\CSLLeftMargin{52. }%
 \CSLRightInline{Nikiforakis, N.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Balafoutas2014}{}}%
-\CSLLeftMargin{51. }%
+\CSLLeftMargin{53. }%
 \CSLRightInline{Balafoutas, L., Nikiforakis, N.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Balafoutas2016}{}}%
-\CSLLeftMargin{52. }%
+\CSLLeftMargin{54. }%
 \CSLRightInline{Balafoutas, L., Nikiforakis, N.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Baumard2010}{}}%
-\CSLLeftMargin{53. }%
+\CSLLeftMargin{55. }%
 \CSLRightInline{Baumard, N.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Fitouchi2023}{}}%
-\CSLLeftMargin{54. }%
+\CSLLeftMargin{56. }%
 \CSLRightInline{Fitouchi, L.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Singh2022}{}}%
-\CSLLeftMargin{55. }%
+\CSLLeftMargin{57. }%
 \CSLRightInline{Singh, M. \& Garfield, Z. H. Evidence for third-party mediation but not punishment in {M}entawai justice. \emph{Nature Human Behaviour} \textbf{6}, 930--940 (2022).}
 
+\leavevmode\vadjust pre{\hypertarget{ref-Molleman2019}{}}%
+\CSLLeftMargin{58. }%
+\CSLRightInline{Molleman, L., Kölle, F., Starmer, C. \& Gächter, S. People prefer coordinated punishment in cooperative interactions. \emph{Nature Human Behaviour} \textbf{3}, 1145--1153 (2019).}
+
 \leavevmode\vadjust pre{\hypertarget{ref-Adler2000}{}}%
-\CSLLeftMargin{56. }%
+\CSLLeftMargin{59. }%
 \CSLRightInline{Adler, N.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Sibley2011}{}}%
-\CSLLeftMargin{57. }%
+\CSLLeftMargin{60. }%
 \CSLRightInline{Sibley, C.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Murphy2011}{}}%
-\CSLLeftMargin{58. }%
+\CSLLeftMargin{61. }%
 \CSLRightInline{Murphy, R.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Ho2015}{}}%
-\CSLLeftMargin{59. }%
+\CSLLeftMargin{62. }%
 \CSLRightInline{Ho, A.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Bizumic2018}{}}%
-\CSLLeftMargin{60. }%
+\CSLLeftMargin{63. }%
 \CSLRightInline{Bizumic, B.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Laurin2012}{}}%
-\CSLLeftMargin{61. }%
+\CSLLeftMargin{64. }%
 \CSLRightInline{Laurin, K., Shariff, A.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-McElreath2020}{}}%
-\CSLLeftMargin{62. }%
+\CSLLeftMargin{65. }%
 \CSLRightInline{McElreath, R. \emph{\href{http://xcelab.net/rm/statistical-rethinking/}{Statistical rethinking: A {Bayesian} course with examples in {R} and {Stan}, 2nd edition}}. (CRC Press, 2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Stan2020}{}}%
-\CSLLeftMargin{63. }%
+\CSLLeftMargin{66. }%
 \CSLRightInline{Stan Development Team. \href{http://mc-stan.org/}{{RStan}: The {R} interface to {Stan}}. (2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
-\CSLLeftMargin{64. }%
+\CSLLeftMargin{67. }%
 \CSLRightInline{R Core Team. \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}}. (R Foundation for Statistical Computing, 2022).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
-\CSLLeftMargin{65. }%
+\CSLLeftMargin{68. }%
 \CSLRightInline{Wickham, H. \emph{\href{https://ggplot2.tidyverse.org}{{ggplot2}: Elegant graphics for data analysis}}. (Springer-Verlag New York, 2016).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
-\CSLLeftMargin{66. }%
+\CSLLeftMargin{69. }%
 \CSLRightInline{Wilke, C. O. \emph{\href{https://CRAN.R-project.org/package=cowplot}{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}}. (2020).}
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-\CSLLeftMargin{67. }%
+\CSLLeftMargin{70. }%
 \CSLRightInline{Landau, W.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
-\CSLLeftMargin{68. }%
+\CSLLeftMargin{71. }%
 \CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}}. (2022).}
@@
 implementation error rate of 5\%. Points represent posterior medians, line ranges
-represent 50\% and 95\% credible intervals.}
+represent 50\% and 95\% credible intervals. AI = advantageous inequity, DI =
+disadvantageous inequity.}
 \end{figure}
@@
 dataset with pre-registered exclusions.} Each row represents a separate model.
-Points represent posterior medians, line ranges represent 95\% credible intervals.}
+Points represent posterior medians, line ranges represent 95\% credible
+intervals. AI = advantageous inequity, DI = disadvantageous inequity.}
 \end{figure}
@@
 without pre-registered exclusions.} Each row represents a separate model. Points
-represent posterior medians, line ranges represent 95\% credible intervals.}
+represent posterior medians, line ranges represent 95\% credible intervals. AI =
+advantageous inequity, DI = disadvantageous inequity.}
 \end{figure}
@@
 control for one another within the same model. Points represent posterior
-medians, line ranges represent 95\% credible intervals.}
+medians, line ranges represent 95\% credible intervals. AI = advantageous
+inequity, DI = disadvantageous inequity.}
 \end{figure}
@@
 competitive strategy. Points represent posterior medians, line ranges represent
-95\% credible intervals.}
+95\% credible intervals. AI = advantageous inequity, DI = disadvantageous
+inequity.}
 \end{figure}
@@
 \caption{\label{fig:plotSampleUS}\emph{Sample characteristics in the United States.}}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotModel5-1.pdf}
+\caption{\label{fig:plotModel5}\emph{Posterior estimates of the probabilities of following
+different punishment strategies from the Bayesian latent state model that
+estimates the implementation error rate as a free parameter.} The model
+estimated the implementation error rate to be 0.03
+(95\% CI {[}0.00
+0.06{]}). Points represent posterior medians, line
+ranges represent 50\% and 95\% credible intervals. AI = advantageous inequity, DI
+= disadvantageous inequity.}
 \end{figure}
@@
 represent known strategy frequencies, grey densities represent posterior
-estimates of strategy frequencies.}
+estimates of strategy frequencies. AI = advantageous inequity, DI =
+disadvantageous inequity.}
 \end{figure}
@@
 000000001010 & \textit{Egalitarian strategy (exact)} & 65 & 0.064 & 71 & 0.071\\
-000000000010 & Punish when take in Game F & 55 & 0.054 & 49 & 0.049\\
-001000001000 & Punish when take in Games B and E & 14 & 0.014 & 11 & 0.011\\
-101000001010 & Punish when take in Games A, B, E, and F & 11 & 0.011 & 4 & 0.004\\
-100000000000 & Punish when take in Game A & 10 & 0.010 & 2 & 0.002\\
-000000100000 & Punish when take in Game D & 9 & 0.009 & 3 & 0.003\\
-001000001010 & Punish when take in Games B, E, and F & 9 & 0.009 & 17 & 0.017\\
+000000000010 & Punish when steal in Game F & 55 & 0.054 & 49 & 0.049\\
+001000001000 & Punish when steal in Games B and E & 14 & 0.014 & 11 & 0.011\\
+101000001010 & Punish when steal in Games A, B, E, and F & 11 & 0.011 & 4 & 0.004\\
+100000000000 & Punish when steal in Game A & 10 & 0.010 & 2 & 0.002\\
+000000100000 & Punish when steal in Game D & 9 & 0.009 & 3 & 0.003\\
+001000001010 & Punish when steal in Games B, E, and F & 9 & 0.009 & 17 & 0.017\\
 101000101000 & \textit{Deterrent strategy (exact)} & 9 & 0.009 & 6 & 0.006\\
-101010101010 & Punish when take in all games & 9 & 0.009 & 15 & 0.015\\
+101010101010 & Punish when steal in all games & 9 & 0.009 & 15 & 0.015\\
 101000101010 & \textit{Norm-enforcing strategy (exact)} & 8 & 0.008 & 16 & 0.016\\
-001000000000 & Punish when take in Game B & 7 & 0.007 & 4 & 0.004\\
-001010101000 & Punish when take in Games B, C, D, and E & 7 & 0.007 & 0 & 0.000\\
-100000001000 & Punish when take in Games A and E & 6 & 0.006 & 5 & 0.005\\
-101000001000 & Punish when take in Games A, B, and E & 6 & 0.006 & 7 & 0.007\\
+001000000000 & Punish when steal in Game B & 7 & 0.007 & 4 & 0.004\\
+001010101000 & Punish when steal in Games B, C, D, and E & 7 & 0.007 & 0 & 0.000\\
+100000001000 & Punish when steal in Games A and E & 6 & 0.006 & 5 & 0.005\\
+101000001000 & Punish when steal in Games A, B, and E & 6 & 0.006 & 7 & 0.007\\
 101010101000 & \textit{Retributive strategy (exact)} & 6 & 0.006 & 5 & 0.005\\
 111111111111 & Always punish & 6 & 0.006 & 16 & 0.016\\
-000000101000 & Punish when take in Games D and E & 5 & 0.005 & 2 & 0.002\\
-000000101010 & Punish when take in Games D, E, and F & 5 & 0.005 & 3 & 0.003\\
-101010001010 & Punish when take in all games except Game D & 5 & 0.005 & 2 & 0.002\\
-001000101000 & Punish when take in Games B, D, and E & 4 & 0.004 & 2 & 0.002\\
-001000101010 & Punish when take in Games B, D, E, and F & 4 & 0.004 & 6 & 0.006\\
-101000000000 & Punish when take in Games A and B & 4 & 0.004 & 2 & 0.002\\
-101010001000 & Punish when take in Games A, B, C, and E & 4 & 0.004 & 0 & 0.000\\
+000000101000 & Punish when steal in Games D and E & 5 & 0.005 & 2 & 0.002\\
+000000101010 & Punish when steal in Games D, E, and F & 5 & 0.005 & 3 & 0.003\\
+101010001010 & Punish when steal in all games except Game D & 5 & 0.005 & 2 & 0.002\\
+001000101000 & Punish when steal in Games B, D, and E & 4 & 0.004 & 2 & 0.002\\
+001000101010 & Punish when steal in Games B, D, E, and F & 4 & 0.004 & 6 & 0.006\\
+101000000000 & Punish when steal in Games A and B & 4 & 0.004 & 2 & 0.002\\
+101010001000 & Punish when steal in Games A, B, C, and E & 4 & 0.004 & 0 & 0.000\\
 \bottomrule

### 2023-12-13 R/functions_modelFitting.R

@@ -197,3 +197,39 @@ fitMLM2 <- function(d) {
       family = binomial
     )
 }
+
+# fit stan model without predictor (estimate implementation error)
+fitModel5 <- function(d, compiledModel5,
+                      iter = 2000, warmup = 1000, chains = 4, cores = 4) {
+  # list for stan
+  dataList <-
+    list(
+      N = nrow(d),
+      Nc = length(unique(d$Country)),
+      countryID = ifelse(d$Country == "United Kingdom", 1L, 2L),
+      pun1_1 = ifelse(!is.na(d$NoDI1_Take),    d$pun1_1, -999),
+      pun1_2 = ifelse(!is.na(d$NoDI1_Nothing), d$pun1_2, -999),
+      pun2_1 = ifelse(!is.na(d$NoDI2_Take),    d$pun2_1, -999),
+      pun2_2 = ifelse(!is.na(d$NoDI2_Nothing), d$pun2_2, -999),
+      pun3_1 = ifelse(!is.na(d$NoDI3_Take),    d$pun3_1, -999),
+      pun3_2 = ifelse(!is.na(d$NoDI3_Nothing), d$pun3_2, -999),
+      pun4_1 = ifelse(!is.na(d$NoDI4_Take),    d$pun4_1, -999),
+      pun4_2 = ifelse(!is.na(d$NoDI4_Nothing), d$pun4_2, -999),
+      pun5_1 = ifelse(!is.na(d$DI_Take),       d$pun5_1, -999),
+      pun5_2 = ifelse(!is.na(d$DI_Nothing),    d$pun5_2, -999),
+      pun6_1 = ifelse(!is.na(d$`3PP_Take`),    d$pun6_1, -999),
+      pun6_2 = ifelse(!is.na(d$`3PP_Nothing`), d$pun6_2, -999)
+    )
+  # fit model
+  out <- 
+    sampling(
+      compiledModel5, 
+      data = dataList, 
+      iter = iter,
+      warmup = warmup,
+      chains = chains,
+      cores = cores,
+      seed = 2113
+    )
+  return(out)
+}

### 2023-12-13 R/functions_simulation.R

@@ -91,7 +91,7 @@ plotSimResults1 <- function(simAlphas, simPost) {
   # strategy vector
   strategies <- c("Competitive", "Avoid DI", "Egalitarian", "Seek AI",
                   "Retributive", "Deterrent", "Norm-enforcing", 
-                  "Antisocial", "Random choice", "Anti-punish")
+                  "Antisocial", "Random choice", "Never punish")
   # get simulated probabilities (intercepts)
   simProbs <- softmax(simAlphas)
   # calculate probabilities
@@ -104,7 +104,10 @@ plotSimResults1 <- function(simAlphas, simPost) {
             P[,1,6], P[,1,7], P[,1,8], P[,1,9], P[,1,10]),
       strategy = rep(strategies, each = length(P[,1,1]))
     ) %>%
-    mutate(strategy = factor(strategy, levels = strategies)) %>%
+    mutate(strategy = factor(strategy, levels = c("Deterrent", "Norm-enforcing", "Retributive", 
+                                                  "Avoid DI", "Egalitarian", "Seek AI",
+                                                  "Competitive", "Antisocial", 
+                                                  "Never punish", "Random choice"))) %>%
     ggplot(aes(x = p, y = fct_rev(strategy))) +
     geom_density_ridges(rel_min_height = 0.01) +
     geom_point(data = tibble(strategy = strategies, p = simProbs), colour = "blue") +
@@ -120,7 +123,7 @@ plotSimResults2 <- function(dSim, simPost) {
   # strategy vector
   strategies <- c("Competitive", "Avoid DI", "Egalitarian", "Seek AI",
                   "Retributive", "Deterrent", "Norm-enforcing", 
-                  "Antisocial", "Random choice", "Anti-punish")
+                  "Antisocial", "Random choice", "Never punish")
   # x sequence
   x <- seq(min(dSim$x), max(dSim$x), length.out = 100)
   # post
@@ -137,7 +140,10 @@ plotSimResults2 <- function(dSim, simPost) {
         post, 
         tibble(
           x = x[i],
-          strategy = factor(strategies, levels = strategies),
+          strategy = factor(strategies, levels = c("Deterrent", "Norm-enforcing", "Retributive", 
+                                                   "Avoid DI", "Egalitarian", "Seek AI",
+                                                   "Competitive", "Antisocial", 
+                                                   "Never punish", "Random choice")),
           med = apply(p, 2, median),
           lower = apply(p, 2, quantile, 0.025),
           upper = apply(p, 2, quantile, 0.975)

### 2023-12-13 R/functions_summariseData.R

@@ -395,28 +395,28 @@ makePatternsTable <- function(d) {
     "000000000000" = "\\textit{Never punish strategy (exact)}",
     "000000001000" = "\\textit{Avoid DI strategy (exact)}",
     "000000001010" = "\\textit{Egalitarian strategy (exact)}",
-    "000000000010" = "Punish when take in Game F",
-    "001000001000" = "Punish when take in Games B and E",
-    "101000001010" = "Punish when take in Games A, B, E, and F",
-    "100000000000" = "Punish when take in Game A",
-    "000000100000" = "Punish when take in Game D",
-    "001000001010" = "Punish when take in Games B, E, and F",
+    "000000000010" = "Punish when steal in Game F",
+    "001000001000" = "Punish when steal in Games B and E",
+    "101000001010" = "Punish when steal in Games A, B, E, and F",
+    "100000000000" = "Punish when steal in Game A",
+    "000000100000" = "Punish when steal in Game D",
+    "001000001010" = "Punish when steal in Games B, E, and F",
     "101000101000" = "\\textit{Deterrent strategy (exact)}",
-    "101010101010" = "Punish when take in all games",
+    "101010101010" = "Punish when steal in all games",
     "101000101010" = "\\textit{Norm-enforcing strategy (exact)}",
-    "001000000000" = "Punish when take in Game B",
-    "001010101000" = "Punish when take in Games B, C, D, and E",
-    "100000001000" = "Punish when take in Games A and E",
-    "101000001000" = "Punish when take in Games A, B, and E",
+    "001000000000" = "Punish when steal in Game B",
+    "001010101000" = "Punish when steal in Games B, C, D, and E",
+    "100000001000" = "Punish when steal in Games A and E",
+    "101000001000" = "Punish when steal in Games A, B, and E",
     "101010101000" = "\\textit{Retributive strategy (exact)}",
     "111111111111" = "Always punish",
-    "000000101000" = "Punish when take in Games D and E",
-    "000000101010" = "Punish when take in Games D, E, and F",
-    "101010001010" = "Punish when take in all games except Game D",
-    "001000101000" = "Punish when take in Games B, D, and E",
-    "001000101010" = "Punish when take in Games B, D, E, and F",
-    "101000000000" = "Punish when take in Games A and B",
-    "101010001000" = "Punish when take in Games A, B, C, and E"
+    "000000101000" = "Punish when steal in Games D and E",
[…]

### 2023-12-13 _targets.R

@@ -53,11 +53,13 @@ list(
   tar_target(fileModel2, "stan/punishStrat_withPredictor.stan", format = "file"),
   tar_target(fileModel3, "stan/punishStrat_withTwoPredictors.stan", format = "file"),
   tar_target(fileModel4, "stan/punishStrat_withCategoricalPredictor.stan", format = "file"),
+  tar_target(fileModel5, "stan/punishStrat_noPredictor_estimateError.stan", format = "file"),
   # compile models
   tar_target(compiledModel1, stan_model(fileModel1), deployment = "worker"),
   tar_target(compiledModel2, stan_model(fileModel2), deployment = "worker"),
   tar_target(compiledModel3, stan_model(fileModel3), deployment = "worker"),
   tar_target(compiledModel4, stan_model(fileModel4), deployment = "worker"),
+  tar_target(compiledModel5, stan_model(fileModel5), deployment = "worker"),
   
   #### Simulation ####
   
@@ -143,7 +145,6 @@ list(
   tar_target(plot1.2, plotModel1(post1.2, file = "figures/modelResults/withExclusions/model1.pdf")),
   # trace plot
   tar_target(plotTrace1.2, plotTraceMCMC(m1.2)),
-  
   
   #### Fit models with continuous predictor ####
   
@@ -236,6 +237,15 @@ list(
     tar_target(plot4.2, plotModel4(d, post4.2, pred, xlab, 
                                    file = paste0("figures/modelResults/withExclusions/model3_", pred, ".pdf")))
   ),
+  
+  #### Fit model of initial frequencies (with implementation error estimation) ####
+  
+  # model fitting
+  tar_target(m5, fitModel5(dExc, compiledModel5), deployment = "worker"),
+  # posterior samples
+  tar_target(post5, extract(m5)),
+  # plot
+  tar_target(plot5, plotModel1(post5, file = "figures/modelResults/withExclusions/model5.pdf")),
   
   #### Summary plots of model results ####
   

### 2023-12-13 manuscript.Rmd
update word count

@@
 keywords          : "punishment; cooperation; economic games"
-wordcount         : "xxxx words"
+wordcount         : "5657 words"
 bibliography      : ["r-references.bib"]
@@
 
-\linenumbers
+\nolinenumbers
 

### 2024-01-04 manuscript.tex
update word count
update manuscript

@@
 {}{\PatchFailed}
-\keywords{punishment; cooperation; economic games\newline\indent Word count: xxxx words}
+\keywords{punishment; cooperation; economic games\newline\indent Word count: 5657 words}
 \usepackage{lineno}
@@
 
-Ethical approval was granted by the University College London Ethics Board
-(project: 3720/002). The study was performed in accordance with all the relevant
-guidelines and regulations. Informed consent was obtained from all participants
-prior to the study and participation was voluntary.
+This research has been approved by the UCL Department of Psychology Ethics
+Committee (project: 3720/002) and ratified by the University of Auckland Human
+Participants Ethics Committee. The study was performed in accordance with all
+the relevant guidelines and regulations. Informed consent was obtained from all
+participants prior to the study and participation was voluntary.
 

### 2024-01-08 appendix.Rmd
prepare to submit pre-print

@@
 \noindent \hspace*{10mm} \small Why do people punish? Evidence for a range of strategic concerns \newline
-\hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, Nichola Raihani\textsuperscript{1,2} \newline
+\hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, & Nichola Raihani\textsuperscript{1,2} \newline
 \raggedright
@@
 \noindent \footnotesize \textsuperscript{2} Department of Experimental Psychology, University College London, London, United Kingdom \newline
+\par
+\noindent{\bfseries{Contents}}
+\contentsline {subsection}{Supplementary Figures S1-S13}{2}{}
+\contentsline {subsection}{Supplementary Tables S1-S4}{15}{}
 \normalsize
@@
 the full dataset without pre-registered exclusions.* The model assumes an 
-implementation error rate of 5%. Points represent posterior medians, line ranges 
+implementation error rate of 5%. Figure \@ref(fig:plotModel1b) in the main text
+shows the same result, but from a model fitted to the reduced dataset with 
+pre-registered exclusions. Points represent posterior medians, line ranges 
 represent 50% and 95% credible intervals. AI = advantageous inequity, DI = 
@@
 personality dimensions and Social Value Orientation, fitted to the full dataset 
-without pre-registered exclusions.* Each row represents a separate model. Points 
-represent posterior medians, line ranges represent 95% credible intervals. AI = 
+without pre-registered exclusions.* Each row represents a separate model. Figure 
+\@ref(fig:plotAllPers2) in the main text shows the same results, but from models
+fitted to the reduced dataset with pre-registered exclusions. Points represent
+posterior medians, line ranges represent 95% credible intervals. AI = 
 advantageous inequity, DI = disadvantageous inequity.
@@
 aside from Social Dominance Orientation and Right Wing Authoritarianism, which 
-control for one another within the same model. Points represent posterior 
-medians, line ranges represent 95% credible intervals. AI = advantageous 
-inequity, DI = disadvantageous inequity.
+control for one another within the same model. Figure \@ref(fig:plotAllPolRel2)
+in the main text shows the same results, but from models fitted to the reduced
+dataset with pre-registered exclusions. Points represent posterior medians, line 
+ranges represent 95% credible intervals. AI = advantageous inequity, DI = 
+disadvantageous inequity.
 
@@
 represent combinations where the self-report slider matched the behavioural 
-strategy. Each strategy had an associated self-report slider except for the 
-competitive strategy. Points represent posterior medians, line ranges represent 
-95% credible intervals. AI = advantageous inequity, DI = disadvantageous 
-inequity.
+strategy. Figure \@ref(fig:plotAllSliders2) in the main text shows the same 
+results, but from models fitted to the reduced dataset with pre-registered 
+exclusions. Points represent posterior medians, line ranges represent 95% 
+credible intervals. AI = advantageous inequity, DI = disadvantageous inequity.
 
@@
 
-(ref:plotTraceCaption) *Trace plots for parameter values from the Bayesian
-latent state model fitted to data with exclusions.*
+(ref:plotTraceCaption) *MCMC trace plots for parameter values from the Bayesian
+latent state model fitted to data with exclusions.* These trace plots suggest
+that the different chains mixed well and the model converged normally.
 

### 2024-01-08 manuscript.Rmd
update manuscript
prepare to submit pre-print

@@
 
-(ref:plotGamesCaption) *Visual summary of the six economic games.* In all
+(ref:plotGamesCaption) *Visual summary of the six economic games.* In the
 games, Player 2 either steals £0.20 from Player 1 (the focal player) or does 
@@
 decisions in total. Each behavioural strategy implies a unique pattern of 
-punishment across all decisions. Green ticks reflect decisions to punish, red 
-crosses reflect decisions to not punish. In column headers, payoffs at the first 
-stage (above) and the second stage (below) are denoted as P1-P2 (or P2-P3 [P1] 
-for Game F) where participants take the role of P1 and P2 is the target of 
-punishment. AI = advantageous inequity, DI = disadvantageous inequity.*
+punishment across all decisions (see Methods for detailed explanation of 
+strategies). Green ticks reflect decisions to punish, red crosses reflect 
+decisions to not punish. In column headers, payoffs at the first stage (above) 
+and the second stage (below) are denoted as P1-P2 (or P2-P3 [P1] for Game F) 
+where participants take the role of P1 and P2 is the target of punishment. AI = 
+advantageous inequity, DI = disadvantageous inequity.*
 
@@
 
-Ethical approval was granted by the University College London Ethics Board 
-(project: 3720/002). The study was performed in accordance with all the relevant 
-guidelines and regulations. Informed consent was obtained from all participants 
-prior to the study and participation was voluntary.
+This research has been approved by the UCL Department of Psychology Ethics 
+Committee (project: 3720/002) and ratified by the University of Auckland Human 
+Participants Ethics Committee. The study was performed in accordance with all 
+the relevant guidelines and regulations. Informed consent was obtained from all 
+participants prior to the study and participation was voluntary.
 

### 2024-01-08 manuscript.tex
prepare to submit pre-print

@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{images/games} \caption{\emph{Visual summary of the six economic games.} In all
+\includegraphics[width=1\linewidth]{images/games} \caption{\emph{Visual summary of the six economic games.} In the
 games, Player 2 either steals £0.20 from Player 1 (the focal player) or does
@@
 decisions in total. Each behavioural strategy implies a unique pattern of
-punishment across all decisions. Green ticks reflect decisions to punish, red
-crosses reflect decisions to not punish. In column headers, payoffs at the first
-stage (above) and the second stage (below) are denoted as P1-P2 (or P2-P3 {[}P1{]}
-for Game F) where participants take the role of P1 and P2 is the target of
-punishment. AI = advantageous inequity, DI = disadvantageous inequity.}}\\
+punishment across all decisions (see Methods for detailed explanation of
+strategies). Green ticks reflect decisions to punish, red crosses reflect
+decisions to not punish. In column headers, payoffs at the first stage (above)
+and the second stage (below) are denoted as P1-P2 (or P2-P3 {[}P1{]} for Game F)
+where participants take the role of P1 and P2 is the target of punishment. AI =
+advantageous inequity, DI = disadvantageous inequity.}}\\
 \toprule
@@
 C.
-\href{https://doi.org/10.1098/rspb.2012.0615}{Outsourcing punishment to god: Beliefs in divine control reduce earthly punishment}.
+\href{https://doi.org/10.1098/rspb.2012.0615}{Outsourcing punishment to {God}: Beliefs in divine control reduce earthly punishment}.
 \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{279}, 3272--3281 (2012).}
@@
 \noindent \hspace*{10mm} \small Why do people punish? Evidence for a range of strategic concerns \newline
-\hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, Nichola Raihani\textsuperscript{1,2} \newline
+\hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, \& Nichola Raihani\textsuperscript{1,2} \newline
 
@@
 \noindent \footnotesize \textsuperscript{2} Department of Experimental Psychology, University College London, London, United Kingdom \newline
+
+\par
+
+\noindent{\bfseries{Contents}}
+\contentsline {subsection}{Supplementary Figures S1-S13}{2}{}
+\contentsline {subsection}{Supplementary Tables S1-S4}{15}{}
 \normalsize
@@
 the full dataset without pre-registered exclusions.} The model assumes an
-implementation error rate of 5\%. Points represent posterior medians, line ranges
+implementation error rate of 5\%. Figure \ref{fig:plotModel1b} in the main text
+shows the same result, but from a model fitted to the reduced dataset with
+pre-registered exclusions. Points represent posterior medians, line ranges
 represent 50\% and 95\% credible intervals. AI = advantageous inequity, DI =
@@
 personality dimensions and Social Value Orientation, fitted to the full dataset
-without pre-registered exclusions.} Each row represents a separate model. Points
-represent posterior medians, line ranges represent 95\% credible intervals. AI =
+without pre-registered exclusions.} Each row represents a separate model. Figure
+\ref{fig:plotAllPers2} in the main text shows the same results, but from models
+fitted to the reduced dataset with pre-registered exclusions. Points represent
+posterior medians, line ranges represent 95\% credible intervals. AI =
 advantageous inequity, DI = disadvantageous inequity.}
@@
 aside from Social Dominance Orientation and Right Wing Authoritarianism, which
-control for one another within the same model. Points represent posterior
-medians, line ranges represent 95\% credible intervals. AI = advantageous
-inequity, DI = disadvantageous inequity.}
+control for one another within the same model. Figure \ref{fig:plotAllPolRel2}
+in the main text shows the same results, but from models fitted to the reduced
+dataset with pre-registered exclusions. Points represent posterior medians, line
+ranges represent 95\% credible intervals. AI = advantageous inequity, DI =
+disadvantageous inequity.}
 \end{figure}
@@
 represent combinations where the self-report slider matched the behavioural
-strategy. Each strategy had an associated self-report slider except for the
-competitive strategy. Points represent posterior medians, line ranges represent
-95\% credible intervals. AI = advantageous inequity, DI = disadvantageous
-inequity.}
+strategy. Figure \ref{fig:plotAllSliders2} in the main text shows the same
+results, but from models fitted to the reduced dataset with pre-registered
+exclusions. Points represent posterior medians, line ranges represent 95\%
+credible intervals. AI = advantageous inequity, DI = disadvantageous inequity.}
 \end{figure}
@@
 \includegraphics{manuscript_files/figure-latex/plotTrace-1.pdf}
-\caption{\label{fig:plotTrace}\emph{Trace plots for parameter values from the Bayesian
-latent state model fitted to data with exclusions.}}
+\caption{\label{fig:plotTrace}\emph{MCMC trace plots for parameter values from the Bayesian
+latent state model fitted to data with exclusions.} These trace plots suggest
+that the different chains mixed well and the model converged normally.}
 \end{figure}

### 2024-05-22 manuscript.Rmd
submit manuscript

@@
 
-keywords          : "punishment; cooperation; economic games"
-wordcount         : "5657 words"
 bibliography      : ["r-references.bib"]
@@
 documentclass     : "apa6"
-classoption       : "man"
+classoption       : "man, donotrepeattitle"
 output            : papaja::apa6_pdf
@@
   - \note{\raggedright * Correspondence concerning this article should be addressed to Scott Claessens, Level 2, Building 302, 23 Symonds Street, Auckland, New Zealand.
-\text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This working paper has not yet been peer-reviewed.}
+\text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This manuscript has not yet been peer-reviewed.
+\par This manuscript was published as a pre-print on PsyArXiv under a CC BY 4.0 \text{license:} \url{https://doi.org/10.31234/osf.io/ys6rm} \par \text{Classification:} Social Sciences; Psychological and Cognitive Sciences \par \text{Keywords:} punishment; cooperation; economic games}
 appendix          : "appendix.Rmd"
-csl               : nature.csl
+csl               : pnas.csl
 ---
@@
 
-\nolinenumbers
+# Significance Statement
+
+Research has shown that humans are willing to punish others at a personal cost,
+but the motives for this behaviour remain unclear. Using a suite of economic 
+games, we show that punitive behaviour is consistent with a variety of strategic
+concerns. While most people choose not to punish in our games, people who do 
+punish tend to be motivated by egalitarianism rather than deterrence or 
+norm enforcement. We also find that different punishment strategies are linked 
+with different character traits, including personality, prosocial preferences, 
+and political and religious views. This research highlights the multifaceted
+nature of human punishment.
+
+\newpage
+\linenumbers
 
@@
 mechanism thought to sustain this level of cooperation is costly punishment, 
-whereby individuals harm others at a personal cost[@CluttonBrock1995], 
-ostensibly encouraging cooperative behaviour from the target (or bystanders 
-[@DosSantos2011;@DosSantos2013;@Raihani2015]) in the future. Punishment offers a 
+whereby individuals harm others at a personal cost [@CluttonBrock1995], 
+ostensibly encouraging cooperative behaviour from the target, or bystanders 
+[@DosSantos2011;@DosSantos2013;@Raihani2015], in the future. Punishment offers a 
 route to maintaining or increasing cooperation by changing the payoff structure 
 of social interactions such that it no longer pays to cheat or exploit social 
-partners[@CluttonBrock1995;@Ostrom1990]. Yet, despite its theoretical 
+partners [@CluttonBrock1995;@Ostrom1990]. Yet, despite its theoretical 
 importance, the question of why people choose to punish others is still hotly 
-contested[@Raihani2019]. In this study, we use a battery of economic games to 
+contested [@Raihani2019]. In this study, we use a battery of economic games to 
 disentangle the different motives underpinning punishment and explore how these 
@@
 In humans, many studies of punishment have been carried out in laboratory 
-settings using economic games[@Balliet2011;@Balliet2013;@Chaudhuri2011;
+settings using economic games [@Balliet2011;@Balliet2013;@Chaudhuri2011;
 @Dreber2008;@Fehr2000;@Fehr2002;@Henrich2006;@Nikiforakis2008a;@Raihani2012a;
@@
 'fine' on their co-players. Several lines of experimental evidence indicate that 
-people use this punishment option[@Henrich2006], that they enjoy 
-punishing[@DeQuervain2004], and that they frequently, though not 
-always[@Herrmann2008], punish cheating or exploitative 
-co-players[@Fehr2000;@Fehr2002].
+people use this punishment option [@Henrich2006], that they enjoy 
+punishing [@DeQuervain2004], and that they frequently, though not 
+always [@Herrmann2008], punish cheating or exploitative 
+co-players [@Fehr2000;@Fehr2002].
 
@@
 typically observed in the Ultimatum Game (where punishment is possible) compared 
-to the structurally-similar Dictator Game (where it is not)[@Camerer2003]. This 
+to the structurally-similar Dictator Game (where it is not) [@Camerer2003]. This 
 typical cooperation-enhancing effect of punishment has also been observed across 
-societies[@Balliet2013], leading some to suggest that costly punishment has 
+societies [@Balliet2013], leading some to suggest that costly punishment has 
 played a key role in the cultural evolution of cooperation in humans
@@
 achieve other ends. Some have argued that punishment is primarily used to shape 
-future behaviour, either to deter personal harm[@Delton2017;@Dreber2008;
+future behaviour, either to deter personal harm [@Delton2017;@Dreber2008;
 @DosSantos2013] or to uphold normative standards of cooperative behaviour 
@@
 the _enactment_ of this punishment does not consistently deter targets from 
-cheating in the future[@Raihani2019]. This calls into question whether 
+cheating in the future [@Raihani2019]. This calls into question whether 
 punishment primarily operates as a behaviour-change tool or whether it is used 
@@
 for retribution rather than deterrence, punishing in proportion to the amount 
-of harm that was personally caused[@Carlsmith2002]. Punishment might be driven 
+of harm that was personally caused [@Carlsmith2002]. Punishment might be driven 
 by concerns about relative payoffs, such as disadvantageous inequity aversion 
-(i.e., avoiding having less than others[@Raihani2012b;@Raihani2019]) and/or 
+(i.e., avoiding having less than others) [@Raihani2012b;@Raihani2019] and/or 
 general egalitarian preferences (i.e., wanting all participants to receive the 
-same payoffs[@Dawes2007]). Such concerns about relative payoffs may be activated 
+same payoffs) [@Dawes2007]. Such concerns about relative payoffs may be activated 
 when participants earn less than cheats in economic games or when there are 
@@
 competitive purposes, seeking advantageous inequity for themselves (i.e., having 
-more than others) and/or improving their relative position[@Raihani2019].
+more than others) and/or improving their relative position [@Raihani2019].
 
@@
 people use punishment in seemingly disparate ways: punishing when no behaviour 
-change is possible, such as in one-shot games[@Bone2015;@Henrich2006;
+change is possible, such as in one-shot games [@Bone2015;@Henrich2006;
 @Raihani2012b;@Walker2004], on the very last round of repeated games
 [@Barclay2016], or in games where the target never learns about the 
-punishment[@Crockett2014]; punishing those who did not cheat or who 
-over-contributed to collective action (antisocial punishment
-[@Herrmann2008;@Sylwester2013]); punishing in scenarios where they were not 
-personally harmed (third-party punishment[@Fehr2004]); and punishing in 
+punishment [@Crockett2014]; punishing those who did not cheat or who 
+over-contributed to collective action (antisocial punishment) 
+[@Herrmann2008;@Sylwester2013]; punishing in scenarios where they were not 
+personally harmed (third-party punishment) [@Fehr2004]; and punishing in 
 scenarios where disparities in payoffs did not arise from participants' 
-actions[@Dawes2007;@Johnson2009;@Fowler2005].
+actions [@Dawes2007;@Johnson2009;@Fowler2005].
 
@@
 of as a flexible behavioural tool that serves a variety of functions that are 
-not mutually exclusive[@Raihani2019]. Due to its multipurpose nature, we should 
+not mutually exclusive [@Raihani2019]. Due to its multipurpose nature, we should 
 therefore expect variation in punishment strategies in the population, much like 
-the observed variation in social learning strategies[@Molleman2014]. Some 
+the observed variation in social learning strategies [@Molleman2014]. Some 
 individuals may use punishment as a behaviour shaping tool, for example, while 
@@
 adherence to a particular strategy? Previous work has reported that personality 
-is related to cooperative behaviour[@Thielmann2020] and demographics, political 
-ideology, and religiosity are related to punitive behaviour[@Hofmann2018], but 
+is related to cooperative behaviour [@Thielmann2020] and demographics, political 
+ideology, and religiosity are related to punitive behaviour [@Hofmann2018], but 
 no research has related these variables to specific punishment strategies. 
@@
 reputational functions of punishment in this table, such as signalling 
-trustworthiness[@Barclay2006;@Batistoni2022;@Jordan2016;@Jordan2017;@Raihani2015], 
+trustworthiness [@Barclay2006;@Batistoni2022;@Jordan2016;@Jordan2017;@Raihani2015], 
 because our focus is on punishment strategies in anonymous economic games without 
-reputational incentives (but see ref[@Jordan2020]).
+reputational incentives [but see [@Jordan2020]].
 
-Building on previous designs[@Bone2015;@Deutchman2021;@Marczyk2017;
+Building on previous designs [@Bone2015;@Deutchman2021;@Marczyk2017;
 @Raihani2012b], we employ a suite of one-shot economic games where individuals 
@@
 we have shown that punishment does not serve just one function, but instead is 
-a flexible tool that can be and is used for different purposes[@Raihani2019]. 
+a flexible tool that can be and is used for different purposes [@Raihani2019]. 
 While some use punishment to enforce norms of cooperation, others use it to 
@@
 contrary to the view that people are often unable to articulate the reasons for 
-their punitive behaviour[@Carlsmith2002], people seem to have some degree of 
+their punitive behaviour [@Carlsmith2002], people seem to have some degree of 
 insight into the strategy they are using. Despite small differences, these 
@@
 provide insight into the motives that may have driven higher levels of peer 
-punishment among conservatives in previous work[@Hofmann2018].
+punishment among conservatives in previous work [@Hofmann2018].
 
@@
 behaviour was intentional or not (Game C), an approach which has been used in 
-previous vignette studies to identify behaviour shaping motives[@Carlsmith2002].
+previous vignette studies to identify behaviour shaping motives [@Carlsmith2002].
 The lower prevalence of behaviour shaping strategies in our study is consistent 
@@
 behaviour shaping is impossible, such as when the target will never find out 
-that they have been punished[@Crockett2014] or on the last round of repeated 
-interactions[@Barclay2016]. We found that demographic, personality, political, 
+that they have been punished [@Crockett2014] or on the last round of repeated 
+interactions [@Barclay2016]. We found that demographic, personality, political, 
 and religious variables tended to be unrelated to behaviour shaping strategies, 
@@
 struggle to accurately report the deterrent motivations for their punitive 
-behaviour[@Carlsmith2002].
+behaviour [@Carlsmith2002].
 
@@
 games suggests that the antisocial punishment that has been observed in other 
-studies[@Herrmann2008;@Pleasant2018] is not aimed at harming cooperators 
-specifically, as has been previously suggested[@Herrmann2008]. Instead, 
+studies [@Herrmann2008;@Pleasant2018] is not aimed at harming cooperators 
+specifically, as has been previously suggested [@Herrmann2008]. Instead, 
 antisocial punishment is more likely to be motivated by improving one's 
@@
 disappears when relative payoffs cannot be changed (e.g., when punishment is 
-only available with a 1:1 fee-fine ratio[@Sylwester2013]).
+only available with a 1:1 fee-fine ratio) [@Sylwester2013].
 
@@
 game studies. In these studies, a common assumption is that participants punish 
-to change the behaviour of cheats[@Fehr2000;@Fehr2002]. But in reality, people 
+to change the behaviour of cheats [@Fehr2000;@Fehr2002]. But in reality, people 
 may be choosing the punishment option to achieve a variety of different goals. 
@@
 interpret punishment as serving a competitive motive, it may elicit retaliation
-rather than encourage cooperation[@Dreber2008;@Nikiforakis2008b;@Raihani2019].
+rather than encourage cooperation [@Dreber2008;@Nikiforakis2008b;@Raihani2019].
 As punishers’ motives must be inferred (and such inferences likely depend on 
@@
 might help to explain the mixed findings in the field as to whether punishment 
-actually motivates cheating targets to cooperate in the future[@Raihani2019].
+actually motivates cheating targets to cooperate in the future [@Raihani2019].
 
@@
 in both WEIRD (Western, educated, industrialised, rich, and democratic) samples
-[@Balafoutas2014;@Balafoutas2016] and in small-scale societies[@Baumard2010]. 
+[@Balafoutas2014;@Balafoutas2016] and in small-scale societies [@Baumard2010]. 
 One reason that people may be averse to peer punishment is that, because it is 
@@
 people might therefore refrain from punishing others to avoid reputational 
-damage[@Raihani2015]. People frequently avoid taking actions that could harm 
+damage [@Raihani2015]. People frequently avoid taking actions that could harm 
 their reputation, even when they don’t know if reputation is at stake (as in the 
 one-shot anonymous settings used here). Another reason that people may be averse
-to peer punishment is that it can trigger retaliation[@Raihani2019]. This may be
+to peer punishment is that it can trigger retaliation [@Raihani2019]. This may be
 especially likely in situations that lack clear institutional norms to 
@@
 [@Fitouchi2023;@Singh2022]. Future research should uncover whether people are 
-more willing to punish in these conventionalised contexts (e.g., see
-[@Molleman2019]).
+more willing to punish in these conventionalised contexts [e.g., see
+[@Molleman2019]].
 
@@
 unable to study other potential reputational strategies underlying punishment, 
-such as signalling trustworthiness[@Jordan2020]. Moreover, stealing may be 
+such as signalling trustworthiness [@Jordan2020]. Moreover, stealing may be 
 evaluated differently to other forms of cheating, such as not contributing to 
@@
 shaping, egalitarianism, and competition. Much like the observed variation in 
-human social learning strategies[@Molleman2014], humans thus also exhibit 
+human social learning strategies [@Molleman2014], humans thus also exhibit 
 variation in their punishment strategies. These individual differences map onto 
@@
   (e.g., age, gender, student status).
-- *Personality*. We used the Mini-IPIP scale[@Sibley2011] to measure the Big 6 
+- *Personality*. We used the Mini-IPIP scale [@Sibley2011] to measure the Big 6 
   personality dimensions of agreeableness ($\alpha$ = 
@@
 - *Social Value Orientation*. We used the Social Value Orientation Slider 
-  Measure to measure other-regarding preferences[@Murphy2011]. Across fifteen 
+  Measure to measure other-regarding preferences [@Murphy2011]. Across fifteen 
   items, participants made decisions on how to allocate different amounts of 
@@
   measure of their other-regarding preference, following the steps outlined in 
-  ref[@Murphy2011].
+  ref [@Murphy2011].
 - *Political ideology*. We included several measures of political ideology, 
-  including left-right conservatism, Social Dominance Orientation[@Ho2015] 
+  including left-right conservatism, Social Dominance Orientation [@Ho2015] 
   ($\alpha$ = `r printnum(tar_read(alphaSDO)$alpha)`; eight items), and Right 
-  Wing Authoritarianism[@Bizumic2018] ($\alpha$ = 
+  Wing Authoritarianism [@Bizumic2018] ($\alpha$ = 
   `r printnum(tar_read(alphaRWA)$alpha)`; six items). We also probed 
@@
   themselves and whether they believe that God or another spiritual non-human 
-  entity controls the events in the world[@Laurin2012].
+  entity controls the events in the world [@Laurin2012].
 
@@
 We pre-registered that we would use a Bayesian latent state model to infer 
-unobserved punishment strategies from the observed data (for a similar version 
-of this model, see ref[@McElreath2020]). In this model, participants $i$ in 
+unobserved punishment strategies from the observed data [for a similar version 
+of this model, see ref [@McElreath2020]]. In this model, participants $i$ in 
 countries $c$ make binary punishment decisions across twelve decisions $j$. We 
@@
 We estimated the posterior distributions of these models using Hamiltonian Monte 
-Carlo as implemented in Stan version 2.26.1[@Stan2020]. We ran each model for 
+Carlo as implemented in Stan version 2.26.1 [@Stan2020]. We ran each model for 
 2000 samples, with 1000 warmup samples. R-hat values and effective sample sizes 
@@
 https://github.com/ScottClaessens/punishStrategies. All analyses were conducted 
-in R version 4.2.1[@RCoreTeam]. Visualisations were created with the *ggplot2*
-[@Wickham2016] and *cowplot*[@Wilke2020] R packages. We used the *targets*
+in R version 4.2.1 [@RCoreTeam]. Visualisations were created with the *ggplot2*
+[@Wickham2016] and *cowplot* [@Wilke2020] R packages. We used the *targets*
 [@Landau2021] R package to create a reproducible data analysis pipeline and the 
-*papaja*[@Aust2022] R package to reproducibly generate the manuscript.
+*papaja* [@Aust2022] R package to reproducibly generate the manuscript.
 

### 2024-05-22 manuscript.tex
submit manuscript

@@
 \documentclass[
-  man,floatsintext]{apa6}
+  man, donotrepeattitle,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
 {}{\PatchFailed}
-\keywords{punishment; cooperation; economic games\newline\indent Word count: 5657 words}
 \usepackage{lineno}
@@
 \note{\raggedright * Correspondence concerning this article should be addressed to Scott Claessens, Level 2, Building 302, 23 Symonds Street, Auckland, New Zealand.
-\text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This working paper has not yet been peer-reviewed.}
+\text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This manuscript has not yet been peer-reviewed.
+\par This manuscript was published as a pre-print on PsyArXiv under a CC BY 4.0 \text{license:} \url{https://doi.org/10.31234/osf.io/ys6rm} \par \text{Classification:} Social Sciences; Psychological and Cognitive Sciences \par \text{Keywords:} punishment; cooperation; economic games}
 \ifLuaTeX
@@
   pdflang={en-EN},
-  pdfkeywords={punishment; cooperation; economic games},
   hidelinks,
@@
 
-\nolinenumbers
+\hypertarget{significance-statement}{%
+\section{Significance Statement}\label{significance-statement}}
+
+Research has shown that humans are willing to punish others at a personal cost,
+but the motives for this behaviour remain unclear. Using a suite of economic
+games, we show that punitive behaviour is consistent with a variety of strategic
+concerns. While most people choose not to punish in our games, people who do
+punish tend to be motivated by egalitarianism rather than deterrence or
+norm enforcement. We also find that different punishment strategies are linked
+with different character traits, including personality, prosocial preferences,
+and political and religious views. This research highlights the multifaceted
+nature of human punishment.
+
+\newpage
+\linenumbers
 
@@
 mechanism thought to sustain this level of cooperation is costly punishment,
-whereby individuals harm others at a personal cost\textsuperscript{1},
-ostensibly encouraging cooperative behaviour from the target (or bystanders\textsuperscript{2--4}) in the future. Punishment offers a
+whereby individuals harm others at a personal cost (1),
+ostensibly encouraging cooperative behaviour from the target, or bystanders
+(2--4), in the future. Punishment offers a
 route to maintaining or increasing cooperation by changing the payoff structure
 of social interactions such that it no longer pays to cheat or exploit social
-partners\textsuperscript{1,5}. Yet, despite its theoretical
+partners (1, 5). Yet, despite its theoretical
 importance, the question of why people choose to punish others is still hotly
-contested\textsuperscript{6}. In this study, we use a battery of economic games to
+contested (6). In this study, we use a battery of economic games to
 disentangle the different motives underpinning punishment and explore how these
@@
 In humans, many studies of punishment have been carried out in laboratory
-settings using economic games\textsuperscript{6--15}. In these games, participants are usually given a sum of money
+settings using economic games (6--15). In these games, participants are usually given a sum of money
 that they can use to invest in collective action or to help others.
@@
 `fine' on their co-players. Several lines of experimental evidence indicate that
-people use this punishment option\textsuperscript{13}, that they enjoy
-punishing\textsuperscript{16}, and that they frequently, though not
-always\textsuperscript{17}, punish cheating or exploitative
-co-players\textsuperscript{11,12}.
+people use this punishment option (13), that they enjoy
+punishing (16), and that they frequently, though not
+always (17), punish cheating or exploitative
+co-players (11, 12).
 
@@
 plays an important role in promoting human cooperation. People tend to cooperate
-more in games where punishment is possible compared to those where it is not\textsuperscript{6--8}. The effect that the threat of
+more in games where punishment is possible compared to those where it is not
+(6--8). The effect that the threat of
 punishment has on cooperation is also evident in the higher contributions
 typically observed in the Ultimatum Game (where punishment is possible) compared
-to the structurally-similar Dictator Game (where it is not)\textsuperscript{18}. This
+to the structurally-similar Dictator Game (where it is not) (18). This
 typical cooperation-enhancing effect of punishment has also been observed across
-societies\textsuperscript{8}, leading some to suggest that costly punishment has
-played a key role in the cultural evolution of cooperation in humans\textsuperscript{19--22}.
+societies (8), leading some to suggest that costly punishment has
+played a key role in the cultural evolution of cooperation in humans
+(19--22).
 
@@
 achieve other ends. Some have argued that punishment is primarily used to shape
-future behaviour, either to deter personal harm\textsuperscript{3,10,23} or to uphold normative standards of cooperative behaviour\textsuperscript{20,21,24--27}.
+future behaviour, either to deter personal harm (3, 10, 23) or to uphold normative standards of cooperative behaviour
+(20, 21, 24--27).
 But while the \emph{threat} of punishment can have a cooperation-enhancing effect,
 the \emph{enactment} of this punishment does not consistently deter targets from
-cheating in the future\textsuperscript{6}. This calls into question whether
+cheating in the future (6). This calls into question whether
 punishment primarily operates as a behaviour-change tool or whether it is used
@@
 for retribution rather than deterrence, punishing in proportion to the amount
-of harm that was personally caused\textsuperscript{28}. Punishment might be driven
+of harm that was personally caused (28). Punishment might be driven
 by concerns about relative payoffs, such as disadvantageous inequity aversion
-(i.e., avoiding having less than others\textsuperscript{6,29}) and/or
+(i.e., avoiding having less than others) (6, 29) and/or
 general egalitarian preferences (i.e., wanting all participants to receive the
-same payoffs\textsuperscript{30}). Such concerns about relative payoffs may be activated
+same payoffs) (30). Such concerns about relative payoffs may be activated
 when participants earn less than cheats in economic games or when there are
@@
 competitive purposes, seeking advantageous inequity for themselves (i.e., having
-more than others) and/or improving their relative position\textsuperscript{6}.
+more than others) and/or improving their relative position (6).
 
@@
 people use punishment in seemingly disparate ways: punishing when no behaviour
-change is possible, such as in one-shot games\textsuperscript{13,29,31,32}, on the very last round of repeated games\textsuperscript{33}, or in games where the target never learns about the
-punishment\textsuperscript{34}; punishing those who did not cheat or who
-over-contributed to collective action (antisocial punishment\textsuperscript{17,35}); punishing in scenarios where they were not
-personally harmed (third-party punishment\textsuperscript{36}); and punishing in
+change is possible, such as in one-shot games (13, 29, 31, 32), on the very last round of repeated games
+(33), or in games where the target never learns about the
+punishment (34); punishing those who did not cheat or who
+over-contributed to collective action (antisocial punishment)
+(17, 35); punishing in scenarios where they were not
+personally harmed (third-party punishment) (36); and punishing in
 scenarios where disparities in payoffs did not arise from participants'
-actions\textsuperscript{30,37,38}.
+actions (30, 37, 38).
 
@@
 of as a flexible behavioural tool that serves a variety of functions that are
-not mutually exclusive\textsuperscript{6}. Due to its multipurpose nature, we should
+not mutually exclusive (6). Due to its multipurpose nature, we should
 therefore expect variation in punishment strategies in the population, much like
-the observed variation in social learning strategies\textsuperscript{39}. Some
+the observed variation in social learning strategies (39). Some
 individuals may use punishment as a behaviour shaping tool, for example, while
@@
 adherence to a particular strategy? Previous work has reported that personality
-is related to cooperative behaviour\textsuperscript{40} and demographics, political
-ideology, and religiosity are related to punitive behaviour\textsuperscript{41}, but
+is related to cooperative behaviour (40) and demographics, political
+ideology, and religiosity are related to punitive behaviour (41), but
 no research has related these variables to specific punishment strategies.
@@
 has argued that people are often unaware of the underlying function of their
-punitive behaviour, yet they feel compelled to enact it anyway\textsuperscript{28,42}.
+punitive behaviour, yet they feel compelled to enact it anyway
+(28, 42).
 
@@
 reputational functions of punishment in this table, such as signalling
-trustworthiness\textsuperscript{4,43--46},
+trustworthiness (4, 43--46),
 because our focus is on punishment strategies in anonymous economic games without
-reputational incentives (but see ref\textsuperscript{47}).
-
-Building on previous designs\textsuperscript{29,31,48,49}, we employ a suite of one-shot economic games where individuals
+reputational incentives {[}but see (47){]}.
+
+Building on previous designs (29, 31, 48, 49), we employ a suite of one-shot economic games where individuals
 are given the opportunity to punish targets at a personal cost (Figure
@@
 we have shown that punishment does not serve just one function, but instead is
-a flexible tool that can be and is used for different purposes\textsuperscript{6}.
+a flexible tool that can be and is used for different purposes (6).
 While some use punishment to enforce norms of cooperation, others use it to
@@
 contrary to the view that people are often unable to articulate the reasons for
-their punitive behaviour\textsuperscript{28}, people seem to have some degree of
+their punitive behaviour (28), people seem to have some degree of
 insight into the strategy they are using. Despite small differences, these
@@
 line with previous studies which have highlighted inequity aversion as an
-important motivation for punishment in economic games\textsuperscript{29--31,37,48,50}.
+important motivation for punishment in economic games
+(29--31, 37, 48, 50).
 Personality, social preferences, and political and religious views were related
@@
 provide insight into the motives that may have driven higher levels of peer
-punishment among conservatives in previous work\textsuperscript{41}.
+punishment among conservatives in previous work (41).
 
@@
 behaviour was intentional or not (Game C), an approach which has been used in
-previous vignette studies to identify behaviour shaping motives\textsuperscript{28}.
+previous vignette studies to identify behaviour shaping motives (28).
 The lower prevalence of behaviour shaping strategies in our study is consistent
@@
 behaviour shaping is impossible, such as when the target will never find out
-that they have been punished\textsuperscript{34} or on the last round of repeated
-interactions\textsuperscript{33}. We found that demographic, personality, political,
+that they have been punished (34) or on the last round of repeated
+interactions (33). We found that demographic, personality, political,
 and religious variables tended to be unrelated to behaviour shaping strategies,
@@
 struggle to accurately report the deterrent motivations for their punitive
-behaviour\textsuperscript{28}.
+behaviour (28).
 
@@
 games suggests that the antisocial punishment that has been observed in other
-studies\textsuperscript{17,51} is not aimed at harming cooperators
-specifically, as has been previously suggested\textsuperscript{17}. Instead,
+studies (17, 51) is not aimed at harming cooperators
+specifically, as has been previously suggested (17). Instead,
 antisocial punishment is more likely to be motivated by improving one's
@@
 disappears when relative payoffs cannot be changed (e.g., when punishment is
-only available with a 1:1 fee-fine ratio\textsuperscript{35}).
+only available with a 1:1 fee-fine ratio) (35).
 
@@
 game studies. In these studies, a common assumption is that participants punish
-to change the behaviour of cheats\textsuperscript{11,12}. But in reality, people
+to change the behaviour of cheats (11, 12). But in reality, people
 may be choosing the punishment option to achieve a variety of different goals.
@@
 interpret punishment as serving a competitive motive, it may elicit retaliation
-rather than encourage cooperation\textsuperscript{6,10,52}.
+rather than encourage cooperation (6, 10, 52).
 As punishers' motives must be inferred (and such inferences likely depend on
@@
 might help to explain the mixed findings in the field as to whether punishment
-actually motivates cheating targets to cooperate in the future\textsuperscript{6}.
+actually motivates cheating targets to cooperate in the future (6).
 
@@
 more general aversion to peer punishment, an aversion that has been highlighted
-in both WEIRD (Western, educated, industrialised, rich, and democratic) samples\textsuperscript{53,54} and in small-scale societies\textsuperscript{55}.
+in both WEIRD (Western, educated, industrialised, rich, and democratic) samples
+(53, 54) and in small-scale societies (55).
 One reason that people may be averse to peer punishment is that, because it is
@@
 people might therefore refrain from punishing others to avoid reputational
-damage\textsuperscript{4}. People frequently avoid taking actions that could harm
+damage (4). People frequently avoid taking actions that could harm
 their reputation, even when they don't know if reputation is at stake (as in the
 one-shot anonymous settings used here). Another reason that people may be averse
-to peer punishment is that it can trigger retaliation\textsuperscript{6}. This may be
+to peer punishment is that it can trigger retaliation (6). This may be
 especially likely in situations that lack clear institutional norms to
@@
 punishment in small-scale societies often functions to compensate victims while
-limiting the potential for feuds and cycles of retaliation\textsuperscript{56,57}. Future research should uncover whether people are
-more willing to punish in these conventionalised contexts (e.g., see\textsuperscript{58}).
+limiting the potential for feuds and cycles of retaliation
+(56, 57). Future research should uncover whether people are
+more willing to punish in these conventionalised contexts {[}e.g., see
+(58){]}.
 
@@
 unable to study other potential reputational strategies underlying punishment,
-such as signalling trustworthiness\textsuperscript{47}. Moreover, stealing may be
+such as signalling trustworthiness (47). Moreover, stealing may be
 evaluated differently to other forms of cheating, such as not contributing to
@@
 shaping, egalitarianism, and competition. Much like the observed variation in
-human social learning strategies\textsuperscript{39}, humans thus also exhibit
+human social learning strategies (39), humans thus also exhibit
 variation in their punishment strategies. These individual differences map onto
@@
   \emph{Demographics}. In the survey, we collected information on participants'
-  education level and self-reported socio-economic status (MacArthur ladder\textsuperscript{59}). We also collected additional demographic data from Prolific
+  education level and self-reported socio-economic status (MacArthur ladder
+  (59)). We also collected additional demographic data from Prolific
   (e.g., age, gender, student status).
 \item
-  \emph{Personality}. We used the Mini-IPIP scale\textsuperscript{60} to measure the Big 6
+  \emph{Personality}. We used the Mini-IPIP scale (60) to measure the Big 6
   personality dimensions of agreeableness (\(\alpha\) =
@@
   \emph{Social Value Orientation}. We used the Social Value Orientation Slider
-  Measure to measure other-regarding preferences\textsuperscript{61}. Across fifteen
+  Measure to measure other-regarding preferences (61). Across fifteen
   items, participants made decisions on how to allocate different amounts of
@@
   measure of their other-regarding preference, following the steps outlined in
-  ref\textsuperscript{61}.
+  ref (61).
 \item
   \emph{Political ideology}. We included several measures of political ideology,
-  including left-right conservatism, Social Dominance Orientation\textsuperscript{62}
+  including left-right conservatism, Social Dominance Orientation (62)
   (\(\alpha\) = 0.91; eight items), and Right
-  Wing Authoritarianism\textsuperscript{63} (\(\alpha\) =
+  Wing Authoritarianism (63) (\(\alpha\) =
   0.82; six items). We also probed
@@
   themselves and whether they believe that God or another spiritual non-human
-  entity controls the events in the world\textsuperscript{64}.
+  entity controls the events in the world (64).
 \end{itemize}
@@
 We pre-registered that we would use a Bayesian latent state model to infer
-unobserved punishment strategies from the observed data (for a similar version
-of this model, see ref\textsuperscript{65}). In this model, participants \(i\) in
+unobserved punishment strategies from the observed data {[}for a similar version
+of this model, see ref (65){]}. In this model, participants \(i\) in
 countries \(c\) make binary punishment decisions across twelve decisions \(j\). We
@@
 We estimated the posterior distributions of these models using Hamiltonian Monte
-Carlo as implemented in Stan version 2.26.1\textsuperscript{66}. We ran each model for
+Carlo as implemented in Stan version 2.26.1 (66). We ran each model for
 2000 samples, with 1000 warmup samples. R-hat values and effective sample sizes
@@
 \url{https://github.com/ScottClaessens/punishStrategies}. All analyses were conducted
-in R version 4.2.1\textsuperscript{67}.
-Visualisations were created with the \emph{ggplot2}\textsuperscript{68} and \emph{cowplot}\textsuperscript{69} R packages.
-We used the \emph{targets}\textsuperscript{70} R package to create a reproducible data analysis pipeline and the
-\emph{papaja}\textsuperscript{71} R package to reproducibly generate the manuscript.
+in R version 4.2.1 (67). Visualisations were created with the \emph{ggplot2}
+(68) and \emph{cowplot} (69) R packages. We used the \emph{targets}
+(70) R package to create a reproducible data analysis pipeline and the
+\emph{papaja} (71) R package to reproducibly generate the manuscript.
 
@@
 \CSLLeftMargin{1. }%
-\CSLRightInline{Clutton-Brock, T. H. \& Parker, G. A. \href{https://doi.org/10.1038/373209a0}{Punishment in animal societies}. \emph{Nature} \textbf{373}, 209--216 (1995).}
+\CSLRightInline{T. H. Clutton-Brock, G. A. Parker, \href{https://doi.org/10.1038/373209a0}{Punishment in animal societies}. \emph{Nature} \textbf{373}, 209--216 (1995).}
 
@@
 \CSLLeftMargin{2. }%
-\CSLRightInline{dos Santos, M., Rankin, D.
+\CSLRightInline{M.
+dos Santos, D.
 J.
-\& Wedekind, C.
-\href{https://doi.org/10.1098/rspb.2010.1275}{The evolution of punishment through reputation}.
+Rankin, C.
+Wedekind, \href{https://doi.org/10.1098/rspb.2010.1275}{The evolution of punishment through reputation}.
 \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{278}, 371--377 (2011).}
@@
 \CSLLeftMargin{3. }%
-\CSLRightInline{dos Santos, M., Rankin, D.
+\CSLRightInline{M.
+dos Santos, D.
 J.
-\& Wedekind, C.
-\href{https://www.jstor.org/stable/24032701}{Human cooperation based on punishment reputation}.
+Rankin, C.
+Wedekind, \href{https://www.jstor.org/stable/24032701}{Human cooperation based on punishment reputation}.
 \emph{Evolution} \textbf{67}, 2446--2450 (2013).}
@@
 \CSLLeftMargin{4. }%
-\CSLRightInline{Raihani, N. J. \& Bshary, R. \href{https://doi.org/10.1016/j.tree.2014.12.003}{The reputation of punishers}. \emph{Trends in Ecology \& Evolution} \textbf{30}, 98--103 (2015).}
+\CSLRightInline{N. J. Raihani, R. Bshary, \href{https://doi.org/10.1016/j.tree.2014.12.003}{The reputation of punishers}. \emph{Trends in Ecology \& Evolution} \textbf{30}, 98--103 (2015).}
 
@@
 \CSLLeftMargin{5. }%
-\CSLRightInline{Ostrom, E. \emph{Governing the commons: The evolution of institutions for collective action}. (Cambridge University Press, 1990).}
+\CSLRightInline{E. Ostrom, \emph{Governing the commons: The evolution of institutions for collective action} (Cambridge University Press, 1990).}
 
@@
 \CSLLeftMargin{6. }%
-\CSLRightInline{Raihani, N. J. \& Bshary, R. \href{https://doi.org/10.1017/ehs.2019.12}{Punishment: One tool, many uses}. \emph{Evolutionary Human Sciences} \textbf{1}, e12 (2019).}
+\CSLRightInline{N. J. Raihani, R. Bshary, \href{https://doi.org/10.1017/ehs.2019.12}{Punishment: One tool, many uses}. \emph{Evolutionary Human Sciences} \textbf{1}, e12 (2019).}
 
@@
 \CSLLeftMargin{7. }%
-\CSLRightInline{Balliet, D., Mulder, L.
+\CSLRightInline{D.
+Balliet, L.
 B.
-\& Van Lange, P.
+Mulder, P.
 A.
 M.
-\href{https://doi.org/10.1037/a0023489}{Reward, punishment, and cooperation: A meta-analysis}.
+Van Lange, \href{https://doi.org/10.1037/a0023489}{Reward, punishment, and cooperation: A meta-analysis}.
 \emph{Psychological Bulletin} \textbf{137}, 594--615 (2011).}
@@
 \CSLLeftMargin{8. }%
-\CSLRightInline{Balliet, D.
-\& Lange, P.
+\CSLRightInline{D.
+Balliet, P.
 A.
@@
 V.
-\href{https://doi.org/10.1177/1745691613488533}{Trust, punishment, and cooperation across 18 societies: A meta-analysis}.
+Lange, \href{https://doi.org/10.1177/1745691613488533}{Trust, punishment, and cooperation across 18 societies: A meta-analysis}.
 \emph{Perspectives on Psychological Science} \textbf{8}, 363--379 (2013).}
@@
 \CSLLeftMargin{9. }%
-\CSLRightInline{Chaudhuri, A.
-\href{https://doi.org/10.1007/s10683-010-9257-1}{Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature}.
+\CSLRightInline{A.
+Chaudhuri, \href{https://doi.org/10.1007/s10683-010-9257-1}{Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature}.
 \emph{Experimental Economics} \textbf{14}, 47--83 (2011).}
@@
 \CSLLeftMargin{10. }%
-\CSLRightInline{Dreber, A., Rand, D. G., Fudenberg, D. \& Nowak, M. A. \href{https://doi.org/10.1038/nature06723}{Winners don't punish}. \emph{Nature} \textbf{452}, 348--351 (2008).}
+\CSLRightInline{A. Dreber, D. G. Rand, D. Fudenberg, M. A. Nowak, \href{https://doi.org/10.1038/nature06723}{Winners don't punish}. \emph{Nature} \textbf{452}, 348--351 (2008).}
 
@@
 \CSLLeftMargin{11. }%
-\CSLRightInline{Fehr, E.
-\& Gächter, S.
-\href{https://www.jstor.org/stable/117319}{Cooperation and punishment in public goods experiments}.
-\emph{American Economic Review} \textbf{90}, 980--994 (2000).}
+\CSLRightInline{E. Fehr, S. Gächter, \href{https://www.jstor.org/stable/117319}{Cooperation and punishment in public goods experiments}. \emph{American Economic Review} \textbf{90}, 980--994 (2000).}
 
@@
 \CSLLeftMargin{12. }%
-\CSLRightInline{Fehr, E. \& Gächter, S. \href{https://doi.org/10.1038/415137a}{Altruistic punishment in humans}. \emph{Nature} \textbf{415}, 137--140 (2002).}
+\CSLRightInline{E. Fehr, S. Gächter, \href{https://doi.org/10.1038/415137a}{Altruistic punishment in humans}. \emph{Nature} \textbf{415}, 137--140 (2002).}
 
@@
 \CSLLeftMargin{13. }%
-\CSLRightInline{Henrich, J. \emph{et al.} \href{https://doi.org/10.1126/science.1127333}{Costly punishment across human societies}. \emph{Science} \textbf{312}, 1767--1770 (2006).}
+\CSLRightInline{J. Henrich, \emph{et al.}, \href{https://doi.org/10.1126/science.1127333}{Costly punishment across human societies}. \emph{Science} \textbf{312}, 1767--1770 (2006).}
 
@@
 \CSLLeftMargin{14. }%
-\CSLRightInline{Nikiforakis, N. \& Normann, H.-T. A comparative statics analysis of punishment in public-good experiments. \emph{Experimental Economics} \textbf{11}, 358--369 (2008).}
+\CSLRightInline{N. Nikiforakis, H.-T. Normann, A comparative statics analysis of punishment in public-good experiments. \emph{Experimental Economics} \textbf{11}, 358--369 (2008).}
 
@@
 \CSLLeftMargin{15. }%
-\CSLRightInline{Raihani, N.
-J., Thornton, A.
-\& Bshary, R.
-\href{https://doi.org/10.1016/j.tree.2011.12.004}{Punishment and cooperation in nature}.
+\CSLRightInline{N.
+J.
+Raihani, A.
+Thornton, R.
+Bshary, \href{https://doi.org/10.1016/j.tree.2011.12.004}{Punishment and cooperation in nature}.
 \emph{Trends in Ecology \& Evolution} \textbf{27}, 288--295 (2012).}
@@
 \CSLLeftMargin{16. }%
-\CSLRightInline{de Quervain, D. J.-F. \emph{et al.} \href{https://doi.org/10.1126/science.1100735}{The neural basis of altruistic punishment}. \emph{Science} \textbf{305}, 1254--1258 (2004).}
+\CSLRightInline{D. J.-F. de Quervain, \emph{et al.}, \href{https://doi.org/10.1126/science.1100735}{The neural basis of altruistic punishment}. \emph{Science} \textbf{305}, 1254--1258 (2004).}
 
@@
 \CSLLeftMargin{17. }%
-\CSLRightInline{Herrmann, B., Thöni, C. \& Gächter, S. \href{https://doi.org/10.1126/science.1153808}{Antisocial punishment across societies}. \emph{Science} \textbf{319}, 1362--1367 (2008).}
+\CSLRightInline{B. Herrmann, C. Thöni, S. Gächter, \href{https://doi.org/10.1126/science.1153808}{Antisocial punishment across societies}. \emph{Science} \textbf{319}, 1362--1367 (2008).}
 
@@
 \CSLLeftMargin{18. }%
-\CSLRightInline{Camerer, C. F. \emph{Behavioral game theory: Experiments in strategic interaction}. (Russell Sage Foundation, 2003).}
+\CSLRightInline{C. F. Camerer, \emph{Behavioral game theory: Experiments in strategic interaction} (Russell Sage Foundation, 2003).}
 
@@
 \CSLLeftMargin{19. }%
-\CSLRightInline{Bowles, S. \& Gintis, H. \emph{A cooperative species: Human reciprocity and its evolution}. (Princeton University Press, 2013).}
+\CSLRightInline{S. Bowles, H. Gintis, \emph{A cooperative species: Human reciprocity and its evolution} (Princeton University Press, 2013).}
 
@@
 \CSLLeftMargin{20. }%
-\CSLRightInline{Boyd, R., Gintis, H., Bowles, S. \& Richerson, P. J. The evolution of altruistic punishment. \emph{Proceedings of the National Academy of Sciences} \textbf{100}, 3531--3535 (2003).}
+\CSLRightInline{R. Boyd, H. Gintis, S. Bowles, P. J. Richerson, The evolution of altruistic punishment. \emph{Proceedings of the National Academy of Sciences} \textbf{100}, 3531--3535 (2003).}
 
@@
 \CSLLeftMargin{21. }%
-\CSLRightInline{Chudek, M.
-\& Henrich, J.
-\href{https://doi.org/10.1016/j.tics.2011.03.003}{Culture-gene coevolution, norm-psychology and the emergence of human prosociality}.
+\CSLRightInline{M.
+Chudek, J.
+Henrich, \href{https://doi.org/10.1016/j.tics.2011.03.003}{Culture-gene coevolution, norm-psychology and the emergence of human prosociality}.
 \emph{Trends in Cognitive Sciences} \textbf{15}, 218--226 (2011).}
@@
 \CSLLeftMargin{22. }%
-\CSLRightInline{Henrich, J. \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter}. (Princeton University Press, 2017).}
+\CSLRightInline{J. Henrich, \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter} (Princeton University Press, 2017).}
 
@@
 \CSLLeftMargin{23. }%
-\CSLRightInline{Delton, A.
+\CSLRightInline{A.
 W.
-\& Krasnow, M.
+Delton, M.
 M.
-\href{https://doi.org/10.1016/j.evolhumbehav.2017.07.003}{The psychology of deterrence explains why group membership matters for third-party punishment}.
+Krasnow, \href{https://doi.org/10.1016/j.evolhumbehav.2017.07.003}{The psychology of deterrence explains why group membership matters for third-party punishment}.
 \emph{Evolution and Human Behavior} \textbf{38}, 734--743 (2017).}
@@
 \CSLLeftMargin{24. }%
-\CSLRightInline{Fehr, E. \& Schurtenberger, I. \href{https://doi.org/10.1038/s41562-018-0385-5}{Normative foundations of human cooperation}. \emph{Nature Human Behaviour} \textbf{2}, 458--468 (2018).}
+\CSLRightInline{E. Fehr, I. Schurtenberger, \href{https://doi.org/10.1038/s41562-018-0385-5}{Normative foundations of human cooperation}. \emph{Nature Human Behaviour} \textbf{2}, 458--468 (2018).}
 
@@
 \CSLLeftMargin{25. }%
-\CSLRightInline{Mathew, S. \& Boyd, R. Punishment sustains large-scale cooperation in prestate warfare. \emph{Proceedings of the National Academy of Sciences} \textbf{108}, 11375--11380 (2011).}
+\CSLRightInline{S. Mathew, R. Boyd, Punishment sustains large-scale cooperation in prestate warfare. \emph{Proceedings of the National Academy of Sciences} \textbf{108}, 11375--11380 (2011).}
 
@@
 \CSLLeftMargin{26. }%
-\CSLRightInline{Mathew, S.
-\& Boyd, R.
-\href{https://doi.org/10.1016/j.evolhumbehav.2013.10.001}{The cost of cowardice: Punitive sentiments towards free riders in {Turkana} raids}.
+\CSLRightInline{S.
+Mathew, R.
+Boyd, \href{https://doi.org/10.1016/j.evolhumbehav.2013.10.001}{The cost of cowardice: Punitive sentiments towards free riders in {Turkana} raids}.
 \emph{Evolution and Human Behavior} \textbf{35}, 58--64 (2014).}
@@
 \CSLLeftMargin{27. }%
-\CSLRightInline{Richerson, P.
-\emph{et al.} \href{https://doi.org/10.1017/S0140525X1400106X}{Cultural group selection plays an essential role in explaining human cooperation: A sketch of the evidence}.
+\CSLRightInline{P.
+Richerson, \emph{et al.}, \href{https://doi.org/10.1017/S0140525X1400106X}{Cultural group selection plays an essential role in explaining human cooperation: A sketch of the evidence}.
 \emph{Behavioral and Brain Sciences} \textbf{39}, e30 (2016).}
@@
 \CSLLeftMargin{28. }%
-\CSLRightInline{Carlsmith, K.
-M., Darley, J.
+\CSLRightInline{K.
 M.
-\& Robinson, P.
+Carlsmith, J.
+M.
+Darley, P.
 H.
-\href{https://doi.org/10.1037/0022-3514.83.2.284}{Why do we punish?
+Robinson, \href{https://doi.org/10.1037/0022-3514.83.2.284}{Why do we punish?
 Deterrence and just deserts as motives for punishment}.
@@
 \CSLLeftMargin{29. }%
-\CSLRightInline{Raihani, N.
+\CSLRightInline{N.
 J.
-\& McAuliffe, K.
-\href{https://doi.org/10.1098/rsbl.2012.0470}{Human punishment is motivated by inequity aversion, not a desire for reciprocity}.
+Raihani, K.
+McAuliffe, \href{https://doi.org/10.1098/rsbl.2012.0470}{Human punishment is motivated by inequity aversion, not a desire for reciprocity}.
 \emph{Biology Letters} \textbf{8}, 802--804 (2012).}
@@
 \CSLLeftMargin{30. }%
-\CSLRightInline{Dawes, C. T., Fowler, J. H., Johnson, T., McElreath, R. \& Smirnov, O. Egalitarian motives in humans. \emph{Nature} \textbf{446}, 794--796 (2007).}
+\CSLRightInline{C. T. Dawes, J. H. Fowler, T. Johnson, R. McElreath, O. Smirnov, Egalitarian motives in humans. \emph{Nature} \textbf{446}, 794--796 (2007).}
 
@@
 \CSLLeftMargin{31. }%
-\CSLRightInline{Bone, J. E. \& Raihani, N. J. Human punishment is motivated by both a desire for revenge and a desire for equality. \emph{Evolution and Human Behavior} \textbf{36}, 323--330 (2015).}
+\CSLRightInline{J. E. Bone, N. J. Raihani, Human punishment is motivated by both a desire for revenge and a desire for equality. \emph{Evolution and Human Behavior} \textbf{36}, 323--330 (2015).}
 
@@
 \CSLLeftMargin{32. }%
-\CSLRightInline{Walker, J. M. \& Halloran, M. A. Rewards and sanctions and the provision of public goods in one-shot settings. \emph{Experimental Economics} \textbf{7}, 235--247 (2004).}
+\CSLRightInline{J. M. Walker, M. A. Halloran, Rewards and sanctions and the provision of public goods in one-shot settings. \emph{Experimental Economics} \textbf{7}, 235--247 (2004).}
 
@@
 \CSLLeftMargin{33. }%
-\CSLRightInline{Barclay, P.
-\& Raihani, N.
-\href{https://doi.org/10.1016/j.evolhumbehav.2015.12.004}{Partner choice versus punishment in human prisoner's dilemmas}.
+\CSLRightInline{P.
+Barclay, N.
+Raihani, \href{https://doi.org/10.1016/j.evolhumbehav.2015.12.004}{Partner choice versus punishment in human prisoner's dilemmas}.
 \emph{Evolution and Human Behavior} \textbf{37}, 263--271 (2016).}
@@
 \CSLLeftMargin{34. }%
-\CSLRightInline{Crockett, M.
-J., Özdemir, Y.
-\& Fehr, E.
-\href{https://doi.org/10.1037/xge0000018}{The value of vengeance and the demand for deterrence}.
+\CSLRightInline{M.
+J.
+Crockett, Y.
+Özdemir, E.
+Fehr, \href{https://doi.org/10.1037/xge0000018}{The value of vengeance and the demand for deterrence}.
 \emph{Journal of Experimental Psychology: General} \textbf{143}, 2279--2286 (2014).}
@@
 \CSLLeftMargin{35. }%
-\CSLRightInline{Sylwester, K., Herrmann, B.
-\& Bryson, J.
+\CSLRightInline{K.
+Sylwester, B.
+Herrmann, J.
 J.
-\href{https://doi.org/10.1037/npe0000009}{Homo homini lupus?
+Bryson, \href{https://doi.org/10.1037/npe0000009}{Homo homini lupus?
 Explaining antisocial punishment}.
@@
 \CSLLeftMargin{36. }%
-\CSLRightInline{Fehr, E. \& Fischbacher, U. Third-party punishment and social norms. \emph{Evolution and Human Behavior} \textbf{25}, 63--87 (2004).}
+\CSLRightInline{E. Fehr, U. Fischbacher, Third-party punishment and social norms. \emph{Evolution and Human Behavior} \textbf{25}, 63--87 (2004).}
 
@@
 \CSLLeftMargin{37. }%
-\CSLRightInline{Johnson, T., Dawes, C.
-T., Fowler, J.
-H., McElreath, R.
-\& Smirnov, O.
-\href{https://doi.org/10.1016/j.econlet.2009.01.003}{The role of egalitarian motives in altruistic punishment}.
+\CSLRightInline{T.
+Johnson, C.
+T.
+Dawes, J.
+H.
+Fowler, R.
+McElreath, O.
+Smirnov, \href{https://doi.org/10.1016/j.econlet.2009.01.003}{The role of egalitarian motives in altruistic punishment}.
 \emph{Economics Letters} \textbf{102}, 192--194 (2009).}
@@
 \CSLLeftMargin{38. }%
-\CSLRightInline{Fowler, J. H., Johnson, T. \& Smirnov, O. \href{https://doi.org/10.1038/nature03256}{Egalitarian motive and altruistic punishment}. \emph{Nature} \textbf{433}, E1 (2005).}
+\CSLRightInline{J. H. Fowler, T. Johnson, O. Smirnov, \href{https://doi.org/10.1038/nature03256}{Egalitarian motive and altruistic punishment}. \emph{Nature} \textbf{433}, E1 (2005).}
 
@@
 \CSLLeftMargin{39. }%
-\CSLRightInline{Molleman, L., Van den Berg, P.
-\& Weissing, F.
+\CSLRightInline{L.
+Molleman, P.
+Van den Berg, F.
 J.
-\href{https://doi.org/10.1038/ncomms4570}{Consistent individual differences in human social learning strategies}.
+Weissing, \href{https://doi.org/10.1038/ncomms4570}{Consistent individual differences in human social learning strategies}.
 \emph{Nature Communications} \textbf{5}, 3570 (2014).}
@@
 \CSLLeftMargin{40. }%
-\CSLRightInline{Thielmann, I., Spadaro, G.
-\& Balliet, D.
-\href{https://doi.org/10.1037/bul0000217}{Personality and prosocial behavior: A theoretical framework and meta-analysis.} \emph{Psychological Bulletin} \textbf{146}, 30--90 (2020).}
+\CSLRightInline{I.
+Thielmann, G.
+Spadaro, D.
+Balliet, \href{https://doi.org/10.1037/bul0000217}{Personality and prosocial behavior: A theoretical framework and meta-analysis.} \emph{Psychological Bulletin} \textbf{146}, 30--90 (2020).}
 
@@
 \CSLLeftMargin{41. }%
-\CSLRightInline{Hofmann, W., Brandt, M.
-J., Wisneski, D.
-C., Rockenbach, B.
-\& Skitka, L.
+\CSLRightInline{W.
+Hofmann, M.
 J.
-\href{https://doi.org/10.1177/0146167218775075}{Moral punishment in everyday life}.
+Brandt, D.
+C.
+Wisneski, B.
+Rockenbach, L.
+J.
+Skitka, \href{https://doi.org/10.1177/0146167218775075}{Moral punishment in everyday life}.
 \emph{Personality and Social Psychology Bulletin} \textbf{44}, 1697--1711 (2018).}
@@
 \CSLLeftMargin{42. }%
-\CSLRightInline{Nisbett, R.
+\CSLRightInline{R.
 E.
-\& Wilson, T.
+Nisbett, T.
 D.
-\href{https://doi.org/10.1037/0033-295X.84.3.231}{Telling more than we can know: Verbal reports on mental processes}.
+Wilson, \href{https://doi.org/10.1037/0033-295X.84.3.231}{Telling more than we can know: Verbal reports on mental processes}.
 \emph{Psychological Review} \textbf{84}, 231--259 (1977).}
@@
 \CSLLeftMargin{43. }%
-\CSLRightInline{Barclay, P.
-\href{https://doi.org/10.1016/j.evolhumbehav.2006.01.003}{Reputational benefits for altruistic punishment}.
+\CSLRightInline{P.
+Barclay, \href{https://doi.org/10.1016/j.evolhumbehav.2006.01.003}{Reputational benefits for altruistic punishment}.
 \emph{Evolution and Human Behavior} \textbf{27}, 325--344 (2006).}
@@
 \CSLLeftMargin{44. }%
-\CSLRightInline{Batistoni, T., Barclay, P.
-\& Raihani, N.
+\CSLRightInline{T.
+Batistoni, P.
+Barclay, N.
 J.
-\href{https://doi.org/10.1098/rspb.2021.1773}{Third-party punishers do not compete to be chosen as partners in an experimental game}.
+Raihani, \href{https://doi.org/10.1098/rspb.2021.1773}{Third-party punishers do not compete to be chosen as partners in an experimental game}.
 \emph{Proceedings of the Royal Society B} \textbf{289}, 20211773 (2022).}
@@
 \CSLLeftMargin{45. }%
-\CSLRightInline{Jordan, J.
-J., Hoffman, M., Bloom, P.
-\& Rand, D.
+\CSLRightInline{J.
+J.
+Jordan, M.
+Hoffman, P.
+Bloom, D.
 G.
-\href{https://doi.org/10.1038/nature16981}{Third-party punishment as a costly signal of trustworthiness}.
+Rand, \href{https://doi.org/10.1038/nature16981}{Third-party punishment as a costly signal of trustworthiness}.
 \emph{Nature} \textbf{530}, 473--476 (2016).}
@@
 \CSLLeftMargin{46. }%
-\CSLRightInline{Jordan, J.
+\CSLRightInline{J.
 J.
-\& Rand, D.
+Jordan, D.
 G.
-Third-party punishment as a costly signal of high continuation probabilities in repeated games.
+Rand, Third-party punishment as a costly signal of high continuation probabilities in repeated games.
 \emph{Journal of Theoretical Biology} \textbf{421}, 189--202 (2017).}
@@
 \CSLLeftMargin{47. }%
-\CSLRightInline{Jordan, J.
+\CSLRightInline{J.
 J.
-\& Rand, D.
+Jordan, D.
 G.
-\href{https://doi.org/10.1037/pspi0000186}{Signaling when no one is watching: A reputation heuristics account of outrage and punishment in one-shot anonymous interactions}.
+Rand, \href{https://doi.org/10.1037/pspi0000186}{Signaling when no one is watching: A reputation heuristics account of outrage and punishment in one-shot anonymous interactions}.
 \emph{Journal of Personality and Social Psychology} \textbf{118}, 57--88 (2020).}
@@
 \CSLLeftMargin{48. }%
-\CSLRightInline{Deutchman, P., Bračič, M., Raihani, N.
+\CSLRightInline{P.
+Deutchman, M.
+Bračič, N.
 J.
-\& McAuliffe, K.
-\href{https://doi.org/10.1016/j.evolhumbehav.2020.06.001}{Punishment is strongly motivated by revenge and weakly motivated by inequity aversion}.
+Raihani, K.
+McAuliffe, \href{https://doi.org/10.1016/j.evolhumbehav.2020.06.001}{Punishment is strongly motivated by revenge and weakly motivated by inequity aversion}.
 \emph{Evolution and Human Behavior} \textbf{42}, 12--20 (2021).}
@@
 \CSLLeftMargin{49. }%
-\CSLRightInline{Marczyk, J. \href{https://doi.org/10.1371/journal.pone.0171298}{Human punishment is not primarily motivated by inequality}. \emph{PLoS ONE} \textbf{12}, e0171298 (2017).}
+\CSLRightInline{J. Marczyk, \href{https://doi.org/10.1371/journal.pone.0171298}{Human punishment is not primarily motivated by inequality}. \emph{PLoS ONE} \textbf{12}, e0171298 (2017).}
 
@@
 \CSLLeftMargin{50. }%
-\CSLRightInline{Bone, J. E., McAuliffe, K. \& Raihani, N. J. Exploring the motivations for punishment: Framing and country-level effects. \emph{PLoS One} \textbf{11}, e0159769 (2016).}
+\CSLRightInline{J. E. Bone, K. McAuliffe, N. J. Raihani, Exploring the motivations for punishment: Framing and country-level effects. \emph{PLoS One} \textbf{11}, e0159769 (2016).}
 
@@
 \CSLLeftMargin{51. }%
-\CSLRightInline{Pleasant, A.
-\& Barclay, P.
-Why hate the good guy?
+\CSLRightInline{A.
+Pleasant, P.
+Barclay, Why hate the good guy?
 Antisocial punishment of high cooperators is greater when people compete to be chosen.
@@
 \CSLLeftMargin{52. }%
-\CSLRightInline{Nikiforakis, N.
-\href{https://doi.org/10.1016/j.jpubeco.2007.04.008}{Punishment and counter-punishment in public good games: Can we really govern ourselves?} \emph{Journal of Public Economics} \textbf{92}, 91--112 (2008).}
+\CSLRightInline{N.
+Nikiforakis, \href{https://doi.org/10.1016/j.jpubeco.2007.04.008}{Punishment and counter-punishment in public good games: Can we really govern ourselves?} \emph{Journal of Public Economics} \textbf{92}, 91--112 (2008).}
 
@@
 \CSLLeftMargin{53. }%
-\CSLRightInline{Balafoutas, L., Nikiforakis, N.
-\& Rockenbach, B.
-\href{https://doi.org/10.1073/pnas.1413170111}{Direct and indirect punishment among strangers in the field}.
+\CSLRightInline{L.
+Balafoutas, N.
+Nikiforakis, B.
+Rockenbach, \href{https://doi.org/10.1073/pnas.1413170111}{Direct and indirect punishment among strangers in the field}.
 \emph{Proceedings of the National Academy of Sciences} \textbf{111}, 15924--15927 (2014).}
@@
 \CSLLeftMargin{54. }%
-\CSLRightInline{Balafoutas, L., Nikiforakis, N.
-\& Rockenbach, B.
-\href{https://doi.org/10.1038/ncomms13327}{Altruistic punishment does not increase with the severity of norm violations in the field}.
+\CSLRightInline{L.
+Balafoutas, N.
+Nikiforakis, B.
+Rockenbach, \href{https://doi.org/10.1038/ncomms13327}{Altruistic punishment does not increase with the severity of norm violations in the field}.
 \emph{Nature Communications} \textbf{7}, 13327 (2016).}
@@
 \CSLLeftMargin{55. }%
-\CSLRightInline{Baumard, N.
-\href{https://doi.org/10.1007/s11299-010-0079-9}{Has punishment played a role in the evolution of cooperation?
+\CSLRightInline{N.
+Baumard, \href{https://doi.org/10.1007/s11299-010-0079-9}{Has punishment played a role in the evolution of cooperation?
 A critical review}.
@@
 \CSLLeftMargin{56. }%
-\CSLRightInline{Fitouchi, L.
-\& Singh, M.
-\href{https://doi.org/10.1016/j.evolhumbehav.2023.03.001}{Punitive justice serves to restore reciprocal cooperation in three small-scale societies}.
+\CSLRightInline{L.
+Fitouchi, M.
+Singh, \href{https://doi.org/10.1016/j.evolhumbehav.2023.03.001}{Punitive justice serves to restore reciprocal cooperation in three small-scale societies}.
 \emph{Evolution and Human Behavior} \textbf{44}, 502--514 (2023).}
@@
 \CSLLeftMargin{57. }%
-\CSLRightInline{Singh, M. \& Garfield, Z. H. Evidence for third-party mediation but not punishment in {M}entawai justice. \emph{Nature Human Behaviour} \textbf{6}, 930--940 (2022).}
+\CSLRightInline{M. Singh, Z. H. Garfield, Evidence for third-party mediation but not punishment in {M}entawai justice. \emph{Nature Human Behaviour} \textbf{6}, 930--940 (2022).}
 
@@
 \CSLLeftMargin{58. }%
-\CSLRightInline{Molleman, L., Kölle, F., Starmer, C. \& Gächter, S. People prefer coordinated punishment in cooperative interactions. \emph{Nature Human Behaviour} \textbf{3}, 1145--1153 (2019).}
+\CSLRightInline{L. Molleman, F. Kölle, C. Starmer, S. Gächter, People prefer coordinated punishment in cooperative interactions. \emph{Nature Human Behaviour} \textbf{3}, 1145--1153 (2019).}
 
@@
 \CSLLeftMargin{59. }%
-\CSLRightInline{Adler, N.
-E., Epel, E.
-S., Castellazzo, G.
-\& Ickovics, J.
+\CSLRightInline{N.
+E.
+Adler, E.
+S.
+Epel, G.
+Castellazzo, J.
 R.
-\href{https://doi.org/10.1037/0278-6133.19.6.586}{Relationship of subjective and objective social status with psychological and physiological functioning: Preliminary data in healthy, {W}hite women}.
+Ickovics, \href{https://doi.org/10.1037/0278-6133.19.6.586}{Relationship of subjective and objective social status with psychological and physiological functioning: Preliminary data in healthy, {W}hite women}.
 \emph{Health Psychology} \textbf{19}, 586--592 (2000).}
@@
 \CSLLeftMargin{60. }%
-\CSLRightInline{Sibley, C.
+\CSLRightInline{C.
 G.
-\emph{et al.} The {M}ini-{IPIP}6: Validation and extension of a short measure of the {B}ig-{S}ix factors of personality in {N}ew {Z}ealand.
+Sibley, \emph{et al.}, The {M}ini-{IPIP}6: Validation and extension of a short measure of the {B}ig-{S}ix factors of personality in {N}ew {Z}ealand.
 \emph{New Zealand Journal of Psychology (Online)} \textbf{40}, 142 (2011).}
@@
 \CSLLeftMargin{61. }%
-\CSLRightInline{Murphy, R.
-O., Ackermann, K.
+\CSLRightInline{R.
+O.
+Murphy, K.
 A.
-\& Handgraaf, M.
+Ackermann, M.
 J.
 J.
-\href{https://doi.org/10.1017/S1930297500004204}{Measuring social value orientation}.
+Handgraaf, \href{https://doi.org/10.1017/S1930297500004204}{Measuring social value orientation}.
 \emph{Judgment and Decision Making} \textbf{6}, 771--781 (2011).}
@@
 \CSLLeftMargin{62. }%
-\CSLRightInline{Ho, A.
+\CSLRightInline{A.
 K.
-\emph{et al.} \href{https://doi.org/10.1037/pspi0000033}{The nature of social dominance orientation: Theorizing and measuring preferences for intergroup inequality using the new SDO7 scale}.
+Ho, \emph{et al.}, \href{https://doi.org/10.1037/pspi0000033}{The nature of social dominance orientation: Theorizing and measuring preferences for intergroup inequality using the new SDO7 scale}.
 \emph{Journal of Personality and Social Psychology} \textbf{109}, 1003 (2015).}
@@
 \CSLLeftMargin{63. }%
-\CSLRightInline{Bizumic, B.
-\& Duckitt, J.
-\href{https://doi.org/10.5964/jspp.v6i1.835}{Investigating right wing authoritarianism with a {V}ery {S}hort {A}uthoritarianism {S}cale}.
+\CSLRightInline{B.
+Bizumic, J.
+Duckitt, \href{https://doi.org/10.5964/jspp.v6i1.835}{Investigating right wing authoritarianism with a {V}ery {S}hort {A}uthoritarianism {S}cale}.
 \emph{Journal of Social and Political Psychology} \textbf{6}, 129--150 (2018).}
@@
 \CSLLeftMargin{64. }%
-\CSLRightInline{Laurin, K., Shariff, A.
-F., Henrich, J.
-\& Kay, A.
+\CSLRightInline{K.
+Laurin, A.
+F.
+Shariff, J.
+Henrich, A.
 C.
-\href{https://doi.org/10.1098/rspb.2012.0615}{Outsourcing punishment to {God}: Beliefs in divine control reduce earthly punishment}.
+Kay, \href{https://doi.org/10.1098/rspb.2012.0615}{Outsourcing punishment to {God}: Beliefs in divine control reduce earthly punishment}.
 \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{279}, 3272--3281 (2012).}
@@
 \CSLLeftMargin{65. }%
-\CSLRightInline{McElreath, R. \emph{\href{http://xcelab.net/rm/statistical-rethinking/}{Statistical rethinking: A {Bayesian} course with examples in {R} and {Stan}, 2nd edition}}. (CRC Press, 2020).}
+\CSLRightInline{R.
+McElreath, \emph{\href{http://xcelab.net/rm/statistical-rethinking/}{Statistical rethinking: A {Bayesian} course with examples in {R} and {Stan}, 2nd edition}}, 2nd Ed.
+(CRC Press, 2020).}
 
@@
 \CSLLeftMargin{66. }%
-\CSLRightInline{Stan Development Team. \href{http://mc-stan.org/}{{RStan}: The {R} interface to {Stan}}. (2020).}
+\CSLRightInline{Stan Development Team, \href{http://mc-stan.org/}{{RStan}: The {R} interface to {Stan}} (2020).}
 
@@
 \CSLLeftMargin{67. }%
-\CSLRightInline{R Core Team. \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}}. (R Foundation for Statistical Computing, 2022).}
+\CSLRightInline{R Core Team, \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}} (R Foundation for Statistical Computing, 2022).}
 
@@
 \CSLLeftMargin{68. }%
-\CSLRightInline{Wickham, H. \emph{\href{https://ggplot2.tidyverse.org}{{ggplot2}: Elegant graphics for data analysis}}. (Springer-Verlag New York, 2016).}
+\CSLRightInline{H. Wickham, \emph{\href{https://ggplot2.tidyverse.org}{{ggplot2}: Elegant graphics for data analysis}} (Springer-Verlag New York, 2016).}
 
@@
 \CSLLeftMargin{69. }%
-\CSLRightInline{Wilke, C. O. \emph{\href{https://CRAN.R-project.org/package=cowplot}{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}}. (2020).}
+\CSLRightInline{C. O. Wilke, \emph{\href{https://CRAN.R-project.org/package=cowplot}{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}} (2020).}
 
@@
 \CSLLeftMargin{70. }%
-\CSLRightInline{Landau, W.
+\CSLRightInline{W.
 M.
-\href{https://doi.org/10.21105/joss.02959}{The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing}.
+Landau, \href{https://doi.org/10.21105/joss.02959}{The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing}.
 \emph{Journal of Open Source Software} \textbf{6}, 2959 (2021).}
@@
 \CSLLeftMargin{71. }%
-\CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}}. (2022).}
+\CSLRightInline{F. Aust, M. Barth, \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}} (2022).}
 

### 2024-07-24 appendix.Rmd
submit manuscript

@@
 \centering
-\noindent \hspace*{10mm} \small Why do people punish? Evidence for a range of strategic concerns \newline
-\hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, & Nichola Raihani\textsuperscript{1,2} \newline
+\noindent \hspace*{25mm} \small Why do people punish? Evidence for a range of strategic concerns \newline
+\hspace*{30mm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, & Nichola Raihani\textsuperscript{1,2} \newline
 \raggedright
@@
 \noindent{\bfseries{Contents}}
-\contentsline {subsection}{Supplementary Figures S1-S13}{2}{}
-\contentsline {subsection}{Supplementary Tables S1-S4}{15}{}
+\contentsline {subsection}{Supplementary Figures S1-S14}{2}{}
+\contentsline {subsection}{Supplementary Tables S1-S4}{16}{}
 \normalsize
@@
 tar_read(plotAllSliders1)
+```
+
+\newpage
+
+(ref:plotSplineCaption) *The relationships between strategy frequencies and the
+overall costs of strategies across all twelve punishment decisions, in both
+countries.* Each point is a unique strategy that appears in our dataset at
+least once (for ease of presentation, the "never punish" strategy is excluded).
+Lines and shaded areas represent posterior predictions from splines fitted to
+the trend in each country separately.
+
+```{r plotSpline, fig.width=6, fig.height=3.5, fig.cap="(ref:plotSplineCaption)", warning=FALSE}
+tar_read(plotSpline)
 ```
@@
   placement = "H",
-  fontsize = "scriptsize"
+  font_size = "scriptsize"
 )

### 2024-07-24 manuscript.Rmd
submit manuscript

@@
 appendix          : "appendix.Rmd"
-csl               : pnas.csl
+csl               : nature.csl
+notes-after-punctuation: false
 ---
@@
 
-# Significance Statement
-
-Research has shown that humans are willing to punish others at a personal cost,
-but the motives for this behaviour remain unclear. Using a suite of economic 
-games, we show that punitive behaviour is consistent with a variety of strategic
-concerns. While most people choose not to punish in our games, people who do 
-punish tend to be motivated by egalitarianism rather than deterrence or 
-norm enforcement. We also find that different punishment strategies are linked 
-with different character traits, including personality, prosocial preferences, 
-and political and religious views. This research highlights the multifaceted
-nature of human punishment.
-
-\newpage
 \linenumbers
@@
 because our focus is on punishment strategies in anonymous economic games without 
-reputational incentives [but see [@Jordan2020]].
+reputational incentives (but see ref[@Jordan2020]).
 
@@
 [@Fitouchi2023;@Singh2022]. Future research should uncover whether people are 
-more willing to punish in these conventionalised contexts [e.g., see
-[@Molleman2019]].
+more willing to punish in these conventionalised contexts (e.g., see ref
+[@Molleman2019]).
 
-We finally highlight some key limitations that can guide future research. One 
-limitation with our design is that some strategies required more punishment than 
-others to be met. For example, the competitive strategy required punishment in 
-ten of twelve decisions, compared to the avoid disadvantageous inequity strategy
-which required only one instance of punishment (Table \@ref(tab:tableStrategies)). 
-Strategies thus differed in how "expensive" they were to implement, perhaps 
-explaining why some strategies were more common than others. This issue is 
-largely unavoidable in our design since strategies, by their very nature, differ 
-in how punitive they are. To partially mitigate this issue, we employed the 
-strategy method to incentivise participants, such that payoffs were calculated 
-from a randomly chosen game instead of summed across all games.
-
-Another limitation is that our results may be contingent on the particular suite 
-of anonymous stealing games that we used. With our anonymous design, we were 
-unable to study other potential reputational strategies underlying punishment, 
-such as signalling trustworthiness [@Jordan2020]. Moreover, stealing may be 
-evaluated differently to other forms of cheating, such as not contributing to 
-public goods, and other negative behaviours, such as lying or breaking taboos. 
-Finally, we were unable to include all permutations of situational features in 
-the games (e.g., second-party vs. third-party, equal vs. unequal) making it 
-difficult to interpret some patterns of behaviour. For example, many 
-participants punished only in the third-party game (Game F), but it is not clear 
-whether these participants were driven by the third-party nature of the game or 
-simply by the fact that stealing in that game generated inequality. Future work 
-should determine whether different reputational contexts, target behaviours, and 
-combinations of situational features elicit different punishment strategies.
+One potential limitation with our design is that some strategies required more 
+punishment than others, meaning that some strategies were more "expensive" to 
+implement. For example, the competitive strategy required punishment in ten of 
+twelve decisions, compared to the avoid disadvantageous inequity strategy which 
+required only one instance of punishment (Table \@ref(tab:tableStrategies)).
+We employed the strategy method to deal with this, calculating participant 
+payoffs from a randomly chosen game instead of accumulating the costs across 
+all games. This mitigates the concern that the total cost is the most salient
+feature determining participants' decisions. But even accounting for this, 
+one could still argue that the difference in overall costs explains why, for 
+example, the competitive strategy is less common in our dataset than the avoid 
+disadvantageous inequity strategy. We do not think this feature of our design 
+is a concern, however, for a number of reasons. First, we are interested in 
+measuring strategies underlying *costly* punishment. Some of these strategies,
+by their very nature, will be more costly to implement than others. This is 
+reflected in our design. It would not make theoretical sense to use an 
+alternative design where more or less punitive strategies are manipulated to 
+cost participants the same amount. Second, when we asked participants whether
+they would have punished if it were free to do so, we found low agreement 
+with this statement (Supplementary Figure \@ref(fig:plotSliders1)), 
+suggesting that participants were not particularly sensitive to the costs of 
+punishment. In line with this, when we plot the frequencies of different 
+strategies against their expected costs, we find that cost is not a perfect
+predictor of strategy frequency: many "cheap" strategies are rare in our 
+data and some "expensive" strategies are quite common (e.g., always punish;
+Supplementary Figure \@ref(fig:plotSpline)). Third, this argument implies 
+that the high frequency of the never punish strategy is merely an artefact
+of our design, since it is the only strategy that does not cost anything to
+implement. But as we have discussed, many other studies have found similar
+aversions to costly punishment in the lab [@Barclay2016;@Batistoni2022;
+@Bone2015;@Deutchman2021;@Raihani2012b] and in the real world
+[@Balafoutas2014;@Balafoutas2016;@Baumard2010] suggesting that this result is
+not an artefact of our design.
 
@@
 \begingroup
-
+\setlength\parindent{0pt}
 <div id="refs" custom-style="Bibliography"></div>

### 2024-07-24 manuscript.tex
submit manuscript

--- 2024-01-08 manuscript.tex
@@
 \documentclass[
-  man,floatsintext]{apa6}
-\usepackage{amsmath,amssymb}
-\usepackage{iftex}
-\ifPDFTeX
+  english,
+  man, donotrepeattitle,floatsintext]{apa6}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
   \usepackage[T1]{fontenc}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math} % this also loads fontspec
+  \usepackage{unicode-math}
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else
-  % xetex/luatex font selection
 \fi
@@
 \usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Why do people punish? Evidence for a range of strategic concerns},
+  pdfauthor={Scott Claessens*1, Quentin D. Atkinson1, \& Nichola Raihani1,2},
+  pdflang={en-EN},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx}
@@
 \fi
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-\newlength{\cslentryspacingunit} % times entry-spacing
-\setlength{\cslentryspacingunit}{\parskip}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1
-  \let\oldpar\par
-  \def\par{\hangindent=\cslhangindent\oldpar}
-  \fi
-  % set entry spacing
-  \setlength{\parskip}{#2\cslentryspacingunit}
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-\ifLuaTeX
-\usepackage[bidi=basic]{babel}
-\else
-\usepackage[bidi=default]{babel}
-\fi
-\babelprovide[main,import]{english}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
 % Manuscript styling
@@
 \note{\raggedright * Correspondence concerning this article should be addressed to Scott Claessens, Level 2, Building 302, 23 Symonds Street, Auckland, New Zealand.
-\text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This working paper has not yet been peer-reviewed.}
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
+\text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This manuscript has not yet been peer-reviewed.
+\par This manuscript was published as a pre-print on PsyArXiv under a CC BY 4.0 \text{license:} \url{https://doi.org/10.31234/osf.io/ys6rm} \par \text{Classification:} Social Sciences; Psychological and Cognitive Sciences \par \text{Keywords:} punishment; cooperation; economic games}
+\ifxetex
+  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
+  \usepackage{polyglossia}
+  \setmainlanguage[]{english}
+\else
+  \usepackage[shorthands=off,main=english]{babel}
 \fi
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same}
-\hypersetup{
-  pdftitle={Why do people punish? Evidence for a range of strategic concerns},
-  pdfauthor={Scott Claessens*1, Quentin D. Atkinson1, \& Nichola Raihani1,2},
-  pdflang={en-EN},
-  pdfkeywords={punishment; cooperation; economic games},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
 
@@
 
-\nolinenumbers
+\linenumbers
 
@@
 whereby individuals harm others at a personal cost\textsuperscript{1},
-ostensibly encouraging cooperative behaviour from the target (or bystanders\textsuperscript{2--4}) in the future. Punishment offers a
+ostensibly encouraging cooperative behaviour from the target, or bystanders\textsuperscript{2--4}, in the future. Punishment offers a
 route to maintaining or increasing cooperation by changing the payoff structure
@@
 by concerns about relative payoffs, such as disadvantageous inequity aversion
-(i.e., avoiding having less than others\textsuperscript{6,29}) and/or
+(i.e., avoiding having less than others)\textsuperscript{6,29} and/or
 general egalitarian preferences (i.e., wanting all participants to receive the
-same payoffs\textsuperscript{30}). Such concerns about relative payoffs may be activated
+same payoffs)\textsuperscript{30}. Such concerns about relative payoffs may be activated
 when participants earn less than cheats in economic games or when there are
@@
 punishment\textsuperscript{34}; punishing those who did not cheat or who
-over-contributed to collective action (antisocial punishment\textsuperscript{17,35}); punishing in scenarios where they were not
-personally harmed (third-party punishment\textsuperscript{36}); and punishing in
+over-contributed to collective action (antisocial punishment)\textsuperscript{17,35}; punishing in scenarios where they were not
+personally harmed (third-party punishment)\textsuperscript{36}; and punishing in
 scenarios where disparities in payoffs did not arise from participants'
@@
 -0.20, 95\% CI {[}-0.38
--0.02{]}) but this effect was small.
+-0.03{]}) but this effect was small.
 
@@
 disappears when relative payoffs cannot be changed (e.g., when punishment is
-only available with a 1:1 fee-fine ratio\textsuperscript{35}).
+only available with a 1:1 fee-fine ratio)\textsuperscript{35}.
 
@@
 limiting the potential for feuds and cycles of retaliation\textsuperscript{56,57}. Future research should uncover whether people are
-more willing to punish in these conventionalised contexts (e.g., see\textsuperscript{58}).
-
-We finally highlight some key limitations that can guide future research. One
-limitation with our design is that some strategies required more punishment than
-others to be met. For example, the competitive strategy required punishment in
-ten of twelve decisions, compared to the avoid disadvantageous inequity strategy
-which required only one instance of punishment (Table \ref{tab:tableStrategies}).
-Strategies thus differed in how ``expensive'' they were to implement, perhaps
-explaining why some strategies were more common than others. This issue is
-largely unavoidable in our design since strategies, by their very nature, differ
-in how punitive they are. To partially mitigate this issue, we employed the
-strategy method to incentivise participants, such that payoffs were calculated
-from a randomly chosen game instead of summed across all games.
-
-Another limitation is that our results may be contingent on the particular suite
-of anonymous stealing games that we used. With our anonymous design, we were
-unable to study other potential reputational strategies underlying punishment,
-such as signalling trustworthiness\textsuperscript{47}. Moreover, stealing may be
-evaluated differently to other forms of cheating, such as not contributing to
-public goods, and other negative behaviours, such as lying or breaking taboos.
-Finally, we were unable to include all permutations of situational features in
-the games (e.g., second-party vs.~third-party, equal vs.~unequal) making it
-difficult to interpret some patterns of behaviour. For example, many
-participants punished only in the third-party game (Game F), but it is not clear
-whether these participants were driven by the third-party nature of the game or
-simply by the fact that stealing in that game generated inequality. Future work
-should determine whether different reputational contexts, target behaviours, and
-combinations of situational features elicit different punishment strategies.
+more willing to punish in these conventionalised contexts (e.g., see ref\textsuperscript{58}).
+
+One potential limitation with our design is that some strategies required more
+punishment than others, meaning that some strategies were more ``expensive'' to
+implement. For example, the competitive strategy required punishment in ten of
+twelve decisions, compared to the avoid disadvantageous inequity strategy which
+required only one instance of punishment (Table \ref{tab:tableStrategies}).
+We employed the strategy method to deal with this, calculating participant
+payoffs from a randomly chosen game instead of accumulating the costs across
+all games. This mitigates the concern that the total cost is the most salient
+feature determining participants' decisions. But even accounting for this,
+one could still argue that the difference in overall costs explains why, for
+example, the competitive strategy is less common in our dataset than the avoid
+disadvantageous inequity strategy. We do not think this feature of our design
+is a concern, however, for a number of reasons. First, we are interested in
+measuring strategies underlying \emph{costly} punishment. Some of these strategies,
+by their very nature, will be more costly to implement than others. This is
+reflected in our design. It would not make theoretical sense to use an
+alternative design where more or less punitive strategies are manipulated to
+cost participants the same amount. Second, when we asked participants whether
+they would have punished if it were free to do so, we found low agreement
+with this statement (Supplementary Figure \ref{fig:plotSliders1}),
+suggesting that participants were not particularly sensitive to the costs of
+punishment. In line with this, when we plot the frequencies of different
+strategies against their expected costs, we find that cost is not a perfect
+predictor of strategy frequency: many ``cheap'' strategies are rare in our
+data and some ``expensive'' strategies are quite common (e.g., always punish;
+Supplementary Figure \ref{fig:plotSpline}). Third, this argument implies
+that the high frequency of the never punish strategy is merely an artefact
+of our design, since it is the only strategy that does not cost anything to
+implement. But as we have discussed, many other studies have found similar
+aversions to costly punishment in the lab\textsuperscript{29,31,33,44,48} and in the real world\textsuperscript{53--55} suggesting that this result is
+not an artefact of our design.
 
@@
 We pre-registered that we would use a Bayesian latent state model to infer
-unobserved punishment strategies from the observed data (for a similar version
-of this model, see ref\textsuperscript{65}). In this model, participants \(i\) in
+unobserved punishment strategies from the observed data {[}for a similar version
+of this model, see ref\textsuperscript{65}{]}. In this model, participants \(i\) in
 countries \(c\) make binary punishment decisions across twelve decisions \(j\). We
@@
 \begingroup
+\setlength\parindent{0pt}
 
 \hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-CluttonBrock1995}{}}%
-\CSLLeftMargin{1. }%
-\CSLRightInline{Clutton-Brock, T. H. \& Parker, G. A. \href{https://doi.org/10.1038/373209a0}{Punishment in animal societies}. \emph{Nature} \textbf{373}, 209--216 (1995).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-DosSantos2011}{}}%
-\CSLLeftMargin{2. }%
-\CSLRightInline{dos Santos, M., Rankin, D.
-J.
-\& Wedekind, C.
-\href{https://doi.org/10.1098/rspb.2010.1275}{The evolution of punishment through reputation}.
-\emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{278}, 371--377 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-DosSantos2013}{}}%
-\CSLLeftMargin{3. }%
-\CSLRightInline{dos Santos, M., Rankin, D.
-J.
-\& Wedekind, C.
-\href{https://www.jstor.org/stable/24032701}{Human cooperation based on punishment reputation}.
-\emph{Evolution} \textbf{67}, 2446--2450 (2013).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Raihani2015}{}}%
-\CSLLeftMargin{4. }%
-\CSLRightInline{Raihani, N. J. \& Bshary, R. \href{https://doi.org/10.1016/j.tree.2014.12.003}{The reputation of punishers}. \emph{Trends in Ecology \& Evolution} \textbf{30}, 98--103 (2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ostrom1990}{}}%
-\CSLLeftMargin{5. }%
-\CSLRightInline{Ostrom, E. \emph{Governing the commons: The evolution of institutions for collective action}. (Cambridge University Press, 1990).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Raihani2019}{}}%
-\CSLLeftMargin{6. }%
-\CSLRightInline{Raihani, N. J. \& Bshary, R. \href{https://doi.org/10.1017/ehs.2019.12}{Punishment: One tool, many uses}. \emph{Evolutionary Human Sciences} \textbf{1}, e12 (2019).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Balliet2011}{}}%
-\CSLLeftMargin{7. }%
-\CSLRightInline{Balliet, D., Mulder, L.
-B.
-\& Van Lange, P.
-A.
-M.
-\href{https://doi.org/10.1037/a0023489}{Reward, punishment, and cooperation: A meta-analysis}.
-\emph{Psychological Bulletin} \textbf{137}, 594--615 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Balliet2013}{}}%
-\CSLLeftMargin{8. }%
-\CSLRightInline{Balliet, D.
-\& Lange, P.
-A.
-M.
-V.
-\href{https://doi.org/10.1177/1745691613488533}{Trust, punishment, and cooperation across 18 societies: A meta-analysis}.
-\emph{Perspectives on Psychological Science} \textbf{8}, 363--379 (2013).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Chaudhuri2011}{}}%
-\CSLLeftMargin{9. }%
-\CSLRightInline{Chaudhuri, A.
-\href{https://doi.org/10.1007/s10683-010-9257-1}{Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature}.
-\emph{Experimental Economics} \textbf{14}, 47--83 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Dreber2008}{}}%
-\CSLLeftMargin{10. }%
-\CSLRightInline{Dreber, A., Rand, D. G., Fudenberg, D. \& Nowak, M. A. \href{https://doi.org/10.1038/nature06723}{Winners don't punish}. \emph{Nature} \textbf{452}, 348--351 (2008).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fehr2000}{}}%
-\CSLLeftMargin{11. }%
-\CSLRightInline{Fehr, E.
-\& Gächter, S.
-\href{https://www.jstor.org/stable/117319}{Cooperation and punishment in public goods experiments}.
-\emph{American Economic Review} \textbf{90}, 980--994 (2000).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fehr2002}{}}%
-\CSLLeftMargin{12. }%
-\CSLRightInline{Fehr, E. \& Gächter, S. \href{https://doi.org/10.1038/415137a}{Altruistic punishment in humans}. \emph{Nature} \textbf{415}, 137--140 (2002).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Henrich2006}{}}%
-\CSLLeftMargin{13. }%
-\CSLRightInline{Henrich, J. \emph{et al.} \href{https://doi.org/10.1126/science.1127333}{Costly punishment across human societies}. \emph{Science} \textbf{312}, 1767--1770 (2006).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nikiforakis2008a}{}}%
-\CSLLeftMargin{14. }%
-\CSLRightInline{Nikiforakis, N. \& Normann, H.-T. A comparative statics analysis of punishment in public-good experiments. \emph{Experimental Economics} \textbf{11}, 358--369 (2008).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Raihani2012a}{}}%
-\CSLLeftMargin{15. }%
-\CSLRightInline{Raihani, N.
-J., Thornton, A.
-\& Bshary, R.
-\href{https://doi.org/10.1016/j.tree.2011.12.004}{Punishment and cooperation in nature}.
-\emph{Trends in Ecology \& Evolution} \textbf{27}, 288--295 (2012).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-DeQuervain2004}{}}%
-\CSLLeftMargin{16. }%
-\CSLRightInline{de Quervain, D. J.-F. \emph{et al.} \href{https://doi.org/10.1126/science.1100735}{The neural basis of altruistic punishment}. \emph{Science} \textbf{305}, 1254--1258 (2004).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Herrmann2008}{}}%
-\CSLLeftMargin{17. }%
-\CSLRightInline{Herrmann, B., Thöni, C. \& Gächter, S. \href{https://doi.org/10.1126/science.1153808}{Antisocial punishment across societies}. \emph{Science} \textbf{319}, 1362--1367 (2008).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Camerer2003}{}}%
-\CSLLeftMargin{18. }%
-\CSLRightInline{Camerer, C. F. \emph{Behavioral game theory: Experiments in strategic interaction}. (Russell Sage Foundation, 2003).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bowles2013}{}}%
-\CSLLeftMargin{19. }%
-\CSLRightInline{Bowles, S. \& Gintis, H. \emph{A cooperative species: Human reciprocity and its evolution}. (Princeton University Press, 2013).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Boyd2003}{}}%
-\CSLLeftMargin{20. }%
-\CSLRightInline{Boyd, R., Gintis, H., Bowles, S. \& Richerson, P. J. The evolution of altruistic punishment. \emph{Proceedings of the National Academy of Sciences} \textbf{100}, 3531--3535 (2003).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Chudek2011}{}}%
-\CSLLeftMargin{21. }%
-\CSLRightInline{Chudek, M.
-\& Henrich, J.
-\href{https://doi.org/10.1016/j.tics.2011.03.003}{Culture-gene coevolution, norm-psychology and the emergence of human prosociality}.
-\emph{Trends in Cognitive Sciences} \textbf{15}, 218--226 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Henrich2017}{}}%
-\CSLLeftMargin{22. }%
-\CSLRightInline{Henrich, J. \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter}. (Princeton University Press, 2017).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Delton2017}{}}%
-\CSLLeftMargin{23. }%
-\CSLRightInline{Delton, A.
-W.
-\& Krasnow, M.
-M.
-\href{https://doi.org/10.1016/j.evolhumbehav.2017.07.003}{The psychology of deterrence explains why group membership matters for third-party punishment}.
-\emph{Evolution and Human Behavior} \textbf{38}, 734--743 (2017).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fehr2018}{}}%
-\CSLLeftMargin{24. }%
-\CSLRightInline{Fehr, E. \& Schurtenberger, I. \href{https://doi.org/10.1038/s41562-018-0385-5}{Normative foundations of human cooperation}. \emph{Nature Human Behaviour} \textbf{2}, 458--468 (2018).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Mathew2011}{}}%
-\CSLLeftMargin{25. }%
-\CSLRightInline{Mathew, S. \& Boyd, R. Punishment sustains large-scale cooperation in prestate warfare. \emph{Proceedings of the National Academy of Sciences} \textbf{108}, 11375--11380 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Mathew2014}{}}%
-\CSLLeftMargin{26. }%
-\CSLRightInline{Mathew, S.
-\& Boyd, R.
-\href{https://doi.org/10.1016/j.evolhumbehav.2013.10.001}{The cost of cowardice: Punitive sentiments towards free riders in {Turkana} raids}.
-\emph{Evolution and Human Behavior} \textbf{35}, 58--64 (2014).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Richerson2016}{}}%
-\CSLLeftMargin{27. }%
-\CSLRightInline{Richerson, P.
-\emph{et al.} \href{https://doi.org/10.1017/S0140525X1400106X}{Cultural group selection plays an essential role in explaining human cooperation: A sketch of the evidence}.
-\emph{Behavioral and Brain Sciences} \textbf{39}, e30 (2016).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Carlsmith2002}{}}%
-\CSLLeftMargin{28. }%
-\CSLRightInline{Carlsmith, K.
+\begin{cslreferences}
+\leavevmode\hypertarget{ref-CluttonBrock1995}{}%
+1. Clutton-Brock, T. H. \& Parker, G. A. Punishment in animal societies. \emph{Nature} \textbf{373}, 209--216 (1995).
+
+\leavevmode\hypertarget{ref-DosSantos2011}{}%
+2. dos Santos, M., Rankin, D. J. \& Wedekind, C. The evolution of punishment through reputation. \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{278}, 371--377 (2011).
+
+\leavevmode\hypertarget{ref-DosSantos2013}{}%
+3. dos Santos, M., Rankin, D. J. \& Wedekind, C. Human cooperation based on punishment reputation. \emph{Evolution} \textbf{67}, 2446--2450 (2013).
+
+\leavevmode\hypertarget{ref-Raihani2015}{}%
+4. Raihani, N. J. \& Bshary, R. The reputation of punishers. \emph{Trends in Ecology \& Evolution} \textbf{30}, 98--103 (2015).
+
+\leavevmode\hypertarget{ref-Ostrom1990}{}%
+5. Ostrom, E. \emph{Governing the commons: The evolution of institutions for collective action}. (Cambridge University Press, 1990).
+
+\leavevmode\hypertarget{ref-Raihani2019}{}%
+6. Raihani, N. J. \& Bshary, R. Punishment: One tool, many uses. \emph{Evolutionary Human Sciences} \textbf{1}, e12 (2019).
+
+\leavevmode\hypertarget{ref-Balliet2011}{}%
+7. Balliet, D., Mulder, L. B. \& Van Lange, P. A. M. Reward, punishment, and cooperation: A meta-analysis. \emph{Psychological Bulletin} \textbf{137}, 594--615 (2011).
+
+\leavevmode\hypertarget{ref-Balliet2013}{}%
+8. Balliet, D. \& Lange, P. A. M. V. Trust, punishment, and cooperation across 18 societies: A meta-analysis. \emph{Perspectives on Psychological Science} \textbf{8}, 363--379 (2013).
+
+\leavevmode\hypertarget{ref-Chaudhuri2011}{}%
+9. Chaudhuri, A. Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature. \emph{Experimental Economics} \textbf{14}, 47--83 (2011).
+
+\leavevmode\hypertarget{ref-Dreber2008}{}%
+10. Dreber, A., Rand, D. G., Fudenberg, D. \& Nowak, M. A. Winners don't punish. \emph{Nature} \textbf{452}, 348--351 (2008).
+
+\leavevmode\hypertarget{ref-Fehr2000}{}%
+11. Fehr, E. \& Gächter, S. Cooperation and punishment in public goods experiments. \emph{American Economic Review} \textbf{90}, 980--994 (2000).
+
+\leavevmode\hypertarget{ref-Fehr2002}{}%
+12. Fehr, E. \& Gächter, S. Altruistic punishment in humans. \emph{Nature} \textbf{415}, 137--140 (2002).
+
+\leavevmode\hypertarget{ref-Henrich2006}{}%
+13. Henrich, J. \emph{et al.} Costly punishment across human societies. \emph{Science} \textbf{312}, 1767--1770 (2006).
+
+\leavevmode\hypertarget{ref-Nikiforakis2008a}{}%
+14. Nikiforakis, N. \& Normann, H.-T. A comparative statics analysis of punishment in public-good experiments. \emph{Experimental Economics} \textbf{11}, 358--369 (2008).
+
+\leavevmode\hypertarget{ref-Raihani2012a}{}%
+15. Raihani, N. J., Thornton, A. \& Bshary, R. Punishment and cooperation in nature. \emph{Trends in Ecology \& Evolution} \textbf{27}, 288--295 (2012).
+
+\leavevmode\hypertarget{ref-DeQuervain2004}{}%
+16. de Quervain, D. J.-F. \emph{et al.} The neural basis of altruistic punishment. \emph{Science} \textbf{305}, 1254--1258 (2004).
+
+\leavevmode\hypertarget{ref-Herrmann2008}{}%
+17. Herrmann, B., Thöni, C. \& Gächter, S. Antisocial punishment across societies. \emph{Science} \textbf{319}, 1362--1367 (2008).
+
+\leavevmode\hypertarget{ref-Camerer2003}{}%
+18. Camerer, C. F. \emph{Behavioral game theory: Experiments in strategic interaction}. (Russell Sage Foundation, 2003).
+
+\leavevmode\hypertarget{ref-Bowles2013}{}%
+19. Bowles, S. \& Gintis, H. \emph{A cooperative species: Human reciprocity and its evolution}. (Princeton University Press, 2013).
+
+\leavevmode\hypertarget{ref-Boyd2003}{}%
+20. Boyd, R., Gintis, H., Bowles, S. \& Richerson, P. J. The evolution of altruistic punishment. \emph{Proceedings of the National Academy of Sciences} \textbf{100}, 3531--3535 (2003).
+
+\leavevmode\hypertarget{ref-Chudek2011}{}%
+21. Chudek, M. \& Henrich, J. Culture-gene coevolution, norm-psychology and the emergence of human prosociality. \emph{Trends in Cognitive Sciences} \textbf{15}, 218--226 (2011).
+
+\leavevmode\hypertarget{ref-Henrich2017}{}%
+22. Henrich, J. \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter}. (Princeton University Press, 2017).
+
+\leavevmode\hypertarget{ref-Delton2017}{}%
+23. Delton, A. W. \& Krasnow, M. M. The psychology of deterrence explains why group membership matters for third-party punishment. \emph{Evolution and Human Behavior} \textbf{38}, 734--743 (2017).
+
+\leavevmode\hypertarget{ref-Fehr2018}{}%
+24. Fehr, E. \& Schurtenberger, I. Normative foundations of human cooperation. \emph{Nature Human Behaviour} \textbf{2}, 458--468 (2018).
+
+\leavevmode\hypertarget{ref-Mathew2011}{}%
+25. Mathew, S. \& Boyd, R. Punishment sustains large-scale cooperation in prestate warfare. \emph{Proceedings of the National Academy of Sciences} \textbf{108}, 11375--11380 (2011).
+
+\leavevmode\hypertarget{ref-Mathew2014}{}%
+26. Mathew, S. \& Boyd, R. The cost of cowardice: Punitive sentiments towards free riders in Turkana raids. \emph{Evolution and Human Behavior} \textbf{35}, 58--64 (2014).
+
+\leavevmode\hypertarget{ref-Richerson2016}{}%
+27.
+Richerson, P.
+\emph{et al.} Cultural group selection plays an essential role in explaining human cooperation: A sketch of the evidence.
+\emph{Behavioral and Brain Sciences} \textbf{39}, e30 (2016).
+
+\leavevmode\hypertarget{ref-Carlsmith2002}{}%
+28.
+Carlsmith, K.
 M., Darley, J.
@@
 H.
-\href{https://doi.org/10.1037/0022-3514.83.2.284}{Why do we punish?
-Deterrence and just deserts as motives for punishment}.
-\emph{Journal of Personality and Social Psychology} \textbf{83}, 284--299 (2002).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Raihani2012b}{}}%
-\CSLLeftMargin{29. }%
-\CSLRightInline{Raihani, N.
-J.
-\& McAuliffe, K.
-\href{https://doi.org/10.1098/rsbl.2012.0470}{Human punishment is motivated by inequity aversion, not a desire for reciprocity}.
-\emph{Biology Letters} \textbf{8}, 802--804 (2012).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Dawes2007}{}}%
-\CSLLeftMargin{30. }%
-\CSLRightInline{Dawes, C. T., Fowler, J. H., Johnson, T., McElreath, R. \& Smirnov, O. Egalitarian motives in humans. \emph{Nature} \textbf{446}, 794--796 (2007).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bone2015}{}}%
-\CSLLeftMargin{31. }%
-\CSLRightInline{Bone, J. E. \& Raihani, N. J. Human punishment is motivated by both a desire for revenge and a desire for equality. \emph{Evolution and Human Behavior} \textbf{36}, 323--330 (2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Walker2004}{}}%
-\CSLLeftMargin{32. }%
-\CSLRightInline{Walker, J. M. \& Halloran, M. A. Rewards and sanctions and the provision of public goods in one-shot settings. \emph{Experimental Economics} \textbf{7}, 235--247 (2004).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Barclay2016}{}}%
-\CSLLeftMargin{33. }%
-\CSLRightInline{Barclay, P.
-\& Raihani, N.
-\href{https://doi.org/10.1016/j.evolhumbehav.2015.12.004}{Partner choice versus punishment in human prisoner's dilemmas}.
-\emph{Evolution and Human Behavior} \textbf{37}, 263--271 (2016).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Crockett2014}{}}%
-\CSLLeftMargin{34. }%
-\CSLRightInline{Crockett, M.
-J., Özdemir, Y.
-\& Fehr, E.
-\href{https://doi.org/10.1037/xge0000018}{The value of vengeance and the demand for deterrence}.
-\emph{Journal of Experimental Psychology: General} \textbf{143}, 2279--2286 (2014).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Sylwester2013}{}}%
-\CSLLeftMargin{35. }%
-\CSLRightInline{Sylwester, K., Herrmann, B.
-\& Bryson, J.
-J.
-\href{https://doi.org/10.1037/npe0000009}{Homo homini lupus?
-Explaining antisocial punishment}.
-\emph{Journal of Neuroscience, Psychology, and Economics} \textbf{6}, 167--188 (2013).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fehr2004}{}}%
-\CSLLeftMargin{36. }%
-\CSLRightInline{Fehr, E. \& Fischbacher, U. Third-party punishment and social norms. \emph{Evolution and Human Behavior} \textbf{25}, 63--87 (2004).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Johnson2009}{}}%
-\CSLLeftMargin{37. }%
-\CSLRightInline{Johnson, T., Dawes, C.
-T., Fowler, J.
-H., McElreath, R.
-\& Smirnov, O.
-\href{https://doi.org/10.1016/j.econlet.2009.01.003}{The role of egalitarian motives in altruistic punishment}.
-\emph{Economics Letters} \textbf{102}, 192--194 (2009).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fowler2005}{}}%
-\CSLLeftMargin{38. }%
-\CSLRightInline{Fowler, J. H., Johnson, T. \& Smirnov, O. \href{https://doi.org/10.1038/nature03256}{Egalitarian motive and altruistic punishment}. \emph{Nature} \textbf{433}, E1 (2005).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Molleman2014}{}}%
-\CSLLeftMargin{39. }%
-\CSLRightInline{Molleman, L., Van den Berg, P.
-\& Weissing, F.
-J.
-\href{https://doi.org/10.1038/ncomms4570}{Consistent individual differences in human social learning strategies}.
-\emph{Nature Communications} \textbf{5}, 3570 (2014).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Thielmann2020}{}}%
-\CSLLeftMargin{40. }%
-\CSLRightInline{Thielmann, I., Spadaro, G.
-\& Balliet, D.
-\href{https://doi.org/10.1037/bul0000217}{Personality and prosocial behavior: A theoretical framework and meta-analysis.} \emph{Psychological Bulletin} \textbf{146}, 30--90 (2020).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Hofmann2018}{}}%
-\CSLLeftMargin{41. }%
-\CSLRightInline{Hofmann, W., Brandt, M.
-J., Wisneski, D.
-C., Rockenbach, B.
-\& Skitka, L.
-J.
-\href{https://doi.org/10.1177/0146167218775075}{Moral punishment in everyday life}.
-\emph{Personality and Social Psychology Bulletin} \textbf{44}, 1697--1711 (2018).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nisbett1977}{}}%
-\CSLLeftMargin{42. }%
-\CSLRightInline{Nisbett, R.
-E.
-\& Wilson, T.
-D.
-\href{https://doi.org/10.1037/0033-295X.84.3.231}{Telling more than we can know: Verbal reports on mental processes}.
-\emph{Psychological Review} \textbf{84}, 231--259 (1977).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Barclay2006}{}}%
-\CSLLeftMargin{43. }%
-\CSLRightInline{Barclay, P.
-\href{https://doi.org/10.1016/j.evolhumbehav.2006.01.003}{Reputational benefits for altruistic punishment}.
-\emph{Evolution and Human Behavior} \textbf{27}, 325--344 (2006).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Batistoni2022}{}}%
-\CSLLeftMargin{44. }%
-\CSLRightInline{Batistoni, T., Barclay, P.
+Why do we punish?
+Deterrence and just deserts as motives for punishment.
+\emph{Journal of Personality and Social Psychology} \textbf{83}, 284--299 (2002).
+
+\leavevmode\hypertarget{ref-Raihani2012b}{}%
+29. Raihani, N. J. \& McAuliffe, K. Human punishment is motivated by inequity aversion, not a desire for reciprocity. \emph{Biology Letters} \textbf{8}, 802--804 (2012).
+
+\leavevmode\hypertarget{ref-Dawes2007}{}%
+30. Dawes, C. T., Fowler, J. H., Johnson, T., McElreath, R. \& Smirnov, O. Egalitarian motives in humans. \emph{Nature} \textbf{446}, 794--796 (2007).
+
+\leavevmode\hypertarget{ref-Bone2015}{}%
+31. Bone, J. E. \& Raihani, N. J. Human punishment is motivated by both a desire for revenge and a desire for equality. \emph{Evolution and Human Behavior} \textbf{36}, 323--330 (2015).
+
+\leavevmode\hypertarget{ref-Walker2004}{}%
+32. Walker, J. M. \& Halloran, M. A. Rewards and sanctions and the provision of public goods in one-shot settings. \emph{Experimental Economics} \textbf{7}, 235--247 (2004).
+
+\leavevmode\hypertarget{ref-Barclay2016}{}%
+33. Barclay, P. \& Raihani, N. Partner choice versus punishment in human prisoner's dilemmas. \emph{Evolution and Human Behavior} \textbf{37}, 263--271 (2016).
+
+\leavevmode\hypertarget{ref-Crockett2014}{}%
+34. Crockett, M. J., Özdemir, Y. \& Fehr, E. The value of vengeance and the demand for deterrence. \emph{Journal of Experimental Psychology: General} \textbf{143}, 2279--2286 (2014).
+
+\leavevmode\hypertarget{ref-Sylwester2013}{}%
+35. Sylwester, K., Herrmann, B. \& Bryson, J. J. Homo homini lupus? Explaining antisocial punishment. \emph{Journal of Neuroscience, Psychology, and Economics} \textbf{6}, 167--188 (2013).
+
+\leavevmode\hypertarget{ref-Fehr2004}{}%
+36. Fehr, E. \& Fischbacher, U. Third-party punishment and social norms. \emph{Evolution and Human Behavior} \textbf{25}, 63--87 (2004).
+
+\leavevmode\hypertarget{ref-Johnson2009}{}%
+37. Johnson, T., Dawes, C. T., Fowler, J. H., McElreath, R. \& Smirnov, O. The role of egalitarian motives in altruistic punishment. \emph{Economics Letters} \textbf{102}, 192--194 (2009).
+
+\leavevmode\hypertarget{ref-Fowler2005}{}%
+38. Fowler, J. H., Johnson, T. \& Smirnov, O. Egalitarian motive and altruistic punishment. \emph{Nature} \textbf{433}, E1 (2005).
+
+\leavevmode\hypertarget{ref-Molleman2014}{}%
+39. Molleman, L., Van den Berg, P. \& Weissing, F. J. Consistent individual differences in human social learning strategies. \emph{Nature Communications} \textbf{5}, 3570 (2014).
+
+\leavevmode\hypertarget{ref-Thielmann2020}{}%
+40. Thielmann, I., Spadaro, G. \& Balliet, D. Personality and prosocial behavior: A theoretical framework and meta-analysis. \emph{Psychological Bulletin} \textbf{146}, 30--90 (2020).
+
+\leavevmode\hypertarget{ref-Hofmann2018}{}%
+41. Hofmann, W., Brandt, M. J., Wisneski, D. C., Rockenbach, B. \& Skitka, L. J. Moral punishment in everyday life. \emph{Personality and Social Psychology Bulletin} \textbf{44}, 1697--1711 (2018).
+
+\leavevmode\hypertarget{ref-Nisbett1977}{}%
+42. Nisbett, R. E. \& Wilson, T. D. Telling more than we can know: Verbal reports on mental processes. \emph{Psychological Review} \textbf{84}, 231--259 (1977).
+
+\leavevmode\hypertarget{ref-Barclay2006}{}%
+43. Barclay, P. Reputational benefits for altruistic punishment. \emph{Evolution and Human Behavior} \textbf{27}, 325--344 (2006).
+
+\leavevmode\hypertarget{ref-Batistoni2022}{}%
+44.
+Batistoni, T., Barclay, P.
 \& Raihani, N.
 J.
-\href{https://doi.org/10.1098/rspb.2021.1773}{Third-party punishers do not compete to be chosen as partners in an experimental game}.
-\emph{Proceedings of the Royal Society B} \textbf{289}, 20211773 (2022).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Jordan2016}{}}%
-\CSLLeftMargin{45. }%
-\CSLRightInline{Jordan, J.
-J., Hoffman, M., Bloom, P.
-\& Rand, D.
-G.
-\href{https://doi.org/10.1038/nature16981}{Third-party punishment as a costly signal of trustworthiness}.
-\emph{Nature} \textbf{530}, 473--476 (2016).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Jordan2017}{}}%
-\CSLLeftMargin{46. }%
-\CSLRightInline{Jordan, J.
+Third-party punishers do not compete to be chosen as partners in an experimental game.
+\emph{Proceedings of the Royal Society B} \textbf{289}, 20211773 (2022).
+
+\leavevmode\hypertarget{ref-Jordan2016}{}%
+45. Jordan, J. J., Hoffman, M., Bloom, P. \& Rand, D. G. Third-party punishment as a costly signal of trustworthiness. \emph{Nature} \textbf{530}, 473--476 (2016).
+
+\leavevmode\hypertarget{ref-Jordan2017}{}%
+46. Jordan, J. J. \& Rand, D. G. Third-party punishment as a costly signal of high continuation probabilities in repeated games. \emph{Journal of Theoretical Biology} \textbf{421}, 189--202 (2017).
+
+\leavevmode\hypertarget{ref-Jordan2020}{}%
+47.
+Jordan, J.
 J.
@@
 G.
-Third-party punishment as a costly signal of high continuation probabilities in repeated games.
-\emph{Journal of Theoretical Biology} \textbf{421}, 189--202 (2017).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Jordan2020}{}}%
-\CSLLeftMargin{47. }%
-\CSLRightInline{Jordan, J.
-J.
-\& Rand, D.
-G.
-\href{https://doi.org/10.1037/pspi0000186}{Signaling when no one is watching: A reputation heuristics account of outrage and punishment in one-shot anonymous interactions}.
-\emph{Journal of Personality and Social Psychology} \textbf{118}, 57--88 (2020).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Deutchman2021}{}}%
-\CSLLeftMargin{48. }%
-\CSLRightInline{Deutchman, P., Bračič, M., Raihani, N.
+Signaling when no one is watching: A reputation heuristics account of outrage and punishment in one-shot anonymous interactions.
+\emph{Journal of Personality and Social Psychology} \textbf{118}, 57--88 (2020).
+
+\leavevmode\hypertarget{ref-Deutchman2021}{}%
+48.
+Deutchman, P., Bračič, M., Raihani, N.
 J.
 \& McAuliffe, K.
-\href{https://doi.org/10.1016/j.evolhumbehav.2020.06.001}{Punishment is strongly motivated by revenge and weakly motivated by inequity aversion}.
-\emph{Evolution and Human Behavior} \textbf{42}, 12--20 (2021).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Marczyk2017}{}}%
-\CSLLeftMargin{49. }%
-\CSLRightInline{Marczyk, J. \href{https://doi.org/10.1371/journal.pone.0171298}{Human punishment is not primarily motivated by inequality}. \emph{PLoS ONE} \textbf{12}, e0171298 (2017).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bone2016}{}}%
-\CSLLeftMargin{50. }%
-\CSLRightInline{Bone, J. E., McAuliffe, K. \& Raihani, N. J. Exploring the motivations for punishment: Framing and country-level effects. \emph{PLoS One} \textbf{11}, e0159769 (2016).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Pleasant2018}{}}%
-\CSLLeftMargin{51. }%
-\CSLRightInline{Pleasant, A.
-\& Barclay, P.
-Why hate the good guy?
-Antisocial punishment of high cooperators is greater when people compete to be chosen.
-\emph{Psychological Science} \textbf{29}, 868--876 (2018).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Nikiforakis2008b}{}}%
-\CSLLeftMargin{52. }%
-\CSLRightInline{Nikiforakis, N.
-\href{https://doi.org/10.1016/j.jpubeco.2007.04.008}{Punishment and counter-punishment in public good games: Can we really govern ourselves?} \emph{Journal of Public Economics} \textbf{92}, 91--112 (2008).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Balafoutas2014}{}}%
-\CSLLeftMargin{53. }%
-\CSLRightInline{Balafoutas, L., Nikiforakis, N.
+Punishment is strongly motivated by revenge and weakly motivated by inequity aversion.
+\emph{Evolution and Human Behavior} \textbf{42}, 12--20 (2021).
+
+\leavevmode\hypertarget{ref-Marczyk2017}{}%
+49. Marczyk, J. Human punishment is not primarily motivated by inequality. \emph{PLoS ONE} \textbf{12}, e0171298 (2017).
+
+\leavevmode\hypertarget{ref-Bone2016}{}%
+50. Bone, J. E., McAuliffe, K. \& Raihani, N. J. Exploring the motivations for punishment: Framing and country-level effects. \emph{PLoS One} \textbf{11}, e0159769 (2016).
+
+\leavevmode\hypertarget{ref-Pleasant2018}{}%
+51. Pleasant, A. \& Barclay, P. Why hate the good guy? Antisocial punishment of high cooperators is greater when people compete to be chosen. \emph{Psychological Science} \textbf{29}, 868--876 (2018).
+
+\leavevmode\hypertarget{ref-Nikiforakis2008b}{}%
+52. Nikiforakis, N. Punishment and counter-punishment in public good games: Can we really govern ourselves? \emph{Journal of Public Economics} \textbf{92}, 91--112 (2008).
+
+\leavevmode\hypertarget{ref-Balafoutas2014}{}%
+53.
+Balafoutas, L., Nikiforakis, N.
 \& Rockenbach, B.
-\href{https://doi.org/10.1073/pnas.1413170111}{Direct and indirect punishment among strangers in the field}.
-\emph{Proceedings of the National Academy of Sciences} \textbf{111}, 15924--15927 (2014).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Balafoutas2016}{}}%
-\CSLLeftMargin{54. }%
-\CSLRightInline{Balafoutas, L., Nikiforakis, N.
-\& Rockenbach, B.
-\href{https://doi.org/10.1038/ncomms13327}{Altruistic punishment does not increase with the severity of norm violations in the field}.
-\emph{Nature Communications} \textbf{7}, 13327 (2016).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Baumard2010}{}}%
-\CSLLeftMargin{55. }%
-\CSLRightInline{Baumard, N.
-\href{https://doi.org/10.1007/s11299-010-0079-9}{Has punishment played a role in the evolution of cooperation?
-A critical review}.
-\emph{Mind \& Society} \textbf{9}, 171--192 (2010).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Fitouchi2023}{}}%
-\CSLLeftMargin{56. }%
-\CSLRightInline{Fitouchi, L.
-\& Singh, M.
-\href{https://doi.org/10.1016/j.evolhumbehav.2023.03.001}{Punitive justice serves to restore reciprocal cooperation in three small-scale societies}.
-\emph{Evolution and Human Behavior} \textbf{44}, 502--514 (2023).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Singh2022}{}}%
-\CSLLeftMargin{57. }%
-\CSLRightInline{Singh, M. \& Garfield, Z. H. Evidence for third-party mediation but not punishment in {M}entawai justice. \emph{Nature Human Behaviour} \textbf{6}, 930--940 (2022).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Molleman2019}{}}%
-\CSLLeftMargin{58. }%
-\CSLRightInline{Molleman, L., Kölle, F., Starmer, C. \& Gächter, S. People prefer coordinated punishment in cooperative interactions. \emph{Nature Human Behaviour} \textbf{3}, 1145--1153 (2019).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Adler2000}{}}%
-\CSLLeftMargin{59. }%
-\CSLRightInline{Adler, N.
+Direct and indirect punishment among strangers in the field.
+\emph{Proceedings of the National Academy of Sciences} \textbf{111}, 15924--15927 (2014).
+
+\leavevmode\hypertarget{ref-Balafoutas2016}{}%
+54. Balafoutas, L., Nikiforakis, N. \& Rockenbach, B. Altruistic punishment does not increase with the severity of norm violations in the field. \emph{Nature Communications} \textbf{7}, 13327 (2016).
+
+\leavevmode\hypertarget{ref-Baumard2010}{}%
+55. Baumard, N. Has punishment played a role in the evolution of cooperation? A critical review. \emph{Mind \& Society} \textbf{9}, 171--192 (2010).
+
+\leavevmode\hypertarget{ref-Fitouchi2023}{}%
+56. Fitouchi, L. \& Singh, M. Punitive justice serves to restore reciprocal cooperation in three small-scale societies. \emph{Evolution and Human Behavior} \textbf{44}, 502--514 (2023).
+
+\leavevmode\hypertarget{ref-Singh2022}{}%
+57. Singh, M. \& Garfield, Z. H. Evidence for third-party mediation but not punishment in Mentawai justice. \emph{Nature Human Behaviour} \textbf{6}, 930--940 (2022).
+
+\leavevmode\hypertarget{ref-Molleman2019}{}%
+58. Molleman, L., Kölle, F., Starmer, C. \& Gächter, S. People prefer coordinated punishment in cooperative interactions. \emph{Nature Human Behaviour} \textbf{3}, 1145--1153 (2019).
+
+\leavevmode\hypertarget{ref-Adler2000}{}%
+59.
+Adler, N.
 E., Epel, E.
@@
 R.
-\href{https://doi.org/10.1037/0278-6133.19.6.586}{Relationship of subjective and objective social status with psychological and physiological functioning: Preliminary data in healthy, {W}hite women}.
-\emph{Health Psychology} \textbf{19}, 586--592 (2000).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Sibley2011}{}}%
-\CSLLeftMargin{60. }%
-\CSLRightInline{Sibley, C.
+Relationship of subjective and objective social status with psychological and physiological functioning: Preliminary data in healthy, White women.
+\emph{Health Psychology} \textbf{19}, 586--592 (2000).
+
+\leavevmode\hypertarget{ref-Sibley2011}{}%
+60.
+Sibley, C.
 G.
-\emph{et al.} The {M}ini-{IPIP}6: Validation and extension of a short measure of the {B}ig-{S}ix factors of personality in {N}ew {Z}ealand.
-\emph{New Zealand Journal of Psychology (Online)} \textbf{40}, 142 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Murphy2011}{}}%
-\CSLLeftMargin{61. }%
-\CSLRightInline{Murphy, R.
-O., Ackermann, K.
-A.
-\& Handgraaf, M.
-J.
-J.
-\href{https://doi.org/10.1017/S1930297500004204}{Measuring social value orientation}.
-\emph{Judgment and Decision Making} \textbf{6}, 771--781 (2011).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Ho2015}{}}%
-\CSLLeftMargin{62. }%
-\CSLRightInline{Ho, A.
+\emph{et al.} The Mini-IPIP6: Validation and extension of a short measure of the Big-Six factors of personality in New Zealand.
+\emph{New Zealand Journal of Psychology (Online)} \textbf{40}, 142 (2011).
+
+\leavevmode\hypertarget{ref-Murphy2011}{}%
+61. Murphy, R. O., Ackermann, K. A. \& Handgraaf, M. J. J. Measuring social value orientation. \emph{Judgment and Decision Making} \textbf{6}, 771--781 (2011).
+
+\leavevmode\hypertarget{ref-Ho2015}{}%
+62.
+Ho, A.
 K.
-\emph{et al.} \href{https://doi.org/10.1037/pspi0000033}{The nature of social dominance orientation: Theorizing and measuring preferences for intergroup inequality using the new SDO7 scale}.
-\emph{Journal of Personality and Social Psychology} \textbf{109}, 1003 (2015).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Bizumic2018}{}}%
-\CSLLeftMargin{63. }%
-\CSLRightInline{Bizumic, B.
-\& Duckitt, J.
-\href{https://doi.org/10.5964/jspp.v6i1.835}{Investigating right wing authoritarianism with a {V}ery {S}hort {A}uthoritarianism {S}cale}.
-\emph{Journal of Social and Political Psychology} \textbf{6}, 129--150 (2018).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Laurin2012}{}}%
-\CSLLeftMargin{64. }%
-\CSLRightInline{Laurin, K., Shariff, A.
+\emph{et al.} The nature of social dominance orientation: Theorizing and measuring preferences for intergroup inequality using the new sdo7 scale.
+\emph{Journal of Personality and Social Psychology} \textbf{109}, 1003 (2015).
+
+\leavevmode\hypertarget{ref-Bizumic2018}{}%
+63. Bizumic, B. \& Duckitt, J. Investigating right wing authoritarianism with a Very Short Authoritarianism Scale. \emph{Journal of Social and Political Psychology} \textbf{6}, 129--150 (2018).
+
+\leavevmode\hypertarget{ref-Laurin2012}{}%
+64.
+Laurin, K., Shariff, A.
 F., Henrich, J.
@@
 C.
-\href{https://doi.org/10.1098/rspb.2012.0615}{Outsourcing punishment to {God}: Beliefs in divine control reduce earthly punishment}.
-\emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{279}, 3272--3281 (2012).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-McElreath2020}{}}%
-\CSLLeftMargin{65. }%
-\CSLRightInline{McElreath, R. \emph{\href{http://xcelab.net/rm/statistical-rethinking/}{Statistical rethinking: A {Bayesian} course with examples in {R} and {Stan}, 2nd edition}}. (CRC Press, 2020).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Stan2020}{}}%
-\CSLLeftMargin{66. }%
-\CSLRightInline{Stan Development Team. \href{http://mc-stan.org/}{{RStan}: The {R} interface to {Stan}}. (2020).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-RCoreTeam}{}}%
-\CSLLeftMargin{67. }%
-\CSLRightInline{R Core Team. \emph{\href{https://www.R-project.org/}{R: A language and environment for statistical computing}}. (R Foundation for Statistical Computing, 2022).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wickham2016}{}}%
-\CSLLeftMargin{68. }%
-\CSLRightInline{Wickham, H. \emph{\href{https://ggplot2.tidyverse.org}{{ggplot2}: Elegant graphics for data analysis}}. (Springer-Verlag New York, 2016).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wilke2020}{}}%
-\CSLLeftMargin{69. }%
-\CSLRightInline{Wilke, C. O. \emph{\href{https://CRAN.R-project.org/package=cowplot}{{cowplot}: Streamlined plot theme and plot annotations for 'ggplot2'}}. (2020).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Landau2021}{}}%
-\CSLLeftMargin{70. }%
-\CSLRightInline{Landau, W.
+Outsourcing punishment to God: Beliefs in divine control reduce earthly punishment.
+\emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{279}, 3272--3281 (2012).
+
+\leavevmode\hypertarget{ref-McElreath2020}{}%
+65. McElreath, R. \emph{Statistical rethinking: A Bayesian course with examples in R and Stan, 2nd edition}. (CRC Press, 2020).
+
+\leavevmode\hypertarget{ref-Stan2020}{}%
+66. Stan Development Team. RStan: The R interface to Stan. (2020).
+
+\leavevmode\hypertarget{ref-RCoreTeam}{}%
+67. R Core Team. \emph{R: A language and environment for statistical computing}. (R Foundation for Statistical Computing, 2022).
+
+\leavevmode\hypertarget{ref-Wickham2016}{}%
+68. Wickham, H. \emph{ggplot2: Elegant graphics for data analysis}. (Springer-Verlag New York, 2016).
+
+\leavevmode\hypertarget{ref-Wilke2020}{}%
+69. Wilke, C. O. \emph{cowplot: Streamlined plot theme and plot annotations for 'ggplot2'}. (2020).
+
+\leavevmode\hypertarget{ref-Landau2021}{}%
+70.
+Landau, W.
 M.
-\href{https://doi.org/10.21105/joss.02959}{The targets {R} package: A dynamic {M}ake-like function-oriented pipeline toolkit for reproducibility and high-performance computing}.
-\emph{Journal of Open Source Software} \textbf{6}, 2959 (2021).}
-
-\leavevmode\vadjust pre{\hypertarget{ref-Aust2022}{}}%
-\CSLLeftMargin{71. }%
-\CSLRightInline{Aust, F. \& Barth, M. \emph{\href{https://github.com/crsh/papaja}{{papaja}: {Prepare} reproducible {APA} journal articles with {R Markdown}}}. (2022).}
-
-\end{CSLReferences}
+The targets R package: A dynamic Make-like function-oriented pipeline toolkit for reproducibility and high-performance computing.
+\emph{Journal of Open Source Software} \textbf{6}, 2959 (2021).
+
+\leavevmode\hypertarget{ref-Aust2022}{}%
+71. Aust, F. \& Barth, M. \emph{papaja: Prepare reproducible APA journal articles with R Markdown}. (2022).
+\end{cslreferences}
 
@@
 
-\noindent \hspace*{10mm} \small Why do people punish? Evidence for a range of strategic concerns \newline
-\hspace*{1cm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, \& Nichola Raihani\textsuperscript{1,2} \newline
-
+\noindent \hspace*{25mm} \small Why do people punish? Evidence for a range of strategic concerns \newline
+\hspace*{30mm} \small Scott Claessens\textsuperscript{1}, Quentin D. Atkinson\textsuperscript{1}, \& Nichola Raihani\textsuperscript{1,2} \newline
 \raggedright
-
 \noindent \footnotesize \textsuperscript{1} School of Psychology, University of Auckland, Auckland, New Zealand \newline
@@
 credible intervals. AI = advantageous inequity, DI = disadvantageous inequity.}
+\end{figure}
+
+\newpage
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/plotSpline-1.pdf}
+\caption{\label{fig:plotSpline}\emph{The relationships between strategy frequencies and the
+overall costs of strategies across all twelve punishment decisions, in both
+countries.} Each point is a unique strategy that appears in our dataset at
+least once (for ease of presentation, the ``never punish'' strategy is excluded).
+Lines and shaded areas represent posterior predictions from splines fitted to
+the trend in each country separately.}
 \end{figure}
@@
 \begin{lltable}
+
+\scriptsize{
 
@@
 
+}
+
 \end{lltable}

### 2024-07-24 R/functions_loadData.R

@@ -423,3 +423,26 @@ excludeData <- function(d) {
       `3PP_Nothing` = ifelse(fail1, NA, `3PP_Nothing`) 
     )
 }
+
+# wrangle data for strategy frequencies
+wrangleStrategyFreqData <- function(d) {
+  d %>%
+    unite(
+      strategyAll,
+      pun1_1:pun6_2,
+      sep = "",
+      remove = FALSE
+      ) %>%
+    rowwise() %>%
+    mutate(
+      cost = 0.10 * sum(c_across(pun1_1:pun6_2))
+      ) %>%
+    group_by(strategyAll, Country) %>%
+    summarise(
+      n = n(),
+      cost = unique(cost),
+      .groups = "drop"
+      ) %>%
+    # remove never punish
+    filter(strategyAll != "000000000000")
+}

### 2024-07-24 R/functions_modelFitting.R

@@ -168,40 +168,6 @@ fitModel4 <- function(d, compiledModel4, predictor, error = 0,
   return(out)
 }
 
-# fit multilevel model for raw data comparisons
-# take vs. not take
-fitMLM1 <- function(d) {
-  d %>%
-    pivot_longer(
-      starts_with("pun"),
-      names_to = "decision",
-      values_to = "pun"
-      ) %>%
-    mutate(otherTook = ifelse(str_ends(decision, "_1"), 1, 0)) %>%
-    glmer(
-      formula = pun ~ 1 + otherTook + (1 | ID) + (1 | decision),
-      data = .,
-      family = binomial
-    )
-}
-
-# fit multilevel model for raw data comparisons
-# target behaviour produced DI
-fitMLM2 <- function(d) {
-  d %>%
-    pivot_longer(
-      starts_with("pun"),
-      names_to = "decision",
-      values_to = "pun"
-    ) %>%
-    mutate(DI = ifelse(decision %in% c("pun5_1", "pun6_1"), 1, 0)) %>%
-    glmer(
-      formula = pun ~ 1 + DI + (1 | ID) + (1 | decision),
-      data = .,
-      family = binomial
-    )
-}
-
 # fit stan model without predictor (estimate implementation error)
 fitModel5 <- function(d, compiledModel5,
                       iter = 2000, warmup = 1000, chains = 4, cores = 4) {
@@ -237,4 +203,53 @@ fitModel5 <- function(d, compiledModel5,
       pars = c("alpha","error","log_lik")
     )
   return(out)
+}
+
+# fit multilevel model for raw data comparisons
+# take vs. not take
+fitMLM1 <- function(d) {
+  d %>%
+    pivot_longer(
+      starts_with("pun"),
+      names_to = "decision",
+      values_to = "pun"
+      ) %>%
+    mutate(otherTook = ifelse(str_ends(decision, "_1"), 1, 0)) %>%
+    glmer(
+      formula = pun ~ 1 + otherTook + (1 | ID) + (1 | decision),
+      data = .,
+      family = binomial
+    )
+}
+
+# fit multilevel model for raw data comparisons
+# target behaviour produced DI
+fitMLM2 <- function(d) {
+  d %>%
+    pivot_longer(
+      starts_with("pun"),
+      names_to = "decision",
+      values_to = "pun"
+    ) %>%
[…]

### 2024-07-24 R/functions_modelResults.R

@@ -479,3 +479,49 @@ plotTraceMCMC <- function(m1.2) {
   ggsave(out, filename = "figures/modelResults/trace.pdf", width = 7, height = 6)
   return(out)
 }
+
+# plot splines model
+plotSplines <- function(dStrategiesAll, mSpline) {
+  # get posterior predictions
+  costs <- seq(0, 1.20, by = 0.01)
+  newdata <- data.frame(
+    cost = rep(costs, times = 2),
+    Country = rep(unique(dStrategiesAll$Country), each = length(costs))
+  )
+  f <- cbind(newdata, fitted(mSpline, newdata = newdata))
+  # plot
+  out <-
+    ggplot() +
+    geom_jitter(
+      aes(x = cost, y = n),
+      data = dStrategiesAll,
+      width = 0,
+      height = 0.1
+    ) +
+    geom_ribbon(
+      aes(x = cost, ymin = Q2.5 + 1, ymax = Q97.5 + 1),
+      data = f,
+      alpha = 0.2
+    ) +
+    geom_line(
+      aes(x = cost, y = Estimate + 1),
+      data = f
+    ) +
+    facet_wrap(. ~ Country) +
+    scale_y_continuous(
+      name = "Number of participants\nfollowing strategy (logged axis)\n",
+      trans = scales::log1p_trans(),
+      breaks = c(1,2,5,10,20,40,80)
+    ) +
+    scale_x_continuous(
+      name = "Cost to implement strategy (GBP)",
+      limits = c(0, 1.21),
+      breaks = c(0.00, 0.50, 1.00),
+      labels = function(x) sprintf("%.2f", x)
+    ) +
+    theme_classic()
+  # save
+  ggsave(out, filename = "figures/dataSummaries/splines.pdf",
+         width = 6, height = 3.5)
+  return(out)
+}

### 2024-07-24 R/functions_summariseData.R

@@ -287,6 +287,112 @@ plotSurveyCorrelations <- function(d) {
   return(out)
 }
 
+# plot splits graph of behavioural patterns
+plotSplitsGraph <- function(d) {
+  d <-
+    d %>%
+    # subset to behavioural decisions
+    dplyr::select(starts_with("pun"), strategy) %>%
+    # unite behavioural decisions
+    unite(
+      col = "pattern",
+      pun1_1:pun6_2,
+      sep = "",
+      remove = FALSE
+    ) %>%
+    # reduce to unique behavioural patterns with counts
+    group_by_at(vars(starts_with("pun"))) %>%
+    summarise(
+      count = n(),
+      pattern = unique(pattern),
+      strategy = unique(as.character(strategy)),
+      .groups = "drop"
+    ) %>%
+    # add tpp strategy
+    #mutate(strategy = ifelse(pattern == "000000000010", "Third-party", strategy)) %>%
+    # add antisocial strategy
+    bind_rows(
+      tibble(pun1_1 = 0, pun1_2 = 1, pun2_1 = 0, pun2_2 = 1, pun3_1 = 0, pun3_2 = 1,
+             pun4_1 = 0, pun4_2 = 1, pun5_1 = 0, pun5_2 = 1, pun6_1 = 0, pun6_2 = 1,
+             count = 0, pattern = "010101010101", strategy = "Antisocial")
+    ) %>%
+    # keep only other behavioural patterns with certain number of counts
+    filter(strategy != "N/A" | (strategy == "N/A" & count >= 3))
+  # n is number of unique patterns of behaviour
+  n <- nrow(d)
+  # calculate Hamming distance matrix for behavioural patterns
+  # (number of substitutions to get from one to another)
+  # e.g. 000000000000 -> 101000000000 = Hamming distance of 2
+  dist <- matrix(NA, nrow = n, ncol = n)
+  for (i in 1:n) {
+    for (j in 1:n) {
+      # how many elements of the behavioural pattern are different?
+      dist[i,j] <- sum(d[i,1:12] != d[j,1:12])
+    }
+  }
+  rownames(dist) <- colnames(dist) <- 1:n
+  # create neighbour net with phangorn package
+  net <- neighborNet(dist)
+  # plot splits graph
+  p <- 
+    ggsplitnet(
+      net,
+      colour = "grey70",
+      linewidth = 0.05
+    )
+  # labels for strategies
+  labelStrategies <- 
+    c(
[…]

### 2024-07-24 R/packages.R

@@ -1,0 +1,23 @@
+options(tidyverse.quiet = TRUE)
+library(crew.cluster)
+library(targets)
+library(tarchetypes)
+library(tidyverse)
+tar_option_set(
+  packages = c("ape", "brms", "cowplot", "ggcorrplot", "ggrepel",
+               "ggridges", "knitr", "lme4", "loo", "ltm",
+               "papaja", "phangorn", "rethinking", "rstan",
+               "tanggle", "tidybayes", "tidytext", "tidyverse",
+               "wordcloud")#,
+  #controller = crew_controller_slurm(
+  #  name = "punishStrategies",
+  #  workers = 1,
+  #  script_lines = "module load R",
+  #  slurm_time_minutes = 1440,
+  #  slurm_memory_gigabytes_per_cpu = 60,
+  #  slurm_cpus_per_task = 4,
+  #  garbage_collection = TRUE
+  #),
+  #memory = "transient",
+  #garbage_collection = TRUE
+)

### 2024-07-24 _targets.R

@@ -1,24 +1,4 @@
-library(targets)
-library(tarchetypes)
-options(tidyverse.quiet = TRUE)
-library(tidyverse)
-source("R/functions_modelFitting.R")
-source("R/functions_simulation.R")
-source("R/functions_powerAnalysis.R")
-source("R/functions_loadData.R")
-source("R/functions_summariseData.R")
-source("R/functions_modelResults.R")
-tar_option_set(
-  packages = c("ape", "cowplot", "ggcorrplot", "ggrepel", 
-               "ggridges", "knitr", "lme4", "ltm", "papaja", "phangorn", 
-               "rethinking", "rstan", "tanggle", "tidybayes", 
-               "tidytext", "tidyverse", "wordcloud"),
-  deployment = "main"
-)
-options(
-  clustermq.scheduler = "slurm", 
-  clustermq.template = "slurm_clustermq.tmpl"
-)
+targets::tar_source()
 
 ## power analysis static branching (see below)
 #powerMap <-
@@ -55,11 +35,11 @@ list(
   tar_target(fileModel4, "stan/punishStrat_withCategoricalPredictor.stan", format = "file"),
   tar_target(fileModel5, "stan/punishStrat_noPredictor_estimateError.stan", format = "file"),
   # compile models
-  tar_target(compiledModel1, stan_model(fileModel1), deployment = "worker"),
-  tar_target(compiledModel2, stan_model(fileModel2), deployment = "worker"),
-  tar_target(compiledModel3, stan_model(fileModel3), deployment = "worker"),
-  tar_target(compiledModel4, stan_model(fileModel4), deployment = "worker"),
-  tar_target(compiledModel5, stan_model(fileModel5), deployment = "worker"),
+  tar_target(compiledModel1, stan_model(fileModel1)),
+  tar_target(compiledModel2, stan_model(fileModel2)),
+  tar_target(compiledModel3, stan_model(fileModel3)),
+  tar_target(compiledModel4, stan_model(fileModel4)),
+  tar_target(compiledModel5, stan_model(fileModel5)),
   
   #### Simulation ####
   
@@ -68,8 +48,8 @@ list(
   tar_target(simBetas, c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0)),
   tar_target(dSim, simulateData(n = 100, errorRate = 0, alphas = simAlphas, betas = simBetas)),
   # fit model to simulated data
[…]

### 2024-08-05 manuscript.Rmd
submit manuscript

@@
 appendix          : "appendix.Rmd"
-csl               : nature.csl
-notes-after-punctuation: false
+csl               : pnas.csl
 ---

### 2024-08-05 manuscript.tex
submit manuscript

@@
 mechanism thought to sustain this level of cooperation is costly punishment,
-whereby individuals harm others at a personal cost\textsuperscript{1},
-ostensibly encouraging cooperative behaviour from the target, or bystanders\textsuperscript{2--4}, in the future. Punishment offers a
+whereby individuals harm others at a personal cost (1),
+ostensibly encouraging cooperative behaviour from the target, or bystanders
+(2--4), in the future. Punishment offers a
 route to maintaining or increasing cooperation by changing the payoff structure
 of social interactions such that it no longer pays to cheat or exploit social
-partners\textsuperscript{1,5}. Yet, despite its theoretical
+partners (1, 5). Yet, despite its theoretical
 importance, the question of why people choose to punish others is still hotly
-contested\textsuperscript{6}. In this study, we use a battery of economic games to
+contested (6). In this study, we use a battery of economic games to
 disentangle the different motives underpinning punishment and explore how these
@@
 for retribution rather than deterrence, punishing in proportion to the amount
-of harm that was personally caused\textsuperscript{28}. Punishment might be driven
+of harm that was personally caused (28). Punishment might be driven
 by concerns about relative payoffs, such as disadvantageous inequity aversion
-(i.e., avoiding having less than others)\textsuperscript{6,29} and/or
+(i.e., avoiding having less than others) (6, 29) and/or
 general egalitarian preferences (i.e., wanting all participants to receive the
-same payoffs)\textsuperscript{30}. Such concerns about relative payoffs may be activated
+same payoffs) (30). Such concerns about relative payoffs may be activated
 when participants earn less than cheats in economic games or when there are
@@
 people use punishment in seemingly disparate ways: punishing when no behaviour
-change is possible, such as in one-shot games\textsuperscript{13,29,31,32}, on the very last round of repeated games\textsuperscript{33}, or in games where the target never learns about the
-punishment\textsuperscript{34}; punishing those who did not cheat or who
-over-contributed to collective action (antisocial punishment)\textsuperscript{17,35}; punishing in scenarios where they were not
-personally harmed (third-party punishment)\textsuperscript{36}; and punishing in
+change is possible, such as in one-shot games (13, 29, 31, 32), on the very last round of repeated games
+(33), or in games where the target never learns about the
+punishment (34); punishing those who did not cheat or who
+over-contributed to collective action (antisocial punishment)
+(17, 35); punishing in scenarios where they were not
+personally harmed (third-party punishment) (36); and punishing in
 scenarios where disparities in payoffs did not arise from participants'
-actions\textsuperscript{30,37,38}.
+actions (30, 37, 38).
 
@@
 reputational functions of punishment in this table, such as signalling
-trustworthiness\textsuperscript{4,43--46},
+trustworthiness (4, 43--46),
 because our focus is on punishment strategies in anonymous economic games without
-reputational incentives (but see ref\textsuperscript{47}).
-
-Building on previous designs\textsuperscript{29,31,48,49}, we employ a suite of one-shot economic games where individuals
+reputational incentives (but see ref(47)).
+
+Building on previous designs (29, 31, 48, 49), we employ a suite of one-shot economic games where individuals
 are given the opportunity to punish targets at a personal cost (Figure
@@
 disappears when relative payoffs cannot be changed (e.g., when punishment is
-only available with a 1:1 fee-fine ratio)\textsuperscript{35}.
+only available with a 1:1 fee-fine ratio) (35).
 
@@
 punishment in small-scale societies often functions to compensate victims while
-limiting the potential for feuds and cycles of retaliation\textsuperscript{56,57}. Future research should uncover whether people are
-more willing to punish in these conventionalised contexts (e.g., see ref\textsuperscript{58}).
+limiting the potential for feuds and cycles of retaliation
+(56, 57). Future research should uncover whether people are
+more willing to punish in these conventionalised contexts (e.g., see ref
+(58)).
 
@@
 implement. But as we have discussed, many other studies have found similar
-aversions to costly punishment in the lab\textsuperscript{29,31,33,44,48} and in the real world\textsuperscript{53--55} suggesting that this result is
+aversions to costly punishment in the lab (29, 31, 33, 44, 48) and in the real world
+(53--55) suggesting that this result is
 not an artefact of our design.
@@
 unobserved punishment strategies from the observed data {[}for a similar version
-of this model, see ref\textsuperscript{65}{]}. In this model, participants \(i\) in
+of this model, see ref (65){]}. In this model, participants \(i\) in
 countries \(c\) make binary punishment decisions across twelve decisions \(j\). We
@@
 \leavevmode\hypertarget{ref-CluttonBrock1995}{}%
-1. Clutton-Brock, T. H. \& Parker, G. A. Punishment in animal societies. \emph{Nature} \textbf{373}, 209--216 (1995).
+1. T. H. Clutton-Brock, G. A. Parker, Punishment in animal societies. \emph{Nature} \textbf{373}, 209--216 (1995).
 
 \leavevmode\hypertarget{ref-DosSantos2011}{}%
-2. dos Santos, M., Rankin, D. J. \& Wedekind, C. The evolution of punishment through reputation. \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{278}, 371--377 (2011).
+2. M. dos Santos, D. J. Rankin, C. Wedekind, The evolution of punishment through reputation. \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{278}, 371--377 (2011).
 
 \leavevmode\hypertarget{ref-DosSantos2013}{}%
-3. dos Santos, M., Rankin, D. J. \& Wedekind, C. Human cooperation based on punishment reputation. \emph{Evolution} \textbf{67}, 2446--2450 (2013).
+3. M. dos Santos, D. J. Rankin, C. Wedekind, Human cooperation based on punishment reputation. \emph{Evolution} \textbf{67}, 2446--2450 (2013).
 
 \leavevmode\hypertarget{ref-Raihani2015}{}%
-4. Raihani, N. J. \& Bshary, R. The reputation of punishers. \emph{Trends in Ecology \& Evolution} \textbf{30}, 98--103 (2015).
+4. N. J. Raihani, R. Bshary, The reputation of punishers. \emph{Trends in Ecology \& Evolution} \textbf{30}, 98--103 (2015).
 
 \leavevmode\hypertarget{ref-Ostrom1990}{}%
-5. Ostrom, E. \emph{Governing the commons: The evolution of institutions for collective action}. (Cambridge University Press, 1990).
+5. E. Ostrom, \emph{Governing the commons: The evolution of institutions for collective action} (Cambridge University Press, 1990).
 
 \leavevmode\hypertarget{ref-Raihani2019}{}%
-6. Raihani, N. J. \& Bshary, R. Punishment: One tool, many uses. \emph{Evolutionary Human Sciences} \textbf{1}, e12 (2019).
+6. N. J. Raihani, R. Bshary, Punishment: One tool, many uses. \emph{Evolutionary Human Sciences} \textbf{1}, e12 (2019).
 
 \leavevmode\hypertarget{ref-Balliet2011}{}%
-7. Balliet, D., Mulder, L. B. \& Van Lange, P. A. M. Reward, punishment, and cooperation: A meta-analysis. \emph{Psychological Bulletin} \textbf{137}, 594--615 (2011).
+7. D. Balliet, L. B. Mulder, P. A. M. Van Lange, Reward, punishment, and cooperation: A meta-analysis. \emph{Psychological Bulletin} \textbf{137}, 594--615 (2011).
 
 \leavevmode\hypertarget{ref-Balliet2013}{}%
-8. Balliet, D. \& Lange, P. A. M. V. Trust, punishment, and cooperation across 18 societies: A meta-analysis. \emph{Perspectives on Psychological Science} \textbf{8}, 363--379 (2013).
+8. D. Balliet, P. A. M. V. Lange, Trust, punishment, and cooperation across 18 societies: A meta-analysis. \emph{Perspectives on Psychological Science} \textbf{8}, 363--379 (2013).
 
 \leavevmode\hypertarget{ref-Chaudhuri2011}{}%
-9. Chaudhuri, A. Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature. \emph{Experimental Economics} \textbf{14}, 47--83 (2011).
+9. A. Chaudhuri, Sustaining cooperation in laboratory public goods experiments: A selective survey of the literature. \emph{Experimental Economics} \textbf{14}, 47--83 (2011).
 
 \leavevmode\hypertarget{ref-Dreber2008}{}%
-10. Dreber, A., Rand, D. G., Fudenberg, D. \& Nowak, M. A. Winners don't punish. \emph{Nature} \textbf{452}, 348--351 (2008).
+10. A. Dreber, D. G. Rand, D. Fudenberg, M. A. Nowak, Winners don't punish. \emph{Nature} \textbf{452}, 348--351 (2008).
 
 \leavevmode\hypertarget{ref-Fehr2000}{}%
-11. Fehr, E. \& Gächter, S. Cooperation and punishment in public goods experiments. \emph{American Economic Review} \textbf{90}, 980--994 (2000).
+11. E. Fehr, S. Gächter, Cooperation and punishment in public goods experiments. \emph{American Economic Review} \textbf{90}, 980--994 (2000).
 
 \leavevmode\hypertarget{ref-Fehr2002}{}%
-12. Fehr, E. \& Gächter, S. Altruistic punishment in humans. \emph{Nature} \textbf{415}, 137--140 (2002).
+12. E. Fehr, S. Gächter, Altruistic punishment in humans. \emph{Nature} \textbf{415}, 137--140 (2002).
 
 \leavevmode\hypertarget{ref-Henrich2006}{}%
-13. Henrich, J. \emph{et al.} Costly punishment across human societies. \emph{Science} \textbf{312}, 1767--1770 (2006).
+13. J. Henrich\emph{et al.}, Costly punishment across human societies. \emph{Science} \textbf{312}, 1767--1770 (2006).
 
 \leavevmode\hypertarget{ref-Nikiforakis2008a}{}%
-14. Nikiforakis, N. \& Normann, H.-T. A comparative statics analysis of punishment in public-good experiments. \emph{Experimental Economics} \textbf{11}, 358--369 (2008).
+14. N. Nikiforakis, H.-T. Normann, A comparative statics analysis of punishment in public-good experiments. \emph{Experimental Economics} \textbf{11}, 358--369 (2008).
 
 \leavevmode\hypertarget{ref-Raihani2012a}{}%
-15. Raihani, N. J., Thornton, A. \& Bshary, R. Punishment and cooperation in nature. \emph{Trends in Ecology \& Evolution} \textbf{27}, 288--295 (2012).
+15. N. J. Raihani, A. Thornton, R. Bshary, Punishment and cooperation in nature. \emph{Trends in Ecology \& Evolution} \textbf{27}, 288--295 (2012).
 
 \leavevmode\hypertarget{ref-DeQuervain2004}{}%
-16. de Quervain, D. J.-F. \emph{et al.} The neural basis of altruistic punishment. \emph{Science} \textbf{305}, 1254--1258 (2004).
+16. D. J.-F. de Quervain\emph{et al.}, The neural basis of altruistic punishment. \emph{Science} \textbf{305}, 1254--1258 (2004).
 
 \leavevmode\hypertarget{ref-Herrmann2008}{}%
-17. Herrmann, B., Thöni, C. \& Gächter, S. Antisocial punishment across societies. \emph{Science} \textbf{319}, 1362--1367 (2008).
+17. B. Herrmann, C. Thöni, S. Gächter, Antisocial punishment across societies. \emph{Science} \textbf{319}, 1362--1367 (2008).
 
 \leavevmode\hypertarget{ref-Camerer2003}{}%
-18. Camerer, C. F. \emph{Behavioral game theory: Experiments in strategic interaction}. (Russell Sage Foundation, 2003).
+18. C. F. Camerer, \emph{Behavioral game theory: Experiments in strategic interaction} (Russell Sage Foundation, 2003).
 
 \leavevmode\hypertarget{ref-Bowles2013}{}%
-19. Bowles, S. \& Gintis, H. \emph{A cooperative species: Human reciprocity and its evolution}. (Princeton University Press, 2013).
+19. S. Bowles, H. Gintis, \emph{A cooperative species: Human reciprocity and its evolution} (Princeton University Press, 2013).
 
 \leavevmode\hypertarget{ref-Boyd2003}{}%
-20. Boyd, R., Gintis, H., Bowles, S. \& Richerson, P. J. The evolution of altruistic punishment. \emph{Proceedings of the National Academy of Sciences} \textbf{100}, 3531--3535 (2003).
+20. R. Boyd, H. Gintis, S. Bowles, P. J. Richerson, The evolution of altruistic punishment. \emph{Proceedings of the National Academy of Sciences} \textbf{100}, 3531--3535 (2003).
 
 \leavevmode\hypertarget{ref-Chudek2011}{}%
-21. Chudek, M. \& Henrich, J. Culture-gene coevolution, norm-psychology and the emergence of human prosociality. \emph{Trends in Cognitive Sciences} \textbf{15}, 218--226 (2011).
+21. M. Chudek, J. Henrich, Culture-gene coevolution, norm-psychology and the emergence of human prosociality. \emph{Trends in Cognitive Sciences} \textbf{15}, 218--226 (2011).
 
 \leavevmode\hypertarget{ref-Henrich2017}{}%
-22. Henrich, J. \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter}. (Princeton University Press, 2017).
+22. J. Henrich, \emph{The secret of our success: How culture is driving human evolution, domesticating our species, and making us smarter} (Princeton University Press, 2017).
 
 \leavevmode\hypertarget{ref-Delton2017}{}%
-23. Delton, A. W. \& Krasnow, M. M. The psychology of deterrence explains why group membership matters for third-party punishment. \emph{Evolution and Human Behavior} \textbf{38}, 734--743 (2017).
+23. A. W. Delton, M. M. Krasnow, The psychology of deterrence explains why group membership matters for third-party punishment. \emph{Evolution and Human Behavior} \textbf{38}, 734--743 (2017).
 
 \leavevmode\hypertarget{ref-Fehr2018}{}%
-24. Fehr, E. \& Schurtenberger, I. Normative foundations of human cooperation. \emph{Nature Human Behaviour} \textbf{2}, 458--468 (2018).
+24. E. Fehr, I. Schurtenberger, Normative foundations of human cooperation. \emph{Nature Human Behaviour} \textbf{2}, 458--468 (2018).
 
 \leavevmode\hypertarget{ref-Mathew2011}{}%
-25. Mathew, S. \& Boyd, R. Punishment sustains large-scale cooperation in prestate warfare. \emph{Proceedings of the National Academy of Sciences} \textbf{108}, 11375--11380 (2011).
+25. S. Mathew, R. Boyd, Punishment sustains large-scale cooperation in prestate warfare. \emph{Proceedings of the National Academy of Sciences} \textbf{108}, 11375--11380 (2011).
 
 \leavevmode\hypertarget{ref-Mathew2014}{}%
-26. Mathew, S. \& Boyd, R. The cost of cowardice: Punitive sentiments towards free riders in Turkana raids. \emph{Evolution and Human Behavior} \textbf{35}, 58--64 (2014).
+26. S. Mathew, R. Boyd, The cost of cowardice: Punitive sentiments towards free riders in Turkana raids. \emph{Evolution and Human Behavior} \textbf{35}, 58--64 (2014).
 
 \leavevmode\hypertarget{ref-Richerson2016}{}%
-27.
-Richerson, P.
-\emph{et al.} Cultural group selection plays an essential role in explaining human cooperation: A sketch of the evidence.
-\emph{Behavioral and Brain Sciences} \textbf{39}, e30 (2016).
+27. P. Richerson\emph{et al.}, Cultural group selection plays an essential role in explaining human cooperation: A sketch of the evidence. \emph{Behavioral and Brain Sciences} \textbf{39}, e30 (2016).
 
@@
 28.
-Carlsmith, K.
-M., Darley, J.
+K.
 M.
-\& Robinson, P.
+Carlsmith, J.
+M.
+Darley, P.
 H.
-Why do we punish?
+Robinson, Why do we punish?
 Deterrence and just deserts as motives for punishment.
@@
 \leavevmode\hypertarget{ref-Raihani2012b}{}%
-29. Raihani, N. J. \& McAuliffe, K. Human punishment is motivated by inequity aversion, not a desire for reciprocity. \emph{Biology Letters} \textbf{8}, 802--804 (2012).
+29. N. J. Raihani, K. McAuliffe, Human punishment is motivated by inequity aversion, not a desire for reciprocity. \emph{Biology Letters} \textbf{8}, 802--804 (2012).
 
 \leavevmode\hypertarget{ref-Dawes2007}{}%
-30. Dawes, C. T., Fowler, J. H., Johnson, T., McElreath, R. \& Smirnov, O. Egalitarian motives in humans. \emph{Nature} \textbf{446}, 794--796 (2007).
+30. C. T. Dawes, J. H. Fowler, T. Johnson, R. McElreath, O. Smirnov, Egalitarian motives in humans. \emph{Nature} \textbf{446}, 794--796 (2007).
 
 \leavevmode\hypertarget{ref-Bone2015}{}%
-31. Bone, J. E. \& Raihani, N. J. Human punishment is motivated by both a desire for revenge and a desire for equality. \emph{Evolution and Human Behavior} \textbf{36}, 323--330 (2015).
+31. J. E. Bone, N. J. Raihani, Human punishment is motivated by both a desire for revenge and a desire for equality. \emph{Evolution and Human Behavior} \textbf{36}, 323--330 (2015).
 
 \leavevmode\hypertarget{ref-Walker2004}{}%
-32. Walker, J. M. \& Halloran, M. A. Rewards and sanctions and the provision of public goods in one-shot settings. \emph{Experimental Economics} \textbf{7}, 235--247 (2004).
+32. J. M. Walker, M. A. Halloran, Rewards and sanctions and the provision of public goods in one-shot settings. \emph{Experimental Economics} \textbf{7}, 235--247 (2004).
 
 \leavevmode\hypertarget{ref-Barclay2016}{}%
-33. Barclay, P. \& Raihani, N. Partner choice versus punishment in human prisoner's dilemmas. \emph{Evolution and Human Behavior} \textbf{37}, 263--271 (2016).
+33. P. Barclay, N. Raihani, Partner choice versus punishment in human prisoner's dilemmas. \emph{Evolution and Human Behavior} \textbf{37}, 263--271 (2016).
 
 \leavevmode\hypertarget{ref-Crockett2014}{}%
-34. Crockett, M. J., Özdemir, Y. \& Fehr, E. The value of vengeance and the demand for deterrence. \emph{Journal of Experimental Psychology: General} \textbf{143}, 2279--2286 (2014).
+34. M. J. Crockett, Y. Özdemir, E. Fehr, The value of vengeance and the demand for deterrence. \emph{Journal of Experimental Psychology: General} \textbf{143}, 2279--2286 (2014).
 
 \leavevmode\hypertarget{ref-Sylwester2013}{}%
-35. Sylwester, K., Herrmann, B. \& Bryson, J. J. Homo homini lupus? Explaining antisocial punishment. \emph{Journal of Neuroscience, Psychology, and Economics} \textbf{6}, 167--188 (2013).
+35. K. Sylwester, B. Herrmann, J. J. Bryson, Homo homini lupus? Explaining antisocial punishment. \emph{Journal of Neuroscience, Psychology, and Economics} \textbf{6}, 167--188 (2013).
 
 \leavevmode\hypertarget{ref-Fehr2004}{}%
-36. Fehr, E. \& Fischbacher, U. Third-party punishment and social norms. \emph{Evolution and Human Behavior} \textbf{25}, 63--87 (2004).
+36. E. Fehr, U. Fischbacher, Third-party punishment and social norms. \emph{Evolution and Human Behavior} \textbf{25}, 63--87 (2004).
 
 \leavevmode\hypertarget{ref-Johnson2009}{}%
-37. Johnson, T., Dawes, C. T., Fowler, J. H., McElreath, R. \& Smirnov, O. The role of egalitarian motives in altruistic punishment. \emph{Economics Letters} \textbf{102}, 192--194 (2009).
+37. T. Johnson, C. T. Dawes, J. H. Fowler, R. McElreath, O. Smirnov, The role of egalitarian motives in altruistic punishment. \emph{Economics Letters} \textbf{102}, 192--194 (2009).
 
 \leavevmode\hypertarget{ref-Fowler2005}{}%
-38. Fowler, J. H., Johnson, T. \& Smirnov, O. Egalitarian motive and altruistic punishment. \emph{Nature} \textbf{433}, E1 (2005).
+38. J. H. Fowler, T. Johnson, O. Smirnov, Egalitarian motive and altruistic punishment. \emph{Nature} \textbf{433}, E1 (2005).
 
 \leavevmode\hypertarget{ref-Molleman2014}{}%
-39. Molleman, L., Van den Berg, P. \& Weissing, F. J. Consistent individual differences in human social learning strategies. \emph{Nature Communications} \textbf{5}, 3570 (2014).
+39. L. Molleman, P. Van den Berg, F. J. Weissing, Consistent individual differences in human social learning strategies. \emph{Nature Communications} \textbf{5}, 3570 (2014).
 
 \leavevmode\hypertarget{ref-Thielmann2020}{}%
-40. Thielmann, I., Spadaro, G. \& Balliet, D. Personality and prosocial behavior: A theoretical framework and meta-analysis. \emph{Psychological Bulletin} \textbf{146}, 30--90 (2020).
+40. I. Thielmann, G. Spadaro, D. Balliet, Personality and prosocial behavior: A theoretical framework and meta-analysis. \emph{Psychological Bulletin} \textbf{146}, 30--90 (2020).
 
 \leavevmode\hypertarget{ref-Hofmann2018}{}%
-41. Hofmann, W., Brandt, M. J., Wisneski, D. C., Rockenbach, B. \& Skitka, L. J. Moral punishment in everyday life. \emph{Personality and Social Psychology Bulletin} \textbf{44}, 1697--1711 (2018).
+41. W. Hofmann, M. J. Brandt, D. C. Wisneski, B. Rockenbach, L. J. Skitka, Moral punishment in everyday life. \emph{Personality and Social Psychology Bulletin} \textbf{44}, 1697--1711 (2018).
 
 \leavevmode\hypertarget{ref-Nisbett1977}{}%
-42. Nisbett, R. E. \& Wilson, T. D. Telling more than we can know: Verbal reports on mental processes. \emph{Psychological Review} \textbf{84}, 231--259 (1977).
+42. R. E. Nisbett, T. D. Wilson, Telling more than we can know: Verbal reports on mental processes. \emph{Psychological Review} \textbf{84}, 231--259 (1977).
 
 \leavevmode\hypertarget{ref-Barclay2006}{}%
-43. Barclay, P. Reputational benefits for altruistic punishment. \emph{Evolution and Human Behavior} \textbf{27}, 325--344 (2006).
+43. P. Barclay, Reputational benefits for altruistic punishment. \emph{Evolution and Human Behavior} \textbf{27}, 325--344 (2006).
 
@@
 44.
-Batistoni, T., Barclay, P.
-\& Raihani, N.
+T.
+Batistoni, P.
+Barclay, N.
 J.
-Third-party punishers do not compete to be chosen as partners in an experimental game.
+Raihani, Third-party punishers do not compete to be chosen as partners in an experimental game.
 \emph{Proceedings of the Royal Society B} \textbf{289}, 20211773 (2022).
@@
 \leavevmode\hypertarget{ref-Jordan2016}{}%
-45. Jordan, J. J., Hoffman, M., Bloom, P. \& Rand, D. G. Third-party punishment as a costly signal of trustworthiness. \emph{Nature} \textbf{530}, 473--476 (2016).
+45. J. J. Jordan, M. Hoffman, P. Bloom, D. G. Rand, Third-party punishment as a costly signal of trustworthiness. \emph{Nature} \textbf{530}, 473--476 (2016).
 
 \leavevmode\hypertarget{ref-Jordan2017}{}%
-46. Jordan, J. J. \& Rand, D. G. Third-party punishment as a costly signal of high continuation probabilities in repeated games. \emph{Journal of Theoretical Biology} \textbf{421}, 189--202 (2017).
+46. J. J. Jordan, D. G. Rand, Third-party punishment as a costly signal of high continuation probabilities in repeated games. \emph{Journal of Theoretical Biology} \textbf{421}, 189--202 (2017).
 
@@
 47.
-Jordan, J.
 J.
-\& Rand, D.
+J.
+Jordan, D.
 G.
-Signaling when no one is watching: A reputation heuristics account of outrage and punishment in one-shot anonymous interactions.
+Rand, Signaling when no one is watching: A reputation heuristics account of outrage and punishment in one-shot anonymous interactions.
 \emph{Journal of Personality and Social Psychology} \textbf{118}, 57--88 (2020).
@@
 48.
-Deutchman, P., Bračič, M., Raihani, N.
+P.
+Deutchman, M.
+Bračič, N.
 J.
-\& McAuliffe, K.
-Punishment is strongly motivated by revenge and weakly motivated by inequity aversion.
+Raihani, K.
+McAuliffe, Punishment is strongly motivated by revenge and weakly motivated by inequity aversion.
 \emph{Evolution and Human Behavior} \textbf{42}, 12--20 (2021).
@@
 \leavevmode\hypertarget{ref-Marczyk2017}{}%
-49. Marczyk, J. Human punishment is not primarily motivated by inequality. \emph{PLoS ONE} \textbf{12}, e0171298 (2017).
+49. J. Marczyk, Human punishment is not primarily motivated by inequality. \emph{PLoS ONE} \textbf{12}, e0171298 (2017).
 
 \leavevmode\hypertarget{ref-Bone2016}{}%
-50. Bone, J. E., McAuliffe, K. \& Raihani, N. J. Exploring the motivations for punishment: Framing and country-level effects. \emph{PLoS One} \textbf{11}, e0159769 (2016).
+50. J. E. Bone, K. McAuliffe, N. J. Raihani, Exploring the motivations for punishment: Framing and country-level effects. \emph{PLoS One} \textbf{11}, e0159769 (2016).
 
 \leavevmode\hypertarget{ref-Pleasant2018}{}%
-51. Pleasant, A. \& Barclay, P. Why hate the good guy? Antisocial punishment of high cooperators is greater when people compete to be chosen. \emph{Psychological Science} \textbf{29}, 868--876 (2018).
+51. A. Pleasant, P. Barclay, Why hate the good guy? Antisocial punishment of high cooperators is greater when people compete to be chosen. \emph{Psychological Science} \textbf{29}, 868--876 (2018).
 
 \leavevmode\hypertarget{ref-Nikiforakis2008b}{}%
-52. Nikiforakis, N. Punishment and counter-punishment in public good games: Can we really govern ourselves? \emph{Journal of Public Economics} \textbf{92}, 91--112 (2008).
+52. N. Nikiforakis, Punishment and counter-punishment in public good games: Can we really govern ourselves? \emph{Journal of Public Economics} \textbf{92}, 91--112 (2008).
 
 \leavevmode\hypertarget{ref-Balafoutas2014}{}%
-53.
-Balafoutas, L., Nikiforakis, N.
-\& Rockenbach, B.
-Direct and indirect punishment among strangers in the field.
-\emph{Proceedings of the National Academy of Sciences} \textbf{111}, 15924--15927 (2014).
+53. L. Balafoutas, N. Nikiforakis, B. Rockenbach, Direct and indirect punishment among strangers in the field. \emph{Proceedings of the National Academy of Sciences} \textbf{111}, 15924--15927 (2014).
 
 \leavevmode\hypertarget{ref-Balafoutas2016}{}%
-54. Balafoutas, L., Nikiforakis, N. \& Rockenbach, B. Altruistic punishment does not increase with the severity of norm violations in the field. \emph{Nature Communications} \textbf{7}, 13327 (2016).
+54. L. Balafoutas, N. Nikiforakis, B. Rockenbach, Altruistic punishment does not increase with the severity of norm violations in the field. \emph{Nature Communications} \textbf{7}, 13327 (2016).
 
 \leavevmode\hypertarget{ref-Baumard2010}{}%
-55. Baumard, N. Has punishment played a role in the evolution of cooperation? A critical review. \emph{Mind \& Society} \textbf{9}, 171--192 (2010).
+55. N. Baumard, Has punishment played a role in the evolution of cooperation? A critical review. \emph{Mind \& Society} \textbf{9}, 171--192 (2010).
 
 \leavevmode\hypertarget{ref-Fitouchi2023}{}%
-56. Fitouchi, L. \& Singh, M. Punitive justice serves to restore reciprocal cooperation in three small-scale societies. \emph{Evolution and Human Behavior} \textbf{44}, 502--514 (2023).
+56. L. Fitouchi, M. Singh, Punitive justice serves to restore reciprocal cooperation in three small-scale societies. \emph{Evolution and Human Behavior} \textbf{44}, 502--514 (2023).
 
 \leavevmode\hypertarget{ref-Singh2022}{}%
-57. Singh, M. \& Garfield, Z. H. Evidence for third-party mediation but not punishment in Mentawai justice. \emph{Nature Human Behaviour} \textbf{6}, 930--940 (2022).
+57. M. Singh, Z. H. Garfield, Evidence for third-party mediation but not punishment in Mentawai justice. \emph{Nature Human Behaviour} \textbf{6}, 930--940 (2022).
 
 \leavevmode\hypertarget{ref-Molleman2019}{}%
-58. Molleman, L., Kölle, F., Starmer, C. \& Gächter, S. People prefer coordinated punishment in cooperative interactions. \emph{Nature Human Behaviour} \textbf{3}, 1145--1153 (2019).
+58. L. Molleman, F. Kölle, C. Starmer, S. Gächter, People prefer coordinated punishment in cooperative interactions. \emph{Nature Human Behaviour} \textbf{3}, 1145--1153 (2019).
 
@@
 59.
-Adler, N.
-E., Epel, E.
-S., Castellazzo, G.
-\& Ickovics, J.
+N.
+E.
+Adler, E.
+S.
+Epel, G.
+Castellazzo, J.
 R.
-Relationship of subjective and objective social status with psychological and physiological functioning: Preliminary data in healthy, White women.
+Ickovics, Relationship of subjective and objective social status with psychological and physiological functioning: Preliminary data in healthy, White women.
 \emph{Health Psychology} \textbf{19}, 586--592 (2000).
@@
 60.
-Sibley, C.
+C.
 G.
-\emph{et al.} The Mini-IPIP6: Validation and extension of a short measure of the Big-Six factors of personality in New Zealand.
+Sibley\emph{et al.}, The Mini-IPIP6: Validation and extension of a short measure of the Big-Six factors of personality in New Zealand.
 \emph{New Zealand Journal of Psychology (Online)} \textbf{40}, 142 (2011).
@@
 \leavevmode\hypertarget{ref-Murphy2011}{}%
-61. Murphy, R. O., Ackermann, K. A. \& Handgraaf, M. J. J. Measuring social value orientation. \emph{Judgment and Decision Making} \textbf{6}, 771--781 (2011).
+61. R. O. Murphy, K. A. Ackermann, M. J. J. Handgraaf, Measuring social value orientation. \emph{Judgment and Decision Making} \textbf{6}, 771--781 (2011).
 
@@
 62.
-Ho, A.
+A.
 K.
-\emph{et al.} The nature of social dominance orientation: Theorizing and measuring preferences for intergroup inequality using the new sdo7 scale.
+Ho\emph{et al.}, The nature of social dominance orientation: Theorizing and measuring preferences for intergroup inequality using the new sdo7 scale.
 \emph{Journal of Personality and Social Psychology} \textbf{109}, 1003 (2015).
@@
 \leavevmode\hypertarget{ref-Bizumic2018}{}%
-63. Bizumic, B. \& Duckitt, J. Investigating right wing authoritarianism with a Very Short Authoritarianism Scale. \emph{Journal of Social and Political Psychology} \textbf{6}, 129--150 (2018).
+63. B. Bizumic, J. Duckitt, Investigating right wing authoritarianism with a Very Short Authoritarianism Scale. \emph{Journal of Social and Political Psychology} \textbf{6}, 129--150 (2018).
 
@@
 64.
-Laurin, K., Shariff, A.
-F., Henrich, J.
-\& Kay, A.
+K.
+Laurin, A.
+F.
+Shariff, J.
+Henrich, A.
 C.
-Outsourcing punishment to God: Beliefs in divine control reduce earthly punishment.
+Kay, Outsourcing punishment to God: Beliefs in divine control reduce earthly punishment.
 \emph{Proceedings of the Royal Society B: Biological Sciences} \textbf{279}, 3272--3281 (2012).
@@
 \leavevmode\hypertarget{ref-McElreath2020}{}%
-65. McElreath, R. \emph{Statistical rethinking: A Bayesian course with examples in R and Stan, 2nd edition}. (CRC Press, 2020).
+65. R. McElreath, \emph{Statistical rethinking: A Bayesian course with examples in R and Stan, 2nd edition}, 2nd Ed. (CRC Press, 2020).
 
 \leavevmode\hypertarget{ref-Stan2020}{}%
-66. Stan Development Team. RStan: The R interface to Stan. (2020).
+66. Stan Development Team, RStan: The R interface to Stan (2020).
 
 \leavevmode\hypertarget{ref-RCoreTeam}{}%
-67. R Core Team. \emph{R: A language and environment for statistical computing}. (R Foundation for Statistical Computing, 2022).
+67. R Core Team, \emph{R: A language and environment for statistical computing} (R Foundation for Statistical Computing, 2022).
 
 \leavevmode\hypertarget{ref-Wickham2016}{}%
-68. Wickham, H. \emph{ggplot2: Elegant graphics for data analysis}. (Springer-Verlag New York, 2016).
+68. H. Wickham, \emph{ggplot2: Elegant graphics for data analysis} (Springer-Verlag New York, 2016).
 
 \leavevmode\hypertarget{ref-Wilke2020}{}%
-69. Wilke, C. O. \emph{cowplot: Streamlined plot theme and plot annotations for 'ggplot2'}. (2020).
+69. C. O. Wilke, \emph{cowplot: Streamlined plot theme and plot annotations for 'ggplot2'} (2020).
 
@@
 70.
-Landau, W.
+W.
 M.
-The targets R package: A dynamic Make-like function-oriented pipeline toolkit for reproducibility and high-performance computing.
+Landau, The targets R package: A dynamic Make-like function-oriented pipeline toolkit for reproducibility and high-performance computing.
 \emph{Journal of Open Source Software} \textbf{6}, 2959 (2021).
@@
 \leavevmode\hypertarget{ref-Aust2022}{}%
-71. Aust, F. \& Barth, M. \emph{papaja: Prepare reproducible APA journal articles with R Markdown}. (2022).
+71. F. Aust, M. Barth, \emph{papaja: Prepare reproducible APA journal articles with R Markdown} (2022).
 \end{cslreferences}

### 2025-08-01 quarto/appendix.qmd
Update manuscript and submit

# Supplementary Materials

## Supplementary Figures

::: {#suppfig-model1a suppfig-pos="H"}

```{r, fig.height=5, fig.width=7}
tar_read(plot1.1)
```

Posterior estimates of the probabilities of following different punishment 
strategies from the Bayesian latent state model fitted to the full dataset 
without pre-registered exclusions. The model assumes an implementation error 
rate of 5%. @fig-model1b in the main text shows the same result, but from a 
model fitted to the reduced dataset with pre-registered exclusions. Points 
represent posterior medians, line ranges represent 50% and 95% credible 
intervals. AI = advantageous inequity, DI = disadvantageous inequity.

:::

\newpage

::: {#suppfig-allDems2 suppfig-pos="H"}

```{r, fig.height=3, fig.width=7.5, warning=FALSE}
tar_read(plotAllDems2)
```

Posterior slopes from models including age, socio-economic status, gender, and 
student status, fitted to the subsetted dataset with pre-registered exclusions.
Each row represents a separate model. Points represent posterior medians, line 
ranges represent 95% credible intervals. AI = advantageous inequity, DI = 
disadvantageous inequity.

:::

\newpage

::: {#suppfig-allDems1 suppfig-pos="H"}

```{r, fig.height=3, fig.width=7.5, warning=FALSE}
tar_read(plotAllDems1)
```

Posterior slopes from models including age, socio-economic status, gender, and 
student status, fitted to the full dataset without pre-registered exclusions. 
Each row represents a separate model. Points represent posterior medians, line 
ranges represent 95% credible intervals. AI = advantageous inequity, DI = 
disadvantageous inequity.

:::

\newpage

::: {#suppfig-allPers1 suppfig-pos="H"}

```{r, fig.height=3.5, fig.width=7.5, warning=FALSE}
tar_read(plotAllPers1)
```

Posterior slopes from models including Big-6 personality dimensions and Social 
Value Orientation, fitted to the full dataset without pre-registered exclusions. 
Each row represents a separate model. @fig-allPers2 in the main text shows the 
same results, but from models fitted to the reduced dataset with pre-registered 
exclusions. Points represent posterior medians, line ranges represent 95% 
credible intervals. AI = advantageous inequity, DI = disadvantageous inequity.

:::

\newpage

::: {#suppfig-allPolRel1 suppfig-pos="H"}

```{r, fig.height=3.5, fig.width=7.5, warning=FALSE}
tar_read(plotAllPolRel1)
```

Posterior slopes from models including political ideology, views about social 
inequality, and religiosity, fitted to the full dataset without pre-registered 
exclusions. Each row represents a separate model aside from Social Dominance 
Orientation and Right Wing Authoritarianism, which control for one another 
within the same model. @fig-allPolRel2 in the main text shows the same results,
but from models fitted to the reduced dataset with pre-registered exclusions. 
Points represent posterior medians, line ranges represent 95% credible 
intervals. AI = advantageous inequity, DI = disadvantageous inequity.

:::

\newpage

::: {#suppfig-sliders1 suppfig-pos="H"}

```{r, fig.height=4, fig.width=7, warning=FALSE}
tar_read(plotSliders1)
```

Boxplots showing the distribution of responses to each self-report question 
about the reasons for participants' behaviour in the games. Boxplots represent 
medians and interquartile ranges.

:::

\newpage

::: {#suppfig-sliders2 suppfig-pos="H"}

```{r, fig.height=4, fig.width=7, warning=FALSE}
tar_read(plotSliders2)
```

Boxplots showing the distribution of responses to each self-report question 
about the reasons for participants' behaviour in the games, presented as 
deviations from participants' average rating across all questions. Boxplots 
represent medians and interquartile ranges.

:::

\newpage

::: {#suppfig-allSliders1 suppfig-pos="H"}

```{r, fig.height=4, fig.width=7.5, warning=FALSE}
tar_read(plotAllSliders1)
```

Posterior slopes from models including self-reported strategy usage, fitted to 
the full dataset without pre-registered exclusions. Each row represents a 
separate model. Highlighted estimates represent combinations where the 
self-report slider matched the behavioural strategy. @fig-allSliders2 in the 
main text shows the same results, but from models fitted to the reduced dataset
with pre-registered exclusions. Points represent posterior medians, line ranges
represent 95% credible intervals. AI = advantageous inequity, DI = 
disadvantageous inequity.

:::

\newpage

::: {#suppfig-spline suppfig-pos="H"}

```{r, fig.height=3.5, fig.width=6, warning=FALSE}
tar_read(plotSpline)
```

The relationships between strategy frequencies and the overall costs of 
strategies across all twelve punishment decisions, in both countries. Each point
is a unique strategy that appears in our dataset at least once (for ease of 
presentation, the "never punish" strategy is excluded). Lines and shaded areas
represent posterior predictions from splines fitted to the trend in each country 
separately.

:::

\newpage

::: {#suppfig-sampleUK suppfig-pos="H"}

```{r, fig.height=5, fig.width=7, warning=FALSE}
tar_read(plotSample_UK)
```

Sample characteristics in the United Kingdom.

:::

\newpage

::: {#suppfig-sampleUS suppfig-pos="H"}

```{r, fig.height=5, fig.width=7, warning=FALSE}
tar_read(plotSample_US)
```

Sample characteristics in the United States.

:::

\newpage

::: {#suppfig-model5 suppfig-pos="H"}

```{r, fig.height=5, fig.width=7, warning=FALSE}
tar_read(plot5)
```

Posterior estimates of the probabilities of following different punishment 
strategies from the Bayesian latent state model that estimates the 
implementation error rate as a free parameter. The model estimated the 
implementation error rate to be `r papaja::printnum(median(error))`
(95% CI [`r papaja::printnum(quantile(error, 0.025))`
`r papaja::printnum(quantile(error, 0.975))`]). Points represent posterior 
medians, line ranges represent 50% and 95% credible intervals. AI = advantageous 
inequity, DI = disadvantageous inequity.

:::

\newpage

::: {#suppfig-trace suppfig-pos="H"}

```{r, fig.height=6, fig.width=7, warning=FALSE}
tar_read(plotTrace1.2)
```

MCMC trace plots for parameter values from the Bayesian latent state model 
fitted to data with exclusions. These trace plots suggest that the different
chains mixed well and the model converged normally.

:::

\newpage

::: {#suppfig-sim suppfig-pos="H"}

```{r, fig.height=5, fig.width=6, warning=FALSE}
tar_read(plotSim1)
```

Results of Bayesian latent state model fitted to simulated data (n = 100) with 
known strategy frequencies in the population. Blue points represent known 
strategy frequencies, grey densities represent posterior estimates of strategy 
frequencies. AI = advantageous inequity, DI = disadvantageous inequity.

:::

\newpage

## Supplementary Tables

::: {#supptbl-patterns supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(tablePatterns) %>%
  flextable() %>%
   set_header_labels(
    Pattern = "Pattern",
    Explanation = "Explanation",
    a = "N",
    b = "Prop",
    c = "N",
    d = "Prop"
  ) %>%
  add_header_row(
    values = c(
      "",
      "United Kingdom\n(N = 1014)",
      "United States\n(N = 996)"
    ),
    colwidths = c(2, 2, 2)
  ) %>%
  theme_apa() %>%
  line_spacing(part = "all") %>%
  align(i = NULL, j = 1:2, align = "left", part = "all") %>%
  autofit() %>%
  fontsize(size = 9.5, part = "all")
```

Counts and proportions of the 25 most common patterns of punitive behaviour 
across all twelve decisions, split by country. Binary strings represent 
punishment (1) or no punishment (0) in each decision, aligning with the order of
game decision columns in @tbl-strategies.

:::

\newpage

::: {#supptbl-sliderWordings supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(tableSliderWordings) %>%
  flextable() %>%
  theme_apa() %>%
  autofit() %>%
  align(i = NULL, j = 2, align = "left", part = "all") %>%
  line_spacing(part = "all")
```

Wordings for 11 self-report slider questions asking participants to report the 
reasons for their behaviour in the six games. Participants were prompted with 
the following text: "We would now like you to answer a few questions about your
main motivation in the games. Please answer truthfully - there is no right or 
wrong answer and your first answer is probably best. Please rate the extent to
which the following statements apply to your decisions to reduce or not to
reduce other players' bonuses in the games."

:::

\newpage

::: {#supptbl-surveyWordings supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(tableSurveyWordings) %>%
  flextable() %>%
  theme_apa() %>%
  autofit() %>%
  align(align = "left", part = "all") %>%
  line_spacing(part = "all")
```

Wordings for survey questions in the study.

:::

\newpage

::: {#supptbl-comp supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(tableComp) %>%
  flextable() %>%
  theme_apa() %>%
  autofit() %>%
  align(j = 1, align = "left", part = "all")
```

Proportions of correct answers to comprehension questions for all six economic 
games, split by country.

:::

### 2025-08-01 quarto/manuscript.qmd
Update manuscript and submit

--- 2024-08-05 manuscript.Rmd
@@
 ---
-title             : "Why do people punish? Evidence for a range of strategic concerns"
-shorttitle        : "Punishment strategies"
-
+title: "Why Do People Punish? Evidence for a Range of Strategic Concerns"
+shorttitle: "Punishment strategies"
 author:
-  - name          : "Scott Claessens"
-    affiliation   : "*1"
-  - name          : "Quentin D. Atkinson"
-    affiliation   : "1"
-  - name          : "Nichola Raihani"
-    affiliation   : "1,2"
-
-affiliation:
-  - id            : "1"
-    institution   : "\\footnotesize School of Psychology, University of Auckland, Auckland, New Zealand"
-  - id            : "2"
-    institution   : "\\footnotesize Department of Experimental Psychology, University College London, London, United Kingdom"
-
-abstract: |
- Costly punishment is thought to be a key mechanism sustaining human cooperation.
+  - name: Scott Claessens
+    corresponding: true
+    orcid: 0000-0002-3562-6981
+    email: scott.claessens@gmail.com
+    affiliations:
+      - name: University of Kent
+        department: School of Psychology
+        city: Canterbury
+        country: United Kingdom
+      - name: University of Auckland
+        department: School of Psychology
+        city: Auckland
+        country: New Zealand
+    role:
+      - conceptualization
+      - data curation
+      - formal analysis
+      - investigation
+      - methodology
+      - visualization
+      - writing
+      - editing
+  - name: Quentin D Atkinson
+    corresponding: false
+    orcid: 0000-0002-8499-7535
+    affiliations:
+      - name: University of Auckland
+        department: School of Psychology
+        city: Auckland
+        country: New Zealand
+    role:
+      - conceptualization
+      - funding acquisition
+      - methodology
+      - supervision
+      - writing
+      - editing
+  - name: Nichola J Raihani
+    corresponding: false
+    orcid: 0000-0003-2339-9889
+    affiliations:
+      - name: University of Auckland
+        department: School of Psychology
+        city: Auckland
+        country: New Zealand
+      - name: University College London
+        department: Department of Experimental Psychology
+        city: London
+        country: United Kingdom
+    role:
+      - conceptualization
+      - funding acquisition
+      - investigation
+      - methodology
+      - supervision
+      - writing
+      - editing
+author-note:
+  status-changes: 
+    affiliation-change: ~
+    deceased: ~
+  disclosures:
+    study-registration: "This manuscript was published as a pre-print on PsyArXiv under a CC BY 4.0 license: <https://doi.org/10.31234/osf.io/ys6rm>"
+    data-sharing: Word count (excluding abstract, figures, tables, and references) = 6065 words.
+    related-report: ~
+    conflict-of-interest: ~
+    financial-support: ~
+    gratitude: ~
+    authorship-agreements: ~
+abstract: "Costly punishment is thought to be a key mechanism sustaining human cooperation.
 However, the motives for punitive behaviour remain unclear.
@@
 Self-reports of behaviour in the games suggested that people have some insight into their punishment strategy.
-These findings highlight the multipurpose nature of human punishment and show how the different motives underpinning punishment decisions are linked with core character traits.
-
-bibliography      : ["r-references.bib"]
-floatsintext      : yes
-figurelist        : no
-figsintext        : yes
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-numbersections    : false
-documentclass     : "apa6"
-classoption       : "man, donotrepeattitle"
-output            : papaja::apa6_pdf
-header-includes: 
-  - \usepackage{array}
-  - \usepackage{caption}
-  - \usepackage{graphicx}
-  - \usepackage{siunitx}
-  - \usepackage[normalem]{ulem}
-  - \usepackage{colortbl}
-  - \usepackage{multirow}
-  - \usepackage{hhline}
-  - \usepackage{calc}
-  - \usepackage{tabularx}
-  - \usepackage{threeparttable}
-  - \usepackage{wrapfig}
-  - \usepackage{adjustbox}
-  - \usepackage{hyperref}
-  - \usepackage{setspace}
-  - \usepackage{amsmath}
-  - \usepackage{xcolor}
-  - \usepackage{makecell}
-  - \raggedbottom
-  - \AtBeginEnvironment{tabular}{\singlespacing}
-  - \AtBeginEnvironment{lltable}{\singlespacing}
-  - \AtBeginEnvironment{tablenotes}{\doublespacing}
-  - \captionsetup[table]{font={stretch=1,small}}
-  - \captionsetup[figure]{font={stretch=1,small}}
-  - \nolinenumbers
-  - \note{\raggedright * Correspondence concerning this article should be addressed to Scott Claessens, Level 2, Building 302, 23 Symonds Street, Auckland, New Zealand.
-\text{E-mail:} \href{scott.claessens@gmail.com}{\nolinkurl{scott.claessens@gmail.com}} \par This manuscript has not yet been peer-reviewed.
-\par This manuscript was published as a pre-print on PsyArXiv under a CC BY 4.0 \text{license:} \url{https://doi.org/10.31234/osf.io/ys6rm} \par \text{Classification:} Social Sciences; Psychological and Cognitive Sciences \par \text{Keywords:} punishment; cooperation; economic games}
-appendix          : "appendix.Rmd"
-csl               : pnas.csl
+These findings highlight the multipurpose nature of human punishment and show how the different motives underpinning punishment decisions are linked with core character traits."
+keywords: [punishment, cooperation, economic games]
+bibliography: bibliography.bib
+engine: knitr
+format:
+  apaquarto-docx:
+    fig-dpi: 300
+execute:
+  echo: false
+  warning: false
+  error: false
+crossref:
+  custom:
+    - kind: float
+      key: suppfig
+      latex-env: suppfig
+      reference-prefix: Supplementary Figure
+      space-before-numbering: true
+      latex-list-of-description: Supplementary Figure
+      caption-location: bottom
+    - kind: float
+      key: supptbl
+      latex-env: supptbl
+      reference-prefix: Supplementary Table
+      space-before-numbering: true
+      latex-list-of-description: Supplementary Table
+      caption-location: top
+floatsintext: true
+csl: nature.csl
 ---
 
-```{r options, echo=FALSE}
-opts_chunk$set(echo = FALSE)
+```{r}
+library(flextable)
+library(knitr)
+library(targets)
+library(tidyverse)
 ```
-
-\linenumbers
-
-# Introduction
 
@@
 whereby individuals harm others at a personal cost [@CluttonBrock1995], 
-ostensibly encouraging cooperative behaviour from the target, or bystanders 
-[@DosSantos2011;@DosSantos2013;@Raihani2015], in the future. Punishment offers a 
-route to maintaining or increasing cooperation by changing the payoff structure 
-of social interactions such that it no longer pays to cheat or exploit social 
-partners [@CluttonBrock1995;@Ostrom1990]. Yet, despite its theoretical 
-importance, the question of why people choose to punish others is still hotly 
-contested [@Raihani2019]. In this study, we use a battery of economic games to 
-disentangle the different motives underpinning punishment and explore how these 
-motives vary across individuals.
+ostensibly encouraging cooperative behaviour from the target or bystanders 
+[@DosSantos2011; @DosSantos2013; @Raihani2015]. Punishment offers a route to 
+maintaining or increasing cooperation by changing the payoff structure of social 
+interactions such that it no longer pays to cheat or exploit social partners 
+[@CluttonBrock1995; @Ostrom1990]. Yet, despite its theoretical importance, the 
+question of why people choose to punish others is still hotly contested 
+[@Raihani2019]. In this study, we use a battery of economic games to disentangle
+the different motives underpinning punishment and explore how these motives vary 
+across individuals.
 
-In humans, many studies of punishment have been carried out in laboratory 
-settings using economic games [@Balliet2011;@Balliet2013;@Chaudhuri2011;
-@Dreber2008;@Fehr2000;@Fehr2002;@Henrich2006;@Nikiforakis2008a;@Raihani2012a;
-@Raihani2019]. In these games, participants are usually given a sum of money 
-that they can use to invest in collective action or to help others. 
-Alternatively, participants can 'cheat' by keeping the money for themselves or 
-by exploiting the contributions of others. Punishment is introduced into such 
-games by giving participants the option to pay a small 'fee' to impose a greater 
-'fine' on their co-players. Several lines of experimental evidence indicate that 
-people use this punishment option [@Henrich2006], that they enjoy 
-punishing [@DeQuervain2004], and that they frequently, though not 
-always [@Herrmann2008], punish cheating or exploitative 
-co-players [@Fehr2000;@Fehr2002].
+In humans, punishment is often studied in laboratory settings using economic 
+games [@Balliet2011; @Balliet2013; @Chaudhuri2011; @Dreber2008; @Fehr2000; 
+@Fehr2002; @Henrich2006; @Nikiforakis2008a; @Raihani2012a; @Raihani2019]. In 
+some of these games, participants are given a sum of money that they can use to 
+invest in collective action or to help others; conversely, they can 'cheat' by 
+keeping the money for themselves or by exploiting the contributions of others. 
+In other games, participants are endowed with sums of money and can then 'cheat'
+by destroying a partner's payoffs or stealing from them [@Abbink2009; 
+@Raihani2012b]. Punishment is introduced into such games by giving participants 
+the option to pay a small 'fee' to impose a greater 'fine' on their co-players. 
+Several lines of experimental evidence indicate that people use this punishment 
+option [@Henrich2006], that they enjoy punishing [@DeQuervain2004], and that 
+they frequently, though not always [@Herrmann2008], punish those who cheat,
+free-ride, or steal [@Fehr2000; @Fehr2002].
 
@@
 more in games where punishment is possible compared to those where it is not
-[@Balliet2011;@Balliet2013;@Raihani2019]. The effect that the threat of 
-punishment has on cooperation is also evident in the higher contributions 
-typically observed in the Ultimatum Game (where punishment is possible) compared 
-to the structurally-similar Dictator Game (where it is not) [@Camerer2003]. This 
-typical cooperation-enhancing effect of punishment has also been observed across 
+[@Balliet2011; @Balliet2013; @Raihani2019]. Similarly, people make higher offers 
+in the Ultimatum Game (where punishment is possible) compared to the 
+structurally-similar Dictator Game (where it is not) [@Camerer2003]. This 
+cooperation-enhancing effect of punishment has also been observed across 
 societies [@Balliet2013], leading some to suggest that costly punishment has 
 played a key role in the cultural evolution of cooperation in humans
-[@Bowles2013;@Boyd2003;@Chudek2011;@Henrich2017].
+[@Bowles2013; @Boyd2003; @Chudek2011; @Henrich2017].
 
 Nevertheless, it remains unclear whether individuals playing economic games use
-punishment as a behaviour-change tool to enforce cooperation or as a means to 
-achieve other ends. Some have argued that punishment is primarily used to shape 
-future behaviour, either to deter personal harm [@Delton2017;@Dreber2008;
-@DosSantos2013] or to uphold normative standards of cooperative behaviour 
-[@Boyd2003;@Chudek2011;@Fehr2018;@Mathew2011;@Mathew2014;@Richerson2016]. 
-But while the _threat_ of punishment can have a cooperation-enhancing effect, 
-the _enactment_ of this punishment does not consistently deter targets from 
-cheating in the future [@Raihani2019]. This calls into question whether 
-punishment primarily operates as a behaviour-change tool or whether it is used 
-to achieve other goals.
+punishment as a behaviour-change tool to enforce cooperation or to achieve other 
+ends. Some have argued that punishment is primarily used to shape behaviour, 
+either to deter personal harm [@Delton2017; @Dreber2008; @DosSantos2013] or to 
+uphold normative standards of cooperative behaviour [@Boyd2003; @Chudek2011; 
+@Fehr2018; @Mathew2011; @Mathew2014; @Richerson2016]. But while the _threat_ of 
+punishment can have a cooperation-enhancing effect, the _enactment_ of this 
+punishment does not consistently deter targets from cheating in the future 
+[@Raihani2019]. This calls into question whether punishment primarily operates 
+as a behaviour-change tool or whether it is used to achieve other goals.
 
-Beyond behaviour shaping concerns, there are a host of other reasons why people 
-may want to punish in economic games. Punishers might be motivated by a desire 
-for retribution rather than deterrence, punishing in proportion to the amount 
-of harm that was personally caused [@Carlsmith2002]. Punishment might be driven 
-by concerns about relative payoffs, such as disadvantageous inequity aversion 
-(i.e., avoiding having less than others) [@Raihani2012b;@Raihani2019] and/or 
-general egalitarian preferences (i.e., wanting all participants to receive the 
-same payoffs) [@Dawes2007]. Such concerns about relative payoffs may be activated 
+Beyond behaviour shaping concerns, there are other reasons why people punish in 
+economic games. Punishers might be motivated by a desire for retribution rather 
+than deterrence, punishing in proportion to the amount of harm that was 
+personally caused [@Carlsmith2002]. Punishment might be driven by concerns about 
+relative payoffs, such as disadvantageous inequity aversion (i.e., avoiding 
+having less than others) [@Raihani2012b; @Raihani2019] and/or general 
+egalitarian preferences (i.e., wanting all participants to receive the same 
+payoffs) [@Dawes2007]. Such concerns about relative payoffs may be activated 
 when participants earn less than cheats in economic games or when there are 
@@
 people use punishment in seemingly disparate ways: punishing when no behaviour 
-change is possible, such as in one-shot games [@Bone2015;@Henrich2006;
-@Raihani2012b;@Walker2004], on the very last round of repeated games
-[@Barclay2016], or in games where the target never learns about the 
-punishment [@Crockett2014]; punishing those who did not cheat or who 
-over-contributed to collective action (antisocial punishment) 
-[@Herrmann2008;@Sylwester2013]; punishing in scenarios where they were not 
-personally harmed (third-party punishment) [@Fehr2004]; and punishing in 
-scenarios where disparities in payoffs did not arise from participants' 
-actions [@Dawes2007;@Johnson2009;@Fowler2005].
+change is possible, such as in one-shot games [@Bone2015; @Henrich2006;
+@Raihani2012b; @Walker2004], on the very last round of repeated games 
+[@Barclay2016], or in games where the target never learns about the punishment 
+[@Crockett2014]; punishing those who did not cheat or who over-contributed to 
+collective action (sometimes called 'antisocial punishment') [@Herrmann2008; 
+@Sylwester2013]; punishing in scenarios where they were not personally harmed 
+(third-party punishment) [@Fehr2004]; and punishing in scenarios where 
+disparities in payoffs did not arise from participants' actions [@Dawes2007; 
+@Johnson2009; @Fowler2005].
 
@@
 punitive behaviour, yet they feel compelled to enact it anyway
-[@Carlsmith2002;@Nisbett1977].
+[@Carlsmith2002; @Nisbett1977].
 
@@
 so, what other characteristics (personality, social preferences, political 
-orientation, religious views) predict the use of different punishment strategies.
-Table \@ref(tab:tableStrategies) summarises the potential functions for costly 
+orientation, religious views) predict the use of different punishment 
+strategies. @tbl-strategies summarises the potential functions for costly 
 punishment in the economic games that we considered, and the behavioural 
-strategies they predict. Note that Table \@ref(tab:tableStrategies) is not 
-an exhaustive list of all possible punishment strategies: we do not include 
-reputational functions of punishment in this table, such as signalling 
-trustworthiness [@Barclay2006;@Batistoni2022;@Jordan2016;@Jordan2017;@Raihani2015], 
-because our focus is on punishment strategies in anonymous economic games without 
-reputational incentives (but see ref[@Jordan2020]).
+strategies they predict. Note that @tbl-strategies is not an exhaustive list of 
+all possible punishment strategies: we do not include reputational functions of 
+punishment in this table, such as signalling trustworthiness [@Barclay2006; 
+@Batistoni2022; @Jordan2016; @Jordan2017; @Raihani2015], because our focus is on 
+punishment strategies in anonymous economic games without reputational 
+incentives (but see ref [@Jordan2020]).
 
-Building on previous designs [@Bone2015;@Deutchman2021;@Marczyk2017;
-@Raihani2012b], we employ a suite of one-shot economic games where individuals 
-are given the opportunity to punish targets at a personal cost (Figure 
-\@ref(fig:plotGames)). In each game, targets either steal from another 
-individual or do nothing. Representative samples of participants from the United 
-Kingdom (*n* = `r sum(tar_read(dExc)$Country == "United Kingdom")`) and the 
-United States (*n* = `r sum(tar_read(dExc)$Country == "United States")`) 
-completed all six games on the online platform Prolific. We carefully designed 
-the suite of games to tease apart the proposed punishment strategies in Table 
-\@ref(tab:tableStrategies), such that each strategy predicts a different pattern 
-of behaviour across all the games (see Methods for more detail about the six 
-games). We use the resulting behavioural patterns to discern which punishment 
-strategy participants are employing. We then combine these behavioural patterns 
-with data on demographics, personality, social preferences, political ideology,
-religiosity, and self-reported strategy usage.
+Building on previous designs that have used one-shot economic games to explore 
+behaviour-shaping and egalitarian motives for punishment [@Bone2015; 
+@Crockett2014; @Deutchman2021; @Marczyk2017; @Raihani2012b], we employ a suite 
+of one-shot economic games where individuals are given the opportunity to punish 
+targets at a personal cost (@fig-games). In each game, targets either steal from 
+another individual or do nothing. Representative samples of participants from 
+the United Kingdom (*n* = `r sum(tar_read(dExc)$Country == "United Kingdom")`) 
+and the United States (*n* = `r sum(tar_read(dExc)$Country == "United States")`) 
+completed all six games on the online platform Prolific. We designed the suite 
+of games to tease apart the proposed punishment strategies in @tbl-strategies, 
+such that each strategy predicts a different pattern of behaviour across all the 
+games (see Methods for more detail about the six games). We use the resulting 
+behavioural patterns to discern which punishment strategy participants are 
+employing. We then combine these behavioural patterns with data on demographics, 
+personality, social preferences, political ideology, religiosity, and self-reported strategy usage.
 
-(ref:plotGamesCaption) *Visual summary of the six economic games.* In the
-games, Player 2 either steals £0.20 from Player 1 (the focal player) or does 
-nothing. Player 1 is then given the option to punish by paying a certain amount 
-of money to remove money from Player 2 (this money is destroyed). The six games
-are variants on this general setup, creating situations where (A) Player 2 is
-still worse off after stealing, (B) Player 2 creates equality by stealing, (C) 
-the computer "decides" whether Player 2 steals, (D) the fee-fine ratio is 1:1, 
-(E) Player 2 is better off after stealing, and (F) Player 2 steals instead from 
-a third-party.
+![Visual summary of the six economic games.](images/games.png){
+  #fig-games
+  width="80%"
+  apa-note="In the games, Player 2 either steals £0.20 from Player 1 (the focal player) or does nothing.
+Player 1 is then given the option to punish by paying a certain amount of money to remove money from Player 2 (this money is destroyed).
+The six games are variants on this general setup, creating situations where (A) Player 2 is still worse off after stealing, (B) Player 2 creates equality by stealing, (C) the computer 'decides' whether Player 2 steals, (D) the fee-fine ratio is 1:1, (E) Player 2 is better off after stealing, and (F) Player 2 steals instead from a third-party."
+  }
 
-```{r plotGames, out.width="100%", fig.cap="(ref:plotGamesCaption)"}
-include_graphics("images/games.png")
+::: {.landscape}
+
+```{r}
+#| label: tbl-strategies
+#| tbl-cap: Summary of the different functions for punishment and the behavioural strategies they predict.
+#| apa-note: Games A-F are the games employed in the current study (see Methods for more details).
+In each of the six games, participants are given the opportunity to punish players who 'steal' and those who do not, meaning that participants make twelve punishment decisions in total.
+Each behavioural strategy implies a unique pattern of punishment across all decisions (see Methods for detailed explanation of strategies).
+Ticks reflect decisions to punish, crosses reflect decisions to not punish.
+In column headers, payoffs at the first stage (above) and the second stage (below) are denoted as P1-P2 (or P2-P3 [P1] for Game F) where participants take the role of P1 and P2 is the target of punishment.
+AI = advantageous inequity, DI = disadvantageous inequity.
+
+tar_read(tableStrategies) %>%
+  flextable() %>%
+  set_header_labels(
+    Function = "Function",
+    `Behavioural strategy` = "Behavioural strategy",
+    a = "Steal \n 50-30",
+    b = "No steal \n 70-10",
+    c = "Steal \n 50-50",
+    d = "No steal \n 70-30",
+    e = "Steal \n 50-50",
+    f = "No steal \n 70-30",
+    g = "Steal \n 50-50",
+    h = "No steal \n 70-30",
+    i = "Steal \n 50-70",
+    j = "No steal \n 70-50",
+    k = "Steal \n 50-90 [100]",
+    l = "No steal \n 70-70 [100]"
+  ) %>%
+  add_header_row(
+    values = c(
+      "",
+      "Game A \n (AI) \n 70-10",
+      "Game B \n (Equal) \n 70-30",
+      "Game C \n (Computer) \n 70-30",
+      "Game D \n (1:1 Fee Fine) \n 70-30",
+      "Game E \n (DI) \n 70-50",
+      "Game F \n (Third Party) \n 70-70 [100]"
+    ),
+    colwidths = rep(2, 7)
+  ) %>%
+  theme_apa() %>%
+  width(width = c(1, 2, rep(0.6, 12))) %>%
+  fontsize(size = 9, part = "all") %>%
+  line_spacing(part = "all") %>%
+  align(i = NULL, j = 1:2, align = "left", part = "all")
 ```
 
-(ref:tableStrategiesCaption) \scriptsize Summary of the different functions for 
-punishment and the behavioural strategies they predict. *Games A-F are 
-the games employed in the current study (see Methods for more details). In each 
-of the six games, participants are given the opportunity to punish players who 
-"steal" and those who do not, meaning that participants make twelve punishment 
-decisions in total. Each behavioural strategy implies a unique pattern of 
-punishment across all decisions (see Methods for detailed explanation of 
-strategies). Green ticks reflect decisions to punish, red crosses reflect 
-decisions to not punish. In column headers, payoffs at the first stage (above) 
-and the second stage (below) are denoted as P1-P2 (or P2-P3 [P1] for Game F) 
-where participants take the role of P1 and P2 is the target of punishment. AI = 
-advantageous inequity, DI = disadvantageous inequity.*
-
-```{r tableStrategies}
-apa_table(
-  tar_read(tableStrategies),
-  escape = FALSE,
-  landscape = TRUE,
-  align = c("p{1.7cm}", "p{3.3cm}", rep("c", 12)),
-  col_spanners = list(
-    "\\thead{\\scriptsize Game A \\\\ (AI) \\\\ 70-10}"   = 3:4,
-    "\\thead{\\scriptsize Game B \\\\ (Equal) \\\\ 70-30}"   = 5:6,
-    "\\thead{\\scriptsize Game C \\\\ (Computer) \\\\ 70-30}"   = 7:8,
-    "\\thead{\\scriptsize Game D \\\\ (1:1 Fee-Fine) \\\\ 70-30}"   = 9:10,
-    "\\thead{\\scriptsize Game E \\\\ (DI) \\\\ 70-50}"        = 11:12,
-    "\\thead{\\scriptsize Game F \\\\ (Third-Party) \\\\ 70-70 [100]}" = 13:14
-    ),
-  caption = "(ref:tableStrategiesCaption)",
-  font_size = "scriptsize"
-)
-```
+:::
 
@@
 The overall pattern of punitive behaviour in the six economic games was very 
-similar across both countries (Figure \@ref(fig:plotDecisions)). Participants 
-were generally more likely to punish targets who stole compared to targets who 
-did not steal (multilevel logistic regression; _b_ = 
-`r printnum(mlm1["otherTook","Estimate"])`, standard error = 
-`r printnum(mlm1["otherTook","Std. Error"])`, _p_ 
-`r printp(mlm1["otherTook","Pr(>|z|)"])`). Participants were also more likely 
-to punish when targets' stealing behaviour generated inequalities, 
+similar across both countries (@fig-decisions). Participants were generally more 
+likely to punish targets who stole compared to targets who did not steal 
+(multilevel logistic regression; _b_ = 
+`r papaja::printnum(mlm1["otherTook","Estimate"])`, standard error = 
+`r papaja::printnum(mlm1["otherTook","Std. Error"])`, _p_ 
+`r papaja::printp(mlm1["otherTook","Pr(>|z|)"])`). Participants were also more 
+likely to punish when targets' stealing behaviour generated inequalities, 
 specifically in Games E and F (_b_ = 
-`r printnum(mlm2["DI","Estimate"])`, SE = `r printnum(mlm2["DI","Std. Error"])`, 
-_p_ `r printp(mlm2["DI","Pr(>|z|)"])`).
+`r papaja::printnum(mlm2["DI","Estimate"])`, SE = 
+`r papaja::printnum(mlm2["DI","Std. Error"])`, 
+_p_ `r papaja::printp(mlm2["DI","Pr(>|z|)"])`).
 
-(ref:plotDecisionsCaption) *Overall pattern of punitive behaviour across all six 
-economic games, split by country.* AI = advantageous inequity, DI = 
-disadvantageous inequity.
+```{r, fig.height=5, fig.width=8}
+#| label: fig-decisions
+#| fig-cap: Overall pattern of punitive behaviour across all six economic games, split by country.
+#| fig-align: center
+#| apa-note: AI = advantageous inequity, DI = disadvantageous inequity.
 
-```{r plotDecisions, fig.width=8, fig.height=5, fig.cap="(ref:plotDecisionsCaption)"}
 tar_read(plotDecisions)
@@
 
+<br>
+
 We classified participants into a particular strategy if their behaviour across 
-all twelve decisions matched our behavioural predictions shown in Table 
-\@ref(tab:tableStrategies) exactly. Table \@ref(tab:tableStrategyCounts) shows 
-the proportion of participants following each strategy, with N/A used to 
-represent participants who did not fit exactly into any particular strategy type.
-Overall, `r printnum(mean(tar_read(dExc)$strategy != "N/A")*100, digits = 0)`%
+all twelve decisions matched our behavioural predictions shown in 
+@tbl-strategies exactly. @tbl-strategycounts shows the proportion of 
+participants following each strategy, with N/A used to represent participants 
+who did not fit exactly into any particular strategy type. Overall, 
+`r papaja::printnum(mean(tar_read(dExc)$strategy != "N/A")*100, digits = 0)`%
 of our participants could be classified exactly into one of the strategies. The
@@
 inequity, competitive). Although participants often punished targets who did not 
-steal in the six games (Figure \@ref(fig:plotDecisions)), no participants 
-followed the antisocial strategy by exclusively punishing targets who did not 
-steal across _all_ games.
+steal in the six games (@fig-decisions), no participants followed the antisocial
+strategy by exclusively punishing targets who did not steal across _all_ games.
 
-(ref:tableStrategyCountsCaption) Counts and proportions of participants 
-following each punishment strategy exactly, split by country. *N/A implies that
-participants were unable to be classified exactly into any of the punishment
-strategies. AI = advantageous inequity, DI = disadvantageous inequity.*
+```{r}
+#| label: tbl-strategycounts
+#| tbl-cap: Counts and proportions of participants following each punishment strategy exactly, split by country.
+#| apa-note: N/A implies that participants were unable to be classified exactly into any of the punishment strategies. AI = advantageous inequity, DI = disadvantageous inequity.
 
-```{r tableStrategyCounts}
-apa_table(
-  tar_read(tableStrategyCounts),
-  escape = FALSE,
-  align = c("l", rep("c", 4)),
-  col_spanners = list(
-    "\\thead{United Kingdom \\\\ (N = 1014)}" = 2:3,
-    "\\thead{United States \\\\ (N = 996)}" = 4:5
+tar_read(tableStrategyCounts) %>%
+  flextable() %>%
+  set_header_labels(
+    Strategy = "Strategy",
+    a = "N",
+    b = "Prop",
+    c = "N",
+    d = "Prop"
+  ) %>%
+  add_header_row(
+    values = c(
+      "",
+      "United Kingdom\n(N = 1014)",
+      "United States\n(N = 996)"
     ),
-  caption = "(ref:tableStrategyCountsCaption)"
-)
+    colwidths = c(1, 2, 2)
+  ) %>%
+  autofit() %>%
+  theme_apa() %>%
+  line_spacing(part = "all") %>%
+  align(i = NULL, j = 1, align = "left", part = "all")
 ```
+
+<br>
 
@@
 examined the most common patterns of punitive behaviour across all twelve 
-decisions. Supplementary Table \@ref(tab:tablePatterns) shows the proportion of 
-participants following the 25 most common behavioural patterns, including, 
-where appropriate, the predetermined strategies from Table 
-\@ref(tab:tableStrategies). In both countries, a common pattern of behaviour 
-not captured by any of the strategies was punishing only when the target stole
-in the third-party game (Game F). Punishment in this game is consistent with an 
-egalitarian motive, as stealing produces unequal outcomes, but third-party 
-punishment here is also consistent with norm-enforcing and competitive motives 
-(see Table \@ref(tab:tableStrategies)). Other common behavioural patterns not 
+decisions. @supptbl-patterns shows the proportion of participants following the
+25 most common behavioural patterns, including, where appropriate, the 
+predetermined strategies from @tbl-strategies. In both countries, a common 
+pattern of behaviour not captured by any of the strategies was punishing only 
+when the target stole in the third-party game (Game F). Punishment in this game
+is consistent with an egalitarian motive, as stealing produces unequal outcomes,
+but third-party punishment here is also consistent with norm-enforcing and 
+competitive motives (see @tbl-strategies). Other common behavioural patterns not 
 captured by our strategies included punishing whenever the target stole across 
@@
 estimates, we fitted a Bayesian latent state model to the data. This model 
-assumes that the nine strategies in Table \@ref(tab:tableStrategies) (plus a 
-"random choice" strategy that chooses randomly for each decision) are the only 
-latent strategies and that these are instantiated into observed behaviour 
-according to the logic in Table \@ref(tab:tableStrategies) with some 
-probability of implementation error (i.e., an intention to punish is implemented 
-as non-punishment and vice versa). Averaging over all strategies and 
-incorporating the possibility of implementation errors, the model estimates the 
-probability of participants following any particular strategy, given the 
-observed data.
+assumes that the nine strategies in @tbl-strategies (plus a "random choice" 
+strategy that chooses randomly for each decision) are the only latent strategies 
+and that these are instantiated into observed behaviour according to the logic 
+in @tbl-strategies with some probability of implementation error (i.e., an 
+intention to punish is implemented as non-punishment and vice versa). Averaging 
+over all strategies and incorporating the possibility of implementation errors,
+the model estimates the probability of participants following any particular 
+strategy, given the observed data.
 
-The posterior estimates from the model are presented in Figure 
-\@ref(fig:plotModel1b). The posterior probabilities for each strategy did not
-differ between the two countries. In both countries, the never punish strategy
-had the highest probability, followed by the egalitarian strategy. The
-norm-enforcing and seek advantageous inequity strategies were the next most 
-likely, with higher posterior estimates than the competitive and antisocial 
-strategies. None of the other strategies differed in their posterior estimates. 
-The same general pattern emerged when we analysed the full dataset without 
-pre-registered exclusions (Supplementary Figure \@ref(fig:plotModel1a)).
+The posterior estimates from the model are presented in @fig-model1b. The 
+posterior probabilities for each strategy did not differ between the two 
+countries. In both countries, the never punish strategy had the highest 
+probability, followed by the egalitarian strategy. The norm-enforcing and seek 
+advantageous inequity strategies were the next most likely, with higher 
+posterior estimates than the competitive and antisocial strategies. None of the 
+other strategies differed in their posterior estimates. The same general pattern 
+emerged when we analysed the full dataset without pre-registered exclusions 
+(@suppfig-model1a).
 
-(ref:plotModel1bCaption) *Posterior estimates of the probabilities of following
-different punishment strategies from the Bayesian latent state model.* The model 
-assumes an implementation error rate of 5%. Points represent posterior medians, 
-line ranges represent 50% and 95% credible intervals. AI = advantageous 
-inequity, DI = disadvantageous inequity.
+```{r, fig.height=5, fig.width=7}
+#| label: fig-model1b
+#| fig-cap: Posterior estimates of the probabilities of following different punishment strategies from the Bayesian latent state model.
+#| fig-align: center
+#| apa-note: The model assumes an implementation error rate of 5%.
+Points represent posterior medians, line ranges represent 50% and 95% credible intervals.
+AI = advantageous inequity, DI = disadvantageous inequity.
 
-```{r plotModel1b, fig.width=7, fig.height=5, fig.cap="(ref:plotModel1bCaption)", warning=FALSE}
 tar_read(plot1.2)
 ```
+
+<br>
 
@@
 variable in a separate model, predicting all ten punishment strategies 
-(the nine from Table 1, plus the 'random choice' strategy) simultaneously.
+(the nine from @tbl-strategies, plus the 'random choice' strategy) 
+simultaneously.
 
@@
 did not predict adherence to a particular punishment strategy (Supplementary 
-Figures \@ref(fig:plotAllDems2) and \@ref(fig:plotAllDems1)). In the United 
-States, the never punish strategy was slightly more common among participants 
-lower in socio-economic status (median posterior slope = 
-`r printnum(median(ses))`, 95% CI [`r printnum(quantile(ses, 0.025))` 
-`r printnum(quantile(ses, 0.975))`]) but this effect was small.
+Figures -@suppfig-allDems2 and -@suppfig-allDems1). In the United States, the 
+never punish strategy was slightly more common among participants lower in 
+socio-economic status (median posterior slope = 
+`r papaja::printnum(median(ses))`, 95% CI
+[`r papaja::printnum(quantile(ses, 0.025))` 
+`r papaja::printnum(quantile(ses, 0.975))`]) but this effect was small.
 
@@
 egalitarian, never punish, and random choice strategies, with small-to-medium 
-effect sizes (Figure \@ref(fig:plotAllPers2)). More prosocial participants 
-(those with higher SVO scores) were more likely to follow the egalitarian and 
-the never punish strategies, while those with lower SVO scores were more likely 
-to enact the random choice strategy. The personality dimensions of 
-honesty-humility and openness to experience were both positively associated with 
-following the never punish strategy, while extraversion negatively predicted 
-this strategy. The effects were mostly similar across countries, but 
-occasionally differed: for example, in the United States, but not in the United 
-Kingdom, honesty-humility was positively associated with following the 
-egalitarian strategy and negatively associated with following the random 
-choice strategy. Overall, the same pattern of results emerged when analysing the 
-full dataset without exclusions (Supplementary Figure \@ref(fig:plotAllPers1)).
+effect sizes (@fig-allPers2). More prosocial participants (those with higher SVO 
+scores) were more likely to follow the egalitarian and the never punish 
+strategies, while those with lower SVO scores were more likely to enact the 
+random choice strategy. The personality dimensions of honesty-humility and 
+openness to experience were both positively associated with following the never 
+punish strategy, while extraversion negatively predicted this strategy. The 
+effects were mostly similar across countries, but occasionally differed: for 
+example, in the United States, but not in the United Kingdom, honesty-humility 
+was positively associated with following the egalitarian strategy and negatively 
+associated with following the random choice strategy. Overall, the same pattern
+of results emerged when analysing the full dataset without exclusions 
+(@suppfig-allPers1).
 
-(ref:plotAllPers2Caption) *Posterior slopes from Bayesian latent state models 
-including Big-6 personality dimensions and Social Value Orientation.* Each row 
-represents a separate model. All models assume an implementation error rate of 
-5%. Points represent posterior medians, line ranges represent 95% credible 
-intervals. AI = advantageous inequity, DI = disadvantageous inequity.
+```{r, fig.height=3.5, fig.width=7.5, warning=FALSE}
+#| label: fig-allPers2
+#| fig-cap: Posterior slopes from Bayesian latent state models including Big-6 personality dimensions and Social Value Orientation.
+#| fig-align: center
+#| apa-note: Each row represents a separate model.
+All models assume an implementation error rate of 5%.
+Points represent posterior medians, line ranges represent 95% credible intervals.
+AI = advantageous inequity, DI = disadvantageous inequity.
 
-```{r plotAllPers2, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotAllPers2Caption)", warning=FALSE}
 tar_read(plotAllPers2)
@@
 
+<br>
+
 Political and religious variables were also associated with punishment strategy 
-(Figure \@ref(fig:plotAllPolRel2)). These effects were small-to-medium in size 
-and tended to be more pronounced in the United States. Controlling for Social 
-Dominance Orientation, American participants higher in Right Wing 
-Authoritarianism were more likely to follow the strategies avoiding 
-disadvantageous inequity and seeking advantageous inequity. Participants who 
-stated that they would like to "bring those below them [on the socio-economic 
-status ladder] up a peg" were more likely to follow the egalitarian strategy, 
-while American participants higher in Social Dominance Orientation, Right Wing 
-Authoritarianism, and believing that God controls events in the world were less 
-likely to follow the egalitarian strategy. In general, religious and 
-conservative participants were less likely to follow the never punish strategy. 
-This general pattern of results was replicated with the full dataset 
-(Supplementary Figure \@ref(fig:plotAllPolRel1)).
+(@fig-allPolRel2). These effects were small-to-medium in size and tended to be 
+more pronounced in the United States. Controlling for Social Dominance 
+Orientation, American participants higher in Right Wing Authoritarianism were 
+more likely to follow the strategies avoiding disadvantageous inequity and 
+seeking advantageous inequity. Participants who stated that they would like to 
+"bring those below them [on the socio-economic status ladder] up a peg" were 
+more likely to follow the egalitarian strategy, while American participants 
+higher in Social Dominance Orientation, Right Wing Authoritarianism, and 
+believing that God controls events in the world were less likely to follow the 
+egalitarian strategy. In general, religious and conservative participants were 
+less likely to follow the never punish strategy. This general pattern of results 
+was replicated with the full dataset (@suppfig-allPolRel1).
 
-(ref:plotAllPolRel2Caption) *Posterior slopes from Bayesian latent state models 
-including political ideology, views about social inequality, and religiosity.* 
-Each row represents a separate model aside from Social Dominance Orientation and 
-Right Wing Authoritarianism, which control for one another within the same 
-model. All models assume an implementation error rate of 5%. Points represent 
-posterior medians, line ranges represent 95% credible intervals. AI = 
-advantageous inequity, DI = disadvantageous inequity.
+```{r, fig.height=3.5, fig.width=7.5, warning=FALSE}
+#| label: fig-allPolRel2
+#| fig-cap: Posterior slopes from Bayesian latent state models including political ideology, views about social inequality, and religiosity.
+#| fig-align: center
+#| apa-note: Each row represents a separate model aside from Social Dominance Orientation and Right Wing Authoritarianism, which control for one another within the same model.
+All models assume an implementation error rate of 5%.
+Points represent posterior medians, line ranges represent 95% credible intervals.
+AI = advantageous inequity, DI = disadvantageous inequity.
 
-```{r plotAllPolRel2, fig.width=7.5, fig.height=3.5, fig.cap="(ref:plotAllPolRel2Caption)", warning=FALSE}
 tar_read(plotAllPolRel2)
 ```
+
+<br>
 
@@
 with the behavioural strategy that participants employed, with effects ranging 
-from small to large in size (see Supplementary Figures \@ref(fig:plotSliders1) 
-and \@ref(fig:plotSliders2) for the distribution of responses to self-report 
-questions). Figure \@ref(fig:plotAllSliders2) shows the relationships between 
-self-report questions and the different punishment strategies, highlighting the 
-combinations where the question matched the behavioural strategy. We found 
-positive relationships between the self-report questions and strategy usage for 
-the norm-enforcing, egalitarian, seek advantageous inequity, never punish, and 
-random choice strategies. The 95% credible intervals for other estimates 
-included zero, though these estimates often trended in a positive direction. The 
-same pattern of results was found when analysing the full dataset without 
-exclusions (Supplementary Figure \@ref(fig:plotAllSliders1)).
+from small to large in size (see Supplementary Figures -@suppfig-sliders1 and 
+-@suppfig-sliders2 for the distribution of responses to self-report questions).
+@fig-allSliders2 shows the relationships between self-report questions and the 
+different punishment strategies, highlighting the combinations where the 
+question matched the behavioural strategy. We found positive relationships 
+between the self-report questions and strategy usage for the norm-enforcing, 
+egalitarian, seek advantageous inequity, never punish, and random choice 
+strategies. The 95% credible intervals for other estimates included zero, though
+these estimates often trended in a positive direction. The same pattern of
+results was found when analysing the full dataset without exclusions
+(@suppfig-allSliders1).
 
-(ref:plotAllSliders2Caption) *Posterior slopes from models including 
-self-reported strategy usage.* Each row represents a separate model. All models 
-assume an implementation error rate of 5%. Highlighted estimates represent 
-combinations where the self-report question matched the behavioural strategy. 
-Points represent posterior medians, line ranges represent 95% credible 
-intervals. AI = advantageous inequity, DI = disadvantageous inequity.
+```{r, fig.height=4, fig.width=7.5, warning=FALSE}
+#| label: fig-allSliders2
+#| fig-cap: Posterior slopes from models including self-reported strategy usage.
+#| fig-align: center
+#| apa-note: Each row represents a separate model.
+All models assume an implementation error rate of 5%.
+Highlighted estimates represent combinations where the self-report question matched the behavioural strategy.
+Points represent posterior medians, line ranges represent 95% credible intervals.
+AI = advantageous inequity, DI = disadvantageous inequity.
 
-```{r plotAllSliders2, fig.width=7.5, fig.height=4, fig.cap="(ref:plotAllSliders2Caption)", warning=FALSE}
 tar_read(plotAllSliders2)
 ```
+
+<br>
 
@@
 line with previous studies which have highlighted inequity aversion as an 
-important motivation for punishment in economic games 
-[@Bone2015;@Bone2016;@Dawes2007;@Deutchman2021;@Johnson2009;@Raihani2012b]. 
-Personality, social preferences, and political and religious views were related 
-to these strategies. Traits associated with other-regarding concern, such as SVO 
-and honesty-humility, predicted following the egalitarian strategy. By contrast,
+important motivation for punishment in economic games [@Bone2015; @Bone2016; 
+@Dawes2007; @Deutchman2021; @Johnson2009; @Raihani2012b]. Personality, social 
+preferences, and political and religious views were related to these strategies.
+Traits associated with other-regarding concern, such as SVO and 
+honesty-humility, predicted following the egalitarian strategy. By contrast,
 religious and conservative individuals, including individuals higher in SDO and
@@
 common than strategies sensitive to inequality in our set of games. This was 
-reflected both in participants’ elicited punishment behaviour (Figure 
-\@ref(fig:plotModel1b)) and in their self-reports of their own strategy 
-(Supplementary Figures \@ref(fig:plotSliders1) and \@ref(fig:plotSliders2)). 
-Although our design did not explicitly allow for behaviour shaping as the 
-interactions were all one-shot, we did manipulate whether the target's stealing 
-behaviour was intentional or not (Game C), an approach which has been used in 
-previous vignette studies to identify behaviour shaping motives [@Carlsmith2002].
-The lower prevalence of behaviour shaping strategies in our study is consistent 
-with prior work showing that punishment often continues to be used when 
-behaviour shaping is impossible, such as when the target will never find out 
-that they have been punished [@Crockett2014] or on the last round of repeated 
-interactions [@Barclay2016]. We found that demographic, personality, political, 
-and religious variables tended to be unrelated to behaviour shaping strategies, 
-though this could reflect low power to detect such associations given the low 
-prevalence of these strategies in our sample. We also found that participants 
-accurately reported using the norm-enforcing strategy, but not the deterrent 
-strategy. This finding is in line with previous research showing that people 
-struggle to accurately report the deterrent motivations for their punitive 
-behaviour [@Carlsmith2002].
+reflected both in participants’ elicited punishment behaviour (@fig-model1b)
+and in their self-reports of their own strategy (Supplementary Figures 
+-@suppfig-sliders1 and -@suppfig-sliders2). Although our design did not 
+explicitly allow for behaviour shaping as the interactions were all one-shot, we
+did manipulate whether the target's stealing behaviour was intentional or not 
+(Game C), an approach which has been used in previous vignette studies to 
+identify behaviour shaping motives [@Carlsmith2002]. The lower prevalence of 
+behaviour shaping strategies in our study is consistent with prior work showing
+that punishment often continues to be used when behaviour shaping is impossible,
+such as when the target will never find out that they have been punished 
+[@Crockett2014] or on the last round of repeated interactions [@Barclay2016]. We
+found that demographic, personality, political, and religious variables tended
+to be unrelated to behaviour shaping strategies, though this could reflect low
+power to detect such associations given the low prevalence of these strategies
+in our sample. We also found that participants accurately reported using the 
+norm-enforcing strategy, but not the deterrent strategy. This finding is in line
+with previous research showing that people struggle to accurately report the 
+deterrent motivations for their punitive behaviour [@Carlsmith2002].
 
@@
 co-players. Harming non-stealing individuals was also consistent with the 
-competitive strategies, which did appear in our sample albeit at low frequencies.
-The fact that no participants in our sample punished non-stealing across all 
-games suggests that the antisocial punishment that has been observed in other 
-studies [@Herrmann2008;@Pleasant2018] is not aimed at harming cooperators 
-specifically, as has been previously suggested [@Herrmann2008]. Instead, 
-antisocial punishment is more likely to be motivated by improving one's 
+competitive strategies, which did appear in our sample albeit at low 
+frequencies. The fact that no participants in our sample punished non-stealing 
+across all games suggests that the antisocial punishment that has been observed
+in other studies [@Herrmann2008; @Pleasant2018] is not aimed at harming 
+cooperators specifically, as has been previously suggested [@Herrmann2008]. 
+Instead, antisocial punishment is more likely to be motivated by improving one's 
 relative position, which is in line with work showing that antisocial punishment 
@@
 game studies. In these studies, a common assumption is that participants punish 
-to change the behaviour of cheats [@Fehr2000;@Fehr2002]. But in reality, people 
+to change the behaviour of cheats [@Fehr2000; @Fehr2002]. But in reality, people 
 may be choosing the punishment option to achieve a variety of different goals. 
@@
 interpret punishment as serving a competitive motive, it may elicit retaliation
-rather than encourage cooperation [@Dreber2008;@Nikiforakis2008b;@Raihani2019].
-As punishers’ motives must be inferred (and such inferences likely depend on 
-character traits of the target, as well as the context in which punishment 
-occurs), there is likely to be some variation and error in attributing motives 
-to punishers. To the extent that inferred motives affect target responses, this 
-might help to explain the mixed findings in the field as to whether punishment 
-actually motivates cheating targets to cooperate in the future [@Raihani2019].
+rather than encourage cooperation [@Dreber2008; @Nikiforakis2008b; 
+@Raihani2019]. As punishers’ motives must be inferred (and such inferences
+likely depend on character traits of the target, as well as the context in which 
+punishment occurs), there is likely to be some variation and error in 
+attributing motives to punishers. To the extent that inferred motives affect 
+target responses, this might help to explain the mixed findings in the field as
+to whether punishment actually motivates cheating targets to subsequently 
+cooperate [@Raihani2019].
 
@@
 to punish, we may have seen more punishment from participants. Indeed, 
-`r printnum(mean(tar_read(dExc)$SelfRate10[tar_read(dExc)$strategy == "Never punish"] > 0, na.rm = TRUE)*100, digits = 0)`% 
+`r papaja::printnum(mean(tar_read(dExc)$SelfRate10[tar_read(dExc)$strategy == "Never punish"] > 0, na.rm = TRUE)*100, digits = 0)`% 
 of participants following the never punish strategy positively stated that they 
@@
 in both WEIRD (Western, educated, industrialised, rich, and democratic) samples
-[@Balafoutas2014;@Balafoutas2016] and in small-scale societies [@Baumard2010]. 
+[@Balafoutas2014; @Balafoutas2016] and in small-scale societies [@Baumard2010]. 
 One reason that people may be averse to peer punishment is that, because it is 
@@
 one-shot anonymous settings used here). Another reason that people may be averse
-to peer punishment is that it can trigger retaliation [@Raihani2019]. This may be
-especially likely in situations that lack clear institutional norms to 
+to peer punishment is that it can trigger retaliation [@Raihani2019]. This may 
+be especially likely in situations that lack clear institutional norms to 
 legitimise punishment, such as our economic games. As with reputation damage, 
@@
 limiting the potential for feuds and cycles of retaliation
-[@Fitouchi2023;@Singh2022]. Future research should uncover whether people are 
+[@Fitouchi2023; @Singh2022]. Future research should uncover whether people are 
 more willing to punish in these conventionalised contexts (e.g., see ref
@@
 twelve decisions, compared to the avoid disadvantageous inequity strategy which 
-required only one instance of punishment (Table \@ref(tab:tableStrategies)).
-We employed the strategy method to deal with this, calculating participant 
-payoffs from a randomly chosen game instead of accumulating the costs across 
-all games. This mitigates the concern that the total cost is the most salient
-feature determining participants' decisions. But even accounting for this, 
-one could still argue that the difference in overall costs explains why, for 
-example, the competitive strategy is less common in our dataset than the avoid 
+required only one instance of punishment (@tbl-strategies). We employed the 
+strategy method to deal with this, calculating participant payoffs from a 
+randomly chosen game instead of accumulating the costs across all games. This 
+mitigates the concern that the total cost is the most salient feature 
+determining participants' decisions. But even accounting for this, one could 
+still argue that the difference in overall costs explains why, for example, the 
+competitive strategy is less common in our dataset than the avoid 
 disadvantageous inequity strategy. We do not think this feature of our design 
@@
 they would have punished if it were free to do so, we found low agreement 
-with this statement (Supplementary Figure \@ref(fig:plotSliders1)), 
-suggesting that participants were not particularly sensitive to the costs of 
-punishment. In line with this, when we plot the frequencies of different 
-strategies against their expected costs, we find that cost is not a perfect
-predictor of strategy frequency: many "cheap" strategies are rare in our 
-data and some "expensive" strategies are quite common (e.g., always punish;
-Supplementary Figure \@ref(fig:plotSpline)). Third, this argument implies 
-that the high frequency of the never punish strategy is merely an artefact
-of our design, since it is the only strategy that does not cost anything to
-implement. But as we have discussed, many other studies have found similar
-aversions to costly punishment in the lab [@Barclay2016;@Batistoni2022;
-@Bone2015;@Deutchman2021;@Raihani2012b] and in the real world
-[@Balafoutas2014;@Balafoutas2016;@Baumard2010] suggesting that this result is
-not an artefact of our design.
+with this statement (@suppfig-sliders1), suggesting that participants were not 
+particularly sensitive to the costs of punishment. In line with this, when we 
+plot the frequencies of different strategies against their expected costs, we 
+find that cost is not a perfect predictor of strategy frequency: many "cheap" 
+strategies are rare in our data and some "expensive" strategies are quite common
+(e.g., always punish; @suppfig-spline). Third, this argument implies that the 
+high frequency of the never punish strategy is merely an artefact of our design,
+since it is the only strategy that does not cost anything to implement. But as
+we have discussed, many other studies have found similar aversions to costly 
+punishment in the lab [@Barclay2016; @Batistoni2022; @Bone2015; @Deutchman2021; 
+@Raihani2012b] and in the real world [@Balafoutas2014; @Balafoutas2016; 
+@Baumard2010] suggesting that this result is not an artefact of our design.
 
@@
 
-This research has been approved by the UCL Department of Psychology Ethics 
-Committee (project: 3720/002) and ratified by the University of Auckland Human 
+This research was approved by the UCL Department of Psychology Ethics Committee 
+(project: 3720/002) and ratified by the University of Auckland Human 
 Participants Ethics Committee. The study was performed in accordance with all 
@@
 We pre-registered the study on the Open Science Framework before collecting data 
-in the United Kingdom (11^th^ November 2022; https://osf.io/k75fc). We submitted 
-another pre-registration before collecting data in the United States (20^th^
-June 2023; https://osf.io/q4hdy). In the pre-registrations, we outlined our 
-study design, exclusion criteria, and analysis plan. As the study was 
-exploratory, we did not pre-register any explicit hypotheses. We did not deviate 
-from the pre-registrations.
+in the United Kingdom (11^th^ November 2022; <https://osf.io/k75fc>). We 
+submitted another pre-registration before collecting data in the United States 
+(20^th^ June 2023; <https://osf.io/q4hdy>). In the pre-registrations, we 
+outlined our study design, exclusion criteria, and analysis plan. As the study 
+was exploratory, we did not pre-register any explicit hypotheses. We did not 
+deviate from the pre-registrations.
 
@@
 
-We collected a representative sample of `r printnum(uk_n_full)` participants 
-from the United Kingdom through the online platform Prolific 
-(https://www.prolific.com/). All of these participants completed the economic 
-games and `r printnum(uk_n_rtrn)` returned to complete the follow-up survey a 
-week later (`r printnum((uk_n_rtrn / uk_n_full)*100, digits = 0)`% retention 
-rate). After exclusions, we were left with `r printnum(uk_n_excl)` participants 
-overall (see Supplementary Figure \@ref(fig:plotSampleUK) for sample 
-characteristics).
+We collected a representative sample of `r papaja::printnum(uk_n_full)` 
+participants from the United Kingdom through the online platform Prolific 
+(<https://www.prolific.com/>). All of these participants completed the economic 
+games and `r papaja::printnum(uk_n_rtrn)` returned to complete the follow-up 
+survey a week later 
+(`r papaja::printnum((uk_n_rtrn / uk_n_full)*100, digits = 0)`% retention rate).
+After exclusions, we were left with `r papaja::printnum(uk_n_excl)` participants 
+overall (see @suppfig-sampleUK).
 
-We later collected a representative sample of `r printnum(us_n_full)` 
+We later collected a representative sample of `r papaja::printnum(us_n_full)` 
 participants from the United States through Prolific. All of these participants 
-completed the economic games and `r printnum(us_n_rtrn)` returned to complete 
-the follow-up survey (`r printnum((us_n_rtrn / us_n_full) * 100, digits = 0)`% 
-retention rate). After exclusions, we were left with `r printnum(us_n_excl)` 
-participants overall (see Supplementary Figure \@ref(fig:plotSampleUS) for 
-sample characteristics).
+completed the economic games and `r papaja::printnum(us_n_rtrn)` returned to 
+complete the follow-up survey 
+(`r papaja::printnum((us_n_rtrn / us_n_full) * 100, digits = 0)`% retention 
+rate). After exclusions, we were left with `r papaja::printnum(us_n_excl)` 
+participants overall (see @suppfig-sampleUS for sample characteristics).
 
@@
 and adds it to their payoff or (b) does nothing. For each of these cases, 
-participants are asked whether they would like to pay money to reduce P2’s 
+participants are asked whether they would like to pay money to reduce P2's 
 payoff. Games A-E have two players and Game F has three players.
 
-The six games are as follows (variations bolded; see Figure \@ref(fig:plotGames)
-for a visual representation of the games):
+The six games are as follows (variations bolded; see @fig-games for a visual 
+representation of the games):
 
@@
 
-Our punishment strategies make different behavioural predictions across these 
-different games (Table \@ref(tab:tableStrategies)). The deterrent strategy
-punishes whenever it is intentionally personally harmed (when the target steals
-in Games A, B, D, and E). The norm-enforcing strategy is the same, except that 
-it also punishes when _others_ are intentionally harmed (when the target steals
-in Game F). The retributive strategy punishes whenever it is personally harmed,
-whether the harm was intentional or not (when the target steals in Games A-E).
-The avoid disadvantageous inequity strategy only punishes when it currently 
-experiences disadvantageous inequity and can alleviate it by punishing (when the
-target steals in Game E). The egalitarian strategy punishes when punishment
-makes payoffs for all players more equal than they currently are (when the
-target steals in Games E and F). The seek advantageous inequity strategy
-punishes when it currently does not experience advantageous inequity and 
-punishment would produce advantageous inequity for itself (when the target
-steals in Games B and C). The competitive strategy punishes whenever the
-fee-fine ratio is 1:3, as it can improve its relative position by punishing, but
-does not punish when the fee-fine ratio is 1:1 (Game D). The antisocial strategy
-punishes whenever the target does not steal across all games. The never punish
-strategy never punishes in any game.
+<br>
+
+We delineate nine punishment strategies that can be isolated based on players' 
+decisions across the six economic games (@tbl-strategies). In existing 
+literature, it is often unclear whether punishment decisions are sensitive to 
+the previous actions of others or to payoff sensitivity. This is because (*i*) 
+players' decisions typically introduce payoff differentials in economic games 
+and (*ii*) where fine > fee, punishment tools can be used to harm co-players and 
+to change relative payoffs [@Raihani2019]. Of the nine strategies we define, 
+*deterrent*, *norm-enforcing*, *retributive* and *antisocial* are strategies 
+that are only sensitive to the actions of others. Conversely, *avoid 
+disadvantageous inequity*, *egalitarian*, *seek advantageous inequity* and 
+*competitive* are strategies that are sensitive to payoffs. *Never punish* is an 
+unconditional strategy (i.e., no sensitivity to actions or payoffs).
+
+Our strategy definitions sometimes differ from how these strategies are defined 
+and discussed in the literature (e.g., antisocial punishment). This is essential 
+because previous definitions permit many possible motives, while we are aiming 
+to isolate punishment motives. We define our strategies as follows:
+
+- *Deterrent*: Deter others who have harmed you from harming you again in the 
+future (when the target steals in Games A, B, D, and E). We note that while no 
+behaviour change is possible in our one-shot games, deterrent motives can be 
+isolated from retributive motives by asking whether participants are sensitive 
+to the intentionality behind cheating, as in previous work [@Carlsmith2002]. 
+- *Norm-enforcing*: Punish to enforce a shared anti-harm norm and encourage 
+future norm compliance (when the target steals in Games A, B, D, E and F). 
+Though we cannot measure future norm compliance in our one-shot games, these 
+non-repeated (and 'stranger design') games have been used to identify 
+norm-enforcing motives in many previous studies [@Fehr2000; @Fehr2002].
+- *Retributive*: Punish if doing so harms another who harmed you (when the 
+target steals in Games A-E). We note that retribution is also sometimes used to 
+describe punishment of those who harmed others [@Carlsmith2002]. We require this 
+tighter definition because our aim is to precisely identify punishment 
+strategies, and sensitivity to whether self or others were harmed may be an 
+important factor affecting the decision to punish (e.g., ref
+[@Feldmanhall2014]). 
+- *Avoid disadvantageous inequity*: Punish if doing so avoids disadvantageous 
+inequity for the self (when the target steals in Game E). 
+- *Egalitarian*: Punish if doing so makes payoffs for all more equal (when the 
+target steals in Games E and F) (c.f. ref [@Dawes2007]).
+- *Seek advantageous inequity*: Punish if doing so produces advantageous
+inequity for the self (when the target steals in Games B and C). This is one 
+possible motive underlying what others have previously called 'antisocial 
+punishment' or 'spite'; here we isolate it.
+- *Competitive*: Punish if doing so improves your relative position (all games 
+except Game D). 
+- *Antisocial*: Punish exclusively those who do not cause harm (punish 
+non-stealing target in all games). This definition differs from some previous 
+definitions of antisocial punishment, but this is essential to isolate the full 
+range of motives that could be driving punishment and which are not typically 
+disentangled in other settings. 
+- *Never punish*: A person who never incurs a personal cost to punish a 
+co-player (does not punish in any game).
+
+<br>
 
@@
 capturing the different reasons for their decisions (for full wordings, see 
-Supplementary Table \@ref(tab:tableSliderWordings)).
+@supptbl-sliderWordings).
 
@@
 In a follow-up survey, we collected the following data on participants (for 
-wordings of all questions, see Supplementary Table 
-\@ref(tab:tableSurveyWordings)):
+wordings of all questions, see @supptbl-surveyWordings):
 
@@
   personality dimensions of agreeableness ($\alpha$ = 
-  `r printnum(tar_read(alphaAgree)$alpha)`), conscientiousness ($\alpha$ = 
-  `r printnum(tar_read(alphaConsc)$alpha)`), extraversion ($\alpha$ = 
-  `r printnum(tar_read(alphaExtra)$alpha)`), honesty-humility ($\alpha$ = 
-  `r printnum(tar_read(alphaHonest)$alpha)`), openness to experience ($\alpha$ = 
-  `r printnum(tar_read(alphaOpen)$alpha)`), and neuroticism ($\alpha$ = 
-  `r printnum(tar_read(alphaNeur)$alpha)`). Four items were used for each 
+  `r papaja::printnum(tar_read(alphaAgree)$alpha)`), conscientiousness ($\alpha$ = 
+  `r papaja::printnum(tar_read(alphaConsc)$alpha)`), extraversion ($\alpha$ = 
+  `r papaja::printnum(tar_read(alphaExtra)$alpha)`), honesty-humility ($\alpha$ = 
+  `r papaja::printnum(tar_read(alphaHonest)$alpha)`), openness to experience ($\alpha$ = 
+  `r papaja::printnum(tar_read(alphaOpen)$alpha)`), and neuroticism ($\alpha$ = 
+  `r papaja::printnum(tar_read(alphaNeur)$alpha)`). Four items were used for each 
   personality dimension.
@@
   including left-right conservatism, Social Dominance Orientation [@Ho2015] 
-  ($\alpha$ = `r printnum(tar_read(alphaSDO)$alpha)`; eight items), and Right 
-  Wing Authoritarianism [@Bizumic2018] ($\alpha$ = 
-  `r printnum(tar_read(alphaRWA)$alpha)`; six items). We also probed 
+  ($\alpha$ = `r papaja::printnum(tar_read(alphaSDO)$alpha)`; eight items), and 
+  Right Wing Authoritarianism [@Bizumic2018] ($\alpha$ = 
+  `r papaja::printnum(tar_read(alphaRWA)$alpha)`; six items). We also probed 
   participants' views on social inequality by asking them whether they would 
@@
   entity controls the events in the world [@Laurin2012].
+
+<br>
 
@@
 June 2023. Our surveys were designed through the online survey platform 
-Qualtrics (https://www.qualtrics.com/).
+Qualtrics (<https://www.qualtrics.com/>).
 
@@
 to comprehension questions suggested that participants understood the six 
-economic games (Supplementary Table \@ref(tab:tableComp)). In order to partially 
-mitigate the fact that some punishment strategies predict more punishment than 
-others and are thus more "expensive" to implement (Table 
-\@ref(tab:tableStrategies)), we used the strategy method to incentivise the 
-economic games, choosing a random game to determine bonus payments rather than 
-summing participants' earnings across all games. After all games, 
-`r printnum((sum(dExc$DecisionConsequences == "Yes", na.rm = TRUE) / nrow(dExc)) * 100, digits = 0)`% 
+economic games (@supptbl-comp). In order to partially mitigate the fact that 
+some punishment strategies predict more punishment than others and are thus more 
+"expensive" to implement (@tbl-strategies), we used the strategy method to 
+incentivise the economic games, choosing a random game to determine bonus
+payments rather than summing participants' earnings across all games. After all
+games, 
+`r papaja::printnum((sum(dExc$DecisionConsequences == "Yes", na.rm = TRUE) / nrow(dExc)) * 100, digits = 0)`% 
 of participants stated that they believed that their decisions had real 
@@
 We pre-registered that we would use a Bayesian latent state model to infer 
-unobserved punishment strategies from the observed data [for a similar version 
-of this model, see ref [@McElreath2020]]. In this model, participants $i$ in 
+unobserved punishment strategies from the observed data (for a similar version 
+of this model, see ref [@McElreath2020]). In this model, participants $i$ in 
 countries $c$ make binary punishment decisions across twelve decisions $j$. We 
@@
 
+$$
 \begin{align}
@@
 \end{align}
+$$
 
@@
 The conditional probabilities $\text{Pr}(\text{punish}|s,j)$ are hard coded in 
-the model as outlined in Table \@ref(tab:tableStrategies). We incorporate an 
-implementation error rate $\delta$ into these conditional probabilities by coding 
-green ticks in Table \@ref(tab:tableStrategies) with a conditional probability 
-of $1 - \delta$ and coding red crosses with a conditional probability of 
-$0 + \delta$. We set $\delta$ to 0.05 in all models, which is similar to its 
-value when we estimate it as a free parameter in an additional model (median 
-posterior $\delta$ = `r printnum(median(error))`, 95% CI
-[`r printnum(quantile(error, 0.025))` `r printnum(quantile(error, 0.975))`];
-Supplementary Figure \@ref(fig:plotModel5)). The random choice strategy is 
-consistently coded with a conditional probability of ½ across all decisions.
+the model as outlined in @tbl-strategies. We incorporate an implementation error
+rate $\delta$ into these conditional probabilities by coding ticks in 
+@tbl-strategies with a conditional probability of $1 - \delta$ and coding 
+crosses with a conditional probability of $0 + \delta$. We set $\delta$ to 0.05
+in all models, which is similar to its value when we estimate it as a free 
+parameter in an additional model (median posterior $\delta$ = 
+`r papaja::printnum(median(error))`, 95% CI
+[`r papaja::printnum(quantile(error, 0.025))`
+`r papaja::printnum(quantile(error, 0.975))`]; @suppfig-model5). The random 
+choice strategy is consistently coded with a conditional probability of ½ across
+all decisions.
 
@@
 We estimated the posterior distributions of these models using Hamiltonian Monte 
-Carlo as implemented in Stan version 2.26.1 [@Stan2020]. We ran each model for 
+Carlo as implemented in Stan version 2.32.2 [@Stan2020]. We ran each model for 
 2000 samples, with 1000 warmup samples. R-hat values and effective sample sizes 
 suggested that all models converged normally. Trace plots are reported in 
-Supplementary Figure \@ref(fig:plotTrace).
+@suppfig-trace.
 
@@
 frequency of strategies. The model was successfully able to recover the known 
-frequency of strategies from the simulated data (Supplementary Figure 
-\@ref(fig:plotSim)).
+frequency of strategies from the simulated data (@suppfig-sim).
 
@@
 All data and code are accessible on GitHub:
-https://github.com/ScottClaessens/punishStrategies. All analyses were conducted 
-in R version 4.2.1 [@RCoreTeam]. Visualisations were created with the *ggplot2*
-[@Wickham2016] and *cowplot* [@Wilke2020] R packages. We used the *targets*
-[@Landau2021] R package to create a reproducible data analysis pipeline and the 
-*papaja* [@Aust2022] R package to reproducibly generate the manuscript.
+<https://github.com/ScottClaessens/punishStrategies>. All analyses were 
+conducted in R version 4.4.2 [@RCoreTeam]. Visualisations were created with the
+*ggplot2* [@Wickham2016] and *cowplot* [@Wilke2020] R packages. We used the
+*targets* [@Landau2021] R package and *quarto* [@Allaire2024] to reproducibly 
+generate the manuscript.
 
 \newpage
-\nolinenumbers
 
@@
 to Q.D.A and N.R. (ref: ILFUOA2002).
-
-# Author Contributions
-
-All authors (S.C., Q.D.A., N.R.) conceptualised the research, designed the 
-study, and developed the surveys. N.R. conducted data collection on Prolific.
-S.C. conducted all analyses and visualisation of the data. All authors wrote the 
-manuscript.
 
@@
 All data used in this study are publicly available on GitHub: 
-https://github.com/ScottClaessens/punishStrategies
+<https://github.com/ScottClaessens/punishStrategies>
 
@@
 All code to reproduce the analyses in this study are publicly available on 
-GitHub: https://github.com/ScottClaessens/punishStrategies
+GitHub: <https://github.com/ScottClaessens/punishStrategies>
 
@@
 
-\begingroup
-\setlength\parindent{0pt}
-<div id="refs" custom-style="Bibliography"></div>
-\endgroup
+<!-- References will auto-populate in the refs div below -->
+
+::: {#refs}
+:::
 
 \newpage
-\vspace*{60mm}
 
-```{r child = "appendix.Rmd"}
-```
+{{< include appendix.qmd >}}

### 2025-08-01 R/functions_modelResults.R

@@ -211,6 +211,8 @@ plotAllSliders <- function(postList, file) {
       }
     }
   }
+  # save memory
+  rm(post, postList)
   # plot
   out <-
     slopes %>%
@@ -291,6 +293,8 @@ plotAllPers <- function(postList, file) {
       }
     }
   }
+  # save memory
+  rm(post, postList)
   # plot
   out <-
     slopes %>%
@@ -361,6 +365,8 @@ plotAllPolRel <- function(postList, file) {
       }
     }
   }
+  # save memory
+  rm(post, postList)
   # plot
   out <-
     slopes %>%
@@ -426,6 +432,8 @@ plotAllDems <- function(postList, file) {
       }
     }
   }
+  # save memory
+  rm(post, postList)
   # plot
   out <-
     slopes %>%
@@ -463,6 +471,7 @@ plotTraceMCMC <- function(m1.2) {
   # plot
   out <-
     tidy_draws(m1.2, alpha) %>%
+    filter(.draw %% 10 == 0) %>% # get every tenth row
     pivot_longer(cols = starts_with("alpha"),
                  names_to = "parameter",
                  values_to = "value") %>%
@@ -477,6 +486,7 @@ plotTraceMCMC <- function(m1.2) {
           legend.position = "none")
   # save plot
   ggsave(out, filename = "figures/modelResults/trace.pdf", width = 7, height = 6)
+  rm(m1.2)
   return(out)
 }
 

### 2025-08-01 R/functions_summariseData.R

@@ -447,8 +447,7 @@ makeStrategyCountTable <- function(d) {
     mutate(Prop = format(round(Prop, digits = 3), nsmall = 3)) %>%
     pivot_wider(names_from = Country, values_from = c(N, Prop)) %>%
     dplyr::select(c(1,2,4,3,5))
-  colnames(out)[c(2,4)] <- "N"
-  colnames(out)[c(3,5)] <- "Prop"
+  colnames(out)[2:5] <- letters[1:4]
   return(out)
 }
 
@@ -498,23 +497,23 @@ makeStrategyCountTable <- function(d) {
 makePatternsTable <- function(d) {
   # vector of explanations for common strategies
   explanations <- c(
-    "000000000000" = "\\textit{Never punish strategy (exact)}",
-    "000000001000" = "\\textit{Avoid DI strategy (exact)}",
-    "000000001010" = "\\textit{Egalitarian strategy (exact)}",
+    "000000000000" = "Never punish strategy (exact)",
+    "000000001000" = "Avoid DI strategy (exact)",
+    "000000001010" = "Egalitarian strategy (exact)",
     "000000000010" = "Punish when steal in Game F",
     "001000001000" = "Punish when steal in Games B and E",
     "101000001010" = "Punish when steal in Games A, B, E, and F",
     "100000000000" = "Punish when steal in Game A",
     "000000100000" = "Punish when steal in Game D",
     "001000001010" = "Punish when steal in Games B, E, and F",
-    "101000101000" = "\\textit{Deterrent strategy (exact)}",
+    "101000101000" = "Deterrent strategy (exact)",
     "101010101010" = "Punish when steal in all games",
-    "101000101010" = "\\textit{Norm-enforcing strategy (exact)}",
+    "101000101010" = "Norm-enforcing strategy (exact)",
     "001000000000" = "Punish when steal in Game B",
     "001010101000" = "Punish when steal in Games B, C, D, and E",
     "100000001000" = "Punish when steal in Games A and E",
     "101000001000" = "Punish when steal in Games A, B, and E",
-    "101010101000" = "\\textit{Retributive strategy (exact)}",
+    "101010101000" = "Retributive strategy (exact)",
     "111111111111" = "Always punish",
     "000000101000" = "Punish when steal in Games D and E",
[…]

### 2025-08-01 R/packages.R

@@ -1,5 +1,5 @@
 options(tidyverse.quiet = TRUE)
-library(crew.cluster)
+library(crew)
 library(targets)
 library(tarchetypes)
 library(tidyverse)
@@ -9,15 +9,5 @@ tar_option_set(
                "papaja", "phangorn", "rethinking", "rstan",
                "tanggle", "tidybayes", "tidytext", "tidyverse",
                "wordcloud")#,
-  #controller = crew_controller_slurm(
-  #  name = "punishStrategies",
-  #  workers = 1,
-  #  script_lines = "module load R",
-  #  slurm_time_minutes = 1440,
-  #  slurm_memory_gigabytes_per_cpu = 60,
-  #  slurm_cpus_per_task = 4,
-  #  garbage_collection = TRUE
-  #),
-  #memory = "transient",
-  #garbage_collection = TRUE
+  #controller = crew_controller_local(workers = 2)
 )

### 2025-08-01 _targets.R

@@ -279,13 +279,9 @@ list(
   tar_target(plotAllDems2, plotAllDems(list(post2.2_Age, post2.2_SES, post4.2_Gender, post4.2_Student),
                                        file = "figures/modelResults/withExclusions/allDems.pdf")),
   
-  #### Render report ####
-  
-  tar_render(report, "report.Rmd"),
-  
   #### Render manuscript ####
   
-  tar_render(manuscript, "manuscript.Rmd"),
+  tar_quarto(manuscript, "quarto/manuscript.qmd", quiet = F),
   
   #### Session info ####
   
